### 2019-06-13 submission/manuscript.Rmd
Added inital code for manuscript generation

---
title: "**NAME OF THIS STUDY**"
bibliography: references.bib
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: header.tex
csl: mbio.csl #Get themes at https://github.com/citation-style-language/styles
fontsize: 11pt
geometry: margin=1.0in
---

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)

inline_hook <- function(x){
	print(x)

	if(is.list(x)){
		x <- unlist(x)
	}

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)
```

\vspace{35mm}

Running title: Selective extraction of DNA and proteins from epiphytic macrophyte mircobial communities

\vspace{35mm}

Your Name Here^1, Joeseph P. Schmo^2, Sally J. Rivers^1, Patrick D. Schloss^1$\dagger$^

\vspace{40mm}

$\dagger$ To whom correspondence should be addressed: marino.korlevic@irb.hr

1\. Ruđer Bošković Institute, Center for Marine Research, G. Paliaga 5, Rovinj, Croatia

2\. University of Vienna

\newpage
\linenumbers

## Abstract

\newpage

## Introduction

## Results and Discussion

## Conclusions

## Materials and Methods

\newpage

\newpage

## References

### 2019-06-13 submission/manuscript.tex
Added inital code for manuscript generation

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={NAME OF THIS STUDY},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{\textbf{NAME OF THIS STUDY}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}

\begin{document}
\maketitle

\vspace{35mm}

Running title: Selective extraction of DNA and proteins from epiphytic
macrophyte mircobial communities

\vspace{35mm}

Your Name Here\^{}1, Joeseph P. Schmo\^{}2, Sally J. Rivers\^{}1,
Patrick D. Schloss\textsuperscript{1\(\dagger\)}

\vspace{40mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}

1. Ruđer Bošković Institute, Center for Marine Research, G. Paliaga 5,
Rovinj, Croatia

2. University of Vienna

\newpage
\linenumbers

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

\newpage

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

\hypertarget{results-and-discussion}{%
\subsection{Results and Discussion}\label{results-and-discussion}}

\hypertarget{conclusions}{%
\subsection{Conclusions}\label{conclusions}}

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\newpage

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\end{document}

### 2019-06-14 code/plot_calculators.R
Added code to plot calculators

@@ -20,7 +20,7 @@ meta_alpha <- inner_join(metadata, alpha, by=c("ID"="group"))
 
 plot <- meta_alpha %>%
     filter(!is.na(depth)) %>%
-    group_by(depth) %>%
+    group_by(label) %>%
     summarize(mean_sobs=mean(sobs),
               mean_chao=mean(chao),
               mean_ace=mean(ace),

### 2019-09-25 code/format_taxonomy.R
Added code to build the reference database

@@ -42,19 +42,21 @@ map.in$taxout <- gsub(" ", "_", map.in$taxout)
 tax.in <- read.table("data/references/silva.nr_v132.full", header=F, stringsAsFactors=F, sep="\t")
 colnames(tax.in) <- c("taxid", "taxlabel")
 
+# Problem corrections observed in the SILVA ARB file (silva.nr_v132.full)
 # Following line corrects the Bacteria;Bacteroidetes;Bacteroidia;Flavobacteriales;Flavobacteriaceae;Polaribacter;Polaribacter; problem
-tax.in$taxlabel <- gsub("Polaribacter;Polaribacter;", "Polaribacter;", tax.in$taxlabel)
-# In the SILVA ARB files the phylum RsaHF231 is labeled as RsaHf231. The same phylum in the SILVA taxa mapping file
-# is labeled as RsaHF231. To avoid confusion we are replacing the RsaHf231 name with RsaHF231 in the data derived
-# from the SILVA ARB file (silva.nr_v132.full)
+tax.in$taxlabel <- gsub("Polaribacter;Polaribacter", "Polaribacter", tax.in$taxlabel)
+# Following line corrects the Bacteria;RsaHf231; problem
 tax.in$taxlabel <- gsub("RsaHf231;", "RsaHF231;", tax.in$taxlabel)
+# Following line corrects the Bacteria;GBS-1;; problem
+tax.in$taxlabel <- gsub("Bacteria;GBS-1;;", "Bacteria;GBS-1;", tax.in$taxlabel)
+# Following line corrects the Eukaryota;Opisthokonta;Nucletmycea;Fungi;Dikarya;Basidiomycota;Agaricomycotina;Tremellomycetes;Tremellales;Incertae sedis;Sterigmatosporidium; problem
+tax.in$taxlabel <- gsub(";Incertae sedis;", ";Incertae Sedis;", tax.in$taxlabel)
 tax.in$taxlabel <- gsub(";[[:space:]]+$", ";", tax.in$taxlabel)
 
 tax.in$id <- 1:nrow(tax.in)
 
 tax.write <- merge(tax.in, map.in, all.x=T, sort=F)
 tax.write <- tax.write[order(tax.write$id), ]
-
 
 # We want to see whether everything has 6 taxonomic level (kingdom to genus)
 getDepth <- function(taxonString) {

### 2019-09-26 code/plot_pcoa.R
Modified code to generate PCoA

@@ -10,7 +10,7 @@
 #################################################################################################################
 
 library(tidyverse)
-pcoa <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.braycurtis.0.03.lt.ave.pcoa.axes")
+pcoa <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.braycurtis.0.03.lt.ave.pcoa.axes", col_types=cols(group = col_character()))
 metadata <- read_tsv("data/raw/metadata.csv")
 metadata_pcoa <- inner_join(metadata, pcoa, by=c('ID'='group'))
 

### 2019-10-11 submission/manuscript.Rmd
Further optimization of the process

@@
       in_header: header.tex
-csl: mbio.csl #Get themes at https://github.com/citation-style-language/styles
+csl: citation_style.csl #Get themes at https://github.com/citation-style-language/styles
 fontsize: 11pt

### 2019-10-13 submission/manuscript.tex
Optimized pipeline for reduced no. of samples.
Added ATCC reference database
Optimization of pdf generation

@@
 % Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
+\providecommand{\subtitle}[1]{
   \posttitle{
@@
 
-\usepackage{helvet} % Helvetica font
-\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
+\usepackage{times} % Times New Roman font
 \usepackage[T1]{fontenc}
@@
 \newpage
+
 \linenumbers
 
-\hypertarget{abstract}{%
-\subsection{Abstract}\label{abstract}}
+\subsection{Abstract}\label{abstract}
 
@@
 
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+\subsection{Introduction}\label{introduction}
 
-\hypertarget{results-and-discussion}{%
-\subsection{Results and Discussion}\label{results-and-discussion}}
+\subsection{Results and Discussion}\label{results-and-discussion}
 
-\hypertarget{conclusions}{%
-\subsection{Conclusions}\label{conclusions}}
+\subsection{Conclusions}\label{conclusions}
 
-\hypertarget{materials-and-methods}{%
-\subsection{Materials and Methods}\label{materials-and-methods}}
+\subsection{Materials and Methods}\label{materials-and-methods}
 
@@
 
-\hypertarget{references}{%
-\subsection{References}\label{references}}
+\subsection{References}\label{references}
 

### 2019-10-14 submission/manuscript.Rmd
Optimization of pdf generation
Rmarkdown optimization

--- 2019-06-13 submission/manuscript.Rmd
@@
 ---
-title: "**NAME OF THIS STUDY**"
-bibliography: references.bib
+title: "**Selective DNA and Protein Extraction from Marine Macrophyte Surfaces**"
 output:
@@
       in_header: header.tex
-csl: mbio.csl #Get themes at https://github.com/citation-style-language/styles
-fontsize: 11pt
+fontsize: 12pt
 geometry: margin=1.0in
+csl: citation_style.csl #Get themes at https://github.com/citation-style-language/styles
+bibliography: references.bib
 ---
@@
 
-\vspace{35mm}
+\vspace{80mm}
 
-Running title: Selective extraction of DNA and proteins from epiphytic macrophyte mircobial communities
-
-\vspace{35mm}
-
-Your Name Here^1, Joeseph P. Schmo^2, Sally J. Rivers^1, Patrick D. Schloss^1$\dagger$^
+^1$\dagger$^
 
@@
 
-2\. University of Vienna
-
+2\. University of Vienna, Department of Limnology and Bio-Oceanography, Althanstraße 14, Vienna, Austria
 \newpage
 \linenumbers
+\setlength\parindent{24pt}
 
 ## Abstract
+jhfajklsgčklajsdbvlafgčladfhg
+
+lčkhfglashgčlashgčlak
 
@@
 ## Introduction
+
+##Materials and Methods
 

### 2019-10-14 submission/manuscript.tex
pdf generation optimization
Rmarkdown optimization

@@
-\documentclass[11pt,]{article}
+\documentclass[12pt,]{article}
 \usepackage{lmodern}
@@
 \hypersetup{unicode=true,
-            pdftitle={NAME OF THIS STUDY},
+            pdftitle={Selective DNA and Protein Extraction from Marine Macrophyte Surfaces},
             pdfborder={0 0 0},
@@
 
-  \title{\textbf{NAME OF THIS STUDY}}
+  \title{\textbf{Selective DNA and Protein Extraction from Marine Macrophyte
+Surfaces}}
     \pretitle{\vspace{\droptitle}\centering\huge}
@@
 
+\usepackage{indentfirst}
+
+\usepackage[labelsep=period, labelfont=bf]{caption}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\captionsetup{justification=raggedright,singlelinecheck=false}
+
+\usepackage{pdflscape}
+\newcommand{\blandscape}{\begin{landscape}}
+\newcommand{\elandscape}{\end{landscape}}
+
 \begin{document}
@@
 
-\vspace{35mm}
+\vspace{80mm}
 
-Running title: Selective extraction of DNA and proteins from epiphytic
-macrophyte mircobial communities
-
-\vspace{35mm}
-
-Your Name Here\^{}1, Joeseph P. Schmo\^{}2, Sally J. Rivers\^{}1,
-Patrick D. Schloss\textsuperscript{1\(\dagger\)}
+\textsuperscript{1\(\dagger\)}
 
@@
 
-2. University of Vienna
-
-\newpage
-
+2. University of Vienna, Department of Limnology and Bio-Oceanography,
+Althanstraße 14, Vienna, Austria \newpage
 \linenumbers
+\setlength\parindent{24pt}
 
 \subsection{Abstract}\label{abstract}
+
+jhfajklsgčklajsdbvlafgčladfhg
+
+lčkhfglashgčlashgčlak
 
@@
 
+\subsection{Materials and Methods}\label{materials-and-methods}
+
 \subsection{Results and Discussion}\label{results-and-discussion}
@@
 
-\subsection{Materials and Methods}\label{materials-and-methods}
+\subsection{Materials and Methods}\label{materials-and-methods-1}
 

### 2019-10-21 submission/manuscript.tex
Chloroplast seq. selection optimization

@@
 2. University of Vienna, Department of Limnology and Bio-Oceanography,
-Althanstraße 14, Vienna, Austria \newpage
-\linenumbers
+Althanstraße 14, Vienna, Austria \newpage \linenumbers
 \setlength\parindent{24pt}
 
-\subsection{Abstract}\label{abstract}
+\hypertarget{abstract}{%
+\subsection{Abstract}\label{abstract}}
 
@@
 
-\subsection{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\subsection{Introduction}\label{introduction}}
 
-\subsection{Materials and Methods}\label{materials-and-methods}
+\#\#Materials and Methods
 
-\subsection{Results and Discussion}\label{results-and-discussion}
+\hypertarget{results-and-discussion}{%
+\subsection{Results and Discussion}\label{results-and-discussion}}
 
-\subsection{Conclusions}\label{conclusions}
+\hypertarget{conclusions}{%
+\subsection{Conclusions}\label{conclusions}}
 
-\subsection{Materials and Methods}\label{materials-and-methods-1}
+\hypertarget{materials-and-methods}{%
+\subsection{Materials and Methods}\label{materials-and-methods}}
 
@@
 
-\subsection{References}\label{references}
+\hypertarget{references}{%
+\subsection{References}\label{references}}
 

### 2019-11-21 submission/manuscript.Rmd
Introduction and part of M&M added

@@
 ---
-title: "**Selective DNA and Protein Extraction from Marine Macrophyte Surfaces**"
+title: "**Selective DNA and Protein Isolation from Marine Macrophyte Surfaces**"
 output:
@@
 ## Abstract
-jhfajklsgčklajsdbvlafgčladfhg
-
-lčkhfglashgčlashgčlak
 
 \newpage
+## Introduction
+Surfaces of marine macrophytes are inhabited by a diverse microbial community whose structure and function are poorly understood [@Egan2013].
+As less than 1 % of all prokaryotic species are culturable, to study these organisms, molecular methods such as 16S rRNA sequencing, metagenomics and metaproteomics are indispensable [@Amann1995; @Su2012].
+Applying these techniques requires an inital isolation step, with the purpose of obataining high quality DNA and proteins.
 
-## Introduction
+Biological material (i.e.
+proteins and DNA) from pelagic microbial communities is usually isolated by collecting cells onto filters and subsequently isolating the aimed compound [@Gilbert2009].
+If a microbial size fraction is aimed sequential filtration is applied [@Massana1997; @Andersson2010].
+In contrast, obtaining biological materials from microorganism inhabiting surfaces requires either a cell detachment procedure prior to isolation or the host material is coextracted together with the targeted material.
+Methods for separating microbial cells form the host include shaking of host tissue [@Gross2003; @Noges2010], scraping of macrophyte surfaces [@Uku2007] or the application of ultrasonication [@Weidner1996; @Cai2014].
+It was shown that shaking alone is not sufficient to remove microbial cells, at least from plant root surfaces [@Richter-Heitmann2016].
+Manual separation methods, such as scraping and brushing, are time consuming and subjective, as the detachment efficency depends on host tissue and the person performing the procedure [@Cai2014].
+Ultrasonication was proposed as an alternative method as it is providing better results in terms of detachemnt efficency [@Cai2014; @Richter-Heitmann2016].
+The downside of this procedure is that complete cell removal was still not obtained and tissue disruption was observed especially after the application of probe ultrasonication [@Richter-Heitmann2016].
+An alterantive to these cell detachment procedures is the isolation of targeted epiphytic compounds togehter with host materials [@Staufenberger2008; @Jiang2015].
+This procedure can lead to problems in the following processing steps such as mitochondrial and chloroplast 16S rRNA sequence contaminations from the host [@Longford2007; @Staufenberger2008].
+In additon, when performing metagenomics and metaproteomics host material can cause biased results towards more abundand host DNA and proteins.
 
-##Materials and Methods
+An alterantive to these procedures is a direct isolation of the targeted material by incubating macrophyte tissues in an extraction buffer.
+After the incubation is done the undistrupted host tissue is removed and the isolation procedure continues ommiting host material contaminations.
+To our knowledge the only procedure describing a direct and selective epiphytic DNA isolation from the surfaces of marine macrophytes was provided by @Burke2009.
+In contrast to previously described methods this protocol enables an almost complete removal of the surface community and was used for 16S rRNA gene clone libraries construction [@Burke2011b] and metagenomes sequencing [@Burke2011a].
+This method, thought providing a selctive isolation procedure, is using in the extraction buffer a rapid multienzyme cleaner (3M) that is not worldwide availabe and whose composition is not know [@Burke2009].
+Also to our knowledge, no selective isolation protocol for proteins from epihytic communities inhabiting marine macrophytes was established.
+
+In the present study, we adapted a protocol widely used for DNA isolation from filters [@Massana1997] and a protocol used for protein isolation from sediments [@Chourey2010] for selective extractions of DNA and proteins from epiphytic communities inhabiting the surfaces of two marine macrophytes: the seagrass *Cymodocea nodosa* and the alga *Caulerpa cylindracea*.
+In addition, we tested the removal efficiency of the protocol and the suitability of obtained DNA and proteins for 16S rRNA sequencing and metaproteomics.
+
+\newpage
+## Materials and Methods
+### Sampling
+Leaves of *Cymodocea nodosa* were sampled in a *Cymodocea nodosa* meadow in the Bay of Saline (45°7´5˝N; 13°37´20˝E) and in a *Cymodocea nodosa* meadow invaded by *Caulerpa cylindracea* in the proximity of the village of Funtana (45°10´39˝N 13°35´42˝E).
+Thalli of *Caulerpa cylindracea* were sampled in the same *Cymodocea nodosa* invaded meadow in Funtana and on a locality of only *Caulerpa cylindracea* located close to the invaded meadow.
+Seagrasses and algae were collected the same day in two contrasting seasons, on 4 December 2017 and 18 June 2018.
+During spring 2018 the *Cymodocea nodosa* meadow in the Bay of Saline decayed to the extent that no leves could be retrieved (Najdek *et al.*, unpublished data).
+Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater.
+Upon arrival to the laboratory, *Cymodocea nodosa* leaves were cut into sections of 1 – 2 cm, while *Caulerpa cylindracea* thalli were cut into 5 – 8 cm long sections.
+Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.
+
+### DNA isolation
+The DNA was isolated according to the protocol for isolation from filters described in @Massana1997.
+This protocol was modified and adapted for DNA isolation from microbial communites from macrophytes surfaces as described below.
+1 g wet weight of leaves and 2 g wet-weight of thalli were placed into 5 ml of lysis buffer (40 mM EDTA, 50 mM Tris-HCl, 0.75 M sucrose; pH 8.3).
+Lysozyme was added (final concentration 1 mg ml^-1^) and the mixture was incubated at 37 °C for 30 minutes.
+Subsequently, proteinase K (final concentration 0.5 mg ml^-1^) and SDS (finl concentration 1 %) were added and the samples were incubated at 55 °C for 2 hours.
+Following the incubation, tubes were vortexed for 10 minutes and the mixture containign lyzed cells was separated from host leaves or thalli by transferrring the solution into a clean tube.
+The lysate was extracted twice with a mixture of phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with chloroform-isoamyl alcohol (24:1).
+After each organic solvent mixture addition tubes were slightly vortexed and centrifugated at 4,500 x g for 10 minutes.
+Following each centrifugation aqueous phases were retrieved.
+After the final extraction 1/10 of 3 M sodium acetate (ph 5.2) was added.
+DNA was precipitated by adding 1 volumne of isopropanol and by incubating the mixtures at -20 °C overnight.
+
+### Protein isolation
 

### 2019-11-21 submission/manuscript.tex
Introduction and part of M&M added

--- 2019-10-14 submission/manuscript.tex
@@
 \hypersetup{unicode=true,
-            pdftitle={Selective DNA and Protein Extraction from Marine Macrophyte Surfaces},
+            pdftitle={Selective DNA and Protein Isolation from Marine Macrophyte Surfaces},
             pdfborder={0 0 0},
@@
 
-  \title{\textbf{Selective DNA and Protein Extraction from Marine Macrophyte
+  \title{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
 Surfaces}}
@@
 
-jhfajklsgčklajsdbvlafgčladfhg
-
-lčkhfglashgčlashgčlak
-
 \newpage
@@
 
+Surfaces of marine macrophytes are inhabited by a diverse microbial
+community whose structure and function are poorly understood (Egan
+\emph{et al.}, 2013). As less than 1 \% of all prokaryotic species are
+culturable, to study these organisms, molecular methods such as 16S rRNA
+sequencing, metagenomics and metaproteomics are indispensable (Amann
+\emph{et al.}, 1995; Su \emph{et al.}, 2012). Applying these techniques
+requires an inital isolation step, with the purpose of obataining high
+quality DNA and proteins.
+
+Biological material (i.e.~proteins and DNA) from pelagic microbial
+communities is usually isolated by collecting cells onto filters and
+subsequently isolating the aimed compound (Gilbert \emph{et al.}, 2009).
+If a microbial size fraction is aimed sequential filtration is applied
+(Massana \emph{et al.}, 1997; Andersson \emph{et al.}, 2010). In
+contrast, obtaining biological materials from microorganism inhabiting
+surfaces requires either a cell detachment procedure prior to isolation
+or the host material is coextracted together with the targeted material.
+Methods for separating microbial cells form the host include shaking of
+host tissue (Gross \emph{et al.}, 2003; Nõges \emph{et al.}, 2010),
+scraping of macrophyte surfaces (Uku \emph{et al.}, 2007) or the
+application of ultrasonication (Weidner \emph{et al.}, 1996; Cai
+\emph{et al.}, 2014). It was shown that shaking alone is not sufficient
+to remove microbial cells, at least from plant root surfaces
+(Richter-Heitmann \emph{et al.}, 2016). Manual separation methods, such
+as scraping and brushing, are time consuming and subjective, as the
+detachment efficency depends on host tissue and the person performing
+the procedure (Cai \emph{et al.}, 2014). Ultrasonication was proposed as
+an alternative method as it is providing better results in terms of
+detachemnt efficency (Cai \emph{et al.}, 2014; Richter-Heitmann \emph{et
+al.}, 2016). The downside of this procedure is that complete cell
+removal was still not obtained and tissue disruption was observed
+especially after the application of probe ultrasonication
+(Richter-Heitmann \emph{et al.}, 2016). An alterantive to these cell
+detachment procedures is the isolation of targeted epiphytic compounds
+togehter with host materials (Staufenberger \emph{et al.}, 2008; Jiang
+\emph{et al.}, 2015). This procedure can lead to problems in the
+following processing steps such as mitochondrial and chloroplast 16S
+rRNA sequence contaminations from the host (Longford \emph{et al.},
+2007; Staufenberger \emph{et al.}, 2008). In additon, when performing
+metagenomics and metaproteomics host material can cause biased results
+towards more abundand host DNA and proteins.
+
+An alterantive to these procedures is a direct isolation of the targeted
+material by incubating macrophyte tissues in an extraction buffer. After
+the incubation is done the undistrupted host tissue is removed and the
+isolation procedure continues ommiting host material contaminations. To
+our knowledge the only procedure describing a direct and selective
+epiphytic DNA isolation from the surfaces of marine macrophytes was
+provided by Burke \emph{et al.} (2009). In contrast to previously
+described methods this protocol enables an almost complete removal of
+the surface community and was used for 16S rRNA gene clone libraries
+construction (Burke \emph{et al.}, 2011b) and metagenomes sequencing
+(Burke \emph{et al.}, 2011a). This method, thought providing a selctive
+isolation procedure, is using in the extraction buffer a rapid
+multienzyme cleaner (3M) that is not worldwide availabe and whose
+composition is not know (Burke \emph{et al.}, 2009). Also to our
+knowledge, no selective isolation protocol for proteins from epihytic
+communities inhabiting marine macrophytes was established.
+
+In the present study, we adapted a protocol widely used for DNA
+isolation from filters (Massana \emph{et al.}, 1997) and a protocol used
+for protein isolation from sediments (Chourey \emph{et al.}, 2010) for
+selective extractions of DNA and proteins from epiphytic communities
+inhabiting the surfaces of two marine macrophytes: the seagrass
+\emph{Cymodocea nodosa} and the alga \emph{Caulerpa cylindracea}. In
+addition, we tested the removal efficiency of the protocol and the
+suitability of obtained DNA and proteins for 16S rRNA sequencing and
+metaproteomics.
+
+\newpage
+
 \subsection{Materials and Methods}\label{materials-and-methods}
+
+\subsubsection{Sampling}\label{sampling}
+
+Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
+nodosa} meadow in the Bay of Saline (45°7´5˝N; 13°37´20˝E) and in a
+\emph{Cymodocea nodosa} meadow invaded by \emph{Caulerpa cylindracea} in
+the proximity of the village of Funtana (45°10´39˝N 13°35´42˝E). Thalli
+of \emph{Caulerpa cylindracea} were sampled in the same \emph{Cymodocea
+nodosa} invaded meadow in Funtana and on a locality of only
+\emph{Caulerpa cylindracea} located close to the invaded meadow.
+Seagrasses and algae were collected the same day in two contrasting
+seasons, on 4 December 2017 and 18 June 2018. During spring 2018 the
+\emph{Cymodocea nodosa} meadow in the Bay of Saline decayed to the
+extent that no leves could be retrieved (Najdek \emph{et al.},
+unpublished data). Leaves and thalli were collected by diving and
+transported to the laboratory in containers placed on ice and filled
+with site seawater. Upon arrival to the laboratory, \emph{Cymodocea
+nodosa} leaves were cut into sections of 1 -- 2 cm, while \emph{Caulerpa
+cylindracea} thalli were cut into 5 -- 8 cm long sections. Leaves and
+thalli were washed three times with sterile artificial seawater (ASW) to
+remove loosely attached microbial cells.
+
+\subsubsection{DNA isolation}\label{dna-isolation}
+
+The DNA was isolated according to the protocol for isolation from
+filters described in Massana \emph{et al.} (1997). This protocol was
+modified and adapted for DNA isolation from microbial communites from
+macrophytes surfaces as described below. 1 g wet weight of leaves and 2
+g wet-weight of thalli were placed into 5 ml of lysis buffer (40 mM
+EDTA, 50 mM Tris-HCl, 0.75 M sucrose; pH 8.3). Lysozyme was added (final
+concentration 1 mg ml\textsuperscript{-1}) and the mixture was incubated
+at 37 °C for 30 minutes. Subsequently, proteinase K (final concentration
+0.5 mg ml\textsuperscript{-1}) and SDS (finl concentration 1 \%) were
+added and the samples were incubated at 55 °C for 2 hours. Following the
+incubation, tubes were vortexed for 10 minutes and the mixture
+containign lyzed cells was separated from host leaves or thalli by
+transferrring the solution into a clean tube. The lysate was extracted
+twice with a mixture of phenol:chloroform:isoamyl alcohol (25:24:1; pH
+8) and once with chloroform-isoamyl alcohol (24:1). After each organic
+solvent mixture addition tubes were slightly vortexed and centrifugated
+at 4,500 x g for 10 minutes. Following each centrifugation aqueous
+phases were retrieved. After the final extraction 1/10 of 3 M sodium
+acetate (ph 5.2) was added. DNA was precipitated by adding 1 volumne of
+isopropanol and incubating the mixtures at -20 °C overnight.
+
+\subsubsection{Protein isolation}\label{protein-isolation}
 
@@
 
-\subsection{References}\label{references}
+\subsection*{References}\label{references}
+\addcontentsline{toc}{subsection}{References}
+
+\hypertarget{refs}{}
+\hypertarget{ref-Amann1995}{}
+Amann, R.I., Ludwig, W., and Schleifer, K.H. (1995) Phylogenetic
+identification and in situ detection of individual microbial cells
+without cultivation. \emph{Microbiological reviews} \textbf{59}:
+143--169.
+
+\hypertarget{ref-Andersson2010}{}
+Andersson, A.F., Riemann, L., and Bertilsson, S. (2010) Pyrosequencing
+reveals contrasting seasonal dynamics of taxa within Baltic Sea
+bacterioplankton communities. \emph{The ISME journal} \textbf{4}:
+171--181.
+
+\hypertarget{ref-Burke2009}{}
+Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective extraction of
+bacterial DNA from the surfaces of macroalgae. \emph{Applied and
+environmental microbiology} \textbf{75}: 252--256.
+
+\hypertarget{ref-Burke2011a}{}
+Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
+(2011a) Bacterial community assembly based on functional genes rather
+than species. \emph{Proceedings of the National Academy of Sciences of
+the United States of America} \textbf{108}: 14288--14293.
+
+\hypertarget{ref-Burke2011b}{}
+Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
+(2011b) Composition, uniqueness and variability of the epiphytic
+bacterial community of the green alga Ulva australis. \emph{The ISME
+journal} \textbf{5}: 590--600.
+
+\hypertarget{ref-Cai2014}{}
+Cai, X., Gao, G., Yang, J., Tang, X., Dai, J., Chen, D., and Song, Y.
+(2014) An ultrasonic method for separation of epiphytic microbes from
+freshwater submerged macrophytes. \emph{Journal of basic microbiology}
+\textbf{54}: 758--761.
+
+\hypertarget{ref-Chourey2010}{}
+Chourey, K., Jansson, J., VerBerkmoes, N., Shah, M., Chavarria, K.L.,
+Tom, L.M., et al. (2010) Direct Cellular Lysis/Protein Extraction
+Protocol for Soil Metaproteomics. \emph{Journal of Proteome Research}
+\textbf{9}: 6615--6622.
+
+\hypertarget{ref-Egan2013}{}
+Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
+Thomas, T. (2013) The seaweed holobiont: understanding seaweed-bacteria
+interactions. \emph{FEMS microbiology reviews} \textbf{37}: 462--476.
+
+\hypertarget{ref-Gilbert2009}{}
+Gilbert, J.A., Field, D., Swift, P., Newbold, L., Oliver, A., Smyth, T.,
+et al. (2009) The seasonal structure of microbial communities in the
+Western English Channel. \emph{Environmental Microbiology} \textbf{11}:
+3132--3139.
+
+\hypertarget{ref-Gross2003}{}
+Gross, E.M., Feldbaum, C., and Graf, A. (2003) Epiphyte biomass and
+elemental composition on submersed macrophytes in shallow eutrophic
+lakes. \emph{Hydrobiologia} \textbf{506-509}: 559--565.
+
+\hypertarget{ref-Jiang2015}{}
+Jiang, Y.-F., Ling, J., Dong, J.-D., Chen, B., Zhang, Y.-Y., Zhang,
+Y.-Z., and Wang, Y.-S. (2015) Illumina-based analysis the microbial
+diversity associated with Thalassia hemprichii in Xincun Bay, South
+China Sea. \emph{Ecotoxicology (London, England)} \textbf{24}: 1548--56.
+
+\hypertarget{ref-Longford2007}{}
+Longford, S., Tujula, N., Crocetti, G., Holmes, A., Holmström, C.,
+Kjelleberg, S., et al. (2007) Comparisons of diversity of bacterial
+communities associated with three sessile marine eukaryotes.
+\emph{Aquatic Microbial Ecology} \textbf{48}: 217--229.
+
+\hypertarget{ref-Massana1997}{}
+Massana, R., Murray, A.E., Preston, C.M., and DeLong, E.F. (1997)
+Vertical distribution and phylogenetic characterization of marine
+planktonic Archaea in the Santa Barbara Channel. \emph{Applied and
+environmental microbiology} \textbf{63}: 50--56.
+
+\hypertarget{ref-Noges2010}{}
+Nõges, T., Luup, H., and Feldmann, T. (2010) Primary production of
+aquatic macrophytes and their epiphytes in two shallow lakes (Peipsi and
+Võrtsjärv) in Estonia. \emph{Aquatic Ecology} \textbf{44}: 83--92.
+
+\hypertarget{ref-Richter-Heitmann2016}{}
+Richter-Heitmann, T., Eickhorst, T., Knauth, S., Friedrich, M.W., and
+Schmidt, H. (2016) Evaluation of Strategies to Separate Root-Associated
+Microbial Communities: A Crucial Choice in Rhizobiome Research.
+\emph{Frontiers in Microbiology} \textbf{7}: 773.
+
+\hypertarget{ref-Staufenberger2008}{}
+Staufenberger, T., Thiel, V., Wiese, J., and Imhoff, J.F. (2008)
+Phylogenetic analysis of bacteria associated with Laminaria saccharina.
+\emph{FEMS Microbiology Ecology} \textbf{64}: 65--77.
+
+\hypertarget{ref-Su2012}{}
+Su, C., Lei, L., Duan, Y., Zhang, K.-Q., and Yang, J. (2012)
+Culture-independent methods for studying environmental microorganisms:
+methods, application, and perspective. \emph{Applied microbiology and
+biotechnology} \textbf{93}: 993--1003.
+
+\hypertarget{ref-Uku2007}{}
+Uku, J., Björk, M., Bergman, B., and Díez, B. (2007) Characterization
+and comparison of prokaryotic epiphytes associated with three East
+African seagrasses. \emph{Journal of Phycology} \textbf{43}: 768--779.
+
+\hypertarget{ref-Weidner1996}{}
+Weidner, S., Arnold, W., and Puhler, A. (1996) Diversity of uncultured
+microorganisms associated with the seagrass Halophila stipulacea
+estimated by restriction fragment length polymorphism analysis of
+PCR-amplified 16S rRNA genes. \emph{Applied and environmental
+microbiology} \textbf{62}: 766--771.
 

### 2019-11-25 submission/manuscript.Rmd
Paper writing: Intro and M&M

--- 2019-10-14 submission/manuscript.Rmd
@@
 csl: citation_style.csl #Get themes at https://github.com/citation-style-language/styles
-bibliography: references.bib
+bibliography:
+  - references.bib
+  - R_citation.bib
 ---
@@
 \linenumbers
+\sisetup{mode=text}
 \setlength\parindent{24pt}
@@
 ## Abstract
-jhfajklsgčklajsdbvlafgčladfhg
-
-lčkhfglashgčlashgčlak
 
 \newpage
+## Introduction
+Surfaces of marine macrophytes are inhabited by a diverse microbial community whose structure and function are poorly understood [@Egan2013].
+As less than 1 % of all prokaryotic species are culturable, to study these organisms, molecular methods such as 16S rRNA sequencing, metagenomics and metaproteomics are indispensable [@Amann1995; @Su2012].
+Applying these techniques requires an initial isolation step, with the purpose of obtaining high quality DNA and proteins.
 
-## Introduction
+Biological material (i.e.
+proteins and DNA) from pelagic microbial communities is usually isolated by collecting cells onto filters and subsequently isolating the aimed compound [@Gilbert2009].
+If a microbial size fraction is aimed sequential filtration is applied [@Massana1997; @Andersson2010].
+In contrast, obtaining biological materials from microorganism inhabiting surfaces requires either a cell detachment procedure prior to isolation or the host material is coextracted together with the targeted material.
+Methods for separating microbial cells form the host include shaking of host tissue [@Gross2003; @Noges2010], scraping of macrophyte surfaces [@Uku2007] or the application of ultrasonication [@Weidner1996; @Cai2014].
+It was shown that shaking alone is not sufficient to remove microbial cells, at least from plant root surfaces [@Richter-Heitmann2016].
+Manual separation methods, such as scraping and brushing, are time consuming and subjective, as the detachment efficiency depends on host tissue and the person performing the procedure [@Cai2014].
+Ultrasonication was proposed as an alternative method as it is providing better results in terms of detachment efficiency [@Cai2014; @Richter-Heitmann2016].
+The downside of this procedure is that complete cell removal was still not obtained and tissue disruption was observed especially after the application of probe ultrasonication [@Richter-Heitmann2016].
+An alternative to these cell detachment procedures is the isolation of targeted epiphytic compounds together with host materials [@Staufenberger2008; @Jiang2015].
+This procedure can lead to problems in the following processing steps such as mitochondrial and chloroplast 16S rRNA sequence contaminations from the host [@Longford2007; @Staufenberger2008].
+In addition, when performing metagenomics and metaproteomics host material can cause biased results towards more abundant host DNA and proteins.
 
-##Materials and Methods
+An alternative to these procedures is a direct isolation of the targeted material by incubating macrophyte tissues in an extraction buffer.
+After the incubation is done the undisrupted host tissue is removed and the isolation procedure continues omitting host material contaminations.
+To our knowledge the only procedure describing a direct and selective epiphytic DNA isolation from the surfaces of marine macrophytes was provided by @Burke2009.
+In contrast to previously described methods this protocol enables an almost complete removal of the surface community and was used for 16S rRNA gene clone libraries construction [@Burke2011b] and metagenomes sequencing [@Burke2011a].
+This method, thought providing a selective isolation procedure, is using in the extraction buffer a rapid multienzyme cleaner (3M) that is not worldwide available and whose composition is not know [@Burke2009].
+Also to our knowledge, no selective isolation protocol for proteins from epiphytic communities inhabiting marine macrophytes was established.
+
+In the present study, we adapted a protocol widely used for DNA isolation from filters [@Massana1997] and a protocol used for protein isolation from sediments [@Chourey2010; @Hultman2015] for selective extractions of DNA and proteins from epiphytic communities inhabiting the surfaces of two marine macrophytes: the seagrass *Cymodocea nodosa* and the alga *Caulerpa cylindracea*.
+In addition, we tested the removal efficiency of the protocol and the suitability of obtained DNA and proteins for 16S rRNA sequencing and metaproteomics.
+
+\newpage
+## Materials and Methods
+### Sampling
+Leaves of *Cymodocea nodosa* were sampled in a *Cymodocea nodosa* meadow in the Bay of Saline (45°7´5˝N, 13°37´20˝E) and in a *Cymodocea nodosa* meadow invaded by *Caulerpa cylindracea* in the proximity of the village of Funtana (45°10´39˝N, 13°35´42˝E).
+Thalli of *Caulerpa cylindracea* were sampled in the same *Cymodocea nodosa* invaded meadow in Funtana and on a locality of only *Caulerpa cylindracea* located in the proximity of the invaded meadow.
+Leaves and thalli were collected on the same day in two contrasting seasons, on 4 December 2017 and 18 June 2018.
+During spring 2018 the *Cymodocea nodosa* meadow in the Bay of Saline decayed to an extent that no leaves could be retrieved (Najdek *et al.*, unpublished data).
+Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater.
+Upon arrival to the laboratory, *Cymodocea nodosa* leaves were cut into sections of 1 &ndash; 2 \si{\cm}, while *Caulerpa cylindracea* thalli were cut into 5 &ndash; 8 \si{\cm} long sections.
+Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.
+
+### DNA Isolation, Illumina 16S rRNA Sequencing and Sequences Analysis
+The DNA was isolated according to the protocol for isolation from filters described in @Massana1997.
+This protocol was modified and adapted for DNA isolation from microbial communities from macrophyte surfaces as described below.
+5 \si{\ml} of lysis buffer (40 \si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar} sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or \si{\g} wet-weight of thalli.
+Lysozyme was added (final concentration 1 \si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius} for 30 \si{\minute}.
+Subsequently, proteinase K (final concentration 0.5 \si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were added and the samples were incubated at 55 \si{\degreeCelsius} for 2 \si{\hour}.
+Following the incubation, tubes were vortexed for 10 \si{\minute} and the mixture containing lyzed epiphytic cells was separated from host leaves or thalli by transferring the solution into a clean tube.
+The lysate was extracted twice with a mixture of phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with chloroform:isoamyl alcohol (24:1).
+After each organic solvent mixture addition tubes were slightly vortexed and centrifuged at 4,500 × g for 10 \si{\minute}.
+Following each centrifugation aqueous phases were retrieved.
+After the final extraction 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
+DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} °C and centrifuging at 16,000 × g and 4 \si{\degreeCelsius} for 20 \si{\minute}.
+The pellet was washed twice with 1 \si{\ml} of chilled isopropanol and centrifuged after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 10 \si{\minute}.
+After the first washing step duplicate pellets form the same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.
+The dried pellet was resuspended in 100 \si{\ul} of deionized water.
+
+An aliquot of isolated DNA was treated with RNase A (final concentration 200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}.
+The DNA concentration was determined using the Quant-iT PicoGreen dsDNA Assay Kit (Invitrogen, USA) according to the manufacturer’s instructions and diluted to 1 \si{\ng\per\ul}.
+The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
+In the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016].
+These primers contained on their 5' end a tagged sequence.
+Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5 Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template.
+Cycling conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}.
+The four parallel reactions volumes were pooled and PCR products were purified using the GeneJET PCR Purification Kit (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and following the protocol that included isopropanol addition for better small DNA fragment yield.
+The column was eluted in 30 \si{\ul} of deionized water.
+Purified PCR products were sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories, Martinsried, Germany.
+Before sequencing, at IMGM the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR.
+In addition, these primers contained adapter and sample-specific index sequences.
+The second PCR was carried out for 8 cycles.
+Beside samples, a positive and negative control were sequenced.
+A negative control was comprised of four parallel PCR reactions without DNA template, while for a postive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
+The sequences obtained in this study have been submitted to the European Nucleotide Archive (ENA) under accession numbers ?????.
+
+```{r}
+error <- read_tsv("data/mothur/data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary") %>%
+  filter(numparents==1) %>%
+  mutate(mismatches_weight=weight*mismatches)
+```
+```{r}
+files <- list.files(path="data/mothur", pattern=".error.summary", full.names=TRUE)
+
+seq_error <- NULL
+for (i in seq(1 : length(files))) {
+  file <- read_tsv(files[i], na="") %>%
+    filter(numparents==1) %>%
+    mutate(mismatches_weight=weight*mismatches) %>%
+    mutate(total_weight=weight*total)
+  reference_database <- gsub(".error.summary$", "", files[i]) %>%
+    gsub("^data/mothur/", "", .)
+  file_error <- tibble(reference_database, weight=sum(file$weight), error=sum(file$mismatches_weight)/sum(file$total_weight)*100)
+  seq_error <- rbind(seq_error, file_error)
+}
+
+Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0)[@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommandations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
+For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 132; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
+Sequences classified as chloroplsts by mothur were exported, aligned using SILVA incremental aligner (SINA, version 1.6.0) [@Pruesse2012] against the same SILVA SSU Ref NR 99 database (release 132) and imported into ARB (version 6.0.6) [@Ludwig2004] for further phylogentic analysis using the same database.
+Reference sequences close to imported were selected and used to calculate a phylogentic tree using the Maximum Likelihood algorithm RAxML (version 7.0.3) with 1000 tree replicates [@Stamatakis2006].
+Imported partial chloroplast sequences were added to the tree using the maximum parsimony criteria and not allowing changes to tree topology.
+Pipeline data processing and visualizaion was done using R (version 3.6.0) [@RCoreTeam2019], package tidyverse (version 1.2.1) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Zhu2019; @Allaire2019; @Xie2018].
+The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (?????).
+Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error of
+
+### Protein Isolation and Metaproteomics
+Proteins were isolated according to the protocol for isolation from soil described in @Chourey2010 and modified by @Hultman2015.
+These protocols were further modified and adapted for protein isolation from microbial communities form macrophyte surfaces as described below.
+20 \si{\ml} of protein extraction buffer (4 \si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl [pH 8.0]) was added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of thalli.
+The mixture was incubated in boiling water for 5 \si{\minute}, vortexed for 10 \si{\minute} and incubated again in boiling water for 5 \si{\minute}.
+After a brief vortex the lysate was transferred to a clean tube separating the host leaves or thalli from the mixture containing lyzed epiphytic cells.
+Dithiothreitol (DTT; final concentration 24 \si{\milli\Molar}) was added and proteins were precipitated with chilled 100 \si{\percent} trichloroacetic acid (TCA; final concentration 20 \si{\percent}) overnight at \num{-20} °C.
+Precipitated proteins were centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
+The obtained protein pellet was washed three times with chilled acetone.
+During the first washing step the pellet was transferred to a clean 1.5 \si{\ml} tube.
+After each washing step the samples were centrifuged at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}.
+Dried pellets were stored at \num{-80} °C until further analysis.
+
+### Confocal Microscopy
 

### 2019-11-25 submission/manuscript.tex
Paper writing: Intro and M&M

@@
 
+\usepackage{siunitx}
+\DeclareSIUnit\molar{\mole\per\cubic\deci\metre}
+\DeclareSIUnit\Molar{\textsc{m}}
+
 \begin{document}
@@
 \linenumbers
-\setlength\parindent{24pt}
+\sisetup{mode=text} \setlength\parindent{24pt}
 
@@
 \emph{et al.}, 1995; Su \emph{et al.}, 2012). Applying these techniques
-requires an inital isolation step, with the purpose of obataining high
+requires an initial isolation step, with the purpose of obtaining high
 quality DNA and proteins.
@@
 as scraping and brushing, are time consuming and subjective, as the
-detachment efficency depends on host tissue and the person performing
+detachment efficiency depends on host tissue and the person performing
 the procedure (Cai \emph{et al.}, 2014). Ultrasonication was proposed as
 an alternative method as it is providing better results in terms of
-detachemnt efficency (Cai \emph{et al.}, 2014; Richter-Heitmann \emph{et
-al.}, 2016). The downside of this procedure is that complete cell
-removal was still not obtained and tissue disruption was observed
+detachment efficiency (Cai \emph{et al.}, 2014; Richter-Heitmann
+\emph{et al.}, 2016). The downside of this procedure is that complete
+cell removal was still not obtained and tissue disruption was observed
 especially after the application of probe ultrasonication
-(Richter-Heitmann \emph{et al.}, 2016). An alterantive to these cell
+(Richter-Heitmann \emph{et al.}, 2016). An alternative to these cell
 detachment procedures is the isolation of targeted epiphytic compounds
-togehter with host materials (Staufenberger \emph{et al.}, 2008; Jiang
+together with host materials (Staufenberger \emph{et al.}, 2008; Jiang
 \emph{et al.}, 2015). This procedure can lead to problems in the
@@
 rRNA sequence contaminations from the host (Longford \emph{et al.},
-2007; Staufenberger \emph{et al.}, 2008). In additon, when performing
+2007; Staufenberger \emph{et al.}, 2008). In addition, when performing
 metagenomics and metaproteomics host material can cause biased results
-towards more abundand host DNA and proteins.
+towards more abundant host DNA and proteins.
 
-An alterantive to these procedures is a direct isolation of the targeted
+An alternative to these procedures is a direct isolation of the targeted
 material by incubating macrophyte tissues in an extraction buffer. After
-the incubation is done the undistrupted host tissue is removed and the
-isolation procedure continues ommiting host material contaminations. To
+the incubation is done the undisrupted host tissue is removed and the
+isolation procedure continues omitting host material contaminations. To
 our knowledge the only procedure describing a direct and selective
@@
 construction (Burke \emph{et al.}, 2011b) and metagenomes sequencing
-(Burke \emph{et al.}, 2011a). This method, thought providing a selctive
+(Burke \emph{et al.}, 2011a). This method, thought providing a selective
 isolation procedure, is using in the extraction buffer a rapid
-multienzyme cleaner (3M) that is not worldwide availabe and whose
+multienzyme cleaner (3M) that is not worldwide available and whose
 composition is not know (Burke \emph{et al.}, 2009). Also to our
-knowledge, no selective isolation protocol for proteins from epihytic
+knowledge, no selective isolation protocol for proteins from epiphytic
 communities inhabiting marine macrophytes was established.
@@
 isolation from filters (Massana \emph{et al.}, 1997) and a protocol used
-for protein isolation from sediments (Chourey \emph{et al.}, 2010) for
-selective extractions of DNA and proteins from epiphytic communities
-inhabiting the surfaces of two marine macrophytes: the seagrass
-\emph{Cymodocea nodosa} and the alga \emph{Caulerpa cylindracea}. In
-addition, we tested the removal efficiency of the protocol and the
-suitability of obtained DNA and proteins for 16S rRNA sequencing and
-metaproteomics.
+for protein isolation from sediments (Chourey \emph{et al.}, 2010;
+Hultman \emph{et al.}, 2015) for selective extractions of DNA and
+proteins from epiphytic communities inhabiting the surfaces of two
+marine macrophytes: the seagrass \emph{Cymodocea nodosa} and the alga
+\emph{Caulerpa cylindracea}. In addition, we tested the removal
+efficiency of the protocol and the suitability of obtained DNA and
+proteins for 16S rRNA sequencing and metaproteomics.
 
@@
 Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
-nodosa} meadow in the Bay of Saline (45°7´5˝N; 13°37´20˝E) and in a
+nodosa} meadow in the Bay of Saline (45°7´5˝N, 13°37´20˝E) and in a
 \emph{Cymodocea nodosa} meadow invaded by \emph{Caulerpa cylindracea} in
-the proximity of the village of Funtana (45°10´39˝N 13°35´42˝E). Thalli
+the proximity of the village of Funtana (45°10´39˝N, 13°35´42˝E). Thalli
 of \emph{Caulerpa cylindracea} were sampled in the same \emph{Cymodocea
 nodosa} invaded meadow in Funtana and on a locality of only
-\emph{Caulerpa cylindracea} located close to the invaded meadow.
-Seagrasses and algae were collected the same day in two contrasting
-seasons, on 4 December 2017 and 18 June 2018. During spring 2018 the
-\emph{Cymodocea nodosa} meadow in the Bay of Saline decayed to the
-extent that no leves could be retrieved (Najdek \emph{et al.},
+\emph{Caulerpa cylindracea} located in the proximity of the invaded
+meadow. Leaves and thalli were collected on the same day in two
+contrasting seasons, on 4 December 2017 and 18 June 2018. During spring
+2018 the \emph{Cymodocea nodosa} meadow in the Bay of Saline decayed to
+an extent that no leaves could be retrieved (Najdek \emph{et al.},
 unpublished data). Leaves and thalli were collected by diving and
@@
 with site seawater. Upon arrival to the laboratory, \emph{Cymodocea
-nodosa} leaves were cut into sections of 1 -- 2 cm, while \emph{Caulerpa
-cylindracea} thalli were cut into 5 -- 8 cm long sections. Leaves and
-thalli were washed three times with sterile artificial seawater (ASW) to
-remove loosely attached microbial cells.
+nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while
+\emph{Caulerpa cylindracea} thalli were cut into 5 -- 8 \si{\cm} long
+sections. Leaves and thalli were washed three times with sterile
+artificial seawater (ASW) to remove loosely attached microbial cells.
 
-\subsubsection{DNA isolation}\label{dna-isolation}
+\subsubsection{DNA Isolation, Illumina 16S rRNA Sequencing and Sequences
+Analysis}\label{dna-isolation-illumina-16s-rrna-sequencing-and-sequences-analysis}
 
@@
 filters described in Massana \emph{et al.} (1997). This protocol was
-modified and adapted for DNA isolation from microbial communites from
-macrophytes surfaces as described below. 1 g wet weight of leaves and 2
-g wet-weight of thalli were placed into 5 ml of lysis buffer (40 mM
-EDTA, 50 mM Tris-HCl, 0.75 M sucrose; pH 8.3). Lysozyme was added (final
-concentration 1 mg ml\textsuperscript{-1}) and the mixture was incubated
-at 37 °C for 30 minutes. Subsequently, proteinase K (final concentration
-0.5 mg ml\textsuperscript{-1}) and SDS (finl concentration 1 \%) were
-added and the samples were incubated at 55 °C for 2 hours. Following the
-incubation, tubes were vortexed for 10 minutes and the mixture
-containign lyzed cells was separated from host leaves or thalli by
-transferrring the solution into a clean tube. The lysate was extracted
-twice with a mixture of phenol:chloroform:isoamyl alcohol (25:24:1; pH
-8) and once with chloroform-isoamyl alcohol (24:1). After each organic
-solvent mixture addition tubes were slightly vortexed and centrifugated
-at 4,500 x g for 10 minutes. Following each centrifugation aqueous
-phases were retrieved. After the final extraction 1/10 of 3 M sodium
-acetate (ph 5.2) was added. DNA was precipitated by adding 1 volumne of
-isopropanol and incubating the mixtures at -20 °C overnight.
+modified and adapted for DNA isolation from microbial communities from
+macrophyte surfaces as described below. 5 \si{\ml} of lysis buffer (40
+\si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar}
+sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or \si{\g}
+wet-weight of thalli. Lysozyme was added (final concentration 1
+\si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius}
+for 30 \si{\minute}. Subsequently, proteinase K (final concentration 0.5
+\si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were
+added and the samples were incubated at 55 \si{\degreeCelsius} for 2
+\si{\hour}. Following the incubation, tubes were vortexed for 10
+\si{\minute} and the mixture containing lyzed epiphytic cells was
+separated from host leaves or thalli by transferring the solution into a
+clean tube. The lysate was extracted twice with a mixture of
+phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with
+chloroform:isoamyl alcohol (24:1). After each organic solvent mixture
+addition tubes were slightly vortexed and centrifuged at 4,500 × g for
+10 \si{\minute}. Following each centrifugation aqueous phases were
+retrieved. After the final extraction 1/10 of chilled 3 \si{\Molar}
+sodium acetate (pH 5.2) was added. DNA was precipitated by adding 1
+volume of chilled isopropanol, incubating the mixtures overnight at
+\num{-20} °C and centrifuging at 16,000 × g and 4 \si{\degreeCelsius}
+for 20 \si{\minute}. The pellet was washed twice with 1 \si{\ml} of
+chilled isopropanol and centrifuged after each washing step at 20,000 ×
+g and 4 \si{\degreeCelsius} for 10 \si{\minute}. After the first washing
+step duplicate pellets form the same sample were pooled and transferred
+to a clean 1.5 \si{\ml} tube. The dried pellet was resuspended in 100
+\si{\ul} of deionized water.
 
-\subsubsection{Protein isolation}\label{protein-isolation}
+An aliquot of isolated DNA was treated with RNase A (final concentration
+200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
+concentration was determined using the Quant-iT PicoGreen dsDNA Assay
+Kit (Invitrogen, USA) according to the manufacturer's instructions and
+diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
+amplified using a two-step PCR procedure. In the first PCR the 515F
+(5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3')
+primers from the Earth Microbiome Project
+(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
+were used to amplify the target region (Caporaso \emph{et al.}, 2012;
+Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
+contained on their 5' end a tagged sequence. Each sample was amplified
+in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5
+Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml}
+BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse
+primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs,
+USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial
+denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of
+denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
+\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
+\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
+72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
+volumes were pooled and PCR products were purified using the GeneJET PCR
+Purification Kit (Thermo Fisher Scientific, USA) according to the
+manufacturer's instructions and following the protocol that included
+isopropanol addition for better small DNA fragment yield. The column was
+eluted in 30 \si{\ul} of deionized water. Purified PCR products were
+sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
+Martinsried, Germany. Before sequencing, at IMGM the second PCR
+amplification of the two-step PCR procedure was performed using primers
+targeting the tagged region incorporated in the first PCR. In addition,
+these primers contained adapter and sample-specific index sequences. The
+second PCR was carried out for 8 cycles. Beside samples, a positive and
+negative control were sequenced. A negative control was comprised of
+four parallel PCR reactions without DNA template, while for a postive
+control a mock community composed of evenly mixed DNA material
+originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
+used. The sequences obtained in this study have been submitted to the
+European Nucleotide Archive (ENA) under accession numbers ??????????.
+
+Obtained sequences were analyzed on the computer cluster Isabella
+(University Computing Center, University of Zagreb) using mothur
+(version 1.43.0)(Schloss \emph{et al.}, 2009) according to the MiSeq
+Standard Operating Procedure (MiSeq SOP;
+\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
+and recommandations given from the Riffomonas project to enhance data
+reproducibility (\url{http://www.riffomonas.org/}). For alignment and
+classification of sequences the SILVA SSU Ref NR 99 database (release
+132; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
+Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplsts by
+mothur were exported, aligned using SILVA incremental aligner (SINA,
+version 1.6.0) (Pruesse \emph{et al.}, 2012) against the same SILVA SSU
+Ref NR 99 database (release 132) and imported into ARB (version 6.0.6)
+(Ludwig \emph{et al.}, 2004) for further phylogentic analysis using the
+same database. Reference sequences close to imported were selected and
+used to calculate a phylogentic tree using the Maximum Likelihood
+algorithm RAxML (version 7.0.3) with 1000 tree replicates (Stamatakis,
+2006). Imported partial chloroplast sequences were added to the tree
+using the maximum parsimony criteria and not allowing changes to tree
+topology. Pipeline data processing and visualizaion was done using R
+(version 3.6.0) (R Core Team, 2019), package tidyverse (version 1.2.1)
+(Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
+2015, 2019a, 2019b; Xie \emph{et al.}, 2018; Zhu, 2019; Allaire \emph{et
+al.}, 2019). The detailed analysis procedure including the R Markdown
+file for this paper are available as a GitHub repository (??????).
+
+\subsubsection{Protein Isolation and
+Metaproteomics}\label{protein-isolation-and-metaproteomics}
+
+Proteins were isolated according to the protocol for isolation from soil
+described in Chourey \emph{et al.} (2010) and modified by Hultman
+\emph{et al.} (2015). These protocols were further modified and adapted
+for protein isolation from microbial communities form macrophyte
+surfaces as described below. 20 \si{\ml} of protein extraction buffer (4
+\si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was
+added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of
+thalli. The mixture was incubated in boiling water for 5 \si{\minute},
+vortexed for 10 \si{\minute} and incubated again in boiling water for 5
+\si{\minute}. After a brief vortex the lysate was transferred to a clean
+tube separating the host leaves or thalli from the mixture containing
+lyzed epiphytic cells. Dithiothreitol (DTT; final concentration 24
+\si{\milli\Molar}) was added and proteins were precipitated with chilled
+100 \si{\percent} trichloroacetic acid (TCA; final concentration 20
+\si{\percent}) overnight at \num{-20} °C. Precipitated proteins were
+centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
+The obtained protein pellet was washed three times with chilled acetone.
+During the first washing step the pellet was transferred to a clean 1.5
+\si{\ml} tube. After each washing step the samples were centrifuged at
+20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets
+were stored at \num{-80} °C until further analysis.
+
+\subsubsection{Confocal Microscopy}\label{confocal-microscopy}
 
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Allaire2019}{}
+Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
+A., et al. (2019) rmarkdown: Dynamic Documents for R.
+
 \hypertarget{ref-Amann1995}{}
@@
 171--181.
+
+\hypertarget{ref-Apprill2015}{}
+Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015) Minor
+revision to V4 region SSU rRNA 806R gene primer greatly increases
+detection of SAR11 bacterioplankton. \emph{Aquatic Microbial Ecology}
+\textbf{75}: 129--137.
 
@@
 
+\hypertarget{ref-Caporaso2012}{}
+Caporaso, J.G., Lauber, C.L., Walters, W.A., Berg-Lyons, D., Huntley,
+J., Fierer, N., et al. (2012) Ultra-high-throughput microbial community
+analysis on the Illumina HiSeq and MiSeq platforms. \emph{The ISME
+Journal} \textbf{6}: 1621--1624.
+
 \hypertarget{ref-Chourey2010}{}
@@
 
+\hypertarget{ref-Hultman2015}{}
+Hultman, J., Waldrop, M.P., Mackelprang, R., David, M.M., McFarland, J.,
+Blazewicz, S.J., et al. (2015) Multi-omics of permafrost, active layer
+and thermokarst bog soil microbiomes. \emph{Nature} \textbf{521}:
+208--212.
+
 \hypertarget{ref-Jiang2015}{}
@@
 
+\hypertarget{ref-Kozich2013}{}
+Kozich, J.J., Westcott, S.L., Baxter, N.T., Highlander, S.K., and
+Schloss, P.D. (2013) Development of a dual-index sequencing strategy and
+curation pipeline for analyzing amplicon sequence data on the MiSeq
+Illumina sequencing platform. \emph{Applied and environmental
+microbiology} \textbf{79}: 5112--5120.
+
 \hypertarget{ref-Longford2007}{}
@@
 \emph{Aquatic Microbial Ecology} \textbf{48}: 217--229.
+
+\hypertarget{ref-Ludwig2004}{}
+Ludwig, W., Strunk, O., Westram, R., Richter, L., Meier, H., Yadhukumar,
+et al. (2004) ARB: a software environment for sequence data.
+\emph{Nucleic Acids Research} \textbf{32}: 1363--1371.
 
@@
 
+\hypertarget{ref-Parada2016}{}
+Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
+matters: assessing small subunit rRNA primers for marine microbiomes
+with mock communities, time series and global field samples.
+\emph{Environmental Microbiology} \textbf{18}: 1403--1414.
+
+\hypertarget{ref-Pruesse2012}{}
+Pruesse, E., Peplies, J., and Glöckner, F.O. (2012) SINA: accurate
+high-throughput multiple sequence alignment of ribosomal RNA genes.
+\emph{Bioinformatics (Oxford, England)} \textbf{28}: 1823--1829.
+
+\hypertarget{ref-Quast2013}{}
+Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
+et al. (2013) The SILVA ribosomal RNA gene database project: improved
+data processing and web-based tools. \emph{Nucleic acids research}
+\textbf{41}: D590--6.
+
+\hypertarget{ref-RCoreTeam2019}{}
+R Core Team (2019) R: A Language and Environment for Statistical
+Computing, Vienna, Austria: R Foundation for Statistical Computing.
+
 \hypertarget{ref-Richter-Heitmann2016}{}
@@
 \emph{Frontiers in Microbiology} \textbf{7}: 773.
+
+\hypertarget{ref-Schloss2009}{}
+Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
+Hollister, E.B., et al. (2009) Introducing mothur: open-source,
+platform-independent, community-supported software for describing and
+comparing microbial communities. \emph{Applied and environmental
+microbiology} \textbf{75}: 7537--7541.
+
+\hypertarget{ref-Stamatakis2006}{}
+Stamatakis, A. (2006) RAxML-VI-HPC: maximum likelihood-based
+phylogenetic analyses with thousands of taxa and mixed models.
+\emph{Bioinformatics} \textbf{22}: 2688--2690.
 
@@
 
+\hypertarget{ref-Wickham2019}{}
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
+R., et al. (2019) Welcome to the tidyverse. \emph{Journal of Open Source
+Software} \textbf{4}: 1686.
+
+\hypertarget{ref-Xie2015}{}
+Xie, Y. (2015) Dynamic Documents with \{R\} and knitr, 2nd ed. Boca
+Raton, Florida: Chapman; Hall/CRC.
+
+\hypertarget{ref-Xie2014}{}
+Xie, Y. (2014) knitr: A Comprehensive Tool for Reproducible Research in
+\{R\}. In, Stodden,V., Leisch,F., and Peng,R.D. (eds),
+\emph{Implementing reproducible computational research}. Chapman;
+Hall/CRC.
+
+\hypertarget{ref-Xie2019a}{}
+Xie, Y. (2019a) knitr: A General-Purpose Package for Dynamic Report
+Generation in R.
+
+\hypertarget{ref-Xie2019}{}
+Xie, Y. (2019b) tinytex: Helper Functions to Install and Maintain 'TeX
+Live', and Compile 'LaTeX' Documents.
+
+\hypertarget{ref-Xie2018}{}
+Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
+Definitive Guide, Boca Raton, Florida: Chapman; Hall/CRC.
+
+\hypertarget{ref-Yilmaz2014}{}
+Yilmaz, P., Parfrey, L.W., Yarza, P., Gerken, J., Pruesse, E., Quast,
+C., et al. (2014) The SILVA and ``All-species Living Tree Project
+(LTP)'' taxonomic frameworks. \emph{Nucleic acids research} \textbf{42}:
+D643--8.
+
+\hypertarget{ref-Zhu2019}{}
+Zhu, H. (2019) kableExtra: Construct Complex Table with 'kable' and Pipe
+Syntax.
+
 \end{document}

### 2019-11-25 submission/manuscript.Rmd
Added text to M&M

@@
 csl: citation_style.csl #Get themes at https://github.com/citation-style-language/styles
-bibliography:
-  - references.bib
-  - R_citation.bib
+bibliography: references.bib
 ---
@@
 ```{r}
-error <- read_tsv("data/mothur/data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary") %>%
-  filter(numparents==1) %>%
-  mutate(mismatches_weight=weight*mismatches)
-```
-```{r}
-files <- list.files(path="data/mothur", pattern=".error.summary", full.names=TRUE)
-
-seq_error <- NULL
-for (i in seq(1 : length(files))) {
-  file <- read_tsv(files[i], na="") %>%
+# Calculation of the error rate based on the ATCC MSA-1002 mock community
+error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
   filter(numparents==1) %>%
@@
   mutate(total_weight=weight*total)
-  reference_database <- gsub(".error.summary$", "", files[i]) %>%
-    gsub("^data/mothur/", "", .)
-  file_error <- tibble(reference_database, weight=sum(file$weight), error=sum(file$mismatches_weight)/sum(file$total_weight)*100)
-  seq_error <- rbind(seq_error, file_error)
-}
+error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
+# Calculation of the number of sequences in the negative control
+nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
+  filter(taxon=="Root") %>%
+  select(ATCC_1)
+```
 
@@
 Sequences classified as chloroplsts by mothur were exported, aligned using SILVA incremental aligner (SINA, version 1.6.0) [@Pruesse2012] against the same SILVA SSU Ref NR 99 database (release 132) and imported into ARB (version 6.0.6) [@Ludwig2004] for further phylogentic analysis using the same database.
-Reference sequences close to imported were selected and used to calculate a phylogentic tree using the Maximum Likelihood algorithm RAxML (version 7.0.3) with 1000 tree replicates [@Stamatakis2006].
+Reference sequences close to imported ones were selected and used to calculate a phylogentic tree using the Maximum Likelihood algorithm RAxML (version 7.0.3) with 1000 tree replicates [@Stamatakis2006].
 Imported partial chloroplast sequences were added to the tree using the maximum parsimony criteria and not allowing changes to tree topology.
@@
 The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (?????).
-Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error of
+Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error rate of `r error` \si{\percent} was detemined that is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
+In addition, the negative control processed together with the samples yielded only `r nc` sequences.
 

### 2019-11-25 submission/manuscript.tex
Added text to M&M

@@
 used. The sequences obtained in this study have been submitted to the
-European Nucleotide Archive (ENA) under accession numbers ??????????.
+European Nucleotide Archive (ENA) under accession numbers ?????.
+
+\begin{verbatim}
+## Parsed with column specification:
+## cols(
+##   .default = col_double(),
+##   query = col_character(),
+##   reference = col_character()
+## )
+\end{verbatim}
+
+\begin{verbatim}
+## See spec(...) for full column specifications.
+\end{verbatim}
+
+\begin{verbatim}
+## Parsed with column specification:
+## cols(
+##   taxlevel = col_double(),
+##   rankID = col_character(),
+##   taxon = col_character(),
+##   daughterlevels = col_double(),
+##   total = col_double(),
+##   `40` = col_double(),
+##   `41` = col_double(),
+##   `42` = col_double(),
+##   `43` = col_double(),
+##   `61` = col_double(),
+##   `62` = col_double(),
+##   `63` = col_double(),
+##   ATCC_1 = col_double()
+## )
+\end{verbatim}
 
@@
 al.}, 2019). The detailed analysis procedure including the R Markdown
-file for this paper are available as a GitHub repository (??????).
+file for this paper are available as a GitHub repository (?????). Based
+on the ATCC MSA-1002 mock community included in the analysis a
+sequencing error of 0.009 \si{\percent} was detemined that is in line
+with previously reported values for next-generation sequencing data
+(Kozich \emph{et al.}, 2013; Schloss \emph{et al.}, 2016). In addition,
+the negative control processed together with the samples yielded only
+44,056 sequences.
 
@@
 
+\hypertarget{ref-Schloss2016}{}
+Schloss, P.D., Girard, R.A., Martin, T., Edwards, J., and Thrash, J.C.
+(2016) Status of the Archaeal and Bacterial Census: an Update.
+\emph{mBio} \textbf{7}: e00201--16.
+
 \hypertarget{ref-Schloss2009}{}

### 2019-11-25 submission/manuscript.Rmd
NC added to analysis

@@
 
-Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0)[@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommandations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
+Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommandations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
 For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 132; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].

### 2019-11-25 submission/manuscript.tex
NC added to analysis

--- 2019-11-25 submission/manuscript.tex
@@
 2. University of Vienna, Department of Limnology and Bio-Oceanography,
-Althanstraße 14, Vienna, Austria \newpage
-\linenumbers
+Althanstraße 14, Vienna, Austria \newpage \linenumbers
 \sisetup{mode=text} \setlength\parindent{24pt}
 
-\subsection{Abstract}\label{abstract}
+\hypertarget{abstract}{%
+\subsection{Abstract}\label{abstract}}
 
@@
 
-\subsection{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\subsection{Introduction}\label{introduction}}
 
@@
 
-\subsection{Materials and Methods}\label{materials-and-methods}
+\hypertarget{materials-and-methods}{%
+\subsection{Materials and Methods}\label{materials-and-methods}}
 
-\subsubsection{Sampling}\label{sampling}
+\hypertarget{sampling}{%
+\subsubsection{Sampling}\label{sampling}}
 
@@
 
+\hypertarget{dna-isolation-illumina-16s-rrna-sequencing-and-sequences-analysis}{%
 \subsubsection{DNA Isolation, Illumina 16S rRNA Sequencing and Sequences
-Analysis}\label{dna-isolation-illumina-16s-rrna-sequencing-and-sequences-analysis}
+Analysis}\label{dna-isolation-illumina-16s-rrna-sequencing-and-sequences-analysis}}
 
@@
 used. The sequences obtained in this study have been submitted to the
-European Nucleotide Archive (ENA) under accession numbers ??????????.
+European Nucleotide Archive (ENA) under accession numbers ?????.
+
+\begin{verbatim}
+## Parsed with column specification:
+## cols(
+##   .default = col_double(),
+##   query = col_character(),
+##   reference = col_character()
+## )
+\end{verbatim}
+
+\begin{verbatim}
+## See spec(...) for full column specifications.
+\end{verbatim}
+
+\begin{verbatim}
+## Parsed with column specification:
+## cols(
+##   taxlevel = col_double(),
+##   rankID = col_character(),
+##   taxon = col_character(),
+##   daughterlevels = col_double(),
+##   total = col_double(),
+##   `40` = col_double(),
+##   `41` = col_double(),
+##   `42` = col_double(),
+##   `43` = col_double(),
+##   `61` = col_double(),
+##   `62` = col_double(),
+##   `63` = col_double(),
+##   ATCC_1 = col_double(),
+##   NC_1 = col_double()
+## )
+\end{verbatim}
 
@@
 (University Computing Center, University of Zagreb) using mothur
-(version 1.43.0)(Schloss \emph{et al.}, 2009) according to the MiSeq
+(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
 Standard Operating Procedure (MiSeq SOP;
@@
 (Ludwig \emph{et al.}, 2004) for further phylogentic analysis using the
-same database. Reference sequences close to imported were selected and
-used to calculate a phylogentic tree using the Maximum Likelihood
+same database. Reference sequences close to imported ones were selected
+and used to calculate a phylogentic tree using the Maximum Likelihood
 algorithm RAxML (version 7.0.3) with 1000 tree replicates (Stamatakis,
@@
 al.}, 2019). The detailed analysis procedure including the R Markdown
-file for this paper are available as a GitHub repository (??????).
+file for this paper are available as a GitHub repository (?????). Based
+on the ATCC MSA-1002 mock community included in the analysis a
+sequencing error rate of 0.009 \si{\percent} was detemined that is in
+line with previously reported values for next-generation sequencing data
+(Kozich \emph{et al.}, 2013; Schloss \emph{et al.}, 2016). In addition,
+the negative control processed together with the samples yielded only
+44,050 sequences.
 
+\hypertarget{protein-isolation-and-metaproteomics}{%
 \subsubsection{Protein Isolation and
-Metaproteomics}\label{protein-isolation-and-metaproteomics}
+Metaproteomics}\label{protein-isolation-and-metaproteomics}}
 
@@
 
-\subsubsection{Confocal Microscopy}\label{confocal-microscopy}
+\hypertarget{confocal-microscopy}{%
+\subsubsection{Confocal Microscopy}\label{confocal-microscopy}}
 
-\subsection{Results and Discussion}\label{results-and-discussion}
+\hypertarget{results-and-discussion}{%
+\subsection{Results and Discussion}\label{results-and-discussion}}
 
-\subsection{Conclusions}\label{conclusions}
+\hypertarget{conclusions}{%
+\subsection{Conclusions}\label{conclusions}}
 
-\subsection{Materials and Methods}\label{materials-and-methods-1}
+\hypertarget{materials-and-methods-1}{%
+\subsection{Materials and Methods}\label{materials-and-methods-1}}
 
@@
 
-\subsection*{References}\label{references}
+\hypertarget{references}{%
+\subsection*{References}\label{references}}
 \addcontentsline{toc}{subsection}{References}
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Allaire2019}{}
+\leavevmode\hypertarget{ref-Allaire2019}{}%
 Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
@@
 
-\hypertarget{ref-Amann1995}{}
+\leavevmode\hypertarget{ref-Amann1995}{}%
 Amann, R.I., Ludwig, W., and Schleifer, K.H. (1995) Phylogenetic
@@
 
-\hypertarget{ref-Andersson2010}{}
+\leavevmode\hypertarget{ref-Andersson2010}{}%
 Andersson, A.F., Riemann, L., and Bertilsson, S. (2010) Pyrosequencing
@@
 
-\hypertarget{ref-Apprill2015}{}
+\leavevmode\hypertarget{ref-Apprill2015}{}%
 Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015) Minor
@@
 
-\hypertarget{ref-Burke2009}{}
+\leavevmode\hypertarget{ref-Burke2009}{}%
 Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective extraction of
@@
 
-\hypertarget{ref-Burke2011a}{}
+\leavevmode\hypertarget{ref-Burke2011a}{}%
 Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
@@
 
-\hypertarget{ref-Burke2011b}{}
+\leavevmode\hypertarget{ref-Burke2011b}{}%
 Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
@@
 
-\hypertarget{ref-Cai2014}{}
+\leavevmode\hypertarget{ref-Cai2014}{}%
 Cai, X., Gao, G., Yang, J., Tang, X., Dai, J., Chen, D., and Song, Y.
@@
 
-\hypertarget{ref-Caporaso2012}{}
+\leavevmode\hypertarget{ref-Caporaso2012}{}%
 Caporaso, J.G., Lauber, C.L., Walters, W.A., Berg-Lyons, D., Huntley,
@@
 
-\hypertarget{ref-Chourey2010}{}
+\leavevmode\hypertarget{ref-Chourey2010}{}%
 Chourey, K., Jansson, J., VerBerkmoes, N., Shah, M., Chavarria, K.L.,
@@
 
-\hypertarget{ref-Egan2013}{}
+\leavevmode\hypertarget{ref-Egan2013}{}%
 Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
@@
 
-\hypertarget{ref-Gilbert2009}{}
+\leavevmode\hypertarget{ref-Gilbert2009}{}%
 Gilbert, J.A., Field, D., Swift, P., Newbold, L., Oliver, A., Smyth, T.,
@@
 
-\hypertarget{ref-Gross2003}{}
+\leavevmode\hypertarget{ref-Gross2003}{}%
 Gross, E.M., Feldbaum, C., and Graf, A. (2003) Epiphyte biomass and
@@
 
-\hypertarget{ref-Hultman2015}{}
+\leavevmode\hypertarget{ref-Hultman2015}{}%
 Hultman, J., Waldrop, M.P., Mackelprang, R., David, M.M., McFarland, J.,
@@
 
-\hypertarget{ref-Jiang2015}{}
+\leavevmode\hypertarget{ref-Jiang2015}{}%
 Jiang, Y.-F., Ling, J., Dong, J.-D., Chen, B., Zhang, Y.-Y., Zhang,
@@
 
-\hypertarget{ref-Kozich2013}{}
+\leavevmode\hypertarget{ref-Kozich2013}{}%
 Kozich, J.J., Westcott, S.L., Baxter, N.T., Highlander, S.K., and
@@
 
-\hypertarget{ref-Longford2007}{}
+\leavevmode\hypertarget{ref-Longford2007}{}%
 Longford, S., Tujula, N., Crocetti, G., Holmes, A., Holmström, C.,
@@
 
-\hypertarget{ref-Ludwig2004}{}
+\leavevmode\hypertarget{ref-Ludwig2004}{}%
 Ludwig, W., Strunk, O., Westram, R., Richter, L., Meier, H., Yadhukumar,
@@
 
-\hypertarget{ref-Massana1997}{}
+\leavevmode\hypertarget{ref-Massana1997}{}%
 Massana, R., Murray, A.E., Preston, C.M., and DeLong, E.F. (1997)
@@
 
-\hypertarget{ref-Noges2010}{}
+\leavevmode\hypertarget{ref-Noges2010}{}%
 Nõges, T., Luup, H., and Feldmann, T. (2010) Primary production of
@@
 
-\hypertarget{ref-Parada2016}{}
+\leavevmode\hypertarget{ref-Parada2016}{}%
 Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
@@
 
-\hypertarget{ref-Pruesse2012}{}
+\leavevmode\hypertarget{ref-Pruesse2012}{}%
 Pruesse, E., Peplies, J., and Glöckner, F.O. (2012) SINA: accurate
@@
 
-\hypertarget{ref-Quast2013}{}
+\leavevmode\hypertarget{ref-Quast2013}{}%
 Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
@@
 
-\hypertarget{ref-RCoreTeam2019}{}
+\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
 R Core Team (2019) R: A Language and Environment for Statistical
@@
 
-\hypertarget{ref-Richter-Heitmann2016}{}
+\leavevmode\hypertarget{ref-Richter-Heitmann2016}{}%
 Richter-Heitmann, T., Eickhorst, T., Knauth, S., Friedrich, M.W., and
@@
 
-\hypertarget{ref-Schloss2009}{}
+\leavevmode\hypertarget{ref-Schloss2016}{}%
+Schloss, P.D., Girard, R.A., Martin, T., Edwards, J., and Thrash, J.C.
+(2016) Status of the Archaeal and Bacterial Census: an Update.
+\emph{mBio} \textbf{7}: e00201--16.
+
+\leavevmode\hypertarget{ref-Schloss2009}{}%
 Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
@@
 
-\hypertarget{ref-Stamatakis2006}{}
+\leavevmode\hypertarget{ref-Stamatakis2006}{}%
 Stamatakis, A. (2006) RAxML-VI-HPC: maximum likelihood-based
@@
 
-\hypertarget{ref-Staufenberger2008}{}
+\leavevmode\hypertarget{ref-Staufenberger2008}{}%
 Staufenberger, T., Thiel, V., Wiese, J., and Imhoff, J.F. (2008)
@@
 
-\hypertarget{ref-Su2012}{}
+\leavevmode\hypertarget{ref-Su2012}{}%
 Su, C., Lei, L., Duan, Y., Zhang, K.-Q., and Yang, J. (2012)
@@
 
-\hypertarget{ref-Uku2007}{}
+\leavevmode\hypertarget{ref-Uku2007}{}%
 Uku, J., Björk, M., Bergman, B., and Díez, B. (2007) Characterization
@@
 
-\hypertarget{ref-Weidner1996}{}
+\leavevmode\hypertarget{ref-Weidner1996}{}%
 Weidner, S., Arnold, W., and Puhler, A. (1996) Diversity of uncultured
@@
 
-\hypertarget{ref-Wickham2019}{}
+\leavevmode\hypertarget{ref-Wickham2019}{}%
 Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
@@
 
-\hypertarget{ref-Xie2015}{}
+\leavevmode\hypertarget{ref-Xie2015}{}%
 Xie, Y. (2015) Dynamic Documents with \{R\} and knitr, 2nd ed. Boca
@@
 
-\hypertarget{ref-Xie2014}{}
+\leavevmode\hypertarget{ref-Xie2014}{}%
 Xie, Y. (2014) knitr: A Comprehensive Tool for Reproducible Research in
@@
 
-\hypertarget{ref-Xie2019a}{}
+\leavevmode\hypertarget{ref-Xie2019a}{}%
 Xie, Y. (2019a) knitr: A General-Purpose Package for Dynamic Report
@@
 
-\hypertarget{ref-Xie2019}{}
+\leavevmode\hypertarget{ref-Xie2019}{}%
 Xie, Y. (2019b) tinytex: Helper Functions to Install and Maintain 'TeX
@@
 
-\hypertarget{ref-Xie2018}{}
+\leavevmode\hypertarget{ref-Xie2018}{}%
 Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
@@
 
-\hypertarget{ref-Yilmaz2014}{}
+\leavevmode\hypertarget{ref-Yilmaz2014}{}%
 Yilmaz, P., Parfrey, L.W., Yarza, P., Gerken, J., Pruesse, E., Quast,
-C., et al. (2014) The SILVA and ``All-species Living Tree Project
-(LTP)'' taxonomic frameworks. \emph{Nucleic acids research} \textbf{42}:
+C., et al. (2014) The SILVA and "All-species Living Tree Project (LTP)"
+taxonomic frameworks. \emph{Nucleic acids research} \textbf{42}:
 D643--8.
 
-\hypertarget{ref-Zhu2019}{}
+\leavevmode\hypertarget{ref-Zhu2019}{}%
 Zhu, H. (2019) kableExtra: Construct Complex Table with 'kable' and Pipe

### 2019-11-28 submission/manuscript.Rmd
Added community barplot figure.

--- 2019-11-25 submission/manuscript.Rmd
@@
 opts_chunk$set("cache" = FALSE)
+opts_chunk$set("message" = FALSE)
 
@@
 Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater.
-Upon arrival to the laboratory, *Cymodocea nodosa* leaves were cut into sections of 1 &ndash; 2 \si{\cm}, while *Caulerpa cylindracea* thalli were cut into 5 &ndash; 8 \si{\cm} long sections.
+Upon arrival to the laboratory, *Cymodocea nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *Caulerpa cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections.
 Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.
 
-### DNA Isolation, Illumina 16S rRNA Sequencing and Sequences Analysis
+### DNA Isolation
 The DNA was isolated according to the protocol for isolation from filters described in @Massana1997.
@@
 
+### Illumina 16S rRNA Sequencing
 An aliquot of isolated DNA was treated with RNase A (final concentration 200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}.
@@
 Beside samples, a positive and negative control were sequenced.
-A negative control was comprised of four parallel PCR reactions without DNA template, while for a postive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
-The sequences obtained in this study have been submitted to the European Nucleotide Archive (ENA) under accession numbers ?????.
+A negative control was comprised of four parallel PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
+The sequences obtained in this study have been submitted to the European Nucleotide Archive (ENA) under accession numbers **TO BE ADDED LATER!**.
 
+### Sequence Analysis
 ```{r}
@@
   filter(taxon=="Root") %>%
-  select(ATCC_1)
+  select(NC_1)
 ```
+Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
+For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 132; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
+Sequences classified as chloroplasts by mothur were exported, aligned using SILVA incremental aligner (SINA, version 1.6.0) [@Pruesse2012] against the same SILVA SSU Ref NR 99 database (release 132) and imported into ARB (version 6.0.6) [@Ludwig2004] for further phylogenetic analysis using the same database.
+Reference sequences close to imported ones were selected and used to calculate a phylogenetic tree using the Maximum Likelihood algorithm RAxML (version 7.0.3) with 1000 tree replicates [@Stamatakis2006].
+Imported partial chloroplast sequences were added to the tree using the maximum parsimony criteria and not allowing changes to tree topology.
+Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019], package tidyverse (version 1.2.1) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Zhu2019; @Allaire2019; @Xie2018].
+The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (**TO BE ADDED LATER!**).
+Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
+In addition, the negative control processed together with the samples yielded only `r nc` sequences after sequence quality curation.
 
-Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0)[@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommandations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
-For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 132; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Sequences classified as chloroplsts by mothur were exported, aligned using SILVA incremental aligner (SINA, version 1.6.0) [@Pruesse2012] against the same SILVA SSU Ref NR 99 database (release 132) and imported into ARB (version 6.0.6) [@Ludwig2004] for further phylogentic analysis using the same database.
-Reference sequences close to imported ones were selected and used to calculate a phylogentic tree using the Maximum Likelihood algorithm RAxML (version 7.0.3) with 1000 tree replicates [@Stamatakis2006].
-Imported partial chloroplast sequences were added to the tree using the maximum parsimony criteria and not allowing changes to tree topology.
-Pipeline data processing and visualizaion was done using R (version 3.6.0) [@RCoreTeam2019], package tidyverse (version 1.2.1) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Zhu2019; @Allaire2019; @Xie2018].
-The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (?????).
-Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error rate of `r error` \si{\percent} was detemined that is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
-In addition, the negative control processed together with the samples yielded only `r nc` sequences.
-
-### Protein Isolation and Metaproteomics
+### Protein Isolation
 Proteins were isolated according to the protocol for isolation from soil described in @Chourey2010 and modified by @Hultman2015.
@@
 During the first washing step the pellet was transferred to a clean 1.5 \si{\ml} tube.
-After each washing step the samples were centrifuged at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}.
+After each washing step samples were centrifuged at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}.
 Dried pellets were stored at \num{-80} °C until further analysis.
 
+### Metaproteomics
+Isolated proteins were whole trypsin digested using the FASP (filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK) according to the manufacturer's instructions [@Wisniewski2009] with small modifications.
+Before loading the solution to the column, protein pellets were solubilized in a urea sample buffer included in the kit amended with DTT (final concentration 100 \si{\milli\Molar}) for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to remove larger particles.
+The first washing step after protein solution loading was repeated twice.
+In addition, centrifugation steps were prolonged if the column was clogged.
+Trypsin digestion was performed on column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
+The final filtrate containing digested proteins was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid, freezed at \num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent to the Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis.
+**TO BE ADDED-VIENNA PART**.
+
 ### Confocal Microscopy
+Host leaves and thalli from DNA and protein isolation steps were washed seven times in deionized water and fixed with formaldehyde (final concentration ~ 3 \si{\percent}).
+In addition, nontreated leaves and thalli, washed three times in ASW to remove loosely attached microbial cells, were fixed in the same concentration of formaldehyde and used as a positive control.
+For long therm storage, fixed leaves and thalli were immersed in a mixture of phosphate-buffered saline (PBS) and ethanol (1:1) and stored at \num{-20} \si{\degreeCelsius}.
+Treated and untreated leaves and thalli segments were stained in a 2 × solution of SYBR Green I and examined under a Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
 
-## Results and Discussion
+## Results
 
-## Conclusions
+## Discussion
 
-## Materials and Methods
+## Acknowledgements
 
 \newpage
+## References
+<div id="refs"></div>
 
 \newpage 
+## Figure Captions
 
-## References
+## Figures
+
+```{r out.width="100%", fig.align="center", fig.cap="Bacteril Community\\label{community}", fig.pos="h"}
+include_graphics("../results/figures/community_barplot.jpg")
+```

### 2019-11-28 submission/manuscript.tex
Added community barplot figure.

--- 2019-11-25 submission/manuscript.tex
@@
 \usepackage[labelsep=period, labelfont=bf]{caption}
-\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thefigure}{\arabic{figure}}
 \captionsetup{justification=raggedright,singlelinecheck=false}
@@
 
-\subsubsection{DNA Isolation, Illumina 16S rRNA Sequencing and Sequences
-Analysis}\label{dna-isolation-illumina-16s-rrna-sequencing-and-sequences-analysis}
+\subsubsection{DNA Isolation}\label{dna-isolation}
 
@@
 \si{\ul} of deionized water.
+
+\subsubsection{Illumina 16S rRNA
+Sequencing}\label{illumina-16s-rrna-sequencing}
 
@@
 negative control were sequenced. A negative control was comprised of
-four parallel PCR reactions without DNA template, while for a postive
+four parallel PCR reactions without DNA template, while for a positive
 control a mock community composed of evenly mixed DNA material
@@
 used. The sequences obtained in this study have been submitted to the
-European Nucleotide Archive (ENA) under accession numbers ??????????.
+European Nucleotide Archive (ENA) under accession numbers \textbf{TO BE
+ADDED LATER!}.
+
+\subsubsection{Sequence Analysis}\label{sequence-analysis}
 
@@
 (University Computing Center, University of Zagreb) using mothur
-(version 1.43.0)(Schloss \emph{et al.}, 2009) according to the MiSeq
+(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
 Standard Operating Procedure (MiSeq SOP;
 \url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
-and recommandations given from the Riffomonas project to enhance data
+and recommendations given from the Riffomonas project to enhance data
 reproducibility (\url{http://www.riffomonas.org/}). For alignment and
@@
 132; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
-Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplsts by
+Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
 mothur were exported, aligned using SILVA incremental aligner (SINA,
@@
 Ref NR 99 database (release 132) and imported into ARB (version 6.0.6)
-(Ludwig \emph{et al.}, 2004) for further phylogentic analysis using the
-same database. Reference sequences close to imported were selected and
-used to calculate a phylogentic tree using the Maximum Likelihood
+(Ludwig \emph{et al.}, 2004) for further phylogenetic analysis using the
+same database. Reference sequences close to imported ones were selected
+and used to calculate a phylogenetic tree using the Maximum Likelihood
 algorithm RAxML (version 7.0.3) with 1000 tree replicates (Stamatakis,
@@
 using the maximum parsimony criteria and not allowing changes to tree
-topology. Pipeline data processing and visualizaion was done using R
+topology. Pipeline data processing and visualization was done using R
 (version 3.6.0) (R Core Team, 2019), package tidyverse (version 1.2.1)
@@
 al.}, 2019). The detailed analysis procedure including the R Markdown
-file for this paper are available as a GitHub repository (??????).
+file for this paper are available as a GitHub repository (\textbf{TO BE
+ADDED LATER!}). Based on the ATCC MSA-1002 mock community included in
+the analysis a sequencing error rate of 0.009 \si{\percent} was
+determined, which is in line with previously reported values for
+next-generation sequencing data (Kozich \emph{et al.}, 2013; Schloss
+\emph{et al.}, 2016). In addition, the negative control processed
+together with the samples yielded only 2 sequences after sequence
+quality curation.
 
-\subsubsection{Protein Isolation and
-Metaproteomics}\label{protein-isolation-and-metaproteomics}
+\subsubsection{Protein Isolation}\label{protein-isolation}
 
@@
 During the first washing step the pellet was transferred to a clean 1.5
-\si{\ml} tube. After each washing step the samples were centrifuged at
+\si{\ml} tube. After each washing step samples were centrifuged at
 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets
@@
 
+\subsubsection{Metaproteomics}\label{metaproteomics}
+
+Isolated proteins were whole trypsin digested using the FASP
+(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
+according to the manufacturer's instructions (Wiśniewski \emph{et al.},
+2009) with small modifications. Before loading the solution to the
+column, protein pellets were solubilized in a urea sample buffer
+included in the kit amended with DTT (final concentration 100
+\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
+centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
+remove larger particles. The first washing step after protein solution
+loading was repeated twice. In addition, centrifugation steps were
+prolonged if the column was clogged. Trypsin digestion was performed on
+column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
+The final filtrate containing digested proteins was acidified with 1
+\si{\percent} (final concentration) trifluoroacetic acid, freezed at
+\num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent
+to the Vienna Metabolomics Center (University of Vienna) for
+metaproteomic analysis. \textbf{TO BE ADDED-VIENNA PART}.
+
 \subsubsection{Confocal Microscopy}\label{confocal-microscopy}
 
-\subsection{Results and Discussion}\label{results-and-discussion}
+Host leaves and thalli from DNA and protein isolation steps were washed
+seven times in deionized water and fixed with formaldehyde (final
+concentration \textasciitilde{} 3 \si{\percent}). In addition,
+nontreated leaves and thalli, washed three times in ASW to remove
+loosely attached microbial cells, were fixed in the same concentration
+of formaldehyde and used as a positive control. For long therm storage,
+fixed leaves and thalli were immersed in a mixture of phosphate-buffered
+saline (PBS) and ethanol (1:1) and stored at \num{-20}
+\si{\degreeCelsius}. Treated and untreated leaves and thalli segments
+were stained in a 2 × solution of SYBR Green I and examined under a
+Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
 
-\subsection{Conclusions}\label{conclusions}
+\subsection{Results}\label{results}
 
-\subsection{Materials and Methods}\label{materials-and-methods-1}
+\subsection{Discussion}\label{discussion}
+
+\subsection{Acknowledgements}\label{acknowledgements}
 
@@
 
-\newpage
-
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
+\subsection{References}\label{references}
 
@@
 
+\hypertarget{ref-Wisniewski2009}{}
+Wiśniewski, J.R., Zougman, A., Nagaraj, N., and Mann, M. (2009)
+Universal sample preparation method for proteome analysis. \emph{Nature
+Methods} \textbf{6}: 359--362.
+
 \hypertarget{ref-Xie2015}{}
@@
 
+\newpage
+
+\subsection{Figure Captions}\label{figure-captions}
+
+\subsection{Figures}\label{figures}
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\linewidth]{../results/figures/community_barplot}
+
+}
+
+\caption{Bacteril Community\label{community}}\label{fig:unnamed-chunk-2}
+\end{figure}
+
 \end{document}

### 2019-11-28 code/plot_community_barplot.R

@@ -1,0 +1,84 @@
+#################################################################################################################
+# plot_community_barplot.R
+# 
+# A script to plot the community structure of each sample.
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+# Produces: results/figures/community_barplot.jpg
+#
+#################################################################################################################
+
+# Loading input data containing sequence abundances and subsequent input data customization
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
+  filter(!str_detect(taxon, "^Eukaryota")) %>%
+  filter(taxon!="Root") %>%
+  filter(!str_detect(taxon, "^Mitochondria")) %>%
+  select(-ATCC_1, -NC_1) %>%
+  group_by(taxlevel) %>%
+  mutate_at(5:ncol(.), funs(. / sum(.) * 100)) %>%
+  ungroup()
+
+# Loading metadata of each sample and selection of desired information of each sample
+metadata <- read_tsv("data/raw/metadata.csv") %>%
+  filter(ID %in% colnames(community[, 6:ncol(community)])) %>%
+  select(ID, label) %>%
+  deframe()
+
+# Selection of groups for plotting
+plot <- filter(community, taxlevel==2 | str_detect(taxon, "Chloroplast$") | str_detect(taxon, "proteobacteria$")) %>%
+  filter_at(6:ncol(.), any_vars(. >= 1)) %>%
+  mutate_at(5:ncol(.), funs(case_when(taxon=="Cyanobacteria" ~ . - .[taxon=="Chloroplast"], TRUE ~ .))) %>%
+  filter(!str_detect(taxon, "^Proteobacteria$")) %>%
+  mutate(taxon=str_replace_all(taxon, c("unknown_unclassified"="No Relative", "unknown"="No Relative"))) %>%
+  filter_at(6:ncol(.), any_vars(. >= 1)) %>%
+  bind_rows(summarise_all(., funs(ifelse(is.numeric(.), 100-sum(.), paste("Other"))))) %>%
+  rename_at(names(metadata), ~unname(metadata)) %>%
+  arrange(-row_number())
+
+# Loading colors for each group on the plot
[…]

### 2019-11-29 submission/manuscript.Rmd
Optimized taxonomy plot.

@@
 
-```{r out.width="100%", fig.align="center", fig.cap="Bacteril Community\\label{community}", fig.pos="h"}
+```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant bacterial sequences on the surfaces of two marine macrophytes (\\textit{Cymodocea nodosa} and \\textit{Caulerpa cylindracea}) from two locations (Saline and Funtana) and in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{community}", fig.pos="h"}
 include_graphics("../results/figures/community_barplot.jpg")

### 2019-11-29 submission/manuscript.tex
Optimized taxonomy plot.

@@
 
-\caption{Bacteril Community\label{community}}\label{fig:unnamed-chunk-2}
+\caption{Taxonomic classification and relative contribution of the most abundant bacterial sequences on the surfaces of two marine macrophytes (\textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea}) from two locations (Saline and Funtana) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-2}
 \end{figure}

### 2019-11-29 code/plot_community_barplot.R

@@ -24,18 +24,21 @@ metadata <- read_tsv("data/raw/metadata.csv") %>%
   deframe()
 
 # Selection of groups for plotting
-plot <- filter(community, taxlevel==2 | str_detect(taxon, "Chloroplast$") | str_detect(taxon, "proteobacteria$")) %>%
+plot <- filter(community, taxlevel==2 |
+                 (taxlevel==4 & str_detect(taxon, "^Chloroplast$")) |
+                 (taxlevel==3 & str_detect(rankID, filter(community, str_detect(taxon, "^Proteobacteria$"))[[2]]))) %>%
   filter_at(6:ncol(.), any_vars(. >= 1)) %>%
   mutate_at(5:ncol(.), funs(case_when(taxon=="Cyanobacteria" ~ . - .[taxon=="Chloroplast"], TRUE ~ .))) %>%
-  filter(!str_detect(taxon, "^Proteobacteria$")) %>%
+  mutate_at(5:ncol(.), funs(case_when(taxon=="Proteobacteria" ~ . - sum(.[taxlevel==3 & str_detect(rankID, filter(community, str_detect(taxon, "^Proteobacteria$"))[[2]])]), TRUE ~ .))) %>%
+  mutate(taxon=str_replace(taxon, "Proteobacteria", "Other Proteobacteria")) %>%
   mutate(taxon=str_replace_all(taxon, c("unknown_unclassified"="No Relative", "unknown"="No Relative"))) %>%
   filter_at(6:ncol(.), any_vars(. >= 1)) %>%
   bind_rows(summarise_all(., funs(ifelse(is.numeric(.), 100-sum(.), paste("Other"))))) %>%
   rename_at(names(metadata), ~unname(metadata)) %>%
-  arrange(-row_number())
+  arrange(taxon %in% "No Relative")
 
 # Loading colors for each group on the plot
-color <- read_tsv("data/raw/group_colors.csv") %>%
+color <- read_tsv("data/raw/group_colors.csv", col_types=list(Taxlevel=col_skip())) %>%
   deframe()
 
 # Generation of italic names for groups
@@ -49,8 +52,8 @@ gather(plot, key="sample", value="abundance", 6:(ncol(plot))) %>%
   mutate(taxon=factor(taxon, levels=unique(plot$taxon))) %>%
   mutate(sample=factor(sample, levels=metadata)) %>%
   ggplot() +
[…]

### 2019-12-02 code/plot_community_barplot_taxlevel.R
Customizing .Rprofile

@@ -1,62 +1,0 @@
-#################################################################################################################
-# plot_community_barplot_taxlevel.R
-# 
-# A script to plot the community structure of each sample at every taxonomic level (Domain-Genus).
-# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
-# Produces: results/figures/community_barplot_domain.jpg
-#           results/figures/community_barplot_phylum.jpg
-#           results/figures/community_barplot_class.jpg
-#           results/figures/community_barplot_order.jpg
-#           results/figures/community_barplot_family.jpg
-#           results/figures/community_barplot_genus.jpg
-#################################################################################################################
-
-library(tidyverse)
-
-community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
-  filter(!str_detect(taxon, "^Eukaryota")) %>%
-  filter(taxon!="Root") %>%
-  filter(!str_detect(taxon, "^Mitochondria")) %>%
-  group_by(taxlevel) %>%
-  mutate_at(5:ncol(.), funs(. / sum(.) * 100)) %>%
-  ungroup()
-
-metadata <- read_tsv("data/raw/metadata.csv") %>%
-  filter(ID %in% colnames(community[, 6:ncol(community)])) %>%
-  select(ID, label) %>%
-  deframe()
-
-taxlevel <- c("domain", "phylum", "class", "order", "family", "genus")
-
-for (i in seq(1:6)) {
-plot <- filter(community, taxlevel==i) %>%
-  filter_at(6:ncol(.), any_vars(. >= 2)) %>%
-  mutate(taxon=str_replace_all(taxon, c("unknown_unclassified"="No Relative", "unknown"="No Relative"))) %>%
-  filter_at(6:ncol(.), any_vars(. >= 2)) %>%
-  bind_rows(summarise_all(., funs(ifelse(is.numeric(.), 100-sum(.), paste("Other"))))) %>%
-  rename_at(names(metadata), ~unname(metadata)) %>%
-  arrange(-row_number())
-
-color <- read_tsv("data/raw/group_colors.csv") %>%
-  deframe()
-
-gather(plot, key="sample", value="abundance", 6:(ncol(plot))) %>%
-  mutate(taxon=factor(taxon, levels=unique(plot$taxon))) %>%
[…]

### 2019-12-02 submission/manuscript.Rmd
Rendering optimization

--- 2019-11-28 submission/manuscript.Rmd
@@
 ## Figures
-
-```{r out.width="100%", fig.align="center", fig.cap="Bacteril Community\\label{community}", fig.pos="h"}
-include_graphics("../results/figures/community_barplot.jpg")
+```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant bacterial sequences on the surfaces of two marine macrophytes (\\textit{Cymodocea nodosa} and \\textit{Caulerpa cylindracea}) from two locations (Saline and Funtana) and in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{community}", fig.pos="h"}
+include_graphics("../results/figures/community_bar_plot.jpg")
 ```

### 2019-12-02 submission/manuscript.tex
Customizing .Rprofile
Rendering optimization

--- 2019-11-25 submission/manuscript.tex
@@
 \DeclareSIUnit\Molar{\textsc{m}}
+
+\usepackage{caption}
+\captionsetup{justification=justified}
 
@@
 
-\hypertarget{dna-isolation-illumina-16s-rrna-sequencing-and-sequences-analysis}{%
-\subsubsection{DNA Isolation, Illumina 16S rRNA Sequencing and Sequences
-Analysis}\label{dna-isolation-illumina-16s-rrna-sequencing-and-sequences-analysis}}
+\hypertarget{dna-isolation}{%
+\subsubsection{DNA Isolation}\label{dna-isolation}}
 
@@
 \si{\ul} of deionized water.
+
+\hypertarget{illumina-16s-rrna-sequencing}{%
+\subsubsection{Illumina 16S rRNA
+Sequencing}\label{illumina-16s-rrna-sequencing}}
 
@@
 used. The sequences obtained in this study have been submitted to the
-European Nucleotide Archive (ENA) under accession numbers ?????.
+European Nucleotide Archive (ENA) under accession numbers \textbf{TO BE
+ADDED LATER!}.
 
-\begin{verbatim}
-## Parsed with column specification:
-## cols(
-##   .default = col_double(),
-##   query = col_character(),
-##   reference = col_character()
-## )
-\end{verbatim}
-
-\begin{verbatim}
-## See spec(...) for full column specifications.
-\end{verbatim}
-
-\begin{verbatim}
-## Parsed with column specification:
-## cols(
-##   taxlevel = col_double(),
-##   rankID = col_character(),
-##   taxon = col_character(),
-##   daughterlevels = col_double(),
-##   total = col_double(),
-##   `40` = col_double(),
-##   `41` = col_double(),
-##   `42` = col_double(),
-##   `43` = col_double(),
-##   `61` = col_double(),
-##   `62` = col_double(),
-##   `63` = col_double(),
-##   ATCC_1 = col_double(),
-##   NC_1 = col_double()
-## )
-\end{verbatim}
+\hypertarget{sequence-analysis}{%
+\subsubsection{Sequence Analysis}\label{sequence-analysis}}
 
@@
 \url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
-and recommandations given from the Riffomonas project to enhance data
+and recommendations given from the Riffomonas project to enhance data
 reproducibility (\url{http://www.riffomonas.org/}). For alignment and
@@
 Ref NR 99 database (release 132) and imported into ARB (version 6.0.6)
-(Ludwig \emph{et al.}, 2004) for further phylogentic analysis using the
+(Ludwig \emph{et al.}, 2004) for further phylogenetic analysis using the
 same database. Reference sequences close to imported ones were selected
-and used to calculate a phylogentic tree using the Maximum Likelihood
+and used to calculate a phylogenetic tree using the Maximum Likelihood
 algorithm RAxML (version 7.0.3) with 1000 tree replicates (Stamatakis,
@@
 al.}, 2019). The detailed analysis procedure including the R Markdown
-file for this paper are available as a GitHub repository (?????). Based
-on the ATCC MSA-1002 mock community included in the analysis a
-sequencing error rate of 0.009 \si{\percent} was detemined that is in
-line with previously reported values for next-generation sequencing data
-(Kozich \emph{et al.}, 2013; Schloss \emph{et al.}, 2016). In addition,
-the negative control processed together with the samples yielded only
-44,050 sequences.
+file for this paper are available as a GitHub repository (\textbf{TO BE
+ADDED LATER!}). Based on the ATCC MSA-1002 mock community included in
+the analysis a sequencing error rate of 0.009 \si{\percent} was
+determined, which is in line with previously reported values for
+next-generation sequencing data (Kozich \emph{et al.}, 2013; Schloss
+\emph{et al.}, 2016). In addition, the negative control processed
+together with the samples yielded only 2 sequences after sequence
+quality curation.
 
-\hypertarget{protein-isolation-and-metaproteomics}{%
-\subsubsection{Protein Isolation and
-Metaproteomics}\label{protein-isolation-and-metaproteomics}}
+\hypertarget{protein-isolation}{%
+\subsubsection{Protein Isolation}\label{protein-isolation}}
 
@@
 During the first washing step the pellet was transferred to a clean 1.5
-\si{\ml} tube. After each washing step the samples were centrifuged at
+\si{\ml} tube. After each washing step samples were centrifuged at
 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets
 were stored at \num{-80} °C until further analysis.
+
+\hypertarget{metaproteomics}{%
+\subsubsection{Metaproteomics}\label{metaproteomics}}
+
+Isolated proteins were whole trypsin digested using the FASP
+(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
+according to the manufacturer's instructions (Wiśniewski \emph{et al.},
+2009) with small modifications. Before loading the solution to the
+column, protein pellets were solubilized in a urea sample buffer
+included in the kit amended with DTT (final concentration 100
+\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
+centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
+remove larger particles. The first washing step after protein solution
+loading was repeated twice. In addition, centrifugation steps were
+prolonged if the column was clogged. Trypsin digestion was performed on
+column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
+The final filtrate containing digested proteins was acidified with 1
+\si{\percent} (final concentration) trifluoroacetic acid, freezed at
+\num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent
+to the Vienna Metabolomics Center (University of Vienna) for
+metaproteomic analysis. \textbf{TO BE ADDED-VIENNA PART}.
 
@@
 
-\hypertarget{results-and-discussion}{%
-\subsection{Results and Discussion}\label{results-and-discussion}}
+Host leaves and thalli from DNA and protein isolation steps were washed
+seven times in deionized water and fixed with formaldehyde (final
+concentration \textasciitilde{} 3 \si{\percent}). In addition,
+nontreated leaves and thalli, washed three times in ASW to remove
+loosely attached microbial cells, were fixed in the same concentration
+of formaldehyde and used as a positive control. For long therm storage,
+fixed leaves and thalli were immersed in a mixture of phosphate-buffered
+saline (PBS) and ethanol (1:1) and stored at \num{-20}
+\si{\degreeCelsius}. Treated and untreated leaves and thalli segments
+were stained in a 2 × solution of SYBR Green I and examined under a
+Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
 
-\hypertarget{conclusions}{%
-\subsection{Conclusions}\label{conclusions}}
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
 
-\hypertarget{materials-and-methods-1}{%
-\subsection{Materials and Methods}\label{materials-and-methods-1}}
+\hypertarget{discussion}{%
+\subsection{Discussion}\label{discussion}}
 
-\newpage
+\hypertarget{acknowledgements}{%
+\subsection{Acknowledgements}\label{acknowledgements}}
 
@@
 \hypertarget{references}{%
-\subsection*{References}\label{references}}
-\addcontentsline{toc}{subsection}{References}
+\subsection{References}\label{references}}
 
@@
 
+\leavevmode\hypertarget{ref-Wisniewski2009}{}%
+Wiśniewski, J.R., Zougman, A., Nagaraj, N., and Mann, M. (2009)
+Universal sample preparation method for proteome analysis. \emph{Nature
+Methods} \textbf{6}: 359--362.
+
 \leavevmode\hypertarget{ref-Xie2015}{}%
@@
 
+\newpage
+
+\hypertarget{figure-captions}{%
+\subsection{Figure Captions}\label{figure-captions}}
+
+\hypertarget{figures}{%
+\subsection{Figures}\label{figures}}
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\linewidth]{../results/figures/community_bar_plot}
+
+}
+
+\caption{Taxonomic classification and relative contribution of the most abundant bacterial sequences on the surfaces of two marine macrophytes (\textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea}) from two locations (Saline and Funtana) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-2}
+\end{figure}
+
 \end{document}

### 2019-12-02 code/plot_community_bar_plot.R

@@ -1,9 +1,9 @@
 #################################################################################################################
-# plot_community_barplot.R
+# plot_community_bar_plot.R
 # 
 # A script to plot the community structure of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
-# Produces: results/figures/community_barplot.jpg
+# Produces: results/figures/community_bar_plot.jpg
 #
 #################################################################################################################
 
@@ -90,4 +90,4 @@ gather(plot, key="sample", value="abundance", 6:(ncol(plot))) %>%
  
 
 # Plot saving
-ggsave("results/figures/community_barplot.jpg", width=297, height=210, units="mm")
+ggsave("results/figures/community_bar_plot.jpg", width=297, height=210, units="mm")

### 2019-12-09 submission/manuscript.Rmd
Added confocal images.

--- 2019-11-28 submission/manuscript.Rmd
@@
 The final filtrate containing digested proteins was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid, freezed at \num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent to the Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis.
-**TO BE ADDED-VIENNA PART**.
+Peptides were resuspended in 1 \si{\percent} (final concentration) trifluoroacetic acid, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) accorind to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
+Obtaied MS/MS spectra were searched against a protein database from metagenomic assembly published in @Burke2011a using SEQUEST-HIT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
+The target-decoy approach was used to reduce the probability of false peptide identification.
+Results whose false discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
+For protein identification a minimum of of two peptides and one unique peptide were required.
+For protein quantification, a chromatographic peak area-based free quantitative method was applied.
 
@@
 \newpage 
+\setlength\parindent{0pt}
+
 ## Figure Captions
+**\autoref{micro_dec}.** \nameref{micro_dec}
 
+**\autoref{micro_jun}.** \nameref{micro_jun}
+
+**\autoref{community}.** \nameref{community}
+
+\newpage
 ## Figures
-
-```{r out.width="100%", fig.align="center", fig.cap="Bacteril Community\\label{community}", fig.pos="h"}
-include_graphics("../results/figures/community_barplot.jpg")
+```{r out.width="85%", fig.align="center", fig.cap="Confocal microscope images of \\textit{Cymodocea nodosa} and \\textit{Caulerpa cylindracea} surfaces from different locations (Saline, Funtana-Invaded and Funtana-Noninvaded) sampled on 4 December 2017 and stained with SYBR Green I.
+Scale bar in all images is 60 \\si{\\um}.\\label{micro_dec}", fig.pos="ht"}
+include_graphics("../results/images/confocal_december.jpg")
 ```
+\newpage
+```{r out.width="85%", fig.align="center", fig.cap="Confocal microscope images of \\textit{Cymodocea nodosa} and \\textit{Caulerpa cylindracea} surfaces from different locations (Funtana-Invaded and Funtana-Noninvaded) sampled on 19 June 2018 and stained with SYBR Green I.
+Scale bar in all images is 60 \\si{\\um}.\\label{micro_jun}", fig.pos="ht"}
+include_graphics("../results/images/confocal_june.jpg")
+```
+\newpage
+```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant bacterial sequences from surfaces of two marine macrophytes (\\textit{Cymodocea nodosa} and \\textit{Caulerpa cylindracea}) sampled in different locations (Saline, Funtana-Invaded and Funtana-Noninvaded) and in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{community}", fig.pos="ht"}
+include_graphics("../results/figures/community_bar_plot.jpg")
+```

### 2019-12-09 submission/manuscript.tex
Added confocal images.

--- 2019-11-28 submission/manuscript.tex
@@
 to the Vienna Metabolomics Center (University of Vienna) for
-metaproteomic analysis. \textbf{TO BE ADDED-VIENNA PART}.
+metaproteomic analysis. Peptides were resuspended in 1 \si{\percent}
+(final concentration) trifluoroacetic acid, desalted using the Pierce
+C18 Tips (Thermo Fisher Scientific, USA) accorind to the manufacturer's
+instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap
+Mass Spectrometer (Thermo Fisher Scientific, USA). Obtaied MS/MS spectra
+were searched against a protein database from metagenomic assembly
+published in Burke \emph{et al.} (2011a) using SEQUEST-HIT engines and
+validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher
+Scientific, USA). The target-decoy approach was used to reduce the
+probability of false peptide identification. Results \SI{< 1}{\percent}
+false discovery rate at the peptide level were kept. For protein
+identification a minimum of of two peptides and one unique peptide were
+required. For protein quantification, a chromatographic peak area-based
+free quantitative method was applied.
 
@@
 
+\setlength\parindent{0pt}
+
 \subsection{Figure Captions}\label{figure-captions}
+
+\textbf{\autoref{micro_dec}.} \nameref{micro_dec}
+
+\textbf{\autoref{micro_jun}.} \nameref{micro_jun}
+
+\textbf{\autoref{community}.} \nameref{community}
+
+\newpage
 
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 
-{\centering \includegraphics[width=1\linewidth]{../results/figures/community_barplot}
+{\centering \includegraphics[width=0.85\linewidth]{../results/images/confocal_december}
 
@@
 
-\caption{Bacteril Community\label{community}}\label{fig:unnamed-chunk-2}
+\caption{Confocal microscope images of \textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea} surfaces from different locations (Saline, Funtana-Invaded and Funtana-Noninvaded) sampled on 4 December 2017 and stained with SYBR Green I.
+Scale bar in all images is 60 \si{\um}.\label{micro_dec}}\label{fig:unnamed-chunk-2}
+\end{figure}\newpage
+
+\begin{figure}[ht]
+
+{\centering \includegraphics[width=0.85\linewidth]{../results/images/confocal_june}
+
+}
+
+\caption{Confocal microscope images of \textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea} surfaces from different locations (Funtana-Invaded and Funtana-Noninvaded) sampled on 19 June 2018 and stained with SYBR Green I.
+Scale bar in all images is 60 \si{\um}.\label{micro_jun}}\label{fig:unnamed-chunk-3}
+\end{figure}\newpage
+
+\begin{figure}[ht]
+
+{\centering \includegraphics[width=1\linewidth]{../results/figures/community_bar_plot}
+
+}
+
+\caption{Taxonomic classification and relative contribution of the most abundant bacterial sequences from surfaces of two marine macrophytes (\textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea}) sampled in different locations (Saline, Funtana-Invaded and Funtana-Noninvaded) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-4}
 \end{figure}

### 2019-12-09 code/plot_confocal_images.R

@@ -1,0 +1,99 @@
+#################################################################################################################
+# plot_confocal_images.R
+# 
+# A script to arrange and plot confocal images.
+# Dependencies: results/images/
+# Produces: results/images/confocal_images.jpg
+#
+#################################################################################################################
+
+library(gridExtra)
+library(magick)
+
+files <-c(scy_n_dec="14_03_2018_1_SCy-N-02a_ch00.tif",
+           scy_dna_dec="16_03_2018_1-SCy-dna1-02a_ch00.tif",
+           scy_prot_dec="14_03_2018_1_SCy-prot1-01a_ch00.tif",
+           fcym_n_dec="15_03_2018_1-FCyM-N-03a_ch00.tif",
+           fcym_dna_dec="15_03_2018_1-FCyM-dna1-ponovo-03a_ch00.tif",
+           fcym_prot_dec="15_03_2018_1-FCyM-prot1-02a_ch00.tif",
+           fcam_n_dec="14_03_2018_01_FCaM-N-02a_ch00.tif",
+           fcam_dna_dec="14_03_2018_01_FCaM-dna1-03a_ch00.tif",
+           fcam_prot_dec="16_03_2018_01_FCaM-prot1-02a_ch00.tif",
+           fca_n_dec="15_03_2018_01-FCa-N-01a_ch00.tif",
+           fca_dna_dec="16_03_2018_01_FCa-dna1-01a_ch00.tif",
+           fca_prot_dec="15_03_2018_01-FCa-prot1-02a_ch00.tif",
+           fcym_n_jun="30_08_2018_7mj-FCyM-N-2-03a_ch00.tif",
+           fcym_dna_jun="29_08_2018_FCyM-dna-1-01a_ch00.tif",
+           fcym_prot_jun="29_08_2018_FCyM-prot-1-0a_ch00.tif",
+           fcam_n_jun="29_08_2018_FCaM-N-02a_ch00.tif",
+           fcam_dna_jun="29_08_2018_FCaM-dna-01a_ch00.tif",
+           fcam_prot_jun="31_08_2018_7mj-FCaM-prot-03a_ch00.tif",
+           fca_n_jun="31_08_2018_7mj-FCa-N-01a_ch00.tif",
+           fca_dna_jun="31_08_2018_7mj-FCa-dna-01a_ch00.tif",
+           fca_dna_jun="31_08_2018_7mj-FCa-prot-01a_ch00.tif")
+
[…]

### 2019-12-17 submission/manuscript.Rmd
Paper optimization

@@
 The final filtrate containing digested proteins was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid, freezed at \num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent to the Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis.
-Peptides were resuspended in 1 \si{\percent} (final concentration) trifluoroacetic acid, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) accorind to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
-Obtaied MS/MS spectra were searched against a protein database from metagenomic assembly published in @Burke2011a using SEQUEST-HIT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
+Peptides were resuspended in 1 \si{\percent} (final concentration) trifluoroacetic acid, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
+Obtained MS/MS spectra were searched against a protein database from metagenomic assembly published in @Burke2011a using SEQUEST-HIT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
 The target-decoy approach was used to reduce the probability of false peptide identification.

### 2019-12-17 submission/manuscript.tex
Paper optimization

@@
 (final concentration) trifluoroacetic acid, desalted using the Pierce
-C18 Tips (Thermo Fisher Scientific, USA) accorind to the manufacturer's
+C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's
 instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap
-Mass Spectrometer (Thermo Fisher Scientific, USA). Obtaied MS/MS spectra
-were searched against a protein database from metagenomic assembly
-published in Burke \emph{et al.} (2011a) using SEQUEST-HIT engines and
-validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher
-Scientific, USA). The target-decoy approach was used to reduce the
-probability of false peptide identification. Results \SI{< 1}{\percent}
-false discovery rate at the peptide level were kept. For protein
-identification a minimum of of two peptides and one unique peptide were
-required. For protein quantification, a chromatographic peak area-based
-free quantitative method was applied.
+Mass Spectrometer (Thermo Fisher Scientific, USA). Obtained MS/MS
+spectra were searched against a protein database from metagenomic
+assembly published in Burke \emph{et al.} (2011a) using SEQUEST-HIT
+engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
+Fisher Scientific, USA). The target-decoy approach was used to reduce
+the probability of false peptide identification. Results whose false
+discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
+For protein identification a minimum of of two peptides and one unique
+peptide were required. For protein quantification, a chromatographic
+peak area-based free quantitative method was applied.
 

### 2019-12-17 code/plot_confocal_images.R
Code for confocal removed.

@@ -1,99 +1,0 @@
-#################################################################################################################
-# plot_confocal_images.R
-# 
-# A script to arrange and plot confocal images.
-# Dependencies: results/images/
-# Produces: results/images/confocal_images.jpg
-#
-#################################################################################################################
-
-library(gridExtra)
-library(magick)
-
-files <-c(scy_n_dec="14_03_2018_1_SCy-N-02a_ch00.tif",
-           scy_dna_dec="16_03_2018_1-SCy-dna1-02a_ch00.tif",
-           scy_prot_dec="14_03_2018_1_SCy-prot1-01a_ch00.tif",
-           fcym_n_dec="15_03_2018_1-FCyM-N-03a_ch00.tif",
-           fcym_dna_dec="15_03_2018_1-FCyM-dna1-ponovo-03a_ch00.tif",
-           fcym_prot_dec="15_03_2018_1-FCyM-prot1-02a_ch00.tif",
-           fcam_n_dec="14_03_2018_01_FCaM-N-02a_ch00.tif",
-           fcam_dna_dec="14_03_2018_01_FCaM-dna1-03a_ch00.tif",
-           fcam_prot_dec="16_03_2018_01_FCaM-prot1-02a_ch00.tif",
-           fca_n_dec="15_03_2018_01-FCa-N-01a_ch00.tif",
-           fca_dna_dec="16_03_2018_01_FCa-dna1-01a_ch00.tif",
-           fca_prot_dec="15_03_2018_01-FCa-prot1-02a_ch00.tif",
-           fcym_n_jun="30_08_2018_7mj-FCyM-N-2-03a_ch00.tif",
-           fcym_dna_jun="29_08_2018_FCyM-dna-1-01a_ch00.tif",
-           fcym_prot_jun="29_08_2018_FCyM-prot-1-0a_ch00.tif",
-           fcam_n_jun="29_08_2018_FCaM-N-02a_ch00.tif",
-           fcam_dna_jun="29_08_2018_FCaM-dna-01a_ch00.tif",
-           fcam_prot_jun="31_08_2018_7mj-FCaM-prot-03a_ch00.tif",
-           fca_n_jun="31_08_2018_7mj-FCa-N-01a_ch00.tif",
-           fca_dna_jun="31_08_2018_7mj-FCa-dna-01a_ch00.tif",
-           fca_dna_jun="31_08_2018_7mj-FCa-prot-01a_ch00.tif")
-
[…]

### 2020-02-04 submission/manuscript.Rmd
Silva datatbase switch to 138.

@@
 # Calculation of the number of sequences in the negative control
-nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
+nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
   filter(taxon=="Root") %>%
@@
 Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
-For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 132; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Sequences classified as chloroplasts by mothur were exported, aligned using SILVA incremental aligner (SINA, version 1.6.0) [@Pruesse2012] against the same SILVA SSU Ref NR 99 database (release 132) and imported into ARB (version 6.0.6) [@Ludwig2004] for further phylogenetic analysis using the same database.
+For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
+Sequences classified as chloroplasts by mothur were exported, aligned using SILVA incremental aligner (SINA, version 1.6.0) [@Pruesse2012] against the same SILVA SSU Ref NR 99 database (release 138) and imported into ARB (version 6.0.6) [@Ludwig2004] for further phylogenetic analysis using the same database.
 Reference sequences close to imported ones were selected and used to calculate a phylogenetic tree using the Maximum Likelihood algorithm RAxML (version 7.0.3) with 1000 tree replicates [@Stamatakis2006].

### 2020-02-04 submission/manuscript.tex
Silva datatbase switch to 138.

--- 2019-12-02 submission/manuscript.tex
@@
 classification of sequences the SILVA SSU Ref NR 99 database (release
-132; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
+138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
 Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
@@
 version 1.6.0) (Pruesse \emph{et al.}, 2012) against the same SILVA SSU
-Ref NR 99 database (release 132) and imported into ARB (version 6.0.6)
+Ref NR 99 database (release 138) and imported into ARB (version 6.0.6)
 (Ludwig \emph{et al.}, 2004) for further phylogenetic analysis using the
@@
 to the Vienna Metabolomics Center (University of Vienna) for
-metaproteomic analysis. \textbf{TO BE ADDED-VIENNA PART}.
+metaproteomic analysis. Peptides were resuspended in 1 \si{\percent}
+(final concentration) trifluoroacetic acid, desalted using the Pierce
+C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's
+instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap
+Mass Spectrometer (Thermo Fisher Scientific, USA). Obtained MS/MS
+spectra were searched against a protein database from metagenomic
+assembly published in Burke \emph{et al.} (2011a) using SEQUEST-HIT
+engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
+Fisher Scientific, USA). The target-decoy approach was used to reduce
+the probability of false peptide identification. Results whose false
+discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
+For protein identification a minimum of of two peptides and one unique
+peptide were required. For protein quantification, a chromatographic
+peak area-based free quantitative method was applied.
 
@@
 \newpage 
+\setlength\parindent{0pt}
 
@@
 
+\textbf{\autoref{micro_dec}.} \nameref{micro_dec}
+
+\textbf{\autoref{micro_jun}.} \nameref{micro_jun}
+
+\textbf{\autoref{community}.} \nameref{community}
+
+\newpage
+
 \hypertarget{figures}{%
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
+
+{\centering \includegraphics[width=0.85\linewidth]{../results/images/confocal_december}
+
+}
+
+\caption{Confocal microscope images of \textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea} surfaces from different locations (Saline, Funtana-Invaded and Funtana-Noninvaded) sampled on 4 December 2017 and stained with SYBR Green I.
+Scale bar in all images is 60 \si{\um}.\label{micro_dec}}\label{fig:unnamed-chunk-2}
+\end{figure}
+\newpage
+\begin{figure}[ht]
+
+{\centering \includegraphics[width=0.85\linewidth]{../results/images/confocal_june}
+
+}
+
+\caption{Confocal microscope images of \textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea} surfaces from different locations (Funtana-Invaded and Funtana-Noninvaded) sampled on 19 June 2018 and stained with SYBR Green I.
+Scale bar in all images is 60 \si{\um}.\label{micro_jun}}\label{fig:unnamed-chunk-3}
+\end{figure}
+\newpage
+\begin{figure}[ht]
 
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant bacterial sequences on the surfaces of two marine macrophytes (\textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea}) from two locations (Saline and Funtana) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-2}
+\caption{Taxonomic classification and relative contribution of the most abundant bacterial sequences from surfaces of two marine macrophytes (\textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea}) sampled in different locations (Saline, Funtana-Invaded and Funtana-Noninvaded) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-4}
 \end{figure}

### 2020-02-04 code/format_taxonomy.R

@@ -1,4 +1,4 @@
-map.in <- read.table("data/references/tax_slv_ssu_132.txt", header=F,sep="\t", stringsAsFactors=F)
+map.in <- read.table("data/references/tax_slv_ssu_138.txt", header=F,sep="\t", stringsAsFactors=F)
 map.in <- map.in[, c(1,3)]
 colnames(map.in) <- c("taxlabel", "taxlevel")
 taxlevels <- c("root", "domain", "major_clade", "superkingdom", "kingdom", "subkingdom", "infrakingdom", "superphylum", "phylum", "subphylum", "infraphylum", "superclass", "class", "subclass", "infraclass", "superorder", "order", "suborder", "superfamily", "family", "subfamily", "genus")
@@ -39,19 +39,12 @@ for(i in 1:nrow(tax.mat)) {
 map.in$taxout <- gsub(" ", "_", map.in$taxout)
 
 # Bring in the old taxonomic levels from SILVA and remap them using the new levels
-tax.in <- read.table("data/references/silva.nr_v132.full", header=F, stringsAsFactors=F, sep="\t")
+tax.in <- read.table("data/references/silva.nr_v138.full", header=F, stringsAsFactors=F, sep="\t")
 colnames(tax.in) <- c("taxid", "taxlabel")
 
-# Problem corrections observed in the SILVA ARB file (silva.nr_v132.full)
-# Following line corrects the Bacteria;Bacteroidetes;Bacteroidia;Flavobacteriales;Flavobacteriaceae;Polaribacter;Polaribacter; problem
-tax.in$taxlabel <- gsub("Polaribacter;Polaribacter", "Polaribacter", tax.in$taxlabel)
-# Following line corrects the Bacteria;RsaHf231; problem
-tax.in$taxlabel <- gsub("RsaHf231;", "RsaHF231;", tax.in$taxlabel)
-# Following line corrects the Bacteria;GBS-1;; problem
-tax.in$taxlabel <- gsub("Bacteria;GBS-1;;", "Bacteria;GBS-1;", tax.in$taxlabel)
-# Following line corrects the Eukaryota;Opisthokonta;Nucletmycea;Fungi;Dikarya;Basidiomycota;Agaricomycotina;Tremellomycetes;Tremellales;Incertae sedis;Sterigmatosporidium; problem
[…]

### 2020-02-04 code/plot_calculators.R

@@ -1,55 +1,0 @@
-#################################################################################################################
-# plot_calculators.R
-# 
-# A script to plot richness and diversity calculators.
-# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.groups.ave-std.summary
-# Produces: results/figures/calculators.jpg
-#           
-#################################################################################################################
-
-library(tidyverse)
-library(gridExtra)
-
-alpha <- read_tsv(file="data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.groups.ave-std.summary",
-                 col_types=cols(group=col_character())) %>%
-  filter(method=='ave')
-  
-metadata <- read_tsv("data/raw/metadata.csv")
-
-meta_alpha <- inner_join(metadata, alpha, by=c("ID"="group"))
-
-plot <- meta_alpha %>%
-    filter(!is.na(depth)) %>%
-    group_by(label) %>%
-    summarize(mean_sobs=mean(sobs),
-              mean_chao=mean(chao),
-              mean_ace=mean(ace),
-              mean_shannon=mean(shannon),
-              mean_invsimpson=mean(invsimpson))
-
-sobs <- ggplot(plot, aes(x=depth, y=mean_sobs)) +
-  geom_boxplot(notch=TRUE) +
-  labs(x="Depth", y="Number of OTUs") +
-  theme_classic()
-chao <- ggplot(plot, aes(x=depth, y=mean_chao)) +
-  geom_boxplot(notch=TRUE) +
-  labs(x="Depth", y="Number of OTUs") +
-  theme_classic()
-ace <- ggplot(plot, aes(x=depth, y=mean_ace)) +
-  geom_boxplot(notch=TRUE) +
-  labs(x="Depth", y="Number of OTUs") +
-  theme_classic()
-shannon <- ggplot(plot, aes(x=depth, y=mean_shannon)) +
-  geom_boxplot(notch=TRUE) +
-  labs(x="Depth", y="Shannon Diversity Index") +
-  theme_classic()
-invsimpson <- ggplot(plot, aes(x=depth, y=mean_invsimpson)) +
-  geom_boxplot(notch=TRUE) +
-  labs(x="Depth", y="Inverse Simpson Diversity Index") +
-  theme_classic()
[…]

### 2020-02-04 code/plot_community_bar_plot.R

@@ -2,7 +2,7 @@
 # plot_community_bar_plot.R
 # 
 # A script to plot the community structure of each sample.
-# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary
+# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary
 #               data/raw/metadata.csv
 #               data/raw/group_colors.csv
 # Produces: results/figures/community_bar_plot.jpg
@@ -10,13 +10,13 @@
 #################################################################################################################
 
 # Loading input data containing sequence abundances and subsequent input data customization
-community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v132.wang.tax.summary") %>%
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
   filter(!str_detect(taxon, "^Eukaryota")) %>%
   filter(taxon!="Root")
 
 # Remove Mitochondria sequences and subtract their number from higher taxonomic levels to which they belong
 mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
-community <- mutate_at(community, 5:ncol(community), funs(case_when(
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
     rankID==str_extract(mitochondria, "(\\d+\\.){4}\\d+") ~ . - .[taxon=="Mitochondria"],
     rankID==str_extract(mitochondria, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Mitochondria"],
     rankID==str_extract(mitochondria, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Mitochondria"],
@@ -25,7 +25,7 @@ community <- mutate_at(community, 5:ncol(community), funs(case_when(
   filter(!str_detect(taxon, "^Mitochondria")) %>%
   select(-ATCC_1, -NC_1) %>%
   group_by(taxlevel) %>%
-  mutate_at(5:ncol(.), funs(. / sum(.) * 100)) %>%
[…]

### 2020-02-04 code/plot_pcoa.R

@@ -1,27 +1,0 @@
-#################################################################################################################
-# plot_pcoa.R
-# 
-# A script to generate the PCoA figure.
-# Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.braycurtis.0.03.lt.ave.pcoa.axes
-#               data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.braycurtis.0.03.lt.ave.pcoa.loadings
-#               data/raw/metadata.csv
-# Produces: results/figures/pcoa_figure.jpg
-#
-#################################################################################################################
-
-library(tidyverse)
-pcoa <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.braycurtis.0.03.lt.ave.pcoa.axes", col_types=cols(group = col_character()))
-metadata <- read_tsv("data/raw/metadata.csv")
-metadata_pcoa <- inner_join(metadata, pcoa, by=c('ID'='group'))
-
-pcoa_loadings <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.braycurtis.0.03.lt.ave.pcoa.loadings")
-
-ggplot(metadata_pcoa, aes(x=axis1, y=axis2)) +
-  geom_point(shape=19, size=2) +
-  geom_text(aes(label=label), nudge_x=0.07) +
-  coord_fixed() + 
-  labs(x=paste("PCoA 1 (", format(round(pcoa_loadings$loading[1], 2), nsmall=2), " %)"),
-       y=paste("PCoA 2 (", format(round(pcoa_loadings$loading[2], 2), nsmall=2), " %)")) +
-  theme_classic()
-
-ggsave("results/figures/pcoa_figure.jpg", width=297, height=210, units="mm")

### 2020-02-06 submission/manuscript.Rmd
Geographic coordinates setup.

@@
 ### Sampling
-Leaves of *Cymodocea nodosa* were sampled in a *Cymodocea nodosa* meadow in the Bay of Saline (45°7´5˝N, 13°37´20˝E) and in a *Cymodocea nodosa* meadow invaded by *Caulerpa cylindracea* in the proximity of the village of Funtana (45°10´39˝N, 13°35´42˝E).
+Leaves of *Cymodocea nodosa* were sampled in a *Cymodocea nodosa* meadow in the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E) and in a *Cymodocea nodosa* meadow invaded by *Caulerpa cylindracea* in the proximity of the village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E).
 Thalli of *Caulerpa cylindracea* were sampled in the same *Cymodocea nodosa* invaded meadow in Funtana and on a locality of only *Caulerpa cylindracea* located in the proximity of the invaded meadow.
@@
 DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} °C and centrifuging at 16,000 × g and 4 \si{\degreeCelsius} for 20 \si{\minute}.
-The pellet was washed twice with 1 \si{\ml} of chilled isopropanol and centrifuged after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 10 \si{\minute}.
+The pellet was washed twice with 1 \si{\ml} of chilled 70 \si{\percent} ethanol and centrifuged after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 10 \si{\minute}.
 After the first washing step duplicate pellets form the same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.

### 2020-02-06 submission/manuscript.tex
Geographic coordinates setup.

--- 2019-12-17 submission/manuscript.tex
@@
 Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
-nodosa} meadow in the Bay of Saline (45°7´5˝N, 13°37´20˝E) and in a
-\emph{Cymodocea nodosa} meadow invaded by \emph{Caulerpa cylindracea} in
-the proximity of the village of Funtana (45°10´39˝N, 13°35´42˝E). Thalli
-of \emph{Caulerpa cylindracea} were sampled in the same \emph{Cymodocea
-nodosa} invaded meadow in Funtana and on a locality of only
-\emph{Caulerpa cylindracea} located in the proximity of the invaded
-meadow. Leaves and thalli were collected on the same day in two
-contrasting seasons, on 4 December 2017 and 18 June 2018. During spring
-2018 the \emph{Cymodocea nodosa} meadow in the Bay of Saline decayed to
-an extent that no leaves could be retrieved (Najdek \emph{et al.},
-unpublished data). Leaves and thalli were collected by diving and
+nodosa} meadow in the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E)
+and in a \emph{Cymodocea nodosa} meadow invaded by \emph{Caulerpa
+cylindracea} in the proximity of the village of Funtana (\ang{45;10;39}
+N, \ang{13;35;42} E). Thalli of \emph{Caulerpa cylindracea} were sampled
+in the same \emph{Cymodocea nodosa} invaded meadow in Funtana and on a
+locality of only \emph{Caulerpa cylindracea} located in the proximity of
+the invaded meadow. Leaves and thalli were collected on the same day in
+two contrasting seasons, on 4 December 2017 and 18 June 2018. During
+spring 2018 the \emph{Cymodocea nodosa} meadow in the Bay of Saline
+decayed to an extent that no leaves could be retrieved (Najdek \emph{et
+al.}, unpublished data). Leaves and thalli were collected by diving and
 transported to the laboratory in containers placed on ice and filled
@@
 for 20 \si{\minute}. The pellet was washed twice with 1 \si{\ml} of
-chilled isopropanol and centrifuged after each washing step at 20,000 ×
-g and 4 \si{\degreeCelsius} for 10 \si{\minute}. After the first washing
-step duplicate pellets form the same sample were pooled and transferred
-to a clean 1.5 \si{\ml} tube. The dried pellet was resuspended in 100
-\si{\ul} of deionized water.
+chilled 70 \si{\percent} ethanol and centrifuged after each washing step
+at 20,000 × g and 4 \si{\degreeCelsius} for 10 \si{\minute}. After the
+first washing step duplicate pellets form the same sample were pooled
+and transferred to a clean 1.5 \si{\ml} tube. The dried pellet was
+resuspended in 100 \si{\ul} of deionized water.
 

### 2020-04-01 submission/manuscript.tex
Process optimization.

--- 2020-02-04 submission/manuscript.tex
@@
 \renewcommand{\thefigure}{\arabic{figure}}
+\renewcommand{\figurename}{Fig.}
 \captionsetup{justification=raggedright,singlelinecheck=false}
@@
 \DeclareSIUnit\Molar{\textsc{m}}
+\DeclareSIUnit\cells{\text{cells}}
 
@@
 \captionsetup{justification=justified}
+
+\usepackage{float}
+
+\usepackage{xr}
+\externaldocument[supp-]{supplementary}
+
+\usepackage{txfonts}
+
+\renewcommand{\figureautorefname}{Fig.}
 
@@
 the surface community and was used for 16S rRNA gene clone libraries
-construction (Burke \emph{et al.}, 2011b) and metagenomes sequencing
-(Burke \emph{et al.}, 2011a). This method, thought providing a selective
-isolation procedure, is using in the extraction buffer a rapid
-multienzyme cleaner (3M) that is not worldwide available and whose
-composition is not know (Burke \emph{et al.}, 2009). Also to our
-knowledge, no selective isolation protocol for proteins from epiphytic
-communities inhabiting marine macrophytes was established.
+construction (Burke and Thomas \emph{et al.}, 2011) and metagenomes
+sequencing (Burke and Peter Steinberg \emph{et al.}, 2011). This method,
+thought providing a selective isolation procedure, is using in the
+extraction buffer a rapid multienzyme cleaner (3M) that is not worldwide
+available and whose composition is not know (Burke \emph{et al.}, 2009).
+Also to our knowledge, no selective isolation protocol for proteins from
+epiphytic communities inhabiting marine macrophytes was established.
 
@@
 spectra were searched against a protein database from metagenomic
-assembly published in Burke \emph{et al.} (2011a) using SEQUEST-HIT
-engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
-Fisher Scientific, USA). The target-decoy approach was used to reduce
-the probability of false peptide identification. Results whose false
-discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
-For protein identification a minimum of of two peptides and one unique
-peptide were required. For protein quantification, a chromatographic
-peak area-based free quantitative method was applied.
+assembly published in Burke and Peter Steinberg \emph{et al.} (2011)
+using SEQUEST-HIT engines and validated with Percolator in Proteome
+Discoverer 2.1 (Thermo Fisher Scientific, USA). The target-decoy
+approach was used to reduce the probability of false peptide
+identification. Results whose false discovery rate at the peptide level
+was \SI{< 1}{\percent} were kept. For protein identification a minimum
+of of two peptides and one unique peptide were required. For protein
+quantification, a chromatographic peak area-based free quantitative
+method was applied.
 
@@
 Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
-(2011a) Bacterial community assembly based on functional genes rather
+(2011) Bacterial community assembly based on functional genes rather
 than species. \emph{Proceedings of the National Academy of Sciences of
@@
 Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
-(2011b) Composition, uniqueness and variability of the epiphytic
+(2011) Composition, uniqueness and variability of the epiphytic
 bacterial community of the green alga Ulva australis. \emph{The ISME
@@
 Xie, Y. (2014) knitr: A Comprehensive Tool for Reproducible Research in
-\{R\}. In, Stodden,V., Leisch,F., and Peng,R.D. (eds),
-\emph{Implementing reproducible computational research}. Chapman;
-Hall/CRC.
+\{R\}. In \emph{Implementing reproducible computational research}.
+Stodden, V., Leisch, F., and Peng, R.D. (eds). Chapman; Hall/CRC.
 

### 2020-04-06 submission/manuscript.Rmd
Added supplementary information.

@@
 
-\vspace{80mm}
+\vspace{20mm}
+Marino Korlević^1$*$^, Marsej Markovski^1^, Zihao Zhao^2^, Gerhard J. Herndl^2,3^, Mirjana Najdek^1^
 
-^1$\dagger$^
+1\. Center for Marine Research, Ruđer Bošković Institute, Croatia
 
-\vspace{40mm}
+2\. Department of Functional and Evolutionary Ecology, University of Vienna, Austria
 
-$\dagger$ To whom correspondence should be addressed: marino.korlevic@irb.hr
+3\. Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands
 
-1\. Ruđer Bošković Institute, Center for Marine Research, G. Paliaga 5, Rovinj, Croatia
+^$*$^To whom correspondence should be addressed:
 
-2\. University of Vienna, Department of Limnology and Bio-Oceanography, Althanstraße 14, Vienna, Austria
+Marino Korlević
+
+G. Paliaga 5, 52210 Rovinj, Croatia
+
+Tel.: +385 52 804 768
+
+Fax: +385 52 804 780
+
+e-mail: marino.korlevic@irb.hr
+
+Running title: DNA and protein isolation from macrophyte surfaces
+
 \newpage
@@
 
-## Abstract
+## Summary
 
@@
 
+## Results
+
+## Discussion
+
 \newpage
-## Materials and Methods
+## Experimental procedures
 ### Sampling
@@
 
-### DNA Isolation
+### DNA isolation
 The DNA was isolated according to the protocol for isolation from filters described in @Massana1997.
@@
 
-### Illumina 16S rRNA Sequencing
+### Illumina 16S rRNA sequencing
 An aliquot of isolated DNA was treated with RNase A (final concentration 200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}.
@@
 
-### Sequence Analysis
+### Sequence analysis
 ```{r}
@@
 
-### Protein Isolation
+### Protein isolation
 Proteins were isolated according to the protocol for isolation from soil described in @Chourey2010 and modified by @Hultman2015.
@@
 
-### Confocal Microscopy
+### Confocal microscopy
 Host leaves and thalli from DNA and protein isolation steps were washed seven times in deionized water and fixed with formaldehyde (final concentration ~ 3 \si{\percent}).
@@
 Treated and untreated leaves and thalli segments were stained in a 2 × solution of SYBR Green I and examined under a Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
-
-## Results
-
-## Discussion
 
@@
 ## Figures
-```{r out.width="85%", fig.align="center", fig.cap="Confocal microscope images of \\textit{Cymodocea nodosa} and \\textit{Caulerpa cylindracea} surfaces from different locations (Saline, Funtana-Invaded and Funtana-Noninvaded) sampled on 4 December 2017 and stained with SYBR Green I.
+```{r out.width="85%", fig.align="center", fig.cap="Confocal microscope images of \\textit{C.
+nodosa} and \\textit{C.
+cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 4 December 2017 and stained with SYBR Green I.
 Scale bar in all images is 60 \\si{\\um}.\\label{micro_dec}", fig.pos="ht"}
@@
 ```
+
 \newpage
-```{r out.width="85%", fig.align="center", fig.cap="Confocal microscope images of \\textit{Cymodocea nodosa} and \\textit{Caulerpa cylindracea} surfaces from different locations (Funtana-Invaded and Funtana-Noninvaded) sampled on 19 June 2018 and stained with SYBR Green I.
+```{r out.width="85%", fig.align="center", fig.cap="Confocal microscope images of \\textit{C.
+nodosa} and \\textit{C.
+cylindracea} surfaces from the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 19 June 2018 and stained with SYBR Green I.
 Scale bar in all images is 60 \\si{\\um}.\\label{micro_jun}", fig.pos="ht"}
@@
 ```
+
 \newpage
-```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant bacterial sequences from surfaces of two marine macrophytes (\\textit{Cymodocea nodosa} and \\textit{Caulerpa cylindracea}) sampled in different locations (Saline, Funtana-Invaded and Funtana-Noninvaded) and in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{community}", fig.pos="ht"}
+```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\\textit{C.
+nodosa} and \\textit{C.
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{community}", fig.pos="ht"}
 include_graphics("../results/figures/community_bar_plot.jpg")

### 2020-04-06 submission/manuscript.tex
Added supplementary information.

@@
 
-\vspace{80mm}
+\vspace{20mm}
 
-\textsuperscript{1\(\dagger\)}
+Marino Korlević\textsuperscript{1\(*\)}, Marsej
+Markovski\textsuperscript{1}, Zihao Zhao\textsuperscript{2}, Gerhard J.
+Herndl\textsuperscript{2,3}, Mirjana Najdek\textsuperscript{1}
 
-\vspace{40mm}
+1. Center for Marine Research, Ruđer Bošković Institute, Croatia
 
-\(\dagger\) To whom correspondence should be addressed:
+2. Department of Functional and Evolutionary Ecology, University of
+Vienna, Austria
+
+3. Department of Marine Microbiology and Biogeochemistry, Royal
+Netherlands Institute for Sea Research, Utrecht University, The
+Netherlands
+
+\textsuperscript{\(*\)}To whom correspondence should be addressed:
+
+Marino Korlević
+
+G. Paliaga 5, 52210 Rovinj, Croatia
+
+Tel.: +385 52 804 768
+
+Fax: +385 52 804 780
+
+e-mail:
 \href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}
 
-1. Ruđer Bošković Institute, Center for Marine Research, G. Paliaga 5,
-Rovinj, Croatia
+Running title: DNA and protein isolation from macrophyte surfaces
 
-2. University of Vienna, Department of Limnology and Bio-Oceanography,
-Althanstraße 14, Vienna, Austria \newpage \linenumbers
-\sisetup{mode=text} \setlength\parindent{24pt}
+\newpage
+\linenumbers
+\sisetup{mode=text}
+\setlength\parindent{24pt}
 
-\hypertarget{abstract}{%
-\subsection{Abstract}\label{abstract}}
+\hypertarget{summary}{%
+\subsection{Summary}\label{summary}}
 
@@
 
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
+
+\hypertarget{discussion}{%
+\subsection{Discussion}\label{discussion}}
+
 \newpage
 
-\hypertarget{materials-and-methods}{%
-\subsection{Materials and Methods}\label{materials-and-methods}}
+\hypertarget{experimental-procedures}{%
+\subsection{Experimental procedures}\label{experimental-procedures}}
 
@@
 \hypertarget{dna-isolation}{%
-\subsubsection{DNA Isolation}\label{dna-isolation}}
+\subsubsection{DNA isolation}\label{dna-isolation}}
 
@@
 \subsubsection{Illumina 16S rRNA
-Sequencing}\label{illumina-16s-rrna-sequencing}}
+sequencing}\label{illumina-16s-rrna-sequencing}}
 
@@
 \hypertarget{sequence-analysis}{%
-\subsubsection{Sequence Analysis}\label{sequence-analysis}}
+\subsubsection{Sequence analysis}\label{sequence-analysis}}
 
@@
 \hypertarget{protein-isolation}{%
-\subsubsection{Protein Isolation}\label{protein-isolation}}
+\subsubsection{Protein isolation}\label{protein-isolation}}
 
@@
 \hypertarget{confocal-microscopy}{%
-\subsubsection{Confocal Microscopy}\label{confocal-microscopy}}
+\subsubsection{Confocal microscopy}\label{confocal-microscopy}}
 
@@
 Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
-
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
-
-\hypertarget{discussion}{%
-\subsection{Discussion}\label{discussion}}
 
@@
 
-\caption{Confocal microscope images of \textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea} surfaces from different locations (Saline, Funtana-Invaded and Funtana-Noninvaded) sampled on 4 December 2017 and stained with SYBR Green I.
+\caption{Confocal microscope images of \textit{C.
+nodosa} and \textit{C.
+cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 4 December 2017 and stained with SYBR Green I.
 Scale bar in all images is 60 \si{\um}.\label{micro_dec}}\label{fig:unnamed-chunk-2}
 \end{figure}
+
 \newpage
@@
 
-\caption{Confocal microscope images of \textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea} surfaces from different locations (Funtana-Invaded and Funtana-Noninvaded) sampled on 19 June 2018 and stained with SYBR Green I.
+\caption{Confocal microscope images of \textit{C.
+nodosa} and \textit{C.
+cylindracea} surfaces from the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 19 June 2018 and stained with SYBR Green I.
 Scale bar in all images is 60 \si{\um}.\label{micro_jun}}\label{fig:unnamed-chunk-3}
 \end{figure}
+
 \newpage
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant bacterial sequences from surfaces of two marine macrophytes (\textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea}) sampled in different locations (Saline, Funtana-Invaded and Funtana-Noninvaded) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-4}
+\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\textit{C.
+nodosa} and \textit{C.
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-4}
 \end{figure}

### 2020-04-06 submission/supplementary.Rmd
Added supplementary information.

---
title: "**Supplementary Information**"
subtitle: "**Selective DNA and Protein Isolation from Marine Macrophyte Surfaces**"
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: header_supplementary.tex
fontsize: 12pt
geometry: margin=1.0in
csl: citation_style.csl #Get themes at https://github.com/citation-style-language/styles
bibliography: references.bib
---

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE, message=FALSE}
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)
opts_chunk$set("message" = FALSE)

inline_hook <- function(x){
	print(x)

	if(is.list(x)){
		x <- unlist(x)
	}

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)

library("kableExtra")
```

\vspace{60mm}
Marino Korlević^1$*$^, Marsej Markovski^1^, Zihao Zhao^2^, Gerhard J. Herndl^2,3^, Mirjana Najdek^1^

\vspace{40mm}

^$*$^To whom correspondence should be addressed: marino.korlevic@irb.hr

1\. Center for Marine Research, Ruđer Bošković Institute, Croatia

2\. Department of Functional and Evolutionary Ecology, University of Vienna, Austria

3\. Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands

\sisetup{mode=text}
\setlength\parindent{24pt}

## Supplementary Figures

```{r out.width="80%", fig.align="center", fig.cap="Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes sampled in the Bay of Saline (\\textit{C. nodosa} [Monospecific Settlement]) and the Bay of Funatana (\\textit{C. nodosa} [Mixed Settlement] and \\textit{C. cylindracea} [Mixed and Monospecific Settlement]) in two contrasting seasons.\\label{rarefaction}", fig.pos="H"}
include_graphics("../results/figures/rarefaction.jpg")
```

\newpage
## Supplementary Table

```{r nseq_notus}
# Loading OTU/sample data
shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")

# Calculating the number of sequences and OTUs in each sample
nseq_notus <- shared %>%
  mutate(nseq=rowSums(select(., starts_with("Otu")))) %>%
  mutate(notus=rowSums(select(., starts_with("Otu"))!=0)) %>%
  select(Group, nseq, notus) %>%
  mutate(Group=as.character(Group))

# Loading metadata 
metadata <- read_tsv("data/raw/metadata.csv")

# Joining metadata with the number of sequences and OTUs
invisible(Sys.setlocale(locale="en_GB.utf8"))
metadata_nseq_notus <- inner_join(metadata, nseq_notus, by=c("ID"="Group")) %>%
  select(ID, station, date, nseq, notus) %>%
  mutate(type=station) %>%
  select(ID, station, type, everything()) %>%
  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
  mutate(date=format(date, "%d %B %Y")) %>%
  mutate(date=str_replace(date, "^0", "")) %>%
  arrange(ID) %>%
  mutate(type=str_replace(type, "^SCy$", "\\\\textit{Cymodocea nodosa} (Monospecific)")) %>%
  mutate(type=str_replace(type, "^FCyM$", "\\\\textit{Cymodocea nodosa} (Mixed)")) %>%
  mutate(type=str_replace(type, "^FCaM$", "\\\\textit{Caulerpa cylindracea} (Mixed)")) %>%
  mutate(type=str_replace(type, "^FCa$", "\\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
  mutate(station=str_replace(station, "^SCy$", "Saline")) %>%
  mutate(station=str_replace(station, "^FCyM$", "Funtana")) %>%
  mutate(station=str_replace(station, "^FCaM$", "Funtana")) %>%
  mutate(station=str_replace(station, "^FCa$", "Funtana")) %>%
  mutate(nseq=format(nseq, big.mark=","), notus=format(notus, big.mark=","))

# Generating the table
kable(metadata_nseq_notus, format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=c("Sample ID", "Station", "Community Type", "Date", "No. of Sequences", "No. of OTUs"), caption="Sample ID, Station, Community Type, Sampling Date, No. of Sequences and No. of OTUs of each sample. No. of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\\label{nseq_notus}", align="c") %>%
  kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
  row_spec(0, bold=TRUE) %>%
  column_spec(1, width="6em")
```

### 2020-04-06 submission/supplementary.tex
Added supplementary information.

--- 2020-02-06 submission/manuscript.tex
@@
 \hypersetup{unicode=true,
-            pdftitle={Selective DNA and Protein Isolation from Marine Macrophyte Surfaces},
+            pdftitle={Supplementary Information},
             pdfborder={0 0 0},
@@
 
-  \title{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
-Surfaces}}
+  \title{\textbf{Supplementary Information}}
     \pretitle{\vspace{\droptitle}\centering\huge}
   \posttitle{\par}
+  \subtitle{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
+Surfaces}}
   \author{}
@@
 \usepackage[labelsep=period, labelfont=bf]{caption}
-\renewcommand{\thefigure}{\arabic{figure}}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\figurename}{Fig.}
+\renewcommand{\thetable}{S\arabic{table}}
 \captionsetup{justification=raggedright,singlelinecheck=false}
@@
 
+\usepackage{float}
+
+\renewcommand{\figureautorefname}{Fig.}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+
 \begin{document}
@@
 
-\vspace{80mm}
+\vspace{60mm}
 
-\textsuperscript{1\(\dagger\)}
+Marino Korlević\textsuperscript{1\(*\)}, Marsej
+Markovski\textsuperscript{1}, Zihao Zhao\textsuperscript{2}, Gerhard J.
+Herndl\textsuperscript{2,3}, Mirjana Najdek\textsuperscript{1}
 
@@
 
-\(\dagger\) To whom correspondence should be addressed:
+\textsuperscript{\(*\)}To whom correspondence should be addressed:
 \href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}
 
-1. Ruđer Bošković Institute, Center for Marine Research, G. Paliaga 5,
-Rovinj, Croatia
+1. Center for Marine Research, Ruđer Bošković Institute, Croatia
 
-2. University of Vienna, Department of Limnology and Bio-Oceanography,
-Althanstraße 14, Vienna, Austria \newpage
-\linenumbers
-\sisetup{mode=text} \setlength\parindent{24pt}
+2. Department of Functional and Evolutionary Ecology, University of
+Vienna, Austria
 
-\subsection{Abstract}\label{abstract}
+3. Department of Marine Microbiology and Biogeochemistry, Royal
+Netherlands Institute for Sea Research, Utrecht University, The
+Netherlands
+
+\sisetup{mode=text}
+\setlength\parindent{24pt}
+
+\hypertarget{supplementary-figures}{%
+\subsection{Supplementary Figures}\label{supplementary-figures}}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth]{../results/figures/rarefaction}
+
+}
+
+\caption{Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes sampled in the Bay of Saline (\textit{C.
+nodosa} [Monospecific Settlement]) and the Bay of Funatana (\textit{C.
+nodosa} [Mixed Settlement] and \textit{C.
+cylindracea} [Mixed and Monospecific Settlement]) in two contrasting seasons.\label{rarefaction}}\label{fig:unnamed-chunk-1}
+\end{figure}
 
@@
 
-\subsection{Introduction}\label{introduction}
+\hypertarget{supplementary-table}{%
+\subsection{Supplementary Table}\label{supplementary-table}}
 
-Surfaces of marine macrophytes are inhabited by a diverse microbial
-community whose structure and function are poorly understood (Egan
-\emph{et al.}, 2013). As less than 1 \% of all prokaryotic species are
-culturable, to study these organisms, molecular methods such as 16S rRNA
-sequencing, metagenomics and metaproteomics are indispensable (Amann
-\emph{et al.}, 1995; Su \emph{et al.}, 2012). Applying these techniques
-requires an initial isolation step, with the purpose of obtaining high
-quality DNA and proteins.
+\begingroup\fontsize{9}{11}\selectfont
 
-Biological material (i.e.~proteins and DNA) from pelagic microbial
-communities is usually isolated by collecting cells onto filters and
-subsequently isolating the aimed compound (Gilbert \emph{et al.}, 2009).
-If a microbial size fraction is aimed sequential filtration is applied
-(Massana \emph{et al.}, 1997; Andersson \emph{et al.}, 2010). In
-contrast, obtaining biological materials from microorganism inhabiting
-surfaces requires either a cell detachment procedure prior to isolation
-or the host material is coextracted together with the targeted material.
-Methods for separating microbial cells form the host include shaking of
-host tissue (Gross \emph{et al.}, 2003; Nõges \emph{et al.}, 2010),
-scraping of macrophyte surfaces (Uku \emph{et al.}, 2007) or the
-application of ultrasonication (Weidner \emph{et al.}, 1996; Cai
-\emph{et al.}, 2014). It was shown that shaking alone is not sufficient
-to remove microbial cells, at least from plant root surfaces
-(Richter-Heitmann \emph{et al.}, 2016). Manual separation methods, such
-as scraping and brushing, are time consuming and subjective, as the
-detachment efficiency depends on host tissue and the person performing
-the procedure (Cai \emph{et al.}, 2014). Ultrasonication was proposed as
-an alternative method as it is providing better results in terms of
-detachment efficiency (Cai \emph{et al.}, 2014; Richter-Heitmann
-\emph{et al.}, 2016). The downside of this procedure is that complete
-cell removal was still not obtained and tissue disruption was observed
-especially after the application of probe ultrasonication
-(Richter-Heitmann \emph{et al.}, 2016). An alternative to these cell
-detachment procedures is the isolation of targeted epiphytic compounds
-together with host materials (Staufenberger \emph{et al.}, 2008; Jiang
-\emph{et al.}, 2015). This procedure can lead to problems in the
-following processing steps such as mitochondrial and chloroplast 16S
-rRNA sequence contaminations from the host (Longford \emph{et al.},
-2007; Staufenberger \emph{et al.}, 2008). In addition, when performing
-metagenomics and metaproteomics host material can cause biased results
-towards more abundant host DNA and proteins.
-
-An alternative to these procedures is a direct isolation of the targeted
-material by incubating macrophyte tissues in an extraction buffer. After
-the incubation is done the undisrupted host tissue is removed and the
-isolation procedure continues omitting host material contaminations. To
-our knowledge the only procedure describing a direct and selective
-epiphytic DNA isolation from the surfaces of marine macrophytes was
-provided by Burke \emph{et al.} (2009). In contrast to previously
-described methods this protocol enables an almost complete removal of
-the surface community and was used for 16S rRNA gene clone libraries
-construction (Burke \emph{et al.}, 2011b) and metagenomes sequencing
-(Burke \emph{et al.}, 2011a). This method, thought providing a selective
-isolation procedure, is using in the extraction buffer a rapid
-multienzyme cleaner (3M) that is not worldwide available and whose
-composition is not know (Burke \emph{et al.}, 2009). Also to our
-knowledge, no selective isolation protocol for proteins from epiphytic
-communities inhabiting marine macrophytes was established.
-
-In the present study, we adapted a protocol widely used for DNA
-isolation from filters (Massana \emph{et al.}, 1997) and a protocol used
-for protein isolation from sediments (Chourey \emph{et al.}, 2010;
-Hultman \emph{et al.}, 2015) for selective extractions of DNA and
-proteins from epiphytic communities inhabiting the surfaces of two
-marine macrophytes: the seagrass \emph{Cymodocea nodosa} and the alga
-\emph{Caulerpa cylindracea}. In addition, we tested the removal
-efficiency of the protocol and the suitability of obtained DNA and
-proteins for 16S rRNA sequencing and metaproteomics.
-
-\newpage
-
-\subsection{Materials and Methods}\label{materials-and-methods}
-
-\subsubsection{Sampling}\label{sampling}
-
-Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
-nodosa} meadow in the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E)
-and in a \emph{Cymodocea nodosa} meadow invaded by \emph{Caulerpa
-cylindracea} in the proximity of the village of Funtana (\ang{45;10;39}
-N, \ang{13;35;42} E). Thalli of \emph{Caulerpa cylindracea} were sampled
-in the same \emph{Cymodocea nodosa} invaded meadow in Funtana and on a
-locality of only \emph{Caulerpa cylindracea} located in the proximity of
-the invaded meadow. Leaves and thalli were collected on the same day in
-two contrasting seasons, on 4 December 2017 and 18 June 2018. During
-spring 2018 the \emph{Cymodocea nodosa} meadow in the Bay of Saline
-decayed to an extent that no leaves could be retrieved (Najdek \emph{et
-al.}, unpublished data). Leaves and thalli were collected by diving and
-transported to the laboratory in containers placed on ice and filled
-with site seawater. Upon arrival to the laboratory, \emph{Cymodocea
-nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while
-\emph{Caulerpa cylindracea} thalli were cut into 5 -- 8 \si{\cm} long
-sections. Leaves and thalli were washed three times with sterile
-artificial seawater (ASW) to remove loosely attached microbial cells.
-
-\subsubsection{DNA Isolation}\label{dna-isolation}
-
-The DNA was isolated according to the protocol for isolation from
-filters described in Massana \emph{et al.} (1997). This protocol was
-modified and adapted for DNA isolation from microbial communities from
-macrophyte surfaces as described below. 5 \si{\ml} of lysis buffer (40
-\si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar}
-sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or \si{\g}
-wet-weight of thalli. Lysozyme was added (final concentration 1
-\si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius}
-for 30 \si{\minute}. Subsequently, proteinase K (final concentration 0.5
-\si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were
-added and the samples were incubated at 55 \si{\degreeCelsius} for 2
-\si{\hour}. Following the incubation, tubes were vortexed for 10
-\si{\minute} and the mixture containing lyzed epiphytic cells was
-separated from host leaves or thalli by transferring the solution into a
-clean tube. The lysate was extracted twice with a mixture of
-phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with
-chloroform:isoamyl alcohol (24:1). After each organic solvent mixture
-addition tubes were slightly vortexed and centrifuged at 4,500 × g for
-10 \si{\minute}. Following each centrifugation aqueous phases were
-retrieved. After the final extraction 1/10 of chilled 3 \si{\Molar}
-sodium acetate (pH 5.2) was added. DNA was precipitated by adding 1
-volume of chilled isopropanol, incubating the mixtures overnight at
-\num{-20} °C and centrifuging at 16,000 × g and 4 \si{\degreeCelsius}
-for 20 \si{\minute}. The pellet was washed twice with 1 \si{\ml} of
-chilled 70 \si{\percent} ethanol and centrifuged after each washing step
-at 20,000 × g and 4 \si{\degreeCelsius} for 10 \si{\minute}. After the
-first washing step duplicate pellets form the same sample were pooled
-and transferred to a clean 1.5 \si{\ml} tube. The dried pellet was
-resuspended in 100 \si{\ul} of deionized water.
-
-\subsubsection{Illumina 16S rRNA
-Sequencing}\label{illumina-16s-rrna-sequencing}
-
-An aliquot of isolated DNA was treated with RNase A (final concentration
-200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
-concentration was determined using the Quant-iT PicoGreen dsDNA Assay
-Kit (Invitrogen, USA) according to the manufacturer's instructions and
-diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
-amplified using a two-step PCR procedure. In the first PCR the 515F
-(5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3')
-primers from the Earth Microbiome Project
-(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
-were used to amplify the target region (Caporaso \emph{et al.}, 2012;
-Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
-contained on their 5' end a tagged sequence. Each sample was amplified
-in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5
-Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml}
-BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse
-primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs,
-USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial
-denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of
-denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
-\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
-\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
-72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
-volumes were pooled and PCR products were purified using the GeneJET PCR
-Purification Kit (Thermo Fisher Scientific, USA) according to the
-manufacturer's instructions and following the protocol that included
-isopropanol addition for better small DNA fragment yield. The column was
-eluted in 30 \si{\ul} of deionized water. Purified PCR products were
-sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
-Martinsried, Germany. Before sequencing, at IMGM the second PCR
-amplification of the two-step PCR procedure was performed using primers
-targeting the tagged region incorporated in the first PCR. In addition,
-these primers contained adapter and sample-specific index sequences. The
-second PCR was carried out for 8 cycles. Beside samples, a positive and
-negative control were sequenced. A negative control was comprised of
-four parallel PCR reactions without DNA template, while for a positive
-control a mock community composed of evenly mixed DNA material
-originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
-used. The sequences obtained in this study have been submitted to the
-European Nucleotide Archive (ENA) under accession numbers \textbf{TO BE
-ADDED LATER!}.
-
-\subsubsection{Sequence Analysis}\label{sequence-analysis}
-
-Obtained sequences were analyzed on the computer cluster Isabella
-(University Computing Center, University of Zagreb) using mothur
-(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
-Standard Operating Procedure (MiSeq SOP;
-\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
-and recommendations given from the Riffomonas project to enhance data
-reproducibility (\url{http://www.riffomonas.org/}). For alignment and
-classification of sequences the SILVA SSU Ref NR 99 database (release
-138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
-Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
-mothur were exported, aligned using SILVA incremental aligner (SINA,
-version 1.6.0) (Pruesse \emph{et al.}, 2012) against the same SILVA SSU
-Ref NR 99 database (release 138) and imported into ARB (version 6.0.6)
-(Ludwig \emph{et al.}, 2004) for further phylogenetic analysis using the
-same database. Reference sequences close to imported ones were selected
-and used to calculate a phylogenetic tree using the Maximum Likelihood
-algorithm RAxML (version 7.0.3) with 1000 tree replicates (Stamatakis,
-2006). Imported partial chloroplast sequences were added to the tree
-using the maximum parsimony criteria and not allowing changes to tree
-topology. Pipeline data processing and visualization was done using R
-(version 3.6.0) (R Core Team, 2019), package tidyverse (version 1.2.1)
-(Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
-2015, 2019a, 2019b; Xie \emph{et al.}, 2018; Zhu, 2019; Allaire \emph{et
-al.}, 2019). The detailed analysis procedure including the R Markdown
-file for this paper are available as a GitHub repository (\textbf{TO BE
-ADDED LATER!}). Based on the ATCC MSA-1002 mock community included in
-the analysis a sequencing error rate of 0.009 \si{\percent} was
-determined, which is in line with previously reported values for
-next-generation sequencing data (Kozich \emph{et al.}, 2013; Schloss
-\emph{et al.}, 2016). In addition, the negative control processed
-together with the samples yielded only 2 sequences after sequence
-quality curation.
-
-\subsubsection{Protein Isolation}\label{protein-isolation}
-
-Proteins were isolated according to the protocol for isolation from soil
-described in Chourey \emph{et al.} (2010) and modified by Hultman
-\emph{et al.} (2015). These protocols were further modified and adapted
-for protein isolation from microbial communities form macrophyte
-surfaces as described below. 20 \si{\ml} of protein extraction buffer (4
-\si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was
-added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of
-thalli. The mixture was incubated in boiling water for 5 \si{\minute},
-vortexed for 10 \si{\minute} and incubated again in boiling water for 5
-\si{\minute}. After a brief vortex the lysate was transferred to a clean
-tube separating the host leaves or thalli from the mixture containing
-lyzed epiphytic cells. Dithiothreitol (DTT; final concentration 24
-\si{\milli\Molar}) was added and proteins were precipitated with chilled
-100 \si{\percent} trichloroacetic acid (TCA; final concentration 20
-\si{\percent}) overnight at \num{-20} °C. Precipitated proteins were
-centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
-The obtained protein pellet was washed three times with chilled acetone.
-During the first washing step the pellet was transferred to a clean 1.5
-\si{\ml} tube. After each washing step samples were centrifuged at
-20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets
-were stored at \num{-80} °C until further analysis.
-
-\subsubsection{Metaproteomics}\label{metaproteomics}
-
-Isolated proteins were whole trypsin digested using the FASP
-(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
-according to the manufacturer's instructions (Wiśniewski \emph{et al.},
-2009) with small modifications. Before loading the solution to the
-column, protein pellets were solubilized in a urea sample buffer
-included in the kit amended with DTT (final concentration 100
-\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
-centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
-remove larger particles. The first washing step after protein solution
-loading was repeated twice. In addition, centrifugation steps were
-prolonged if the column was clogged. Trypsin digestion was performed on
-column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
-The final filtrate containing digested proteins was acidified with 1
-\si{\percent} (final concentration) trifluoroacetic acid, freezed at
-\num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent
-to the Vienna Metabolomics Center (University of Vienna) for
-metaproteomic analysis. Peptides were resuspended in 1 \si{\percent}
-(final concentration) trifluoroacetic acid, desalted using the Pierce
-C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's
-instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap
-Mass Spectrometer (Thermo Fisher Scientific, USA). Obtained MS/MS
-spectra were searched against a protein database from metagenomic
-assembly published in Burke \emph{et al.} (2011a) using SEQUEST-HIT
-engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
-Fisher Scientific, USA). The target-decoy approach was used to reduce
-the probability of false peptide identification. Results whose false
-discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
-For protein identification a minimum of of two peptides and one unique
-peptide were required. For protein quantification, a chromatographic
-peak area-based free quantitative method was applied.
-
-\subsubsection{Confocal Microscopy}\label{confocal-microscopy}
-
-Host leaves and thalli from DNA and protein isolation steps were washed
-seven times in deionized water and fixed with formaldehyde (final
-concentration \textasciitilde{} 3 \si{\percent}). In addition,
-nontreated leaves and thalli, washed three times in ASW to remove
-loosely attached microbial cells, were fixed in the same concentration
-of formaldehyde and used as a positive control. For long therm storage,
-fixed leaves and thalli were immersed in a mixture of phosphate-buffered
-saline (PBS) and ethanol (1:1) and stored at \num{-20}
-\si{\degreeCelsius}. Treated and untreated leaves and thalli segments
-were stained in a 2 × solution of SYBR Green I and examined under a
-Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
-
-\subsection{Results}\label{results}
-
-\subsection{Discussion}\label{discussion}
-
-\subsection{Acknowledgements}\label{acknowledgements}
-
-\newpage
-
-\subsection{References}\label{references}
-
-\hypertarget{refs}{}
-\hypertarget{ref-Allaire2019}{}
-Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
-A., et al. (2019) rmarkdown: Dynamic Documents for R.
-
-\hypertarget{ref-Amann1995}{}
-Amann, R.I., Ludwig, W., and Schleifer, K.H. (1995) Phylogenetic
-identification and in situ detection of individual microbial cells
-without cultivation. \emph{Microbiological reviews} \textbf{59}:
-143--169.
-
-\hypertarget{ref-Andersson2010}{}
-Andersson, A.F., Riemann, L., and Bertilsson, S. (2010) Pyrosequencing
-reveals contrasting seasonal dynamics of taxa within Baltic Sea
-bacterioplankton communities. \emph{The ISME journal} \textbf{4}:
-171--181.
-
-\hypertarget{ref-Apprill2015}{}
-Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015) Minor
-revision to V4 region SSU rRNA 806R gene primer greatly increases
-detection of SAR11 bacterioplankton. \emph{Aquatic Microbial Ecology}
-\textbf{75}: 129--137.
-
-\hypertarget{ref-Burke2009}{}
-Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective extraction of
-bacterial DNA from the surfaces of macroalgae. \emph{Applied and
-environmental microbiology} \textbf{75}: 252--256.
-
-\hypertarget{ref-Burke2011a}{}
-Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
-(2011a) Bacterial community assembly based on functional genes rather
-than species. \emph{Proceedings of the National Academy of Sciences of
-the United States of America} \textbf{108}: 14288--14293.
-
-\hypertarget{ref-Burke2011b}{}
-Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
-(2011b) Composition, uniqueness and variability of the epiphytic
-bacterial community of the green alga Ulva australis. \emph{The ISME
-journal} \textbf{5}: 590--600.
-
-\hypertarget{ref-Cai2014}{}
-Cai, X., Gao, G., Yang, J., Tang, X., Dai, J., Chen, D., and Song, Y.
-(2014) An ultrasonic method for separation of epiphytic microbes from
-freshwater submerged macrophytes. \emph{Journal of basic microbiology}
-\textbf{54}: 758--761.
-
-\hypertarget{ref-Caporaso2012}{}
-Caporaso, J.G., Lauber, C.L., Walters, W.A., Berg-Lyons, D., Huntley,
-J., Fierer, N., et al. (2012) Ultra-high-throughput microbial community
-analysis on the Illumina HiSeq and MiSeq platforms. \emph{The ISME
-Journal} \textbf{6}: 1621--1624.
-
-\hypertarget{ref-Chourey2010}{}
-Chourey, K., Jansson, J., VerBerkmoes, N., Shah, M., Chavarria, K.L.,
-Tom, L.M., et al. (2010) Direct Cellular Lysis/Protein Extraction
-Protocol for Soil Metaproteomics. \emph{Journal of Proteome Research}
-\textbf{9}: 6615--6622.
-
-\hypertarget{ref-Egan2013}{}
-Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
-Thomas, T. (2013) The seaweed holobiont: understanding seaweed-bacteria
-interactions. \emph{FEMS microbiology reviews} \textbf{37}: 462--476.
-
-\hypertarget{ref-Gilbert2009}{}
-Gilbert, J.A., Field, D., Swift, P., Newbold, L., Oliver, A., Smyth, T.,
-et al. (2009) The seasonal structure of microbial communities in the
-Western English Channel. \emph{Environmental Microbiology} \textbf{11}:
-3132--3139.
-
-\hypertarget{ref-Gross2003}{}
-Gross, E.M., Feldbaum, C., and Graf, A. (2003) Epiphyte biomass and
-elemental composition on submersed macrophytes in shallow eutrophic
-lakes. \emph{Hydrobiologia} \textbf{506-509}: 559--565.
-
-\hypertarget{ref-Hultman2015}{}
-Hultman, J., Waldrop, M.P., Mackelprang, R., David, M.M., McFarland, J.,
-Blazewicz, S.J., et al. (2015) Multi-omics of permafrost, active layer
-and thermokarst bog soil microbiomes. \emph{Nature} \textbf{521}:
-208--212.
-
-\hypertarget{ref-Jiang2015}{}
-Jiang, Y.-F., Ling, J., Dong, J.-D., Chen, B., Zhang, Y.-Y., Zhang,
-Y.-Z., and Wang, Y.-S. (2015) Illumina-based analysis the microbial
-diversity associated with Thalassia hemprichii in Xincun Bay, South
-China Sea. \emph{Ecotoxicology (London, England)} \textbf{24}: 1548--56.
-
-\hypertarget{ref-Kozich2013}{}
-Kozich, J.J., Westcott, S.L., Baxter, N.T., Highlander, S.K., and
-Schloss, P.D. (2013) Development of a dual-index sequencing strategy and
-curation pipeline for analyzing amplicon sequence data on the MiSeq
-Illumina sequencing platform. \emph{Applied and environmental
-microbiology} \textbf{79}: 5112--5120.
-
-\hypertarget{ref-Longford2007}{}
-Longford, S., Tujula, N., Crocetti, G., Holmes, A., Holmström, C.,
-Kjelleberg, S., et al. (2007) Comparisons of diversity of bacterial
-communities associated with three sessile marine eukaryotes.
-\emph{Aquatic Microbial Ecology} \textbf{48}: 217--229.
-
-\hypertarget{ref-Ludwig2004}{}
-Ludwig, W., Strunk, O., Westram, R., Richter, L., Meier, H., Yadhukumar,
-et al. (2004) ARB: a software environment for sequence data.
-\emph{Nucleic Acids Research} \textbf{32}: 1363--1371.
-
-\hypertarget{ref-Massana1997}{}
-Massana, R., Murray, A.E., Preston, C.M., and DeLong, E.F. (1997)
-Vertical distribution and phylogenetic characterization of marine
-planktonic Archaea in the Santa Barbara Channel. \emph{Applied and
-environmental microbiology} \textbf{63}: 50--56.
-
-\hypertarget{ref-Noges2010}{}
-Nõges, T., Luup, H., and Feldmann, T. (2010) Primary production of
-aquatic macrophytes and their epiphytes in two shallow lakes (Peipsi and
-Võrtsjärv) in Estonia. \emph{Aquatic Ecology} \textbf{44}: 83--92.
-
-\hypertarget{ref-Parada2016}{}
-Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
-matters: assessing small subunit rRNA primers for marine microbiomes
-with mock communities, time series and global field samples.
-\emph{Environmental Microbiology} \textbf{18}: 1403--1414.
-
-\hypertarget{ref-Pruesse2012}{}
-Pruesse, E., Peplies, J., and Glöckner, F.O. (2012) SINA: accurate
-high-throughput multiple sequence alignment of ribosomal RNA genes.
-\emph{Bioinformatics (Oxford, England)} \textbf{28}: 1823--1829.
-
-\hypertarget{ref-Quast2013}{}
-Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
-et al. (2013) The SILVA ribosomal RNA gene database project: improved
-data processing and web-based tools. \emph{Nucleic acids research}
-\textbf{41}: D590--6.
-
-\hypertarget{ref-RCoreTeam2019}{}
-R Core Team (2019) R: A Language and Environment for Statistical
-Computing, Vienna, Austria: R Foundation for Statistical Computing.
-
-\hypertarget{ref-Richter-Heitmann2016}{}
-Richter-Heitmann, T., Eickhorst, T., Knauth, S., Friedrich, M.W., and
-Schmidt, H. (2016) Evaluation of Strategies to Separate Root-Associated
-Microbial Communities: A Crucial Choice in Rhizobiome Research.
-\emph{Frontiers in Microbiology} \textbf{7}: 773.
-
-\hypertarget{ref-Schloss2016}{}
-Schloss, P.D., Girard, R.A., Martin, T., Edwards, J., and Thrash, J.C.
-(2016) Status of the Archaeal and Bacterial Census: an Update.
-\emph{mBio} \textbf{7}: e00201--16.
-
-\hypertarget{ref-Schloss2009}{}
-Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
-Hollister, E.B., et al. (2009) Introducing mothur: open-source,
-platform-independent, community-supported software for describing and
-comparing microbial communities. \emph{Applied and environmental
-microbiology} \textbf{75}: 7537--7541.
-
-\hypertarget{ref-Stamatakis2006}{}
-Stamatakis, A. (2006) RAxML-VI-HPC: maximum likelihood-based
-phylogenetic analyses with thousands of taxa and mixed models.
-\emph{Bioinformatics} \textbf{22}: 2688--2690.
-
-\hypertarget{ref-Staufenberger2008}{}
-Staufenberger, T., Thiel, V., Wiese, J., and Imhoff, J.F. (2008)
-Phylogenetic analysis of bacteria associated with Laminaria saccharina.
-\emph{FEMS Microbiology Ecology} \textbf{64}: 65--77.
-
-\hypertarget{ref-Su2012}{}
-Su, C., Lei, L., Duan, Y., Zhang, K.-Q., and Yang, J. (2012)
-Culture-independent methods for studying environmental microorganisms:
-methods, application, and perspective. \emph{Applied microbiology and
-biotechnology} \textbf{93}: 993--1003.
-
-\hypertarget{ref-Uku2007}{}
-Uku, J., Björk, M., Bergman, B., and Díez, B. (2007) Characterization
-and comparison of prokaryotic epiphytes associated with three East
-African seagrasses. \emph{Journal of Phycology} \textbf{43}: 768--779.
-
-\hypertarget{ref-Weidner1996}{}
-Weidner, S., Arnold, W., and Puhler, A. (1996) Diversity of uncultured
-microorganisms associated with the seagrass Halophila stipulacea
-estimated by restriction fragment length polymorphism analysis of
-PCR-amplified 16S rRNA genes. \emph{Applied and environmental
-microbiology} \textbf{62}: 766--771.
-
-\hypertarget{ref-Wickham2019}{}
-Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
-R., et al. (2019) Welcome to the tidyverse. \emph{Journal of Open Source
-Software} \textbf{4}: 1686.
-
-\hypertarget{ref-Wisniewski2009}{}
-Wiśniewski, J.R., Zougman, A., Nagaraj, N., and Mann, M. (2009)
-Universal sample preparation method for proteome analysis. \emph{Nature
-Methods} \textbf{6}: 359--362.
-
-\hypertarget{ref-Xie2015}{}
-Xie, Y. (2015) Dynamic Documents with \{R\} and knitr, 2nd ed. Boca
-Raton, Florida: Chapman; Hall/CRC.
-
-\hypertarget{ref-Xie2014}{}
-Xie, Y. (2014) knitr: A Comprehensive Tool for Reproducible Research in
-\{R\}. In, Stodden,V., Leisch,F., and Peng,R.D. (eds),
-\emph{Implementing reproducible computational research}. Chapman;
-Hall/CRC.
-
-\hypertarget{ref-Xie2019a}{}
-Xie, Y. (2019a) knitr: A General-Purpose Package for Dynamic Report
-Generation in R.
-
-\hypertarget{ref-Xie2019}{}
-Xie, Y. (2019b) tinytex: Helper Functions to Install and Maintain 'TeX
-Live', and Compile 'LaTeX' Documents.
-
-\hypertarget{ref-Xie2018}{}
-Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
-Definitive Guide, Boca Raton, Florida: Chapman; Hall/CRC.
-
-\hypertarget{ref-Yilmaz2014}{}
-Yilmaz, P., Parfrey, L.W., Yarza, P., Gerken, J., Pruesse, E., Quast,
-C., et al. (2014) The SILVA and ``All-species Living Tree Project
-(LTP)'' taxonomic frameworks. \emph{Nucleic acids research} \textbf{42}:
-D643--8.
-
-\hypertarget{ref-Zhu2019}{}
-Zhu, H. (2019) kableExtra: Construct Complex Table with 'kable' and Pipe
-Syntax.
-
-\newpage
-
-\setlength\parindent{0pt}
-
-\subsection{Figure Captions}\label{figure-captions}
-
-\textbf{\autoref{micro_dec}.} \nameref{micro_dec}
-
-\textbf{\autoref{micro_jun}.} \nameref{micro_jun}
-
-\textbf{\autoref{community}.} \nameref{community}
-
-\newpage
-
-\subsection{Figures}\label{figures}
-
-\begin{figure}[ht]
-
-{\centering \includegraphics[width=0.85\linewidth]{../results/images/confocal_december}
-
-}
-
-\caption{Confocal microscope images of \textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea} surfaces from different locations (Saline, Funtana-Invaded and Funtana-Noninvaded) sampled on 4 December 2017 and stained with SYBR Green I.
-Scale bar in all images is 60 \si{\um}.\label{micro_dec}}\label{fig:unnamed-chunk-2}
-\end{figure}\newpage
-
-\begin{figure}[ht]
-
-{\centering \includegraphics[width=0.85\linewidth]{../results/images/confocal_june}
-
-}
-
-\caption{Confocal microscope images of \textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea} surfaces from different locations (Funtana-Invaded and Funtana-Noninvaded) sampled on 19 June 2018 and stained with SYBR Green I.
-Scale bar in all images is 60 \si{\um}.\label{micro_jun}}\label{fig:unnamed-chunk-3}
-\end{figure}\newpage
-
-\begin{figure}[ht]
-
-{\centering \includegraphics[width=1\linewidth]{../results/figures/community_bar_plot}
-
-}
-
-\caption{Taxonomic classification and relative contribution of the most abundant bacterial sequences from surfaces of two marine macrophytes (\textit{Cymodocea nodosa} and \textit{Caulerpa cylindracea}) sampled in different locations (Saline, Funtana-Invaded and Funtana-Noninvaded) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-4}
-\end{figure}
+\begin{longtable}[t]{>{\centering\arraybackslash}p{6em}ccccc}
+\caption{\label{tab:nseq_notus}Sample ID, Station, Community Type, Sampling Date, No.
+of Sequences and No.
+of OTUs of each sample.
+No.
+of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\label{nseq_notus}}\\
+\toprule
+\textbf{Sample ID} & \textbf{Station} & \textbf{Community Type} & \textbf{Date} & \textbf{No. of Sequences} & \textbf{No. of OTUs}\\
+\midrule
+\endfirsthead
+\caption[]{Sample ID, Station, Community Type, Sampling Date, No.
+of Sequences and No.
+of OTUs of each sample.
+No.
+of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\label{nseq_notus} \textit{(continued)}}\\
+\toprule
+\textbf{Sample ID} & \textbf{Station} & \textbf{Community Type} & \textbf{Date} & \textbf{No. of Sequences} & \textbf{No. of OTUs}\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\endlastfoot
+40 & Saline & \textit{Cymodocea nodosa} (Monospecific) & 4 December 2017 & 19,178 & 938\\
+41 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 4 December 2017 & 24,248 & 1,119\\
+42 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 4 December 2017 & 28,373 & 3,139\\
+43 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 4 December 2017 & 13,656 & 1,816\\
+61 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 19 June 2018 & 31,728 & 2,045\\
+62 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 19 June 2018 & 36,484 & 3,225\\
+63 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 19 June 2018 & 41,840 & 3,861\\*
+\end{longtable}
+\endgroup{}
 

### 2020-04-06 code/plot_community_bar_plot.R

@@ -91,10 +91,10 @@ gather(plot, key="sample", value="abundance", 6:(ncol(plot))) %>%
   annotation_custom(linesGrob(gp=gpar(lwd=2)), xmin=4.55, xmax=6.45, ymin=-9, ymax=-9) +
   annotation_custom(linesGrob(gp=gpar(lwd=2)), xmin=6.55, xmax=7.45, ymin=-9, ymax=-9) +
   annotation_custom(textGrob(str_wrap("Saline", width=10), gp=gpar(fontsize=14, fontface="bold")), xmin=0.55, xmax=1.45, ymin=-13, ymax=-13) +
-  annotation_custom(textGrob(str_wrap("Funtana-Invaded", width=10), gp=gpar(fontsize=14, fontface="bold")), xmin=1.55, xmax=3.45, ymin=-13, ymax=-13) +
-  annotation_custom(textGrob(str_wrap("Funtana-Noninvaded", width=10), gp=gpar(fontsize=14, fontface="bold")), xmin=3.55, xmax=4.45, ymin=-13, ymax=-13) +
-  annotation_custom(textGrob(str_wrap("Funtana-Invaded", width=10), gp=gpar(fontsize=14, fontface="bold")), xmin=4.55, xmax=6.45, ymin=-13, ymax=-13) +
-  annotation_custom(textGrob(str_wrap("Funtana-Noninvaded", width=10), gp=gpar(fontsize=14, fontface="bold")), xmin=6.55, xmax=7.45, ymin=-13, ymax=-13) +
+  annotation_custom(textGrob(str_wrap("Funtana (Mixed)", width=10), gp=gpar(fontsize=14, fontface="bold")), xmin=1.55, xmax=3.45, ymin=-13, ymax=-13) +
+  annotation_custom(textGrob(str_wrap("Funtana (Monospecific)", width=10), gp=gpar(fontsize=14, fontface="bold")), xmin=3.55, xmax=4.45, ymin=-13, ymax=-13) +
+  annotation_custom(textGrob(str_wrap("Funtana (Mixed)", width=10), gp=gpar(fontsize=14, fontface="bold")), xmin=4.55, xmax=6.45, ymin=-13, ymax=-13) +
[…]

### 2020-04-06 code/plot_rarefaction.R

@@ -3,27 +3,68 @@
 # 
 # A script to plot the rarefaction curve of each sample.
 # Dependencies: data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.groups.rarefaction
+#               data/raw/metadata.csv
 # Produces: results/figures/rarefaction.jpg
 #
 #################################################################################################################
 
-library(tidyverse)
-
-rarefy <- read_tsv(file="data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.groups.rarefaction") %>%
+# Loading input data and selection of values for plotting
+rarefaction <- read_tsv(file="data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.groups.rarefaction") %>%
   select(-contains("lci-"), -contains("hci-")) %>%
   gather(-numsampled, key=sample, value=sobs) %>%
   mutate(sample=str_replace_all(sample, pattern="0.03-", replacement="")) %>%
   drop_na()
 
-metadata <- read_tsv("data/raw/metadata.csv")
+# Loading metadata
+Sys.setlocale(locale="en_GB.utf8")
+metadata <- read_tsv("data/raw/metadata.csv") %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  arrange(date) %>%
+  mutate(date=format(date, "%d %B %Y")) %>%
+  mutate(date=str_replace(date, "^0", ""))
 
-metadata_rarefy <- inner_join(metadata, rarefy, by=c("ID"="sample"))
+# Joining metadata and input data
+metadata_rarefaction <- inner_join(metadata, rarefaction, by=c("ID"="sample"))
 
-ggplot(metadata_rarefy, aes(x=numsampled, y=sobs, group=ID, color=station)) +
-  geom_text(data=metadata_rarefy %>% group_by(label) %>% filter(sobs==max(sobs)), aes(x=numsampled, y=sobs, label=label), 
-            show.legend=FALSE) +
-  geom_line() +
+# Defining line color and type for each sampling date
+color_type <- tribble(~label, ~color,
+                      "4/12-17 SCy", "#A6CEE3",
+                      "4/12-17 FCyM", "#1F78B4",
+                      "4/12-17 FCaM", "#33A02C",
[…]

### 2020-04-14 submission/manuscript.Rmd
Chloroplast classification.

@@
 Peptides were resuspended in 1 \si{\percent} (final concentration) trifluoroacetic acid, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
-Obtained MS/MS spectra were searched against a protein database from metagenomic assembly published in @Burke2011a using SEQUEST-HIT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
+Obtained MS/MS spectra were searched against a protein database from metagenomic assembly published in @Burke2011a using SEQUEST-HT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
 The target-decoy approach was used to reduce the probability of false peptide identification.

### 2020-04-14 submission/manuscript.tex
Chloroplast classification.

@@
 assembly published in Burke and Peter Steinberg \emph{et al.} (2011)
-using SEQUEST-HIT engines and validated with Percolator in Proteome
+using SEQUEST-HT engines and validated with Percolator in Proteome
 Discoverer 2.1 (Thermo Fisher Scientific, USA). The target-decoy

### 2020-04-14 submission/supplementary.tex
Chloroplast classification.

@@
 \endlastfoot
-40 & Saline & \textit{Cymodocea nodosa} (Monospecific) & 4 December 2017 & 19,178 & 938\\
-41 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 4 December 2017 & 24,248 & 1,119\\
-42 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 4 December 2017 & 28,373 & 3,139\\
-43 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 4 December 2017 & 13,656 & 1,816\\
-61 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 19 June 2018 & 31,728 & 2,045\\
-62 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 19 June 2018 & 36,484 & 3,225\\
-63 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 19 June 2018 & 41,840 & 3,861\\*
+40 & Saline & \textit{Cymodocea nodosa} (Monospecific) & 4 December 2017 & 19,176 & 942\\
+41 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 4 December 2017 & 24,250 & 1,123\\
+42 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 4 December 2017 & 28,387 & 3,131\\
+43 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 4 December 2017 & 13,661 & 1,809\\
+61 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 19 June 2018 & 31,734 & 2,039\\
+62 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 19 June 2018 & 36,488 & 3,211\\
+63 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 19 June 2018 & 41,871 & 3,851\\*
 \end{longtable}

### 2020-04-20 submission/manuscript.Rmd
Added chloroplast plot.

@@
 To our knowledge the only procedure describing a direct and selective epiphytic DNA isolation from the surfaces of marine macrophytes was provided by @Burke2009.
-In contrast to previously described methods this protocol enables an almost complete removal of the surface community and was used for 16S rRNA gene clone libraries construction [@Burke2011b] and metagenomes sequencing [@Burke2011a].
+In contrast to previously described methods this protocol enables an almost complete removal of the surface community and was used for 16S rRNA gene clone libraries construction [@Burke2011a] and metagenomes sequencing [@Burke2011].
 This method, thought providing a selective isolation procedure, is using in the extraction buffer a rapid multienzyme cleaner (3M) that is not worldwide available and whose composition is not know [@Burke2009].
@@
 A negative control was comprised of four parallel PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
-The sequences obtained in this study have been submitted to the European Nucleotide Archive (ENA) under accession numbers **TO BE ADDED LATER!**.
+Sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 -- SAMEA6648811.
 
@@
 For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Sequences classified as chloroplasts by mothur were exported, aligned using SILVA incremental aligner (SINA, version 1.6.0) [@Pruesse2012] against the same SILVA SSU Ref NR 99 database (release 138) and imported into ARB (version 6.0.6) [@Ludwig2004] for further phylogenetic analysis using the same database.
-Reference sequences close to imported ones were selected and used to calculate a phylogenetic tree using the Maximum Likelihood algorithm RAxML (version 7.0.3) with 1000 tree replicates [@Stamatakis2006].
-Imported partial chloroplast sequences were added to the tree using the maximum parsimony criteria and not allowing changes to tree topology.
-Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019], package tidyverse (version 1.2.1) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Zhu2019; @Allaire2019; @Xie2018].
-The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (**TO BE ADDED LATER!**).
+Sequences classified as chloroplasts by SILVA were exported and reclassifed using mothur and the RDP (Ribosomal Database Project; http://rdp.cme.msu.edu/) training set (version 16) reference files adapted for mothur [@Cole2014].
+In comparison to SILVA, RDP allows a more detailed classification of chloroplast sequences.
+Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014].
+The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020).
 Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
@@
 Peptides were resuspended in 1 \si{\percent} (final concentration) trifluoroacetic acid, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
-Obtained MS/MS spectra were searched against a protein database from metagenomic assembly published in @Burke2011a using SEQUEST-HT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
+Obtained MS/MS spectra were searched against a protein database from metagenomic assembly published in @Burke2011 using SEQUEST-HT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
 The target-decoy approach was used to reduce the probability of false peptide identification.
@@
 
+**\autoref{chloroplast}.** \nameref{chloroplast}
+
 \newpage
@@
 ```
+
+\newpage
+```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of chloroplast sequences from surfaces of two marine macrophytes (\\textit{C.
+nodosa} and \\textit{C.
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{chloroplast}", fig.pos="ht"}
+include_graphics("../results/figures/chloroplast.jpg")
+```

### 2020-04-20 submission/manuscript.tex
Added chloroplast plot.

@@
 originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
-used. The sequences obtained in this study have been submitted to the
-European Nucleotide Archive (ENA) under accession numbers \textbf{TO BE
-ADDED LATER!}.
+used. Sequences obtained in this study have been deposited in the
+European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers
+SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 --
+SAMEA6648811.
 
@@
 Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
-mothur were exported, aligned using SILVA incremental aligner (SINA,
-version 1.6.0) (Pruesse \emph{et al.}, 2012) against the same SILVA SSU
-Ref NR 99 database (release 138) and imported into ARB (version 6.0.6)
-(Ludwig \emph{et al.}, 2004) for further phylogenetic analysis using the
-same database. Reference sequences close to imported ones were selected
-and used to calculate a phylogenetic tree using the Maximum Likelihood
-algorithm RAxML (version 7.0.3) with 1000 tree replicates (Stamatakis,
-2006). Imported partial chloroplast sequences were added to the tree
-using the maximum parsimony criteria and not allowing changes to tree
-topology. Pipeline data processing and visualization was done using R
-(version 3.6.0) (R Core Team, 2019), package tidyverse (version 1.2.1)
-(Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
-2015, 2019a, 2019b; Xie \emph{et al.}, 2018; Zhu, 2019; Allaire \emph{et
-al.}, 2019). The detailed analysis procedure including the R Markdown
-file for this paper are available as a GitHub repository (\textbf{TO BE
-ADDED LATER!}). Based on the ATCC MSA-1002 mock community included in
-the analysis a sequencing error rate of 0.009 \si{\percent} was
-determined, which is in line with previously reported values for
-next-generation sequencing data (Kozich \emph{et al.}, 2013; Schloss
-\emph{et al.}, 2016). In addition, the negative control processed
-together with the samples yielded only 2 sequences after sequence
-quality curation.
+SILVA were exported and reclassifed using mothur and the RDP (Ribosomal
+Database Project; \url{http://rdp.cme.msu.edu/}) training set (version
+16) reference files adapted for mothur (Cole \emph{et al.}, 2014). In
+comparison to SILVA, RDP allows a more detailed classification of
+chloroplast sequences. Pipeline data processing and visualization was
+done using R (version 3.6.0) (R Core Team, 2019), package tidyverse
+(version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple other
+packages (Xie, 2014, 2015, 2019b, 2019a, 2020; Neuwirth, 2014; Xie
+\emph{et al.}, 2018; Allaire \emph{et al.}, 2019; Zhu, 2019). The
+detailed analysis procedure including the R Markdown file for this paper
+are available as a GitHub repository
+(\url{https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).
+Based on the ATCC MSA-1002 mock community included in the analysis a
+sequencing error rate of 0.009 \si{\percent} was determined, which is in
+line with previously reported values for next-generation sequencing data
+(Kozich \emph{et al.}, 2013; Schloss \emph{et al.}, 2016). In addition,
+the negative control processed together with the samples yielded only 2
+sequences after sequence quality curation.
 
@@
 Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
-A., et al. (2019) rmarkdown: Dynamic Documents for R.
+A., et al. (2019) Rmarkdown: Dynamic documents for R.
 
@@
 Amann, R.I., Ludwig, W., and Schleifer, K.H. (1995) Phylogenetic
-identification and in situ detection of individual microbial cells
-without cultivation. \emph{Microbiological reviews} \textbf{59}:
-143--169.
+identification and \emph{in situ} detection of individual microbial
+cells without cultivation. \emph{Microbiol Rev} \textbf{59}: 143--169.
 
@@
 reveals contrasting seasonal dynamics of taxa within Baltic Sea
-bacterioplankton communities. \emph{The ISME journal} \textbf{4}:
-171--181.
+bacterioplankton communities. \emph{ISME J} \textbf{4}: 171--181.
 
@@
 revision to V4 region SSU rRNA 806R gene primer greatly increases
-detection of SAR11 bacterioplankton. \emph{Aquatic Microbial Ecology}
+detection of SAR11 bacterioplankton. \emph{Aquat Microb Ecol}
 \textbf{75}: 129--137.
@@
 Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective extraction of
-bacterial DNA from the surfaces of macroalgae. \emph{Applied and
-environmental microbiology} \textbf{75}: 252--256.
+bacterial DNA from the surfaces of macroalgae. \emph{Appl Environ
+Microbiol} \textbf{75}: 252--256.
+
+\leavevmode\hypertarget{ref-Burke2011}{}%
+Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
+(2011) Bacterial community assembly based on functional genes rather
+than species. \emph{Proc Natl Acad Sci U S A} \textbf{108}:
+14288--14293.
 
 \leavevmode\hypertarget{ref-Burke2011a}{}%
-Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
-(2011) Bacterial community assembly based on functional genes rather
-than species. \emph{Proceedings of the National Academy of Sciences of
-the United States of America} \textbf{108}: 14288--14293.
-
-\leavevmode\hypertarget{ref-Burke2011b}{}%
 Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
 (2011) Composition, uniqueness and variability of the epiphytic
-bacterial community of the green alga Ulva australis. \emph{The ISME
-journal} \textbf{5}: 590--600.
+bacterial community of the green alga \emph{Ulva australis}. \emph{ISME
+J} \textbf{5}: 590--600.
 
@@
 (2014) An ultrasonic method for separation of epiphytic microbes from
-freshwater submerged macrophytes. \emph{Journal of basic microbiology}
-\textbf{54}: 758--761.
+freshwater submerged macrophytes. \emph{J Basic Microbiol} \textbf{54}:
+758--761.
 
@@
 J., Fierer, N., et al. (2012) Ultra-high-throughput microbial community
-analysis on the Illumina HiSeq and MiSeq platforms. \emph{The ISME
-Journal} \textbf{6}: 1621--1624.
+analysis on the Illumina HiSeq and MiSeq platforms. \emph{ISME J}
+\textbf{6}: 1621--1624.
 
@@
 Chourey, K., Jansson, J., VerBerkmoes, N., Shah, M., Chavarria, K.L.,
-Tom, L.M., et al. (2010) Direct Cellular Lysis/Protein Extraction
-Protocol for Soil Metaproteomics. \emph{Journal of Proteome Research}
-\textbf{9}: 6615--6622.
+Tom, L.M., et al. (2010) Direct cellular lysis/protein extraction
+protocol for soil metaproteomics. \emph{J Proteome Res} \textbf{9}:
+6615--6622.
+
+\leavevmode\hypertarget{ref-Cole2014}{}%
+Cole, J.R., Wang, Q., Fish, J.A., Chai, B., McGarrell, D.M., Sun, Y., et
+al. (2014) Ribosomal Database Project: Data and tools for high
+throughput rRNA analysis. \emph{Nucleic Acids Res} \textbf{42}:
+D633--D642.
 
@@
 Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
-Thomas, T. (2013) The seaweed holobiont: understanding seaweed-bacteria
-interactions. \emph{FEMS microbiology reviews} \textbf{37}: 462--476.
+Thomas, T. (2013) The seaweed holobiont: Understanding seaweed-bacteria
+interactions. \emph{FEMS Microbiol Rev} \textbf{37}: 462--476.
 
@@
 et al. (2009) The seasonal structure of microbial communities in the
-Western English Channel. \emph{Environmental Microbiology} \textbf{11}:
+Western English Channel. \emph{Environ Microbiol} \textbf{11}:
 3132--3139.
@@
 Y.-Z., and Wang, Y.-S. (2015) Illumina-based analysis the microbial
-diversity associated with Thalassia hemprichii in Xincun Bay, South
-China Sea. \emph{Ecotoxicology (London, England)} \textbf{24}: 1548--56.
+diversity associated with \emph{Thalassia hemprichii} in Xincun Bay,
+South China Sea. \emph{Ecotoxicology} \textbf{24}: 1548--56.
 
@@
 curation pipeline for analyzing amplicon sequence data on the MiSeq
-Illumina sequencing platform. \emph{Applied and environmental
-microbiology} \textbf{79}: 5112--5120.
+Illumina sequencing platform. \emph{Appl Environ Microbiol} \textbf{79}:
+5112--5120.
 
@@
 Kjelleberg, S., et al. (2007) Comparisons of diversity of bacterial
-communities associated with three sessile marine eukaryotes.
-\emph{Aquatic Microbial Ecology} \textbf{48}: 217--229.
-
-\leavevmode\hypertarget{ref-Ludwig2004}{}%
-Ludwig, W., Strunk, O., Westram, R., Richter, L., Meier, H., Yadhukumar,
-et al. (2004) ARB: a software environment for sequence data.
-\emph{Nucleic Acids Research} \textbf{32}: 1363--1371.
+communities associated with three sessile marine eukaryotes. \emph{Aquat
+Microb Ecol} \textbf{48}: 217--229.
 
@@
 Vertical distribution and phylogenetic characterization of marine
-planktonic Archaea in the Santa Barbara Channel. \emph{Applied and
-environmental microbiology} \textbf{63}: 50--56.
+planktonic \emph{Archaea} in the Santa Barbara Channel. \emph{Appl
+Environ Microbiol} \textbf{63}: 50--56.
+
+\leavevmode\hypertarget{ref-Neuwirth2014}{}%
+Neuwirth, E. (2014) RColorBrewer: ColorBrewer palettes.
 
@@
 aquatic macrophytes and their epiphytes in two shallow lakes (Peipsi and
-Võrtsjärv) in Estonia. \emph{Aquatic Ecology} \textbf{44}: 83--92.
+Võrtsjärv) in Estonia. \emph{Aquat Ecol} \textbf{44}: 83--92.
 
@@
 Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
-matters: assessing small subunit rRNA primers for marine microbiomes
+matters: Assessing small subunit rRNA primers for marine microbiomes
 with mock communities, time series and global field samples.
-\emph{Environmental Microbiology} \textbf{18}: 1403--1414.
-
-\leavevmode\hypertarget{ref-Pruesse2012}{}%
-Pruesse, E., Peplies, J., and Glöckner, F.O. (2012) SINA: accurate
-high-throughput multiple sequence alignment of ribosomal RNA genes.
-\emph{Bioinformatics (Oxford, England)} \textbf{28}: 1823--1829.
+\emph{Environ Microbiol} \textbf{18}: 1403--1414.
 
@@
 Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
-et al. (2013) The SILVA ribosomal RNA gene database project: improved
-data processing and web-based tools. \emph{Nucleic acids research}
-\textbf{41}: D590--6.
+et al. (2013) The SILVA ribosomal RNA gene database project: Improved
+data processing and web-based tools. \emph{Nucleic Acids Res}
+\textbf{41}: D590--D596.
 
 \leavevmode\hypertarget{ref-RCoreTeam2019}{}%
-R Core Team (2019) R: A Language and Environment for Statistical
-Computing, Vienna, Austria: R Foundation for Statistical Computing.
+R Core Team (2019) A language and environment for statistical computing,
+Vienna, Austria: R Foundation for Statistical Computing.
 
@@
 Richter-Heitmann, T., Eickhorst, T., Knauth, S., Friedrich, M.W., and
-Schmidt, H. (2016) Evaluation of Strategies to Separate Root-Associated
-Microbial Communities: A Crucial Choice in Rhizobiome Research.
-\emph{Frontiers in Microbiology} \textbf{7}: 773.
+Schmidt, H. (2016) Evaluation of strategies to separate root-associated
+microbial communities: A crucial choice in rhizobiome research.
+\emph{Front Microbiol} \textbf{7}: 773.
 
 \leavevmode\hypertarget{ref-Schloss2016}{}%
-Schloss, P.D., Girard, R.A., Martin, T., Edwards, J., and Thrash, J.C.
-(2016) Status of the Archaeal and Bacterial Census: an Update.
-\emph{mBio} \textbf{7}: e00201--16.
+Schloss, P.D., Jenior, M.L., Koumpouras, C.C., Westcott, S.L., and
+Highlander, S.K. (2016) Sequencing 16S rRNA gene fragments using the
+PacBio SMRT DNA sequencing system. \emph{PeerJ} \textbf{4}: e1869.
 
@@
 Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
-Hollister, E.B., et al. (2009) Introducing mothur: open-source,
+Hollister, E.B., et al. (2009) Introducing mothur: Open-source,
 platform-independent, community-supported software for describing and
-comparing microbial communities. \emph{Applied and environmental
-microbiology} \textbf{75}: 7537--7541.
-
-\leavevmode\hypertarget{ref-Stamatakis2006}{}%
-Stamatakis, A. (2006) RAxML-VI-HPC: maximum likelihood-based
-phylogenetic analyses with thousands of taxa and mixed models.
-\emph{Bioinformatics} \textbf{22}: 2688--2690.
+comparing microbial communities. \emph{Appl Environ Microbiol}
+\textbf{75}: 7537--7541.
 
@@
 Staufenberger, T., Thiel, V., Wiese, J., and Imhoff, J.F. (2008)
-Phylogenetic analysis of bacteria associated with Laminaria saccharina.
-\emph{FEMS Microbiology Ecology} \textbf{64}: 65--77.
+Phylogenetic analysis of bacteria associated with \emph{Laminaria
+saccharina}. \emph{FEMS Microbiol Ecol} \textbf{64}: 65--77.
 
@@
 Culture-independent methods for studying environmental microorganisms:
-methods, application, and perspective. \emph{Applied microbiology and
-biotechnology} \textbf{93}: 993--1003.
+Methods, application, and perspective. \emph{Appl Microbiol Biotechnol}
+\textbf{93}: 993--1003.
 
@@
 and comparison of prokaryotic epiphytes associated with three East
-African seagrasses. \emph{Journal of Phycology} \textbf{43}: 768--779.
+African seagrasses. \emph{J Phycol} \textbf{43}: 768--779.
 
@@
 Weidner, S., Arnold, W., and Puhler, A. (1996) Diversity of uncultured
-microorganisms associated with the seagrass Halophila stipulacea
+microorganisms associated with the seagrass \emph{Halophila stipulacea}
 estimated by restriction fragment length polymorphism analysis of
-PCR-amplified 16S rRNA genes. \emph{Applied and environmental
-microbiology} \textbf{62}: 766--771.
+PCR-amplified 16S rRNA genes. \emph{Appl Environ Microbiol} \textbf{62}:
+766--771.
 
@@
 Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
-R., et al. (2019) Welcome to the tidyverse. \emph{Journal of Open Source
-Software} \textbf{4}: 1686.
+R., et al. (2019) Welcome to the tidyverse. \emph{J Open Source Softw}
+\textbf{4}: 1686.
 
@@
 Wiśniewski, J.R., Zougman, A., Nagaraj, N., and Mann, M. (2009)
-Universal sample preparation method for proteome analysis. \emph{Nature
+Universal sample preparation method for proteome analysis. \emph{Nat
 Methods} \textbf{6}: 359--362.
@@
 \leavevmode\hypertarget{ref-Xie2015}{}%
-Xie, Y. (2015) Dynamic Documents with \{R\} and knitr, 2nd ed. Boca
-Raton, Florida: Chapman; Hall/CRC.
+Xie, Y. (2015) Dynamic Documents with R and knitr, 2nd ed. Boca Raton,
+Florida: Chapman and Hall/CRC.
 
 \leavevmode\hypertarget{ref-Xie2014}{}%
-Xie, Y. (2014) knitr: A Comprehensive Tool for Reproducible Research in
-\{R\}. In \emph{Implementing reproducible computational research}.
-Stodden, V., Leisch, F., and Peng, R.D. (eds). Chapman; Hall/CRC.
+Xie, Y. (2014) Knitr: A comprehensive tool for reproducible research in
+R. In \emph{Implementing Reproducible Computational Research}. Stodden,
+V., Leisch, F., and Peng, R.D. (eds). New York: Chapman and Hall/CRC,
+pp. 3--32.
+
+\leavevmode\hypertarget{ref-Xie2019}{}%
+Xie, Y. (2019a) Knitr: A general-purpose package for dynamic report
+generation in R.
 
 \leavevmode\hypertarget{ref-Xie2019a}{}%
-Xie, Y. (2019a) knitr: A General-Purpose Package for Dynamic Report
-Generation in R.
+Xie, Y. (2019b) TinyTeX: A lightweight, cross-platform, and
+easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
+\textbf{40}: 30--32.
 
-\leavevmode\hypertarget{ref-Xie2019}{}%
-Xie, Y. (2019b) tinytex: Helper Functions to Install and Maintain 'TeX
-Live', and Compile 'LaTeX' Documents.
+\leavevmode\hypertarget{ref-Xie2020}{}%
+Xie, Y. (2020) TinyTex: Helper functions to install and maintain 'TeX
+Live', and compile 'LaTeX' documents.
 
@@
 Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
-Definitive Guide, Boca Raton, Florida: Chapman; Hall/CRC.
+Definitive Guide, 1st ed. Boca Raton, Florida: Chapman and Hall/CRC.
 
@@
 Yilmaz, P., Parfrey, L.W., Yarza, P., Gerken, J., Pruesse, E., Quast,
-C., et al. (2014) The SILVA and "All-species Living Tree Project (LTP)"
-taxonomic frameworks. \emph{Nucleic acids research} \textbf{42}:
-D643--8.
+C., et al. (2014) The SILVA and "All-Species Living Tree Project (LTP)"
+taxonomic frameworks. \emph{Nucleic Acids Res} \textbf{42}: D643--D648.
 
 \leavevmode\hypertarget{ref-Zhu2019}{}%
-Zhu, H. (2019) kableExtra: Construct Complex Table with 'kable' and Pipe
-Syntax.
+Zhu, H. (2019) KableExtra: Construct complex table with 'kable' and pipe
+syntax.
 
@@
 \textbf{\autoref{community}.} \nameref{community}
+
+\textbf{\autoref{chloroplast}.} \nameref{chloroplast}
 
@@
 
+\newpage
+\begin{figure}[ht]
+
+{\centering \includegraphics[width=1\linewidth]{../results/figures/chloroplast}
+
+}
+
+\caption{Taxonomic classification and relative contribution of chloroplast sequences from surfaces of two marine macrophytes (\textit{C.
+nodosa} and \textit{C.
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{chloroplast}}\label{fig:unnamed-chunk-5}
+\end{figure}
+
 \end{document}

### 2020-04-20 code/plot_chloroplast.R

@@ -1,0 +1,91 @@
+#################################################################################################################
+# plot_chloroplast.R
+# 
+# A script to plot the abundance and taxonomy of chloroplast sequences of each sample.
+# Dependencies: data/mothur/chloroplast.pds.wang.tax.summary
+#               data/raw/metadata.csv
+# Produces: results/figures/chloroplast.jpg
+#
+#################################################################################################################
+
+# Loading input data containing sequence abundances and subsequent input data customization
+community <- read_tsv("data/mothur/chloroplast.pds.wang.tax.summary") %>%
+  filter(taxon!="Root")
+
+# Calculate relative abundaces
+community <- group_by(community, taxlevel) %>%
+  mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
+  ungroup()
+
+# Selection of groups for plotting
+plot <- filter(community, taxlevel==5)
+
+# Defying group colors
+color <- c("Bacteria_unclassified"="goldenrod1",
+           "Bacillariophyta"="darkolivegreen3",
+           "Chlorarachniophyceae"="yellow",
+           "Chlorophyta"="aquamarine4",
+           "Chloroplast_unclassified"="green",
+           "Cryptomonadaceae"="orange3",
+           "Streptophyta"="violet",
+           "Cyanobacteria_unclassified"="forestgreen",
+           "Cyanobacteria/Chloroplast_unclassified"="lightgreen")
+
+# Generation of italic names for groups
+names <- parse(text=case_when(plot$taxon=="Bacteria_unclassified" ~ "italic('Bacteria')~plain('(No Relative)')",
+                              plot$taxon=="Chloroplast_unclassified" ~ "plain('Chloroplast')~plain('(No Relative)')",
+                              plot$taxon=="Cyanobacteria_unclassified" ~ "italic('Cyanobacteria')~plain('(No Relative)')",
+                              plot$taxon=="Cyanobacteria/Chloroplast_unclassified" ~ "italic('Cyanobacteria')*plain('/Chloroplast')~'(No Relative)'",
[…]

### 2020-04-20 code/plot_community_bar_plot.R

@@ -90,15 +90,15 @@ gather(plot, key="sample", value="abundance", 6:(ncol(plot))) %>%
   annotation_custom(linesGrob(gp=gpar(lwd=2)), xmin=3.55, xmax=4.45, ymin=-9, ymax=-9) +
   annotation_custom(linesGrob(gp=gpar(lwd=2)), xmin=4.55, xmax=6.45, ymin=-9, ymax=-9) +
   annotation_custom(linesGrob(gp=gpar(lwd=2)), xmin=6.55, xmax=7.45, ymin=-9, ymax=-9) +
-  annotation_custom(textGrob(str_wrap("Saline", width=10), gp=gpar(fontsize=14, fontface="bold")), xmin=0.55, xmax=1.45, ymin=-13, ymax=-13) +
-  annotation_custom(textGrob(str_wrap("Funtana (Mixed)", width=10), gp=gpar(fontsize=14, fontface="bold")), xmin=1.55, xmax=3.45, ymin=-13, ymax=-13) +
-  annotation_custom(textGrob(str_wrap("Funtana (Monospecific)", width=10), gp=gpar(fontsize=14, fontface="bold")), xmin=3.55, xmax=4.45, ymin=-13, ymax=-13) +
-  annotation_custom(textGrob(str_wrap("Funtana (Mixed)", width=10), gp=gpar(fontsize=14, fontface="bold")), xmin=4.55, xmax=6.45, ymin=-13, ymax=-13) +
-  annotation_custom(textGrob(str_wrap("Funtana (Monospecific)", width=10), gp=gpar(fontsize=14, fontface="bold")), xmin=6.55, xmax=7.45, ymin=-13, ymax=-13) +
+  annotation_custom(textGrob(str_wrap("Saline", width=10), gp=gpar(fontsize=14, fontface="bold", fontfamily="Times")), xmin=0.55, xmax=1.45, ymin=-13, ymax=-13) +
+  annotation_custom(textGrob(str_wrap("Funtana (Mixed)", width=10), gp=gpar(fontsize=14, fontface="bold", fontfamily="Times")), xmin=1.55, xmax=3.45, ymin=-13, ymax=-13) +
+  annotation_custom(textGrob(str_wrap("Funtana (Monospecific)", width=10), gp=gpar(fontsize=14, fontface="bold", fontfamily="Times")), xmin=3.55, xmax=4.45, ymin=-13, ymax=-13) +
[…]

### 2020-07-08 submission/manuscript.Rmd
Process optimization.

@@
 
+\newpage
 ## Results
+
+```{r community}
+# Loading input data containing sequence abundances and subsequent input data customization
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
+  filter(!str_detect(taxon, "^Eukaryota")) %>%
+  filter(taxon!="Root")
+
+# Remove Mitochondria sequences and subtract their number from higher taxonomic levels to which they belong
+mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
+    rankID==str_extract(mitochondria, "(\\d+\\.){4}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Mitochondria"],
+    TRUE ~ .))) %>%
+  filter(!str_detect(taxon, "^Mitochondria")) %>%
+  select(-ATCC_1, -NC_1)
+
+# Calculating the number of sequences
+number_sequences <- filter(community, taxon=="Archaea" | taxon=="Bacteria") %>%
+  select(total) %>%
+  sum()
+
+# Calculating the number of chloroplast sequences
+number_chloroplast_sequences <- filter(community, taxon=="Chloroplast") %>%
+  select(total) %>%
+  sum()
+
+# Calculating the number of sequences without chloroplast sequences
+number_sequences_without_chloroplast <- number_sequences - number_chloroplast_sequences
+
+# Calculating the proprotion of chloroplast sequences
+proprotion_chloroplast_sequences <- (number_chloroplast_sequences / number_sequences) * 100
+
+# Calculating the minimal and maximal number of sequences
+min_max <- filter(community, taxon=="Archaea" | taxon=="Bacteria") %>%
+  summarise_at(6:ncol(.), sum) %>%
+  summarise(min=min(.), max=max(.))
+```
+
+```{r shared}
+# Loading OTU/sample data
+shared <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")
+
+# Calculating the number of OTUs before rarefaction
+no_otus <- select(shared, starts_with("Otu")) %>%
+  colnames() %>%
+  length()
+```
+
+```{r taxonomy}
+# Loading input data containing sequence abundances and subsequent input data customization
+community <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
+  filter(!str_detect(taxon, "^Eukaryota")) %>%
+  filter(taxon!="Root")
+
+# Remove Mitochondria sequences and subtract their number from higher taxonomic levels to which they belong
+mitochondria <- filter(community, str_detect(taxon, "^Mitochondria$"))$rankID
+community <- mutate_at(community, 5:ncol(community), list(~case_when(
+    rankID==str_extract(mitochondria, "(\\d+\\.){4}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){3}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){2}\\d+") ~ . - .[taxon=="Mitochondria"],
+    rankID==str_extract(mitochondria, "(\\d+\\.){1}\\d+") ~ . - .[taxon=="Mitochondria"],
+    TRUE ~ .))) %>%
+  filter(!str_detect(taxon, "^Mitochondria")) %>%
+  select(-ATCC_1, -NC_1) %>%
+  group_by(taxlevel) %>%
+  mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
+  ungroup()
+
+# Loading metadata of each sample and selection of desired information of each sample
+metadata <- read_tsv("data/raw/metadata.csv") %>%
+  filter(ID %in% colnames(community[, 6:ncol(community)])) %>%
+  select(ID, label) %>%
+  deframe()
+
+# Selection of groups for plotting
+plot <- filter(community, taxlevel==2 |
+                 (taxlevel==4 & str_detect(taxon, "^Chloroplast$")) |
+                 (taxlevel==3 & str_detect(rankID, filter(community, str_detect(taxon, "^Proteobacteria$"))[[2]]))) %>%
+  filter_at(6:ncol(.), any_vars(. >= 1)) %>%
+  mutate_at(5:ncol(.), list(~case_when(taxon=="Cyanobacteria" ~ . - .[taxon=="Chloroplast"], TRUE ~ .))) %>%
+  mutate_at(5:ncol(.), list(~case_when(taxon=="Proteobacteria" ~ . - sum(.[taxlevel==3 & str_detect(rankID, filter(community, str_detect(taxon, "^Proteobacteria$"))[[2]])]), TRUE ~ .))) %>%
+  mutate(taxon=str_replace(taxon, "Proteobacteria", "Other Proteobacteria")) %>%
+  mutate(taxon=str_replace_all(taxon, c("unknown_unclassified"="No_Relative", "unknown"="No_Relative"))) %>%
+  filter_at(6:ncol(.), any_vars(. >= 1)) %>%
+  bind_rows(summarise_all(., list(~ifelse(is.numeric(.), 100-sum(.), paste("Other"))))) %>%
+  rename_at(names(metadata), ~unname(metadata)) %>%
+  arrange(taxon %in% "No_Relative") %>%
+  mutate(mean=rowMeans(select(., contains(c("SCy", "FCyM", "FCaM", "FCa"))), na.rm=TRUE)) %>%
+  mutate(sd=rowSds(as.matrix(select(., contains(c("SCy", "FCyM", "FCaM", "FCa")))), na.rm=TRUE)) %>%
+  arrange(desc(mean))
+```
+
+```{r chloroplast}
+# Loading input data containing sequence abundances and subsequent input data customization
+community <- read_tsv("data/mothur/chloroplast.pds.wang.tax.summary") %>%
+  filter(taxon!="Root")
+
+# Calculating relative abundaces
+community <- group_by(community, taxlevel) %>%
+  mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
+  ungroup()
+
+# Loading metadata of each sample and selection of desired information of each sample
+metadata <- read_tsv("data/raw/metadata.csv") %>%
+  filter(ID %in% colnames(community[, 6:ncol(community)])) %>%
+  select(ID, label) %>%
+  deframe()
+
+# Selection of groups for plotting
+chloroplast <- filter(community, taxlevel==5) %>%
+  rename_at(names(metadata), ~unname(metadata)) %>%
+  mutate(mean=rowMeans(select(., contains(c("SCy", "FCyM", "FCaM", "FCa"))), na.rm=TRUE)) %>%
+  mutate(sd=rowSds(as.matrix(select(., contains(c("SCy", "FCyM", "FCaM", "FCa")))), na.rm=TRUE)) %>%
+  arrange(desc(mean))
+```
+
+To assess the removal efficiency of DNA and protein isolation procedures leaves and thalli were stained with SYBR Green I and examined under a confocal microscope before and after treatments were performed.
+As it is known that marine macrophytes are harboring more algal epiphytes during Autumn and Winter months  [@Reyes2001], the sampling was performed in two contrasting seasons, in December 2017 when higher abundances of algal epiphytes are expected, and in June 2018, when lower abundances could be expected.
+In addition, to further test the isolation protocols on different macrophyte species, DNA and proteins were isolated from the surfaces of two macrophytes, the seagrass *C.
+nodosa* and the macroalga *C.
+cylindracea*.
+*C.
+nodosa* was sampled in a monospecific meadow (Bay of Saline) and *C.
+cylindracea* invaded meadow (Bay of Funtana), while samples of *C.
+cylindracea* originated from a monospecific settlement and a *C.
+cylindracea* invaded *C.
+nodosa* meadow in the Bay of Funtana.
+*C.
+nodosa* June samples from the Bay of Saline were missing as the meadow went to a decline from Spring 2018 (M.
+Najdek, personal communication).
+Procedures developed for DNA and protein isolation showed an almost complete removal of the surface community on both, samples collected in December 2017 (\autoref{confocal_december}) and June 2018 (\autoref{confocal_june}).
+In addition, a similar removal efficiency was observed for communities associated with *C.
+nodosa* and *C.
+cylindracea*.
+Also, no effect of station, settlement or isolation procedure (DNA or protein) on the removal efficiency was observed (Figs.
+\ref{confocal_december} and \ref{confocal_june}).
+
+To evaluate if the obtained DNA is suitable to determine the microbial community structure an Illumina sequencing of the V4 16S rRNA region was performed.
+Sequencing yielded a total of `r number_sequences` sequences after quality curation and exclusion of eukaryotic, mitochondrial and no relative sequences.
+The number of sequences classified as chloroplast was `r number_chloroplast_sequences` (`r proprotion_chloroplast_sequences` \si{\percent})˛ After the exclusion of these sequences the total number of retrieved reads dropped to `r number_sequences_without_chloroplast`, ranging from `r min_max$min` and `r min_max$max` sequences per sample (\autoref{supp-nseq_notus}).
+Even when the highest sequencing effort was applied the rarefaction curves did not level off that is commonly observed in high-throughput 16S rRNA amplicon sequencing procedures (\autoref{supp-rarefaction}).
+Sequence clustering at a similarity level of 97 \si{\percent} yielded a total of `r no_otus` different OTUs.
+Taxonomic classification of reads allowed the determination of the macrophyte associated epiphytic community that was mainly composed of: *Alphaproteobacteria* (`r filter(plot, taxon=="Alphaproteobacteria")$mean` ± `r filter(plot, taxon=="Alphaproteobacteria")$sd` \si{\percent}), *Bacteroidota* (`r filter(plot, taxon=="Bacteroidota")$mean` ± `r filter(plot, taxon=="Bacteroidota")$sd` \si{\percent}), *Gammaproteobacteria* (`r filter(plot, taxon=="Gammaproteobacteria")$mean` ± `r filter(plot, taxon=="Gammaproteobacteria")$sd` \si{\percent}), *Desulfobacterota* (`r filter(plot, taxon=="Desulfobacterota")$mean` ± `r filter(plot, taxon=="Desulfobacterota")$sd` \si{\percent}), *Cyanobacteria* (`r filter(plot, taxon=="Cyanobacteria")$mean` ± `r filter(plot, taxon=="Cyanobacteria")$sd` \si{\percent}) and *Planctomycetota* (`r filter(plot, taxon=="Planctomycetota")$mean` ± `r filter(plot, taxon=="Planctomycetota")$sd` \si{\percent}) (\autoref{community}).
+
+Primers used in this study, as in many other 16S rRNA amplicon procedures, also amplify chloroplast 16S rRNA genes.
+We observed a high proportion of chloroplast sequences in all analyzed samples (`r filter(plot, taxon=="Chloroplast")$mean` ± `r filter(plot, taxon=="Chloroplast")$sd` \si{\percent}) (\autoref{community}).
+To determine if chloroplast sequences originate from hosts or eukaryotic epiphytic organisms, we exported SILVA classified chloroplast sequences and reclassified them using the RDP (Ribosomal Database Project) training set that allows for a more detailed chloroplast classification.
+The largest proportion of sequences was classified as Bacillariophyta (`r filter(chloroplast, taxon=="Bacillariophyta")$mean` ± `r filter(chloroplast, taxon=="Bacillariophyta")$sd` \si{\percent}) indicating that the DNA removal procedure did not coextract larger quantities of host DNA (\autoref{chloroplast}).
 
@@
 ### Sampling
-Leaves of *Cymodocea nodosa* were sampled in a *Cymodocea nodosa* meadow in the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E) and in a *Cymodocea nodosa* meadow invaded by *Caulerpa cylindracea* in the proximity of the village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E).
-Thalli of *Caulerpa cylindracea* were sampled in the same *Cymodocea nodosa* invaded meadow in Funtana and on a locality of only *Caulerpa cylindracea* located in the proximity of the invaded meadow.
+Leaves of *C.
+nodosa* were sampled in a *C.
+nodosa* meadow in the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E) and in a *C.
+nodosa* meadow invaded by *C.
+cylindracea* in the proximity of the village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E).
+Thalli of *C.
+cylindracea* were sampled in the same *C.
+nodosa* invaded meadow in Funtana and on a locality of only *C.
+cylindracea* located in the proximity of the invaded meadow.
 Leaves and thalli were collected on the same day in two contrasting seasons, on 4 December 2017 and 18 June 2018.
-During spring 2018 the *Cymodocea nodosa* meadow in the Bay of Saline decayed to an extent that no leaves could be retrieved (Najdek *et al.*, unpublished data).
+During spring 2018 the *C.
+nodosa* meadow in the Bay of Saline decayed to an extent that no leaves could be retrieved (Najdek *et al.*, unpublished data).
 Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater.
-Upon arrival to the laboratory, *Cymodocea nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *Caulerpa cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections.
+Upon arrival to the laboratory, *C.
+nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *C.
+cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections.
 Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.
@@
 For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Sequences classified as chloroplasts by SILVA were exported and reclassifed using mothur and the RDP (Ribosomal Database Project; http://rdp.cme.msu.edu/) training set (version 16) reference files adapted for mothur [@Cole2014].
+Sequences classified as chloroplasts by SILVA were exported and reclassifed using mothur and the RDP (http://rdp.cme.msu.edu/) training set (version 16) reference files adapted for mothur [@Cole2014].
 In comparison to SILVA, RDP allows a more detailed classification of chloroplast sequences.
-Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014].
+Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020].
 The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020).
@@
 ## Figure Captions
-**\autoref{micro_dec}.** \nameref{micro_dec}
+**\autoref{confocal_december}.** \nameref{confocal_december}
 
-**\autoref{micro_jun}.** \nameref{micro_jun}
+**\autoref{confocal_june}.** \nameref{confocal_june}
 
@@
 cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 4 December 2017 and stained with SYBR Green I.
-Scale bar in all images is 60 \\si{\\um}.\\label{micro_dec}", fig.pos="ht"}
+Scale bar in all images is 60 \\si{\\um}.\\label{confocal_december}", fig.pos="ht"}
 include_graphics("../results/images/confocal_december.jpg")
@@
 cylindracea} surfaces from the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 19 June 2018 and stained with SYBR Green I.
-Scale bar in all images is 60 \\si{\\um}.\\label{micro_jun}", fig.pos="ht"}
+Scale bar in all images is 60 \\si{\\um}.\\label{confocal_june}", fig.pos="ht"}
 include_graphics("../results/images/confocal_june.jpg")

### 2020-07-08 submission/manuscript.tex
Process optimization.

@@
 
+\newpage
+
 \hypertarget{results}{%
 \subsection{Results}\label{results}}
+
+To assess the removal efficiency of DNA and protein isolation procedures
+leaves and thalli were stained with SYBR Green I and examined under a
+confocal microscope before and after treatments were performed. As it is
+known that marine macrophytes are harboring more algal epiphytes during
+Autumn and Winter months (Reyes and Sansón, 2001), the sampling was
+performed in two contrasting seasons, in December 2017 when higher
+abundances of algal epiphytes are expected, and in June 2018, when lower
+abundances could be expected. In addition, to further test the isolation
+protocols on different macrophyte species, DNA and proteins were
+isolated from the surfaces of two macrophytes, the seagrass \emph{C.
+nodosa} and the macroalga \emph{C. cylindracea}. \emph{C. nodosa} was
+sampled in a monospecific meadow (Bay of Saline) and \emph{C.
+cylindracea} invaded meadow (Bay of Funtana), while samples of \emph{C.
+cylindracea} originated from a monospecific settlement and a \emph{C.
+cylindracea} invaded \emph{C. nodosa} meadow in the Bay of Funtana.
+\emph{C. nodosa} June samples from the Bay of Saline were missing as the
+meadow went to a decline from Spring 2018 (M. Najdek, personal
+communication). Procedures developed for DNA and protein isolation
+showed an almost complete removal of the surface community on both,
+samples collected in December 2017 (\autoref{confocal_december}) and
+June 2018 (\autoref{confocal_june}). In addition, a similar removal
+efficiency was observed for communities associated with \emph{C. nodosa}
+and \emph{C. cylindracea}. Also, no effect of station, settlement or
+isolation procedure (DNA or protein) on the removal efficiency was
+observed (Figs. \ref{confocal_december} and \ref{confocal_june}).
+
+To evaluate if the obtained DNA is suitable to determine the microbial
+community structure an Illumina sequencing of the V4 16S rRNA region was
+performed. Sequencing yielded a total of 336,937 sequences after quality
+curation and exclusion of eukaryotic, mitochondrial and no relative
+sequences. The number of sequences classified as chloroplast was 97,328
+(28.9 \si{\percent})˛ After the exclusion of these sequences the total
+number of retrieved reads dropped to 239,609, ranging from 22,587 and
+52,958 sequences per sample (\autoref{supp-nseq_notus}). Even when the
+highest sequencing effort was applied the rarefaction curves did not
+level off that is commonly observed in high-throughput 16S rRNA amplicon
+sequencing procedures (\autoref{supp-rarefaction}). Sequence clustering
+at a similarity level of 97 \si{\percent} yielded a total of 8,360
+different OTUs. Taxonomic classification of reads allowed the
+determination of the macrophyte associated epiphytic community that was
+mainly composed of: \emph{Alphaproteobacteria} (14.9 ± 3.5
+\si{\percent}), \emph{Bacteroidota} (12.5 ± 2.4 \si{\percent}),
+\emph{Gammaproteobacteria} (11.6 ± 4.3 \si{\percent}),
+\emph{Desulfobacterota} (7.8 ± 7.4 \si{\percent}), \emph{Cyanobacteria}
+(6.5 ± 4.7 \si{\percent}) and \emph{Planctomycetota} (2.9 ± 1.7
+\si{\percent}) (\autoref{community}).
+
+Primers used in this study, as in many other 16S rRNA amplicon
+procedures, also amplify chloroplast 16S rRNA genes. We observed a high
+proportion of chloroplast sequences in all analyzed samples (33.4 ± 9.4
+\si{\percent}) (\autoref{community}). To determine if chloroplast
+sequences originate from hosts or eukaryotic epiphytic organisms, we
+exported SILVA classified chloroplast sequences and reclassified them
+using the RDP (Ribosomal Database Project) training set that allows for
+a more detailed chloroplast classification. The largest proportion of
+sequences was classified as Bacillariophyta (89.7 ± 5.7 \si{\percent})
+indicating that the DNA removal procedure did not coextract larger
+quantities of host DNA (\autoref{chloroplast}).
 
@@
 
-Leaves of \emph{Cymodocea nodosa} were sampled in a \emph{Cymodocea
-nodosa} meadow in the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E)
-and in a \emph{Cymodocea nodosa} meadow invaded by \emph{Caulerpa
-cylindracea} in the proximity of the village of Funtana (\ang{45;10;39}
-N, \ang{13;35;42} E). Thalli of \emph{Caulerpa cylindracea} were sampled
-in the same \emph{Cymodocea nodosa} invaded meadow in Funtana and on a
-locality of only \emph{Caulerpa cylindracea} located in the proximity of
-the invaded meadow. Leaves and thalli were collected on the same day in
-two contrasting seasons, on 4 December 2017 and 18 June 2018. During
-spring 2018 the \emph{Cymodocea nodosa} meadow in the Bay of Saline
-decayed to an extent that no leaves could be retrieved (Najdek \emph{et
-al.}, unpublished data). Leaves and thalli were collected by diving and
-transported to the laboratory in containers placed on ice and filled
-with site seawater. Upon arrival to the laboratory, \emph{Cymodocea
-nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while
-\emph{Caulerpa cylindracea} thalli were cut into 5 -- 8 \si{\cm} long
-sections. Leaves and thalli were washed three times with sterile
-artificial seawater (ASW) to remove loosely attached microbial cells.
+Leaves of \emph{C. nodosa} were sampled in a \emph{C. nodosa} meadow in
+the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E) and in a \emph{C.
+nodosa} meadow invaded by \emph{C. cylindracea} in the proximity of the
+village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E). Thalli of
+\emph{C. cylindracea} were sampled in the same \emph{C. nodosa} invaded
+meadow in Funtana and on a locality of only \emph{C. cylindracea}
+located in the proximity of the invaded meadow. Leaves and thalli were
+collected on the same day in two contrasting seasons, on 4 December 2017
+and 18 June 2018. During spring 2018 the \emph{C. nodosa} meadow in the
+Bay of Saline decayed to an extent that no leaves could be retrieved
+(Najdek \emph{et al.}, unpublished data). Leaves and thalli were
+collected by diving and transported to the laboratory in containers
+placed on ice and filled with site seawater. Upon arrival to the
+laboratory, \emph{C. nodosa} leaves were cut into sections of 1 -- 2
+\si{\cm}, while \emph{C. cylindracea} thalli were cut into 5 -- 8
+\si{\cm} long sections. Leaves and thalli were washed three times with
+sterile artificial seawater (ASW) to remove loosely attached microbial
+cells.
 
@@
 Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
-SILVA were exported and reclassifed using mothur and the RDP (Ribosomal
-Database Project; \url{http://rdp.cme.msu.edu/}) training set (version
-16) reference files adapted for mothur (Cole \emph{et al.}, 2014). In
-comparison to SILVA, RDP allows a more detailed classification of
-chloroplast sequences. Pipeline data processing and visualization was
-done using R (version 3.6.0) (R Core Team, 2019), package tidyverse
-(version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple other
-packages (Xie, 2014, 2015, 2019b, 2019a, 2020; Neuwirth, 2014; Xie
-\emph{et al.}, 2018; Allaire \emph{et al.}, 2019; Zhu, 2019). The
-detailed analysis procedure including the R Markdown file for this paper
-are available as a GitHub repository
+SILVA were exported and reclassifed using mothur and the RDP
+(\url{http://rdp.cme.msu.edu/}) training set (version 16) reference
+files adapted for mothur (Cole \emph{et al.}, 2014). In comparison to
+SILVA, RDP allows a more detailed classification of chloroplast
+sequences. Pipeline data processing and visualization was done using R
+(version 3.6.0) (R Core Team, 2019), package tidyverse (version 1.3.0)
+(Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
+2015, 2019b, 2019a, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018;
+Allaire \emph{et al.}, 2019; Zhu, 2019; Bengtsson, 2020). The detailed
+analysis procedure including the R Markdown file for this paper are
+available as a GitHub repository
 (\url{https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).
@@
 \textbf{75}: 129--137.
+
+\leavevmode\hypertarget{ref-Bengtsson2020}{}%
+Bengtsson, H. (2020) MatrixStats: Functions that apply to rows and
+columns of matrices (and to vectors).
 
@@
 
+\leavevmode\hypertarget{ref-Reyes2001}{}%
+Reyes, J. and Sansón, M. (2001) Biomass and production of the epiphytes
+on the leaves of \emph{Cymodocea nodosa} in the Canary Islands.
+\emph{Bot Mar} \textbf{44}: 307--313.
+
 \leavevmode\hypertarget{ref-Richter-Heitmann2016}{}%
@@
 
-\textbf{\autoref{micro_dec}.} \nameref{micro_dec}
+\textbf{\autoref{confocal_december}.} \nameref{confocal_december}
 
-\textbf{\autoref{micro_jun}.} \nameref{micro_jun}
+\textbf{\autoref{confocal_june}.} \nameref{confocal_june}
 
@@
 cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 4 December 2017 and stained with SYBR Green I.
-Scale bar in all images is 60 \si{\um}.\label{micro_dec}}\label{fig:unnamed-chunk-2}
+Scale bar in all images is 60 \si{\um}.\label{confocal_december}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 cylindracea} surfaces from the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 19 June 2018 and stained with SYBR Green I.
-Scale bar in all images is 60 \si{\um}.\label{micro_jun}}\label{fig:unnamed-chunk-3}
+Scale bar in all images is 60 \si{\um}.\label{confocal_june}}\label{fig:unnamed-chunk-3}
 \end{figure}

### 2020-07-08 code/plot_chloroplast.R

@@ -12,7 +12,7 @@
 community <- read_tsv("data/mothur/chloroplast.pds.wang.tax.summary") %>%
   filter(taxon!="Root")
 
-# Calculate relative abundaces
+# Calculating relative abundaces
 community <- group_by(community, taxlevel) %>%
   mutate_at(5:ncol(.), list(~. / sum(.) * 100)) %>%
   ungroup()

### 2020-07-08 code/plot_rarefaction.R

@@ -59,8 +59,8 @@ p <- metadata_rarefaction %>%
            "19/6-18 FCa"=parse(text="plain('Funtana, ')~plain('19 June 2018, ')~italic('Cymodocea nodosa ')~plain('(Monospecific)')"))) +
   labs(x="Number of Sequences", y="Number of OTUs") +
   theme +
-  theme(legend.position=c(0.67, 0.18), legend.title=element_blank(),
-        legend.text=element_text(size=10, margin=margin(r=0.2, unit="cm")),
+  theme(legend.position=c(0.58, 0.18), legend.title=element_blank(),
+        legend.text=element_text(size=14, margin=margin(r=0.2, unit="cm")),
         legend.key.width=unit(1.4, "cm"), legend.key.height=unit(0.5, "cm"),
         legend.key=element_rect(fill="white"), legend.justification=c("top"),
         legend.text.align=0, legend.spacing.x=unit(0, "cm")) +

### 2020-07-10 submission/manuscript.Rmd
Metagenomic part in M&M done.

@@
   arrange(taxon %in% "No_Relative") %>%
-  mutate(mean=rowMeans(select(., contains(c("SCy", "FCyM", "FCaM", "FCa"))), na.rm=TRUE)) %>%
-  mutate(sd=rowSds(as.matrix(select(., contains(c("SCy", "FCyM", "FCaM", "FCa")))), na.rm=TRUE)) %>%
+  mutate(mean=rowMeans(select(., contains("SCy"), contains("FCyM"), contains("FCaM"), contains("FCa")), na.rm=TRUE)) %>%
+  mutate(sd=rowSds(as.matrix(select(., contains("SCy"), contains("FCyM"), contains("FCaM"), contains("FCa"))), na.rm=TRUE)) %>%
   arrange(desc(mean))
@@
 
-# Calculating relative abundaces
+# Calculating relative abundances
 community <- group_by(community, taxlevel) %>%
@@
   rename_at(names(metadata), ~unname(metadata)) %>%
-  mutate(mean=rowMeans(select(., contains(c("SCy", "FCyM", "FCaM", "FCa"))), na.rm=TRUE)) %>%
-  mutate(sd=rowSds(as.matrix(select(., contains(c("SCy", "FCyM", "FCaM", "FCa")))), na.rm=TRUE)) %>%
+  mutate(mean=rowMeans(select(., contains("SCy"), contains("FCyM"), contains("FCaM"), contains("FCa")), na.rm=TRUE)) %>%
+  mutate(sd=rowSds(as.matrix(select(., contains("SCy"), contains("FCyM"), contains("FCaM"), contains("FCa"))), na.rm=TRUE)) %>%
   arrange(desc(mean))
@@
 
-### Sequence analysis
 ```{r}
@@
 In addition, the negative control processed together with the samples yielded only `r nc` sequences after sequence quality curation.
+
+### Metagenomics
+Four samples were selected and sent on dry ice to IMGM Laboratories, Martinsried, Germany for metagenomic sequencing.
+DNA was purified using AMPure XP Beads (Beckman Coulter, USA) applying a bead:DNA ratio of 1:1 (v/v), quantified with a Qubit dsDNA HS Assay Kit (Thermo Fisher Scientific, USA) and check for integrity on a 1 \si{\percent} agarose gel.
+Metagenomic sequencing libraries were generated from 300 \si{\ng} of input DNA using a NEBNext Ultra II FS DNA Library Prep Kit for Illumina (New England Biolabs, USA) according to the manufacturer's instructions.
+Fragments were selected (500 -- 700 bp) using AMPure XP Beads, PCR enriched for 3 -- 5 cycles and quality controlled.
+Libraries generated from different DNA samples were pooled and sequenced on an Illumina NovaSeq 6000 sequencing system (2 × 250 bp).
+
+Obtained sequences were assembled individually using MEGAHIT (version  1.1.2) [@Li2015] under default settings.
+Putative genes were predicted from contings longer than 200 bp using Prodigal (version  2.6.3) [@Hyatt2010] in metagenome mode (-p meta).
+Abundances of predicted genes were expressed as Reads Per Kilobase Million (RPKM) and calculated using the BWA algorithm (version  0.7.16a) [@Li2009a].
+All predicted genes were functionally annotated using the eggNOG-mapper [@Huerta-Cepas2017] and eggNOG database (version 5.0) [@Huerta-Cepas2019].
+The taxonomic classification of sequences was determined using the lowest common ancestor algorithm adapted from DIAMOND (version 0.8.36) [@Buchfink2015] and by searching against the non-redundant database (NR).
+To determine phylogeny the top 10 \si{\percent} hits with an e-value < 1 × 10^-5^ were used (\-\-top 10).
 

### 2020-07-10 submission/manuscript.tex
Metagenomic part in M&M done.

@@
-\documentclass[12pt,]{article}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  12pt,
+]{article}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Selective DNA and Protein Isolation from Marine Macrophyte Surfaces},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1.0in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
-            pdftitle={Selective DNA and Protein Isolation from Marine Macrophyte Surfaces},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
-Surfaces}}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{}
-    \preauthor{}\postauthor{}
-    \date{}
-    \predate{}\postdate{}
-
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 \usepackage{times} % Times New Roman font
@@
 \renewcommand{\figureautorefname}{Fig.}
+
+\title{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
+Surfaces}}
+\author{}
+\date{\vspace{-2.5em}}
 
@@
 
-\hypertarget{sequence-analysis}{%
-\subsubsection{Sequence analysis}\label{sequence-analysis}}
-
 Obtained sequences were analyzed on the computer cluster Isabella
@@
 (Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
-2015, 2019b, 2019a, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018;
-Allaire \emph{et al.}, 2019; Zhu, 2019; Bengtsson, 2020). The detailed
-analysis procedure including the R Markdown file for this paper are
-available as a GitHub repository
+2015, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018; Y. Xie, 2019b,
+2019a; Allaire \emph{et al.}, 2019; Zhu, 2019; Bengtsson, 2020). The
+detailed analysis procedure including the R Markdown file for this paper
+are available as a GitHub repository
 (\url{https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).
@@
 sequences after sequence quality curation.
+
+\hypertarget{metagenomics}{%
+\subsubsection{Metagenomics}\label{metagenomics}}
+
+Four samples were selected and sent on dry ice to IMGM Laboratories,
+Martinsried, Germany for metagenomic sequencing. DNA was purified using
+AMPure XP Beads (Beckman Coulter, USA) applying a bead:DNA ratio of 1:1
+(v/v), quantified with a Qubit dsDNA HS Assay Kit (Thermo Fisher
+Scientific, USA) and check for integrity on a 1 \si{\percent} agarose
+gel. Metagenomic sequencing libraries were generated from 300 \si{\ng}
+of input DNA using a NEBNext Ultra II FS DNA Library Prep Kit for
+Illumina (New England Biolabs, USA) according to the manufacturer's
+instructions. Fragments were selected (500 -- 700 bp) using AMPure XP
+Beads, PCR enriched for 3 -- 5 cycles and quality controlled. Libraries
+generated from different DNA samples were pooled and sequenced on an
+Illumina NovaSeq 6000 sequencing system (2 × 250 bp).
+
+Obtained sequences were assembled individually using MEGAHIT (version
+1.1.2) (Li \emph{et al.}, 2015) under default settings. Putative genes
+were predicted from contings longer than 200 bp using Prodigal (version
+2.6.3) (Hyatt \emph{et al.}, 2010) in metagenome mode (-p meta).
+Abundances of predicted genes were expressed as Reads Per Kilobase
+Million (RPKM) and calculated using the BWA algorithm (version 0.7.16a)
+(Li and Durbin, 2009). All predicted genes were functionally annotated
+using the eggNOG-mapper (Huerta-Cepas \emph{et al.}, 2017) and eggNOG
+database (version 5.0) (Huerta-Cepas \emph{et al.}, 2019). The taxonomic
+classification of sequences was determined using the lowest common
+ancestor algorithm adapted from DIAMOND (version 0.8.36) (Buchfink
+\emph{et al.}, 2015) and by searching against the non-redundant database
+(NR). To determine phylogeny the top 10 \si{\percent} hits with an
+e-value \textless{} 1 × 10\textsuperscript{-5} were used (-\/-top 10).
 
@@
 Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
-A., et al. (2019) Rmarkdown: Dynamic documents for R.
+A., et al. (2019) rmarkdown: Dynamic documents for R.
 
@@
 columns of matrices (and to vectors).
+
+\leavevmode\hypertarget{ref-Buchfink2015}{}%
+Buchfink, B., Xie, C., and Huson, D.H. (2015) Fast and sensitive protein
+alignment using DIAMOND. \emph{Nat Methods} \textbf{12}: 59--60.
 
@@
 
+\leavevmode\hypertarget{ref-Huerta-Cepas2017}{}%
+Huerta-Cepas, J., Forslund, K., Coelho, L.P., Szklarczyk, D., Jensen,
+L.J., von Mering, C., and Bork, P. (2017) Fast genome-wide functional
+annotation through orthology assignment by eggNOG-mapper. \emph{Mol Biol
+Evol} \textbf{34}: 2115--2122.
+
+\leavevmode\hypertarget{ref-Huerta-Cepas2019}{}%
+Huerta-Cepas, J., Szklarczyk, D., Heller, D., Hernández-Plaza, A.,
+Forslund, S.K., Cook, H., et al. (2019) eggNOG 5.0: A hierarchical,
+functionally and phylogenetically annotated orthology resource based on
+5090 organisms and 2502 viruses. \emph{Nucleic Acids Res} \textbf{47}:
+D309--D314.
+
 \leavevmode\hypertarget{ref-Hultman2015}{}%
@@
 208--212.
+
+\leavevmode\hypertarget{ref-Hyatt2010}{}%
+Hyatt, D., Chen, G.-L., LoCascio, P.F., Land, M.L., Larimer, F.W., and
+Hauser, L.J. (2010) Prodigal: Prokaryotic gene recognition and
+translation initiation site identification. \emph{BMC Bioinformatics}
+\textbf{11}: 119.
 
@@
 5112--5120.
+
+\leavevmode\hypertarget{ref-Li2015}{}%
+Li, D., Liu, C.-M., Luo, R., Sadakane, K., and Lam, T.-W. (2015)
+MEGAHIT: An ultra-fast single-node solution for large and complex
+metagenomics assembly via succinct de Bruijn graph.
+\emph{Bioinformatics} \textbf{31}: 1674--1676.
+
+\leavevmode\hypertarget{ref-Li2009a}{}%
+Li, H. and Durbin, R. (2009) Fast and accurate short read alignment with
+BurrowsWheeler transform. \emph{Bioinformatics} \textbf{25}: 1754--1760.
 
@@
 \leavevmode\hypertarget{ref-RCoreTeam2019}{}%
-R Core Team (2019) A language and environment for statistical computing,
-Vienna, Austria: R Foundation for Statistical Computing.
+R Core Team (2019) R: A language and environment for statistical
+computing, Vienna, Austria: R Foundation for Statistical Computing.
 
@@
 \leavevmode\hypertarget{ref-Xie2014}{}%
-Xie, Y. (2014) Knitr: A comprehensive tool for reproducible research in
+Xie, Y. (2014) knitr: A comprehensive tool for reproducible research in
 R. In \emph{Implementing Reproducible Computational Research}. Stodden,
@@
 \leavevmode\hypertarget{ref-Xie2019}{}%
-Xie, Y. (2019a) Knitr: A general-purpose package for dynamic report
+Xie, Y. (2019a) knitr: A general-purpose package for dynamic report
 generation in R.
@@
 \leavevmode\hypertarget{ref-Xie2020}{}%
-Xie, Y. (2020) TinyTex: Helper functions to install and maintain 'TeX
+Xie, Y. (2020) tinytex: Helper functions to install and maintain 'TeX
 Live', and compile 'LaTeX' documents.
@@
 \leavevmode\hypertarget{ref-Zhu2019}{}%
-Zhu, H. (2019) KableExtra: Construct complex table with 'kable' and pipe
+Zhu, H. (2019) kableExtra: Construct complex table with 'kable' and pipe
 syntax.

### 2020-09-14 submission/manuscript.Rmd
Manuscript done.

@@
 
-3\. Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands
+3\. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands
 
@@
 ## Summary
+Studies of unculturable microbes often combine methods based on DNA, such as 16S rRNA sequencing and metagenomics, with methods that allow an insight into the metabolic status, such as metaproteomics.
+To apply these procedures to the microbial community inhabiting the surfaces of marine macrophytes it is advisable to perform, prior to the analysis, a selective DNA and protein isolation so that the host material, present in higher quantities, is not hampering the analysis.
+Two protocols, for DNA and protein isolation, were adapted for selective extractions of DNA and proteins from epiphytic communities inhabiting the surfaces of two marine macrophytes: the seagrass *Cymodocea nodosa* and the macroalga *Caulerpa cylindracea*.
+Protocols showed an almost complete removal of the epiphytic community regardless of the sampling season, station, settlement or host species.
+The obtained DNA was suitable for metagenomic and 16S rRNA sequencing, while isolated proteins could be identified by mass spectrometry showing that protocols can be used in 16S rRNA, metagenomic and metaproteomic analysis.
+Low presence of host DNA and proteins, observed in isolated samples, indicated a selective nature of protocols.
+Furthermore, the procedures are based on universally available laboratory chemicals making the protocols widely applicable.
+Taken together, the adapted protocols ensure an almost complete removal of the macrophyte epiphytic community, are selective for microbes inhabiting macrophyte surfaces and are providing DNA and proteins applicable in 16S rRNA sequencing, metagenomics and metaproteomics.
 
@@
 proteins and DNA) from pelagic microbial communities is usually isolated by collecting cells onto filters and subsequently isolating the aimed compound [@Gilbert2009].
-If a microbial size fraction is aimed sequential filtration is applied [@Massana1997; @Andersson2010].
+If a specific microbial size fraction is aimed sequential filtration is applied [@Massana1997; @Andersson2010].
 In contrast, obtaining biological materials from microorganism inhabiting surfaces requires either a cell detachment procedure prior to isolation or the host material is coextracted together with the targeted material.
@@
 An alternative to these procedures is a direct isolation of the targeted material by incubating macrophyte tissues in an extraction buffer.
-After the incubation is done the undisrupted host tissue is removed and the isolation procedure continues omitting host material contaminations.
+After the incubation, the undisrupted host tissue is removed and the isolation procedure continues omitting host material contaminations.
 To our knowledge the only procedure describing a direct and selective epiphytic DNA isolation from the surfaces of marine macrophytes was provided by @Burke2009.
-In contrast to previously described methods this protocol enables an almost complete removal of the surface community and was used for 16S rRNA gene clone libraries construction [@Burke2011a] and metagenomes sequencing [@Burke2011].
-This method, thought providing a selective isolation procedure, is using in the extraction buffer a rapid multienzyme cleaner (3M) that is not worldwide available and whose composition is not know [@Burke2009].
+In contrast to previously described methods this protocol enables an almost complete removal of the surface community and it was used for 16S rRNA gene clone library construction [@Burke2011a] and metagenome sequencing [@Burke2011].
+This method, thought providing a selective isolation procedure, uses a rapid multienzyme cleaner (3M) that is not available worldwide and without a known composition [@Burke2009].
 Also to our knowledge, no selective isolation protocol for proteins from epiphytic communities inhabiting marine macrophytes was established.
 
-In the present study, we adapted a protocol widely used for DNA isolation from filters [@Massana1997] and a protocol used for protein isolation from sediments [@Chourey2010; @Hultman2015] for selective extractions of DNA and proteins from epiphytic communities inhabiting the surfaces of two marine macrophytes: the seagrass *Cymodocea nodosa* and the alga *Caulerpa cylindracea*.
-In addition, we tested the removal efficiency of the protocol and the suitability of obtained DNA and proteins for 16S rRNA sequencing and metaproteomics.
+In the present study, we adapted a protocol widely used for DNA isolation from filters [@Massana1997] and a protocol used for protein isolation from sediments [@Chourey2010; @Hultman2015] for selective extractions of DNA and proteins from epiphytic communities inhabiting the surfaces of two marine macrophytes: the seagrass *Cymodocea nodosa* and the macroalga *Caulerpa cylindracea*.
+In addition, we tested the removal efficiency of the protocol and the suitability of obtained DNA and proteins for 16S rRNA sequencing, metagenomics and metaproteomics.
 
@@
 number_sequences_without_chloroplast <- number_sequences - number_chloroplast_sequences
-
-# Calculating the proprotion of chloroplast sequences
-proprotion_chloroplast_sequences <- (number_chloroplast_sequences / number_sequences) * 100
 
@@
 
-# Selection of groups for plotting
+# Selecting groups for plotting
 chloroplast <- filter(community, taxlevel==5) %>%
@@
   arrange(desc(mean))
+
+# Calculating proportion of Streptophyta in C. nodosa associated communities
+streptophyta_16S <- chloroplast %>%
+  filter(taxon=="Streptophyta") %>%
+  select(contains("Cy")) %>%
+  mutate(mean=rowMeans(select(., contains("Cy")))) %>%
+  mutate(sd=rowSds(as.matrix(select(., contains("Cy")))))
+
+# Calculating proportion of Chlorophyta in C. cylindracea associated communities
+chlorophyta_16S <- chloroplast %>%
+  filter(taxon=="Chlorophyta") %>%
+  select(contains("Ca")) %>%
+  mutate(mean=rowMeans(select(., contains("Ca")))) %>%
+  mutate(sd=rowSds(as.matrix(select(., contains("Ca")))))
+```
+
+```{r metagenomic_statistics}
+# Metagenomic statistics
+# Loading raw sequence statistics
+raw_45_R1 <- read_tsv("results/metagenomics/statistics/00raw_45_R1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+raw_45_R2 <- read_tsv("results/metagenomics/statistics/00raw_45_R2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+raw_47_R1 <- read_tsv("results/metagenomics/statistics/00raw_47_R1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+raw_47_R2 <- read_tsv("results/metagenomics/statistics/00raw_47_R2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+raw_61_R1 <- read_tsv("results/metagenomics/statistics/00raw_61_R1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+raw_61_R2 <- read_tsv("results/metagenomics/statistics/00raw_61_R2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+raw_63_R1 <- read_tsv("results/metagenomics/statistics/00raw_63_R1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+raw_63_R2 <- read_tsv("results/metagenomics/statistics/00raw_63_R2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+# Loading sequence statistics following AdapterRemoval
+truncated_45_R1 <- read_tsv("results/metagenomics/statistics/01truncated_45_pair1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+truncated_45_R2 <- read_tsv("results/metagenomics/statistics/01truncated_45_pair2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+truncated_47_R1 <- read_tsv("results/metagenomics/statistics/01truncated_47_pair1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+truncated_47_R2 <- read_tsv("results/metagenomics/statistics/01truncated_47_pair2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+truncated_61_R1 <- read_tsv("results/metagenomics/statistics/01truncated_61_pair1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+truncated_61_R2 <- read_tsv("results/metagenomics/statistics/01truncated_61_pair2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+truncated_63_R1 <- read_tsv("results/metagenomics/statistics/01truncated_63_pair1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+truncated_63_R2 <- read_tsv("results/metagenomics/statistics/01truncated_63_pair2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+# Loading sequence statistics following MEGAHIT
+contigs_45 <- read_tsv("results/metagenomics/statistics/03contigs_45.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:" | X1=="Main genome contig N/L50:") %>%
+  separate(X2, c("X2", "X3"), sep="/", fill="right")
+contigs_47 <- read_tsv("results/metagenomics/statistics/03contigs_47.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:" | X1=="Main genome contig N/L50:") %>%
+  separate(X2, c("X2", "X3"), sep="/", fill="right")
+contigs_61 <- read_tsv("results/metagenomics/statistics/03contigs_61.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:" | X1=="Main genome contig N/L50:") %>%
+  separate(X2, c("X2", "X3"), sep="/", fill="right")
+contigs_63 <- read_tsv("results/metagenomics/statistics/03contigs_63.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:" | X1=="Main genome contig N/L50:") %>%
+  separate(X2, c("X2", "X3"), sep="/", fill="right")
+
+# Loading sequence statistics following Prodigal
+cds_45 <- read_tsv("results/metagenomics/statistics/04cds_nucl_45.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+cds_47 <- read_tsv("results/metagenomics/statistics/04cds_nucl_47.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+cds_61 <- read_tsv("results/metagenomics/statistics/04cds_nucl_61.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+cds_63 <- read_tsv("results/metagenomics/statistics/04cds_nucl_63.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+# Loading sequence statistics following eggNOG mapper
+eggnog_45 <- read_tsv("results/metagenomics/statistics/07eggnog_45.txt", col_names=FALSE)
+eggnog_47 <- read_tsv("results/metagenomics/statistics/07eggnog_47.txt", col_names=FALSE)
+eggnog_61 <- read_tsv("results/metagenomics/statistics/07eggnog_61.txt", col_names=FALSE)
+eggnog_63 <- read_tsv("results/metagenomics/statistics/07eggnog_63.txt", col_names=FALSE)
+
+# Combining statistical data in a table
+metagenomic_statistics <- tibble(
+  ID=c(45, 47, 61, 63),
+  raw_R1=c(raw_45_R1$X2, raw_47_R1$X2, raw_61_R1$X2, raw_63_R1$X2),
+  raw_R2=c(raw_45_R2$X2, raw_47_R2$X2, raw_61_R2$X2, raw_63_R2$X2),
+  truncated_R1=c(truncated_45_R1$X2, truncated_47_R1$X2, truncated_61_R1$X2, truncated_63_R1$X2),
+  truncated_R2=c(truncated_45_R2$X2, truncated_47_R2$X2, truncated_61_R2$X2, truncated_63_R2$X2),
+  contigs=c(contigs_45$X2[1], contigs_47$X2[1], contigs_61$X2[1], contigs_63$X2[1]),
+  N50=c(contigs_45$X2[2], contigs_47$X2[2], contigs_61$X2[2], contigs_63$X2[2]),
+  L50=c(contigs_45$X3[2], contigs_47$X3[2], contigs_61$X3[2], contigs_63$X3[2]),
+  cds=c(cds_45$X2, cds_47$X2, cds_61$X2, cds_63$X2),
+  eggnog=c(eggnog_45$X2, eggnog_47$X2, eggnog_61$X2, eggnog_63$X2)
+  ) %>%
+  mutate(L50=str_replace(L50, " KB", "")) %>%
+  mutate(L50=str_replace(L50, "\\.", ""))
+
+# Loading metagenomic metadata 
+metadata_metagenomics <- read_tsv("data/raw/metadata_metagenomics.csv")
+
+# Joining metagenomic metadata with metagenomic statistical data
+metadata_metagenomic_statistics <- inner_join(metadata_metagenomics, metagenomic_statistics, by=c("ID"="ID")) %>%
+  select(ID, station, date, raw_R1, contigs, N50, L50, cds, eggnog) %>%
+  mutate_at(vars(c(raw_R1, contigs, N50, L50, cds, eggnog)), list(~as.numeric(.)))
+```
+
+```{r cog_metagenomes}
+# Loading COG input data and combining together
+cog_45 <- read_tsv("results/metagenomics/cog/45.cog_categories.tsv", col_names = c("COG Functional Category", "45"), skip = 1)
+cog_47 <- read_tsv("results/metagenomics/cog/47.cog_categories.tsv", col_names = c("COG Functional Category", "47"), skip = 1)
+cog_61 <- read_tsv("results/metagenomics/cog/61.cog_categories.tsv", col_names = c("COG Functional Category", "61"), skip = 1)
+cog_63 <- read_tsv("results/metagenomics/cog/63.cog_categories.tsv", col_names = c("COG Functional Category", "63"), skip = 1)
+
+cog <- full_join(cog_45, cog_47, by = c("COG Functional Category")) %>%
+  full_join(., cog_61, by = c("COG Functional Category")) %>%
+  full_join(., cog_63, by = c("COG Functional Category"))
+
+# Calculating relative abundances of each COG functional category
+cog_relative <- cog %>%
+  mutate(`COG Functional Category`, `COG Functional Category`=if_else(str_length(`COG Functional Category`) > 1,
+         "multiple_COG", `COG Functional Category`)) %>%
+  mutate(`COG Functional Category`, `COG Functional Category`=if_else(is.na(`COG Functional Category`), 
+         "no_COG", `COG Functional Category`)) %>%
+  replace(is.na(.), 0) %>%
+  group_by(`COG Functional Category`) %>%
+  summarize(`14/12-17 FCyM`=sum(`45`),
+            `14/12-17 FCa`=sum(`47`),
+            `19/6-18 FCyM`=sum(`61`),
+            `19/6-18 FCa`=sum(`63`)) %>%
+  mutate_at(2:ncol(.), list(~. / sum(.) * 100)) %>%
+  mutate(mean=rowMeans(select(., contains("FCyM"), contains("FCa")))) %>%
+  arrange(desc(mean))
+
+# Loading COG functional categories names
+names <- read_tsv("data/raw/cog_categories_colors.tsv")
+
+# Adding category description to relative abundances of each COG functional category
+cog_relative <- left_join(cog_relative, names, by=c("COG Functional Category"="Code"))
+```
+
+```{r metagenomic_taxonomy}
+# Metagenomic taxonomy
+# Loading summary taxonomy files (phylum)
+phylum_45 <- read_tsv("results/metagenomics/taxonomy/45.phylum.tsv") %>%
+  add_column(sample = 45, .before = "phylum") %>%
+  filter(!is.na(phylum)) %>%
+  mutate(cds_relative = cds / sum(cds) * 100) %>%
+  mutate(rpkm_relative = rpkm / sum(rpkm) * 100) %>%
+  arrange(desc(rpkm_relative)) %>%
+  top_n(10)
+phylum_47 <- read_tsv("results/metagenomics/taxonomy/47.phylum.tsv") %>%
+  add_column(sample = 47, .before = "phylum") %>%
+  filter(!is.na(phylum)) %>%
+  mutate(cds_relative = cds / sum(cds) * 100) %>%
+  mutate(rpkm_relative = rpkm / sum(rpkm) * 100) %>%
+  arrange(desc(rpkm_relative)) %>%
+  top_n(10)
+phylum_61 <- read_tsv("results/metagenomics/taxonomy/61.phylum.tsv") %>%
+  add_column(sample = 61, .before = "phylum") %>%
+  filter(!is.na(phylum)) %>%
+  mutate(cds_relative = cds / sum(cds) * 100) %>%
+  mutate(rpkm_relative = rpkm / sum(rpkm) * 100) %>%
+  arrange(desc(rpkm_relative)) %>%
+  top_n(10)
+phylum_63 <- read_tsv("results/metagenomics/taxonomy/63.phylum.tsv") %>%
+  add_column(sample = 63, .before = "phylum") %>%
+  filter(!is.na(phylum)) %>%
+  mutate(cds_relative = cds / sum(cds) * 100) %>%
+  mutate(rpkm_relative = rpkm / sum(rpkm) * 100) %>%
+  arrange(desc(rpkm_relative)) %>%
+  top_n(10)
+
+# Combining summary taxonomy files (phylum)
+metagenomic_phylum <- bind_rows(phylum_45, phylum_47, phylum_61, phylum_63)
+
+# Loading metagenomic metadata 
+metadata_metagenomics <- read_tsv("data/raw/metadata_metagenomics.csv")
+
+# Joining metagenomic metadata with metagenomic taxonomy data
+metadata_metagenomic_phylum <- inner_join(metadata_metagenomics, metagenomic_phylum, by=c("ID"="sample")) %>%
+  select(label, phylum, cds, cds_relative, rpkm, rpkm_relative)
+```
+
+```{r cog_metaproteomes}
+###################################################################################
+# Number of identified proteins and their COG functional category in metaproteomes 
+###################################################################################
+
+# Loading metaproteomic COG input data
+metap_cog <- read_tsv("results/metaproteomics/metaproteomics_cog.tsv")
+
+# Adding category "Multiple functional categories", renaming category NA to
+# "No COG category available"
+metap_cog <- metap_cog %>%
+  mutate(`COG Functional Category`= if_else(str_length(`COG Functional Category`) > 1,
+                                            "multiple_COG", `COG Functional Category`)) %>%
+  replace_na(list(`COG Functional Category` = "no_COG")) %>%
+  mutate(cog_40 = if_else(is.na(`Abundance: 40: Sample`), 0, 1),
+         cog_41 = if_else(is.na(`Abundance: 41: Sample`), 0, 1),
+         cog_42 = if_else(is.na(`Abundance: 42: Sample`), 0, 1),
+         cog_43 = if_else(is.na(`Abundance: 43: Sample`), 0, 1),
+         cog_61 = if_else(is.na(`Abundance: 61: Sample`), 0, 1),
+         cog_62 = if_else(is.na(`Abundance: 62: Sample`), 0, 1),
+         cog_63 = if_else(is.na(`Abundance: 63: Sample`), 0, 1))
+
+# Transforming abundance data to NAAF (Normalized Area Abundance Factor)
+metap_cog <- metap_cog %>%
+  mutate_at(vars(contains("Abundance")), list(~. / `# AAs`)) %>%
+  mutate_at(vars(contains("Abundance")), list(~./ sum(., na.rm = TRUE) * 100))
+
+# Calculating the number of proteins in each COG functional category
+metap_cog_abund <- metap_cog %>%
+  group_by(`COG Functional Category`) %>%
+  summarize(`40` = sum(cog_40),
+            `41` = sum(cog_41),
+            `42` = sum(cog_42),
+            `43` = sum(cog_43),
+            `61` = sum(cog_61),
+            `62` = sum(cog_62),
+            `63` = sum(cog_63))
+
+# Calculating the sum of identified proteins in each sample
+metap_cog_num <- metap_cog_abund %>%
+  select_if(is.numeric) %>%
+  colSums()
+
+# Calculating the relative contribution of proteins to each COG functional category
+metap_cog_relative <- metap_cog_abund %>%
+  mutate_at(2:ncol(.), list(~. / sum(.) * 100)) %>%
+  mutate(mean=rowMeans(select_if(., is.numeric))) %>%
+  arrange(desc(mean))
+
+# Loading COG functional categories names
+names <- read_tsv("data/raw/cog_categories_colors.tsv")
+
+# Adding category description to relative abundances of each COG functional category
+metap_cog_relative <- left_join(metap_cog_relative, names, by=c("COG Functional Category"="Code"))
+
+###################################################################################
+# Number and abundance of identified proteins associated with host species
+###################################################################################
+
+# Loading taxa belonging to phyla Streptophyta and Chlorophyta
+streptophyta_meta <- read_tsv("results/metaproteomics/streptophyta_names.tsv")
+chlorophyta_meta <- read_tsv("results/metaproteomics/chlorophyta_names.tsv")
+
+# Calculating the relative contribution of each protein to the total number of proteins
+metap_cog <- metap_cog %>%
+  mutate_at(vars(contains('cog_')), list(relative = ~ . / sum(.) * 100))
+
+# Loading metadata
+metadata <- read_tsv("data/raw/metadata.csv")
+
+# Filtering proteins associated with members of the phylum Streptophyta and calculating their contribution
+metap_cog_streptophyta <- metap_cog %>%
+  filter(`Predicted taxonomic group` %in% streptophyta_meta$names) %>%
+  filter(!is.na(`Predicted taxonomic group`)) %>%
+  summarize(cog_40_relative = sum(cog_40_relative),
+            cog_41_relative = sum(cog_41_relative),
+            cog_42_relative = sum(cog_42_relative),
+            cog_43_relative = sum(cog_43_relative),
+            cog_61_relative = sum(cog_61_relative),
+            cog_62_relative = sum(cog_62_relative),
+            cog_63_relative = sum(cog_63_relative),
+            abundance_40 = sum(`Abundance: 40: Sample`, na.rm = TRUE),
+            abundance_41 = sum(`Abundance: 41: Sample`, na.rm = TRUE),
+            abundance_42 = sum(`Abundance: 42: Sample`, na.rm = TRUE),
+            abundance_43 = sum(`Abundance: 43: Sample`, na.rm = TRUE),
+            abundance_61 = sum(`Abundance: 61: Sample`, na.rm = TRUE),
+            abundance_62 = sum(`Abundance: 62: Sample`, na.rm = TRUE),
+            abundance_63 = sum(`Abundance: 63: Sample`, na.rm = TRUE)) %>%
+  gather(key="sample", value="abundance") %>%
+  mutate(ID=str_replace(sample, "cog_", "")) %>%
+  mutate(ID=str_replace(ID, "abundance_", "")) %>%
+  mutate(ID=str_replace(ID, "_relative", "")) %>%
+  left_join(., metadata, by="ID") %>%
+  mutate(host = if_else(str_detect(station, "Cy"), "Cy", "Ca")) %>%
+  mutate(cog_abundance = if_else(str_detect(sample, "_relative"), "cog", "abundance")) %>%
+  group_by(host, cog_abundance) %>%
+  summarize(mean = mean(abundance), sd = sd(abundance))
+
+# Filtering proteins associated with members of the phylum Chlorophyta and calculating their contribution
+metap_cog_chlorophyta <- metap_cog %>%
+  filter(`Predicted taxonomic group` %in% chlorophyta_meta$names) %>%
+  filter(!is.na(`Predicted taxonomic group`)) %>%
+  summarize(cog_40_relative = sum(cog_40_relative),
+            cog_41_relative = sum(cog_41_relative),
+            cog_42_relative = sum(cog_42_relative),
+            cog_43_relative = sum(cog_43_relative),
+            cog_61_relative = sum(cog_61_relative),
+            cog_62_relative = sum(cog_62_relative),
+            cog_63_relative = sum(cog_63_relative),
+            abundance_40 = sum(`Abundance: 40: Sample`, na.rm = TRUE),
+            abundance_41 = sum(`Abundance: 41: Sample`, na.rm = TRUE),
+            abundance_42 = sum(`Abundance: 42: Sample`, na.rm = TRUE),
+            abundance_43 = sum(`Abundance: 43: Sample`, na.rm = TRUE),
+            abundance_61 = sum(`Abundance: 61: Sample`, na.rm = TRUE),
+            abundance_62 = sum(`Abundance: 62: Sample`, na.rm = TRUE),
+            abundance_63 = sum(`Abundance: 63: Sample`, na.rm = TRUE)) %>%
+  gather(key="sample", value="abundance") %>%
+  mutate(ID=str_replace(sample, "cog_", "")) %>%
+  mutate(ID=str_replace(ID, "abundance_", "")) %>%
+  mutate(ID=str_replace(ID, "_relative", "")) %>%
+  left_join(., metadata, by="ID") %>%
+  mutate(host = if_else(str_detect(station, "Ca"), "Ca", "Cy")) %>%
+  mutate(cog_abundance = if_else(str_detect(sample, "_relative"), "cog", "abundance")) %>%
+  group_by(host, cog_abundance) %>%
+  summarize(mean = mean(abundance), sd = sd(abundance))
 ```
@@
 To assess the removal efficiency of DNA and protein isolation procedures leaves and thalli were stained with SYBR Green I and examined under a confocal microscope before and after treatments were performed.
-As it is known that marine macrophytes are harboring more algal epiphytes during Autumn and Winter months  [@Reyes2001], the sampling was performed in two contrasting seasons, in December 2017 when higher abundances of algal epiphytes are expected, and in June 2018, when lower abundances could be expected.
-In addition, to further test the isolation protocols on different macrophyte species, DNA and proteins were isolated from the surfaces of two macrophytes, the seagrass *C.
+To test the isolation protocols on different macrophyte species, DNA and proteins were isolated from the surfaces of two macrophytes, the seagrass *C.
 nodosa* and the macroalga *C.
@@
 nodosa* meadow in the Bay of Funtana.
-*C.
-nodosa* June samples from the Bay of Saline were missing as the meadow went to a decline from Spring 2018 (M.
-Najdek, personal communication).
-Procedures developed for DNA and protein isolation showed an almost complete removal of the surface community on both, samples collected in December 2017 (\autoref{confocal_december}) and June 2018 (\autoref{confocal_june}).
-In addition, a similar removal efficiency was observed for communities associated with *C.
+In addition, the sampling was performed in two contrasting seasons, in December 2017 and in June 2018, to test the impact of seasonal conditions on the removal efficiency.
+Procedures developed for DNA and protein isolation showed an almost complete removal of the surface community of both, *C.
 nodosa* and *C.
 cylindracea*.
+In addition, a similar removal efficiency was observed for communities sampled in December 2017 (\autoref{confocal_december}) and June 2018 (\autoref{confocal_june}).
 Also, no effect of station, settlement or isolation procedure (DNA or protein) on the removal efficiency was observed (Figs.
@@
 Sequencing yielded a total of `r number_sequences` sequences after quality curation and exclusion of eukaryotic, mitochondrial and no relative sequences.
-The number of sequences classified as chloroplast was `r number_chloroplast_sequences` (`r proprotion_chloroplast_sequences` \si{\percent})˛ After the exclusion of these sequences the total number of retrieved reads dropped to `r number_sequences_without_chloroplast`, ranging from `r min_max$min` and `r min_max$max` sequences per sample (\autoref{supp-nseq_notus}).
-Even when the highest sequencing effort was applied the rarefaction curves did not level off that is commonly observed in high-throughput 16S rRNA amplicon sequencing procedures (\autoref{supp-rarefaction}).
+The number of sequences classified as chloroplast was `r number_chloroplast_sequences`.
+After excluding these sequences the total number of retrieved reads dropped to `r number_sequences_without_chloroplast`, ranging from `r min_max$min` to `r min_max$max` sequences per sample (\autoref{supp-nseq_notus}).
+Even when the highest sequencing effort was applied the rarefaction curves did not level off which is commonly observed in high-throughput 16S rRNA amplicon sequencing procedures (\autoref{supp-rarefaction}).
 Sequence clustering at a similarity level of 97 \si{\percent} yielded a total of `r no_otus` different OTUs.
-Taxonomic classification of reads allowed the determination of the macrophyte associated epiphytic community that was mainly composed of: *Alphaproteobacteria* (`r filter(plot, taxon=="Alphaproteobacteria")$mean` ± `r filter(plot, taxon=="Alphaproteobacteria")$sd` \si{\percent}), *Bacteroidota* (`r filter(plot, taxon=="Bacteroidota")$mean` ± `r filter(plot, taxon=="Bacteroidota")$sd` \si{\percent}), *Gammaproteobacteria* (`r filter(plot, taxon=="Gammaproteobacteria")$mean` ± `r filter(plot, taxon=="Gammaproteobacteria")$sd` \si{\percent}), *Desulfobacterota* (`r filter(plot, taxon=="Desulfobacterota")$mean` ± `r filter(plot, taxon=="Desulfobacterota")$sd` \si{\percent}), *Cyanobacteria* (`r filter(plot, taxon=="Cyanobacteria")$mean` ± `r filter(plot, taxon=="Cyanobacteria")$sd` \si{\percent}) and *Planctomycetota* (`r filter(plot, taxon=="Planctomycetota")$mean` ± `r filter(plot, taxon=="Planctomycetota")$sd` \si{\percent}) (\autoref{community}).
+Taxonomic classification of reads allowed for a macrophyte associated epiphytic community determination that was mainly composed of: *Alphaproteobacteria* (`r filter(plot, taxon=="Alphaproteobacteria")$mean` ± `r filter(plot, taxon=="Alphaproteobacteria")$sd` \si{\percent}), *Bacteroidota* (`r filter(plot, taxon=="Bacteroidota")$mean` ± `r filter(plot, taxon=="Bacteroidota")$sd` \si{\percent}), *Gammaproteobacteria* (`r filter(plot, taxon=="Gammaproteobacteria")$mean` ± `r filter(plot, taxon=="Gammaproteobacteria")$sd` \si{\percent}), *Desulfobacterota* (`r filter(plot, taxon=="Desulfobacterota")$mean` ± `r filter(plot, taxon=="Desulfobacterota")$sd` \si{\percent}), *Cyanobacteria* (`r filter(plot, taxon=="Cyanobacteria")$mean` ± `r filter(plot, taxon=="Cyanobacteria")$sd` \si{\percent}) and *Planctomycetota* (`r filter(plot, taxon=="Planctomycetota")$mean` ± `r filter(plot, taxon=="Planctomycetota")$sd` \si{\percent}) (\autoref{community}).
 
-Primers used in this study, as in many other 16S rRNA amplicon procedures, also amplify chloroplast 16S rRNA genes.
+Primers used in this study, as in many other 16S rRNA amplicon procedures, also amplified chloroplast 16S rRNA genes.
 We observed a high proportion of chloroplast sequences in all analyzed samples (`r filter(plot, taxon=="Chloroplast")$mean` ± `r filter(plot, taxon=="Chloroplast")$sd` \si{\percent}) (\autoref{community}).
 To determine if chloroplast sequences originate from hosts or eukaryotic epiphytic organisms, we exported SILVA classified chloroplast sequences and reclassified them using the RDP (Ribosomal Database Project) training set that allows for a more detailed chloroplast classification.
-The largest proportion of sequences was classified as Bacillariophyta (`r filter(chloroplast, taxon=="Bacillariophyta")$mean` ± `r filter(chloroplast, taxon=="Bacillariophyta")$sd` \si{\percent}) indicating that the DNA removal procedure did not coextract larger quantities of host DNA (\autoref{chloroplast}).
+The largest proportion of sequences were classified as Bacillariophyta (`r filter(chloroplast, taxon=="Bacillariophyta")$mean` ± `r filter(chloroplast, taxon=="Bacillariophyta")$sd` \si{\percent}) indicating that the DNA removal procedure did not coextract larger quantities of host DNA (\autoref{chloroplast}).
+Chloroplast sequences classified as Streptophyta constituted `r streptophyta_16S$mean` ± `r streptophyta_16S$sd` \si{\percent} of all chloroplast sequences originating from *C.
+nodosa* samples, while sequences classified as Chlorophyta comprised only `r chlorophyta_16S$mean` ± `r chlorophyta_16S$sd` \si{\percent} of all chloroplast sequences associated with *C.
+cylindracea* samples.
+
+To determine if the extracted DNA can be used for metagenomic sequencing four samples containing epiphytic DNA were selected and shotgun sequenced using an Illumina platform.
+Metagenomic sequencing yielded from `r format(min(metadata_metagenomic_statistics$raw_R1), big.mark=',')` to `r format(max(metadata_metagenomic_statistics$raw_R1), big.mark=',')` sequence pairs (\autoref{supp-metagenomic_statistics}).
+Obtained sequences were successfully assembled into contigs whose L50 ranged from `r format(min(metadata_metagenomic_statistics$L50), big.mark=',')` to `r format(max(metadata_metagenomic_statistics$L50), big.mark=',')` bp.
+In addition, predicted coding sequences were successfully functionally annotated (`r format(min(metadata_metagenomic_statistics$eggnog), big.mark=',')` -- `r format(max(metadata_metagenomic_statistics$eggnog), big.mark=',')` annotated sequences; \autoref{cog}) and taxonomically classified.
+Functional annotation allowed for an assessment of the relative contribution of each COG (Clusters of Orthologous Groups) functional category to the total number of annotated coding sequences (\autoref{cog}a).
+Functional categories containing the highest number of sequences were: `r cog_relative$"COG Functional Category"[3]` (`r cog_relative$Name[3]`), `r cog_relative$"COG Functional Category"[4]` (`r cog_relative$Name[4]`), `r cog_relative$"COG Functional Category"[5]` (`r cog_relative$Name[5]`), `r cog_relative$"COG Functional Category"[6]` (`r cog_relative$Name[6]`) and `r cog_relative$"COG Functional Category"[7]` (`r cog_relative$Name[7]`).
+If host DNA was coextracted with epiphytic it should be detected in larger proportions in sequenced metagenomes.
+Indeed, no higher proportions of coding sequences classified into phylum Streptophyta or Chlorophyta were detected (\autoref{supp-metagenomic_taxonomy}).
+Sequenced metagenomic DNA originating from the surface of *C.
+nodosa* contained `r filter(metadata_metagenomic_phylum, label=="14/12-17 FCyM" & phylum=="Streptophyta")$cds_relative` \si{\percent} of coding sequences classified into the phylum Streptophyta in December 2017 and `r filter(metadata_metagenomic_phylum, label=="19/6-18 FCyM" & phylum=="Streptophyta")$cds_relative` \si{\percent} in June 2018.
+Furthermore, the summed RPKM (Reads Per Kilobase Million) of these sequences constituted `r filter(metadata_metagenomic_phylum, label=="14/12-17 FCyM" & phylum=="Streptophyta")$rpkm_relative` \si{\percent} of total RPKM of all successfully classified sequences in December 2017 and `r filter(metadata_metagenomic_phylum, label=="19/6-18 FCyM" & phylum=="Streptophyta")$rpkm_relative` \si{\percent} in June 2018.
+Similar low proportions of host coding sequences were detected in metagenomic samples originating from the surfaces of *C.
+cylindracea*.
+Of all successfully classified coding sequences `r filter(metadata_metagenomic_phylum, label=="14/12-17 FCa" & phylum=="Chlorophyta")$cds_relative` \si{\percent} were classified into Chlorophyta in December 2017 and `r filter(metadata_metagenomic_phylum, label=="19/6-18 FCa" & phylum=="Chlorophyta")$cds_relative` \si{\percent} in June 2018.
+A relatively higher proportion of these sequences' RPKM than in the case of *C.
+nodosa* was observed, indicating a higher coextraction of host DNA in *C.
+cylindracea*.
+In December, the proportion of RPKM of sequences classified into Chlorophyta was `r filter(metadata_metagenomic_phylum, label=="14/12-17 FCa" & phylum=="Chlorophyta")$rpkm_relative` \si{\percent}, while in June 2018 it reached `r filter(metadata_metagenomic_phylum, label=="19/6-18 FCa" & phylum=="Chlorophyta")$rpkm_relative` \si{\percent}.
+
+To evaluate if the procedure for protein extraction is suitable for metaproteomic analysis, obtained proteins were trypsin digested and sequenced using a mass spectrometer.
+Obtained MS/MS spectra were searched against a protein database from sequenced metagenomes.
+From `r format(min(metap_cog_num), big.mark=',')` to `r format(max(metap_cog_num), big.mark=',')` proteins were identified in isolated protein samples (\autoref{cog}b).
+In addition, successful identification of proteins allowed for an assessment of the relative contribution of each COG functional category to the total number of identified proteins (\autoref{cog}b).
+Functional categories containing the highest number of identified proteins were: `r metap_cog_relative$"COG Functional Category"[1]` (`r metap_cog_relative$Name[1]`), `r metap_cog_relative$"COG Functional Category"[2]` (`r metap_cog_relative$Name[2]`), `r metap_cog_relative$"COG Functional Category"[3]` (`r metap_cog_relative$Name[3]`), `r metap_cog_relative$"COG Functional Category"[4]` (`r metap_cog_relative$Name[4]`) and `r metap_cog_relative$"COG Functional Category"[5]` (`r metap_cog_relative$Name[5]`).
+Isolated proteins could derive from epiphytic organisms inhabiting the macrophyte surface or from macrophyte tissue underlying them.
+The contribution of proteins originating from host tissues was evaluated by identifying all the proteins predicted to belong to any taxonomic group within the phyla Streptophyta and Chlorophyta and by calculating their contribution to the number and abundance of all identified proteins.
+On average, proteins isolated from the surface of *C.
+nodosa* contained `r filter(metap_cog_streptophyta, cog_abundance=="cog" & host=="Cy")$mean`  ± `r filter(metap_cog_streptophyta, cog_abundance=="cog" & host=="Cy")$sd` \si{\percent} of proteins associated with Streptophyta, contributing to `r filter(metap_cog_streptophyta, cog_abundance=="abundance" & host=="Cy")$mean`  ± `r filter(metap_cog_streptophyta, cog_abundance=="abundance" & host=="Cy")$sd` \si{\percent} of total proteins.
+Similar to metagenomes, proteins associated with Chlorophyta contributed more to *C.
+cylindracea* than proteins associated with Streptophyta to *C.
+nodosa*.
+Chlorophyta associated proteins comprised `r filter(metap_cog_chlorophyta, cog_abundance=="cog" & host=="Ca")$mean`  ± `r filter(metap_cog_chlorophyta, cog_abundance=="cog" & host=="Ca")$sd` \si{\percent} of all identified proteins in *C.
+cylindracea*, contributing to `r filter(metap_cog_chlorophyta, cog_abundance=="abundance" & host=="Ca")$mean`  ± `r filter(metap_cog_chlorophyta, cog_abundance=="abundance" & host=="Ca")$sd` \si{\percent} of all protein abundances.
 
 ## Discussion
+The study of marine macrophyte epiphytic communities using culture independent techniques, such as 16S rRNA analysis, metagenomics and metaproteomics, requires an initial step of biological material isolation.
+Methods that have been developed for selective isolation of epiphytic biological material could be divided into two groups: (i) procedures involving, prior to extraction, a cell detachment step such as shaking [@Gross2003; @Noges2010], scraping [@Uku2007] and ultrasonication [@Weidner1996; @Cai2014] and (ii) procedures involving a host tissue incubation aiming at direct lysis of epiphytic microbial cells [@Burke2009].
+Protocols that include a cell detachment procedure do not provide complete cell removal [@Richter-Heitmann2016; @Cai2014] and in the case of probe ultrasonication tissue disruption can also be observed [@Richter-Heitmann2016].
+To avoid these problems a procedure for DNA isolation involving host tissue incubation in the extraction buffer and direct epiphytic cell lysis on the macrophyte surface was developed [@Burke2009].
+Though providing an almost complete cell removal and selective DNA isolation this protocol uses in the extraction buffer a rapid multienzyme cleaner (3M) not available worldwide and without a known composition [@Burke2009].
+To circumvent these problems we developed and tested two protocols for selective DNA and protein isolation from marine macrophyte epiphytic communities.
+
+To test if the developed DNA and protein isolation procedures could be applied on a variety of macrophyte species we selected *C.
+nodosa* and *C.
+cylindracea* as representatives of seagrass and macroalgal species, on which the procedures were tested.
+These species especially differ morphologically.
+While *C.
+nodosa* leaves are flat, *C.
+cylindracea* thallus is composed of uneven surfaces [@Kuo2001; @Verlaque2003].
+The developed procedures showed an almost complete removal of epiphytic cells from the surfaces of both species comparable to the result of @Burke2009 and indicating that structural differences do not impact the removal efficiency.
+In addition, isolation protocols were tested in two contrasting season as it is known that macrophytes are harboring more algal epiphytes during autumn and winter [@Reyes2001].
+No differences in the removal efficiency was observed between seasons suggesting that protocols can be used on macrophyte samples retrieved throughout the year.
+Also, no removal differences were observed on samples derived from the same host but from different localities or settlements demonstrating the stability of the protocol in cell removal efficiency.
+
+Successful amplification and sequencing of the V4 16S rRNA gene region proved that the isolated DNA can be used to estimate the microbial epiphytic diversity.
+Taxonomic groups detected in this step can also be often found in epiphytic communities associated with other macrophytes [@Morrissey2019; @Burke2011a].
+A problem often encountered in studies focusing on epiphytic communities is the presence of large proportions of chloroplast 16S rRNA sequences in the pool of amplified molecules, especially if the epiphytic DNA was isolated without prior selection [@Staufenberger2008].
+These sequences can derive from host chloroplasts or from eukaryotic epiphytic chloroplast DNA.
+Although the proportion of obtained chloroplast 16S rRNA sequences in our samples was substantial they derived almost exclusively from eukaryotic epiphytes.
+High proportion of chloroplast 16S rRNA sequences in studies applying selective procedures that include direct cellular lysis on host surfaces were observed before [@Michelou2013].
+It is possible that chloroplast specific sequences even in these studies are originating from eukaryotic epiphytic cells and not host chloroplast.
+Indeed, it is common during 16S rRNA profiling of pelagic microbial communities to observe high proportions of chloroplast sequences [@Korlevic2016; @Gilbert2009].
+In addition, a very low proportion of chloroplast 16S rRNA sequences in samples originating from *C.
+cylindracea* in comparison to *C.
+nodosa* could be explained by the presence of three introns in the gene for 16S rRNA in some members of the genus *Caulerpa* that could hamper the amplification process [@Lam2016].
+
+Beside 16S rRNA sequencing high quality DNA is needed for metagenomics.
+The obtained number of metagenomic sequences and assembly statistics were comparable to metagenomes and metatranscriptomes derived from similar surface associated communities [@Crump2018; @Cucio2018].
+In addition, functional annotation of predicted coding sequences to COG functional categories showed that the obtained metagenomes can be used to determine the metabolic capacity of surface associated communities [@Cucio2018; @Leary2014].
+The proportion of coding sequences, including their RPKM, originating from *C.
+nodosa* metagenomes and classified as Streptophyta was low indicating a good selectivity of the isolation procedure towards epiphytic cells.
+In the case of DNA samples isolated from the surface of *C.
+cylindracea* the proportion of Chlorophyta coding sequences was also low but their RPKM was higher than in the case of *C.
+nodosa*.
+One of the causes for this elevated RPKM of Chlorophyta sequences in *C.
+cylindracea* could lay in tissue structure differences between these two host species.
+While *C.
+nodosa* leaves are composed of individual cells, the thallus of *C.
+cylindracea* is, like in other siphonous algal species, composed of a single large multinucleate cell [@Coneva2015].
+The absence of individual cells in *C.
+cylindracea* could cause a leakage of genetic material into the extraction buffer causing an elevated presence of host sequences in metagenomic data.
+
+Metagenomic analysis reveals the potential metabolic capacity of microbial communities, while to obtain an insight into the metabolic status of uncultivated members, a metaproteomic approach is required [@Saito2019].
+The applied protocol for epipyhtic protein isolation followed by a metaproteomic analysis identified between `r format(min(metap_cog_num), big.mark=',')` and `r format(max(metap_cog_num), big.mark=',')` proteins that is higher than previously reported for e.g.
+soils [@Chourey2010; @Hultman2015], seawater [@Williams2012] and biofilms [@Leary2014].
+The functional annotation of identified proteins into COG functional categories showed that the protein isolation procedure can be used to assess the metabolic status of the macrophyte epiphytic community [@Leary2014].
+Similar to metagenomes, the number of identified proteins, including their abundances, associated with Streptophyta in *C.
+nodosa* samples were low indicating that the procedure is selective for epiphytic cell proteins.
+In addition, a higher number of identified proteins, and especially their abundances, associated with Chlorophyta was observed in *C.
+cylindracea* samples.
+The cause of this elevated presence of Chlorophyta associated proteins can be, as in the case of the DNA isolation procedure, explained by the absence of individual cells in this siphonous alga [@Coneva2015].
+
+In conclusion, the developed protocols for DNA and protein isolation from macrophyte surfaces are providing an almost complete removal of the epiphytic community and are shown to ensure removal from both, *C.
+nodosa* and *C.
+cylindracea*, in different seasons.
+Also, the obtained DNA and proteins are suitable for 16S rRNA sequencing, metagenomics and metaproteomics, while the obtained material contains low quantities of host DNA or proteins making the procedures epiphyte selective.
+Furthermore, the procedures are based on universally available laboratory chemicals making the protocols widely applicable.
 
@@
 nodosa* meadow invaded by *C.
-cylindracea* in the proximity of the village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E).
+cylindracea* in the proximity of the village of Funtana (Bay of Funtana; \ang{45;10;39} N, \ang{13;35;42} E).
 Thalli of *C.
 cylindracea* were sampled in the same *C.
-nodosa* invaded meadow in Funtana and on a locality of only *C.
+nodosa* invaded meadow in the Bay Funtana and on a locality of only *C.
 cylindracea* located in the proximity of the invaded meadow.
-Leaves and thalli were collected on the same day in two contrasting seasons, on 4 December 2017 and 18 June 2018.
+Leaves and thalli for 16S rRNA analysis, metagenomics and metaproteomics were collected in two contrasting seasons, on 4 December 2017 (16S rRNA analysis and metaproteomics), 14 December 2017 (metagenomics) and 18 June 2018 (16S rRNA analysis, metagenomics and metaproteomics).
 During spring 2018 the *C.
-nodosa* meadow in the Bay of Saline decayed to an extent that no leaves could be retrieved (Najdek *et al.*, unpublished data).
+nodosa* meadow in the Bay of Saline decayed to an extent that no leaves could be retrieved [@Najdek2020].
+In addition, as not enough DNA for both metagenomic and 16S RNA analysis were obtained during the sampling on 4 December 2017, an additional sampling on 14 December 2017 was carried out in the Bay of Funtana.
 Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater.
@@
 The DNA was isolated according to the protocol for isolation from filters described in @Massana1997.
-This protocol was modified and adapted for DNA isolation from microbial communities from macrophyte surfaces as described below.
-5 \si{\ml} of lysis buffer (40 \si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar} sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or \si{\g} wet-weight of thalli.
+This protocol was modified and adapted for microbial DNA isolation from macrophyte surfaces as described below.
+5 \si{\ml} of lysis buffer (40 \si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar} sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or 2 \si{\g} wet-weight of thalli.
+For every sample duplicate isolations were performed.
 Lysozyme was added (final concentration 1 \si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius} for 30 \si{\minute}.
@@
 Purified PCR products were sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories, Martinsried, Germany.
-Before sequencing, at IMGM the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR.
+Before sequencing at IMGM, the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR.
 In addition, these primers contained adapter and sample-specific index sequences.
@@
 A negative control was comprised of four parallel PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
-Sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 -- SAMEA6648811.
+Partial 16S rRNA sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 -- SAMEA6648811.
 
-```{r}
+```{r error_negative_control}
 # Calculation of the error rate based on the ATCC MSA-1002 mock community
@@
 error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
+
 # Calculation of the number of sequences in the negative control
@@
 ```
+
 Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
@@
 In comparison to SILVA, RDP allows a more detailed classification of chloroplast sequences.
-Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020].
+Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020; @Wilke2018].
 The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020).
@@
 ### Metagenomics
-Four samples were selected and sent on dry ice to IMGM Laboratories, Martinsried, Germany for metagenomic sequencing.
+Four DNA samples were selected and sent on dry ice to IMGM Laboratories, Martinsried, Germany for metagenomic sequencing.
 DNA was purified using AMPure XP Beads (Beckman Coulter, USA) applying a bead:DNA ratio of 1:1 (v/v), quantified with a Qubit dsDNA HS Assay Kit (Thermo Fisher Scientific, USA) and check for integrity on a 1 \si{\percent} agarose gel.
@@
 
-Obtained sequences were assembled individually using MEGAHIT (version  1.1.2) [@Li2015] under default settings.
+Obtained sequences were analyzed on a Life Science Compute Cluster (LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
+Individual sequences were assembled using MEGAHIT (version  1.1.2) [@Li2015] under default settings.
 Putative genes were predicted from contings longer than 200 bp using Prodigal (version  2.6.3) [@Hyatt2010] in metagenome mode (-p meta).
@@
 All predicted genes were functionally annotated using the eggNOG-mapper [@Huerta-Cepas2017] and eggNOG database (version 5.0) [@Huerta-Cepas2019].
-The taxonomic classification of sequences was determined using the lowest common ancestor algorithm adapted from DIAMOND (version 0.8.36) [@Buchfink2015] and by searching against the non-redundant database (NR).
+Sequence taxonomy classification was determined using the lowest common ancestor algorithm adapted from DIAMOND (version 0.8.36) [@Buchfink2015] and by searching against the NCBI non-redundant database (NR).
 To determine phylogeny the top 10 \si{\percent} hits with an e-value < 1 × 10^-5^ were used (\-\-top 10).
+Sequence renaming, coverage information computing and metagenomic statistics calculations were performed using software tools from BBTools (https://jgi.doe.gov/data-and-tools/bbtools).
+Metagenomic sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers **TO BE ADDED**.
 
 ### Protein isolation
-Proteins were isolated according to the protocol for isolation from soil described in @Chourey2010 and modified by @Hultman2015.
-These protocols were further modified and adapted for protein isolation from microbial communities form macrophyte surfaces as described below.
+Proteins were isolated according to the protocol for protein isolation from soil described in @Chourey2010 and modified by @Hultman2015.
+This protocol was further modified and adapted for microbial protein isolation from macrophyte surfaces as described below.
 20 \si{\ml} of protein extraction buffer (4 \si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl [pH 8.0]) was added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of thalli.
@@
 ### Metaproteomics
-Isolated proteins were whole trypsin digested using the FASP (filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK) according to the manufacturer's instructions [@Wisniewski2009] with small modifications.
-Before loading the solution to the column, protein pellets were solubilized in a urea sample buffer included in the kit amended with DTT (final concentration 100 \si{\milli\Molar}) for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to remove larger particles.
+Isolated proteins were whole trypsin digested using the FASP (filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK) according to the manufacturer's instructions with small modifications [@Wisniewski2009].
+Before the solution was loaded to the column, protein pellets were solubilized in a urea sample buffer included in the kit amended with DTT (final concentration 100 \si{\milli\Molar}) for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to remove larger particles.
 The first washing step after protein solution loading was repeated twice.
@@
 Trypsin digestion was performed on column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
-The final filtrate containing digested proteins was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid, freezed at \num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent to the Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis.
-Peptides were resuspended in 1 \si{\percent} (final concentration) trifluoroacetic acid, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
-Obtained MS/MS spectra were searched against a protein database from metagenomic assembly published in @Burke2011 using SEQUEST-HT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
+The final filtrate containing peptides was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid, freezed at \num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent to the Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis.
+Peptides were resuspended in 1 \si{\percent} (final concentration) trifluoroacetic acid (TFA), desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
+Obtained MS/MS spectra were searched against a protein database composed of combined sequenced metagenomes using SEQUEST-HT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
 The target-decoy approach was used to reduce the probability of false peptide identification.
 Results whose false discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
-For protein identification a minimum of of two peptides and one unique peptide were required.
+For protein identification a minimum of two peptides and one unique peptide were required.
 For protein quantification, a chromatographic peak area-based free quantitative method was applied.
@@
 For long therm storage, fixed leaves and thalli were immersed in a mixture of phosphate-buffered saline (PBS) and ethanol (1:1) and stored at \num{-20} \si{\degreeCelsius}.
-Treated and untreated leaves and thalli segments were stained in a 2 × solution of SYBR Green I and examined under a Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
+Treated and untreated segemnts of leaves and thalli were stained in a 2 × solution of SYBR Green I and examined under a Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
 
 ## Acknowledgements
+This work was founded by the Croatian Science Foundation through the MICRO-SEAGRASS project (IP-2016-06-7118).
+We would like to thank Margareta Buterer for technical support, Paolo Paliaga for help during sampling, Lucija Horvat for technical support in confocal microscopy and Dušica Vujaklija for help in accessing the confocal microscope.
+In addition, we would like to thank the University Computing Center of the University of Zagreb for access to the computer cluster Isabella and CUBE -- Computational Systems Biology of the University of Vienna for access to the Life Science Compute Cluster (LiSC).
 
@@
 ```
+
+\newpage
+```{r out.width="100%", fig.align="center", fig.cap="Relative contribution of each COG category to the total number of annotated coding sequences (A) from metagenomes and identified proteins (B) from metaproteomes associated with surfaces of two marine macrophytes (\\textit{C.
+nodosa} and \\textit{C.
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4/14 December 2017 and 19 June 2018).
+Total number of annotated coding sequences and identified proteins is given above the corresponding bar.\\label{cog}", fig.pos="ht"}
+include_graphics("../results/figures/cog.jpg")
+```

### 2020-09-14 submission/manuscript.tex
Manuscript done.

@@
 
-3. Department of Marine Microbiology and Biogeochemistry, Royal
+3. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal
 Netherlands Institute for Sea Research, Utrecht University, The
@@
 \subsection{Summary}\label{summary}}
+
+Studies of unculturable microbes often combine methods based on DNA,
+such as 16S rRNA sequencing and metagenomics, with methods that allow an
+insight into the metabolic status, such as metaproteomics. To apply
+these procedures to the microbial community inhabiting the surfaces of
+marine macrophytes it is advisable to perform, prior to the analysis, a
+selective DNA and protein isolation so that the host material, present
+in higher quantities, is not hampering the analysis. Two protocols, for
+DNA and protein isolation, were adapted for selective extractions of DNA
+and proteins from epiphytic communities inhabiting the surfaces of two
+marine macrophytes: the seagrass \emph{Cymodocea nodosa} and the
+macroalga \emph{Caulerpa cylindracea}. Protocols showed an almost
+complete removal of the epiphytic community regardless of the sampling
+season, station, settlement or host species. The obtained DNA was
+suitable for metagenomic and 16S rRNA sequencing, while isolated
+proteins could be identified by mass spectrometry showing that protocols
+can be used in 16S rRNA, metagenomic and metaproteomic analysis. Low
+presence of host DNA and proteins, observed in isolated samples,
+indicated a selective nature of protocols. Furthermore, the procedures
+are based on universally available laboratory chemicals making the
+protocols widely applicable. Taken together, the adapted protocols
+ensure an almost complete removal of the macrophyte epiphytic community,
+are selective for microbes inhabiting macrophyte surfaces and are
+providing DNA and proteins applicable in 16S rRNA sequencing,
+metagenomics and metaproteomics.
 
@@
 subsequently isolating the aimed compound (Gilbert \emph{et al.}, 2009).
-If a microbial size fraction is aimed sequential filtration is applied
-(Massana \emph{et al.}, 1997; Andersson \emph{et al.}, 2010). In
+If a specific microbial size fraction is aimed sequential filtration is
+applied (Massana \emph{et al.}, 1997; Andersson \emph{et al.}, 2010). In
 contrast, obtaining biological materials from microorganism inhabiting
@@
 material by incubating macrophyte tissues in an extraction buffer. After
-the incubation is done the undisrupted host tissue is removed and the
-isolation procedure continues omitting host material contaminations. To
-our knowledge the only procedure describing a direct and selective
-epiphytic DNA isolation from the surfaces of marine macrophytes was
-provided by Burke \emph{et al.} (2009). In contrast to previously
-described methods this protocol enables an almost complete removal of
-the surface community and was used for 16S rRNA gene clone libraries
-construction (Burke and Thomas \emph{et al.}, 2011) and metagenomes
-sequencing (Burke and Peter Steinberg \emph{et al.}, 2011). This method,
-thought providing a selective isolation procedure, is using in the
-extraction buffer a rapid multienzyme cleaner (3M) that is not worldwide
-available and whose composition is not know (Burke \emph{et al.}, 2009).
-Also to our knowledge, no selective isolation protocol for proteins from
-epiphytic communities inhabiting marine macrophytes was established.
+the incubation, the undisrupted host tissue is removed and the isolation
+procedure continues omitting host material contaminations. To our
+knowledge the only procedure describing a direct and selective epiphytic
+DNA isolation from the surfaces of marine macrophytes was provided by
+Burke \emph{et al.} (2009). In contrast to previously described methods
+this protocol enables an almost complete removal of the surface
+community and it was used for 16S rRNA gene clone library construction
+(Burke and Thomas \emph{et al.}, 2011) and metagenome sequencing (Burke
+and Peter Steinberg \emph{et al.}, 2011). This method, thought providing
+a selective isolation procedure, uses a rapid multienzyme cleaner (3M)
+that is not available worldwide and without a known composition (Burke
+\emph{et al.}, 2009). Also to our knowledge, no selective isolation
+protocol for proteins from epiphytic communities inhabiting marine
+macrophytes was established.
 
@@
 proteins from epiphytic communities inhabiting the surfaces of two
-marine macrophytes: the seagrass \emph{Cymodocea nodosa} and the alga
-\emph{Caulerpa cylindracea}. In addition, we tested the removal
-efficiency of the protocol and the suitability of obtained DNA and
-proteins for 16S rRNA sequencing and metaproteomics.
+marine macrophytes: the seagrass \emph{Cymodocea nodosa} and the
+macroalga \emph{Caulerpa cylindracea}. In addition, we tested the
+removal efficiency of the protocol and the suitability of obtained DNA
+and proteins for 16S rRNA sequencing, metagenomics and metaproteomics.
 
@@
 leaves and thalli were stained with SYBR Green I and examined under a
-confocal microscope before and after treatments were performed. As it is
-known that marine macrophytes are harboring more algal epiphytes during
-Autumn and Winter months (Reyes and Sansón, 2001), the sampling was
-performed in two contrasting seasons, in December 2017 when higher
-abundances of algal epiphytes are expected, and in June 2018, when lower
-abundances could be expected. In addition, to further test the isolation
-protocols on different macrophyte species, DNA and proteins were
-isolated from the surfaces of two macrophytes, the seagrass \emph{C.
-nodosa} and the macroalga \emph{C. cylindracea}. \emph{C. nodosa} was
-sampled in a monospecific meadow (Bay of Saline) and \emph{C.
-cylindracea} invaded meadow (Bay of Funtana), while samples of \emph{C.
-cylindracea} originated from a monospecific settlement and a \emph{C.
-cylindracea} invaded \emph{C. nodosa} meadow in the Bay of Funtana.
-\emph{C. nodosa} June samples from the Bay of Saline were missing as the
-meadow went to a decline from Spring 2018 (M. Najdek, personal
-communication). Procedures developed for DNA and protein isolation
-showed an almost complete removal of the surface community on both,
-samples collected in December 2017 (\autoref{confocal_december}) and
-June 2018 (\autoref{confocal_june}). In addition, a similar removal
-efficiency was observed for communities associated with \emph{C. nodosa}
-and \emph{C. cylindracea}. Also, no effect of station, settlement or
+confocal microscope before and after treatments were performed. To test
+the isolation protocols on different macrophyte species, DNA and
+proteins were isolated from the surfaces of two macrophytes, the
+seagrass \emph{C. nodosa} and the macroalga \emph{C. cylindracea}.
+\emph{C. nodosa} was sampled in a monospecific meadow (Bay of Saline)
+and \emph{C. cylindracea} invaded meadow (Bay of Funtana), while samples
+of \emph{C. cylindracea} originated from a monospecific settlement and a
+\emph{C. cylindracea} invaded \emph{C. nodosa} meadow in the Bay of
+Funtana. In addition, the sampling was performed in two contrasting
+seasons, in December 2017 and in June 2018, to test the impact of
+seasonal conditions on the removal efficiency. Procedures developed for
+DNA and protein isolation showed an almost complete removal of the
+surface community of both, \emph{C. nodosa} and \emph{C. cylindracea}.
+In addition, a similar removal efficiency was observed for communities
+sampled in December 2017 (\autoref{confocal_december}) and June 2018
+(\autoref{confocal_june}). Also, no effect of station, settlement or
 isolation procedure (DNA or protein) on the removal efficiency was
@@
 curation and exclusion of eukaryotic, mitochondrial and no relative
-sequences. The number of sequences classified as chloroplast was 97,328
-(28.9 \si{\percent})˛ After the exclusion of these sequences the total
-number of retrieved reads dropped to 239,609, ranging from 22,587 and
-52,958 sequences per sample (\autoref{supp-nseq_notus}). Even when the
-highest sequencing effort was applied the rarefaction curves did not
-level off that is commonly observed in high-throughput 16S rRNA amplicon
-sequencing procedures (\autoref{supp-rarefaction}). Sequence clustering
-at a similarity level of 97 \si{\percent} yielded a total of 8,360
-different OTUs. Taxonomic classification of reads allowed the
-determination of the macrophyte associated epiphytic community that was
-mainly composed of: \emph{Alphaproteobacteria} (14.9 ± 3.5
-\si{\percent}), \emph{Bacteroidota} (12.5 ± 2.4 \si{\percent}),
+sequences. The number of sequences classified as chloroplast was 97,328.
+After excluding these sequences the total number of retrieved reads
+dropped to 239,609, ranging from 22,587 to 52,958 sequences per sample
+(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
+applied the rarefaction curves did not level off which is commonly
+observed in high-throughput 16S rRNA amplicon sequencing procedures
+(\autoref{supp-rarefaction}). Sequence clustering at a similarity level
+of 97 \si{\percent} yielded a total of 8,360 different OTUs. Taxonomic
+classification of reads allowed for a macrophyte associated epiphytic
+community determination that was mainly composed of:
+\emph{Alphaproteobacteria} (14.9 ± 3.5 \si{\percent}),
+\emph{Bacteroidota} (12.5 ± 2.4 \si{\percent}),
 \emph{Gammaproteobacteria} (11.6 ± 4.3 \si{\percent}),
@@
 Primers used in this study, as in many other 16S rRNA amplicon
-procedures, also amplify chloroplast 16S rRNA genes. We observed a high
-proportion of chloroplast sequences in all analyzed samples (33.4 ± 9.4
-\si{\percent}) (\autoref{community}). To determine if chloroplast
+procedures, also amplified chloroplast 16S rRNA genes. We observed a
+high proportion of chloroplast sequences in all analyzed samples (33.4 ±
+9.4 \si{\percent}) (\autoref{community}). To determine if chloroplast
 sequences originate from hosts or eukaryotic epiphytic organisms, we
@@
 a more detailed chloroplast classification. The largest proportion of
-sequences was classified as Bacillariophyta (89.7 ± 5.7 \si{\percent})
+sequences were classified as Bacillariophyta (89.7 ± 5.7 \si{\percent})
 indicating that the DNA removal procedure did not coextract larger
-quantities of host DNA (\autoref{chloroplast}).
+quantities of host DNA (\autoref{chloroplast}). Chloroplast sequences
+classified as Streptophyta constituted 3.3 ± 2.8 \si{\percent} of all
+chloroplast sequences originating from \emph{C. nodosa} samples, while
+sequences classified as Chlorophyta comprised only 0.02 ± 0.01
+\si{\percent} of all chloroplast sequences associated with \emph{C.
+cylindracea} samples.
+
+To determine if the extracted DNA can be used for metagenomic sequencing
+four samples containing epiphytic DNA were selected and shotgun
+sequenced using an Illumina platform. Metagenomic sequencing yielded
+from 207,149,524 to 624,029,930 sequence pairs
+(\autoref{supp-metagenomic_statistics}). Obtained sequences were
+successfully assembled into contigs whose L50 ranged from 654 to 1,011
+bp. In addition, predicted coding sequences were successfully
+functionally annotated (9,066,667 -- 20,256,215 annotated sequences;
+\autoref{cog}) and taxonomically classified. Functional annotation
+allowed for an assessment of the relative contribution of each COG
+(Clusters of Orthologous Groups) functional category to the total number
+of annotated coding sequences (\autoref{cog}a). Functional categories
+containing the highest number of sequences were: C (Energy production
+and conversion), E (Amino acid transport and metabolism), M (Cell
+wall/membrane/envelope biogenesis), L (Replication, recombination and
+repair) and P (Inorganic ion transport and metabolism). If host DNA was
+coextracted with epiphytic it should be detected in larger proportions
+in sequenced metagenomes. Indeed, no higher proportions of coding
+sequences classified into phylum Streptophyta or Chlorophyta were
+detected (\autoref{supp-metagenomic_taxonomy}). Sequenced metagenomic
+DNA originating from the surface of \emph{C. nodosa} contained 1.3
+\si{\percent} of coding sequences classified into the phylum
+Streptophyta in December 2017 and 0.7 \si{\percent} in June 2018.
+Furthermore, the summed RPKM (Reads Per Kilobase Million) of these
+sequences constituted 1.7 \si{\percent} of total RPKM of all
+successfully classified sequences in December 2017 and 1.1 \si{\percent}
+in June 2018. Similar low proportions of host coding sequences were
+detected in metagenomic samples originating from the surfaces of
+\emph{C. cylindracea}. Of all successfully classified coding sequences
+0.2 \si{\percent} were classified into Chlorophyta in December 2017 and
+0.1 \si{\percent} in June 2018. A relatively higher proportion of these
+sequences' RPKM than in the case of \emph{C. nodosa} was observed,
+indicating a higher coextraction of host DNA in \emph{C. cylindracea}.
+In December, the proportion of RPKM of sequences classified into
+Chlorophyta was 8.2 \si{\percent}, while in June 2018 it reached 13.6
+\si{\percent}.
+
+To evaluate if the procedure for protein extraction is suitable for
+metaproteomic analysis, obtained proteins were trypsin digested and
+sequenced using a mass spectrometer. Obtained MS/MS spectra were
+searched against a protein database from sequenced metagenomes. From
+14,219 to 16,449 proteins were identified in isolated protein samples
+(\autoref{cog}b). In addition, successful identification of proteins
+allowed for an assessment of the relative contribution of each COG
+functional category to the total number of identified proteins
+(\autoref{cog}b). Functional categories containing the highest number of
+identified proteins were: C (Energy production and conversion), G
+(Carbohydrate transport and metabolism), P (Inorganic ion transport and
+metabolism), O (Posttranslational modification, protein turnover,
+chaperones) and E (Amino acid transport and metabolism). Isolated
+proteins could derive from epiphytic organisms inhabiting the macrophyte
+surface or from macrophyte tissue underlying them. The contribution of
+proteins originating from host tissues was evaluated by identifying all
+the proteins predicted to belong to any taxonomic group within the phyla
+Streptophyta and Chlorophyta and by calculating their contribution to
+the number and abundance of all identified proteins. On average,
+proteins isolated from the surface of \emph{C. nodosa} contained 1.8 ±
+0.06 \si{\percent} of proteins associated with Streptophyta,
+contributing to 2.2 ± 0.8 \si{\percent} of total proteins. Similar to
+metagenomes, proteins associated with Chlorophyta contributed more to
+\emph{C. cylindracea} than proteins associated with Streptophyta to
+\emph{C. nodosa}. Chlorophyta associated proteins comprised 5.2 ± 0.06
+\si{\percent} of all identified proteins in \emph{C. cylindracea},
+contributing to 19.2 ± 1.5 \si{\percent} of all protein abundances.
 
@@
 \subsection{Discussion}\label{discussion}}
+
+The study of marine macrophyte epiphytic communities using culture
+independent techniques, such as 16S rRNA analysis, metagenomics and
+metaproteomics, requires an initial step of biological material
+isolation. Methods that have been developed for selective isolation of
+epiphytic biological material could be divided into two groups: (i)
+procedures involving, prior to extraction, a cell detachment step such
+as shaking (Gross \emph{et al.}, 2003; Nõges \emph{et al.}, 2010),
+scraping (Uku \emph{et al.}, 2007) and ultrasonication (Weidner \emph{et
+al.}, 1996; Cai \emph{et al.}, 2014) and (ii) procedures involving a
+host tissue incubation aiming at direct lysis of epiphytic microbial
+cells (Burke \emph{et al.}, 2009). Protocols that include a cell
+detachment procedure do not provide complete cell removal (Cai \emph{et
+al.}, 2014; Richter-Heitmann \emph{et al.}, 2016) and in the case of
+probe ultrasonication tissue disruption can also be observed
+(Richter-Heitmann \emph{et al.}, 2016). To avoid these problems a
+procedure for DNA isolation involving host tissue incubation in the
+extraction buffer and direct epiphytic cell lysis on the macrophyte
+surface was developed (Burke \emph{et al.}, 2009). Though providing an
+almost complete cell removal and selective DNA isolation this protocol
+uses in the extraction buffer a rapid multienzyme cleaner (3M) not
+available worldwide and without a known composition (Burke \emph{et
+al.}, 2009). To circumvent these problems we developed and tested two
+protocols for selective DNA and protein isolation from marine macrophyte
+epiphytic communities.
+
+To test if the developed DNA and protein isolation procedures could be
+applied on a variety of macrophyte species we selected \emph{C. nodosa}
+and \emph{C. cylindracea} as representatives of seagrass and macroalgal
+species, on which the procedures were tested. These species especially
+differ morphologically. While \emph{C. nodosa} leaves are flat, \emph{C.
+cylindracea} thallus is composed of uneven surfaces (Kuo and den Hartog,
+2001; Verlaque \emph{et al.}, 2003). The developed procedures showed an
+almost complete removal of epiphytic cells from the surfaces of both
+species comparable to the result of Burke \emph{et al.} (2009) and
+indicating that structural differences do not impact the removal
+efficiency. In addition, isolation protocols were tested in two
+contrasting season as it is known that macrophytes are harboring more
+algal epiphytes during autumn and winter (Reyes and Sansón, 2001). No
+differences in the removal efficiency was observed between seasons
+suggesting that protocols can be used on macrophyte samples retrieved
+throughout the year. Also, no removal differences were observed on
+samples derived from the same host but from different localities or
+settlements demonstrating the stability of the protocol in cell removal
+efficiency.
+
+Successful amplification and sequencing of the V4 16S rRNA gene region
+proved that the isolated DNA can be used to estimate the microbial
+epiphytic diversity. Taxonomic groups detected in this step can also be
+often found in epiphytic communities associated with other macrophytes
+(Burke and Thomas \emph{et al.}, 2011; Morrissey \emph{et al.}, 2019). A
+problem often encountered in studies focusing on epiphytic communities
+is the presence of large proportions of chloroplast 16S rRNA sequences
+in the pool of amplified molecules, especially if the epiphytic DNA was
+isolated without prior selection (Staufenberger \emph{et al.}, 2008).
+These sequences can derive from host chloroplasts or from eukaryotic
+epiphytic chloroplast DNA. Although the proportion of obtained
+chloroplast 16S rRNA sequences in our samples was substantial they
+derived almost exclusively from eukaryotic epiphytes. High proportion of
+chloroplast 16S rRNA sequences in studies applying selective procedures
+that include direct cellular lysis on host surfaces were observed before
+(Michelou \emph{et al.}, 2013). It is possible that chloroplast specific
+sequences even in these studies are originating from eukaryotic
+epiphytic cells and not host chloroplast. Indeed, it is common during
+16S rRNA profiling of pelagic microbial communities to observe high
+proportions of chloroplast sequences (Gilbert \emph{et al.}, 2009;
+Korlević \emph{et al.}, 2016). In addition, a very low proportion of
+chloroplast 16S rRNA sequences in samples originating from \emph{C.
+cylindracea} in comparison to \emph{C. nodosa} could be explained by the
+presence of three introns in the gene for 16S rRNA in some members of
+the genus \emph{Caulerpa} that could hamper the amplification process
+(Lam and Lopez-Bautista, 2016).
+
+Beside 16S rRNA sequencing high quality DNA is needed for metagenomics.
+The obtained number of metagenomic sequences and assembly statistics
+were comparable to metagenomes and metatranscriptomes derived from
+similar surface associated communities (Crump \emph{et al.}, 2018; Cúcio
+\emph{et al.}, 2018). In addition, functional annotation of predicted
+coding sequences to COG functional categories showed that the obtained
+metagenomes can be used to determine the metabolic capacity of surface
+associated communities (Leary \emph{et al.}, 2014; Cúcio \emph{et al.},
+2018). The proportion of coding sequences, including their RPKM,
+originating from \emph{C. nodosa} metagenomes and classified as
+Streptophyta was low indicating a good selectivity of the isolation
+procedure towards epiphytic cells. In the case of DNA samples isolated
+from the surface of \emph{C. cylindracea} the proportion of Chlorophyta
+coding sequences was also low but their RPKM was higher than in the case
+of \emph{C. nodosa}. One of the causes for this elevated RPKM of
+Chlorophyta sequences in \emph{C. cylindracea} could lay in tissue
+structure differences between these two host species. While \emph{C.
+nodosa} leaves are composed of individual cells, the thallus of \emph{C.
+cylindracea} is, like in other siphonous algal species, composed of a
+single large multinucleate cell (Coneva and Chitwood, 2015). The absence
+of individual cells in \emph{C. cylindracea} could cause a leakage of
+genetic material into the extraction buffer causing an elevated presence
+of host sequences in metagenomic data.
+
+Metagenomic analysis reveals the potential metabolic capacity of
+microbial communities, while to obtain an insight into the metabolic
+status of uncultivated members, a metaproteomic approach is required
+(Saito \emph{et al.}, 2019). The applied protocol for epipyhtic protein
+isolation followed by a metaproteomic analysis identified between 14,219
+and 16,449 proteins that is higher than previously reported for
+e.g.~soils (Chourey \emph{et al.}, 2010; Hultman \emph{et al.}, 2015),
+seawater (Williams \emph{et al.}, 2012) and biofilms (Leary \emph{et
+al.}, 2014). The functional annotation of identified proteins into COG
+functional categories showed that the protein isolation procedure can be
+used to assess the metabolic status of the macrophyte epiphytic
+community (Leary \emph{et al.}, 2014). Similar to metagenomes, the
+number of identified proteins, including their abundances, associated
+with Streptophyta in \emph{C. nodosa} samples were low indicating that
+the procedure is selective for epiphytic cell proteins. In addition, a
+higher number of identified proteins, and especially their abundances,
+associated with Chlorophyta was observed in \emph{C. cylindracea}
+samples. The cause of this elevated presence of Chlorophyta associated
+proteins can be, as in the case of the DNA isolation procedure,
+explained by the absence of individual cells in this siphonous alga
+(Coneva and Chitwood, 2015).
+
+In conclusion, the developed protocols for DNA and protein isolation
+from macrophyte surfaces are providing an almost complete removal of the
+epiphytic community and are shown to ensure removal from both, \emph{C.
+nodosa} and \emph{C. cylindracea}, in different seasons. Also, the
+obtained DNA and proteins are suitable for 16S rRNA sequencing,
+metagenomics and metaproteomics, while the obtained material contains
+low quantities of host DNA or proteins making the procedures epiphyte
+selective. Furthermore, the procedures are based on universally
+available laboratory chemicals making the protocols widely applicable.
 
@@
 nodosa} meadow invaded by \emph{C. cylindracea} in the proximity of the
-village of Funtana (\ang{45;10;39} N, \ang{13;35;42} E). Thalli of
-\emph{C. cylindracea} were sampled in the same \emph{C. nodosa} invaded
-meadow in Funtana and on a locality of only \emph{C. cylindracea}
-located in the proximity of the invaded meadow. Leaves and thalli were
-collected on the same day in two contrasting seasons, on 4 December 2017
-and 18 June 2018. During spring 2018 the \emph{C. nodosa} meadow in the
-Bay of Saline decayed to an extent that no leaves could be retrieved
-(Najdek \emph{et al.}, unpublished data). Leaves and thalli were
-collected by diving and transported to the laboratory in containers
-placed on ice and filled with site seawater. Upon arrival to the
-laboratory, \emph{C. nodosa} leaves were cut into sections of 1 -- 2
-\si{\cm}, while \emph{C. cylindracea} thalli were cut into 5 -- 8
-\si{\cm} long sections. Leaves and thalli were washed three times with
-sterile artificial seawater (ASW) to remove loosely attached microbial
-cells.
+village of Funtana (Bay of Funtana; \ang{45;10;39} N, \ang{13;35;42} E).
+Thalli of \emph{C. cylindracea} were sampled in the same \emph{C.
+nodosa} invaded meadow in the Bay Funtana and on a locality of only
+\emph{C. cylindracea} located in the proximity of the invaded meadow.
+Leaves and thalli for 16S rRNA analysis, metagenomics and metaproteomics
+were collected in two contrasting seasons, on 4 December 2017 (16S rRNA
+analysis and metaproteomics), 14 December 2017 (metagenomics) and 18
+June 2018 (16S rRNA analysis, metagenomics and metaproteomics). During
+spring 2018 the \emph{C. nodosa} meadow in the Bay of Saline decayed to
+an extent that no leaves could be retrieved (Najdek \emph{et al.},
+2020). In addition, as not enough DNA for both metagenomic and 16S RNA
+analysis were obtained during the sampling on 4 December 2017, an
+additional sampling on 14 December 2017 was carried out in the Bay of
+Funtana. Leaves and thalli were collected by diving and transported to
+the laboratory in containers placed on ice and filled with site
+seawater. Upon arrival to the laboratory, \emph{C. nodosa} leaves were
+cut into sections of 1 -- 2 \si{\cm}, while \emph{C. cylindracea} thalli
+were cut into 5 -- 8 \si{\cm} long sections. Leaves and thalli were
+washed three times with sterile artificial seawater (ASW) to remove
+loosely attached microbial cells.
 
@@
 filters described in Massana \emph{et al.} (1997). This protocol was
-modified and adapted for DNA isolation from microbial communities from
-macrophyte surfaces as described below. 5 \si{\ml} of lysis buffer (40
+modified and adapted for microbial DNA isolation from macrophyte
+surfaces as described below. 5 \si{\ml} of lysis buffer (40
 \si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar}
-sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or \si{\g}
-wet-weight of thalli. Lysozyme was added (final concentration 1
-\si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius}
-for 30 \si{\minute}. Subsequently, proteinase K (final concentration 0.5
+sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or 2
+\si{\g} wet-weight of thalli. For every sample duplicate isolations were
+performed. Lysozyme was added (final concentration 1 \si{\mg\per\ml})
+and the mixture was incubated at 37 \si{\degreeCelsius} for 30
+\si{\minute}. Subsequently, proteinase K (final concentration 0.5
 \si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were
@@
 sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
-Martinsried, Germany. Before sequencing, at IMGM the second PCR
+Martinsried, Germany. Before sequencing at IMGM, the second PCR
 amplification of the two-step PCR procedure was performed using primers
@@
 originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
-used. Sequences obtained in this study have been deposited in the
-European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers
-SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 --
-SAMEA6648811.
+used. Partial 16S rRNA sequences obtained in this study have been
+deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under
+accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794,
+SAMEA6648809 -- SAMEA6648811.
 
@@
 (Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
-2015, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018; Y. Xie, 2019b,
-2019a; Allaire \emph{et al.}, 2019; Zhu, 2019; Bengtsson, 2020). The
-detailed analysis procedure including the R Markdown file for this paper
-are available as a GitHub repository
+2015, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018; Wilke, 2018; Y.
+Xie, 2019b, 2019a; Allaire \emph{et al.}, 2019; Zhu, 2019; Bengtsson,
+2020). The detailed analysis procedure including the R Markdown file for
+this paper are available as a GitHub repository
 (\url{https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).
@@
 
-Four samples were selected and sent on dry ice to IMGM Laboratories,
+Four DNA samples were selected and sent on dry ice to IMGM Laboratories,
 Martinsried, Germany for metagenomic sequencing. DNA was purified using
@@
 
-Obtained sequences were assembled individually using MEGAHIT (version
-1.1.2) (Li \emph{et al.}, 2015) under default settings. Putative genes
-were predicted from contings longer than 200 bp using Prodigal (version
+Obtained sequences were analyzed on a Life Science Compute Cluster
+(LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
+Individual sequences were assembled using MEGAHIT (version 1.1.2) (Li
+\emph{et al.}, 2015) under default settings. Putative genes were
+predicted from contings longer than 200 bp using Prodigal (version
 2.6.3) (Hyatt \emph{et al.}, 2010) in metagenome mode (-p meta).
@@
 using the eggNOG-mapper (Huerta-Cepas \emph{et al.}, 2017) and eggNOG
-database (version 5.0) (Huerta-Cepas \emph{et al.}, 2019). The taxonomic
-classification of sequences was determined using the lowest common
-ancestor algorithm adapted from DIAMOND (version 0.8.36) (Buchfink
-\emph{et al.}, 2015) and by searching against the non-redundant database
-(NR). To determine phylogeny the top 10 \si{\percent} hits with an
-e-value \textless{} 1 × 10\textsuperscript{-5} were used (-\/-top 10).
+database (version 5.0) (Huerta-Cepas \emph{et al.}, 2019). Sequence
+taxonomy classification was determined using the lowest common ancestor
+algorithm adapted from DIAMOND (version 0.8.36) (Buchfink \emph{et al.},
+2015) and by searching against the NCBI non-redundant database (NR). To
+determine phylogeny the top 10 \si{\percent} hits with an e-value
+\textless{} 1 × 10\textsuperscript{-5} were used (-\/-top 10). Sequence
+renaming, coverage information computing and metagenomic statistics
+calculations were performed using software tools from BBTools
+(\url{https://jgi.doe.gov/data-and-tools/bbtools}). Metagenomic
+sequences obtained in this study have been deposited in the European
+Nucleotide Archive (ENA) at EMBL-EBI under accession numbers \textbf{TO
+BE ADDED}.
 
@@
 
-Proteins were isolated according to the protocol for isolation from soil
-described in Chourey \emph{et al.} (2010) and modified by Hultman
-\emph{et al.} (2015). These protocols were further modified and adapted
-for protein isolation from microbial communities form macrophyte
-surfaces as described below. 20 \si{\ml} of protein extraction buffer (4
+Proteins were isolated according to the protocol for protein isolation
+from soil described in Chourey \emph{et al.} (2010) and modified by
+Hultman \emph{et al.} (2015). This protocol was further modified and
+adapted for microbial protein isolation from macrophyte surfaces as
+described below. 20 \si{\ml} of protein extraction buffer (4
 \si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was
@@
 (filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
-according to the manufacturer's instructions (Wiśniewski \emph{et al.},
-2009) with small modifications. Before loading the solution to the
+according to the manufacturer's instructions with small modifications
+(Wiśniewski \emph{et al.}, 2009). Before the solution was loaded to the
 column, protein pellets were solubilized in a urea sample buffer
@@
 column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
-The final filtrate containing digested proteins was acidified with 1
+The final filtrate containing peptides was acidified with 1
 \si{\percent} (final concentration) trifluoroacetic acid, freezed at
@@
 metaproteomic analysis. Peptides were resuspended in 1 \si{\percent}
-(final concentration) trifluoroacetic acid, desalted using the Pierce
-C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's
-instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap
-Mass Spectrometer (Thermo Fisher Scientific, USA). Obtained MS/MS
-spectra were searched against a protein database from metagenomic
-assembly published in Burke and Peter Steinberg \emph{et al.} (2011)
-using SEQUEST-HT engines and validated with Percolator in Proteome
-Discoverer 2.1 (Thermo Fisher Scientific, USA). The target-decoy
-approach was used to reduce the probability of false peptide
-identification. Results whose false discovery rate at the peptide level
-was \SI{< 1}{\percent} were kept. For protein identification a minimum
-of of two peptides and one unique peptide were required. For protein
-quantification, a chromatographic peak area-based free quantitative
-method was applied.
+(final concentration) trifluoroacetic acid (TFA), desalted using the
+Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the
+manufacturer's instructions and sequenced on a Q Exactive Hybrid
+Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
+Obtained MS/MS spectra were searched against a protein database composed
+of combined sequenced metagenomes using SEQUEST-HT engines and validated
+with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific,
+USA). The target-decoy approach was used to reduce the probability of
+false peptide identification. Results whose false discovery rate at the
+peptide level was \SI{< 1}{\percent} were kept. For protein
+identification a minimum of two peptides and one unique peptide were
+required. For protein quantification, a chromatographic peak area-based
+free quantitative method was applied.
 
@@
 saline (PBS) and ethanol (1:1) and stored at \num{-20}
-\si{\degreeCelsius}. Treated and untreated leaves and thalli segments
+\si{\degreeCelsius}. Treated and untreated segemnts of leaves and thalli
 were stained in a 2 × solution of SYBR Green I and examined under a
@@
 \subsection{Acknowledgements}\label{acknowledgements}}
+
+This work was founded by the Croatian Science Foundation through the
+MICRO-SEAGRASS project (IP-2016-06-7118). We would like to thank
+Margareta Buterer for technical support, Paolo Paliaga for help during
+sampling, Lucija Horvat for technical support in confocal microscopy and
+Dušica Vujaklija for help in accessing the confocal microscope. In
+addition, we would like to thank the University Computing Center of the
+University of Zagreb for access to the computer cluster Isabella and
+CUBE -- Computational Systems Biology of the University of Vienna for
+access to the Life Science Compute Cluster (LiSC).
 
@@
 D633--D642.
+
+\leavevmode\hypertarget{ref-Coneva2015}{}%
+Coneva, V. and Chitwood, D.H. (2015) Plant architecture without
+multicellularity: Quandaries over patterning and the soma-germline
+divide in siphonous algae. \emph{Front Plant Sci} \textbf{6}: 287.
+
+\leavevmode\hypertarget{ref-Crump2018}{}%
+Crump, B.C., Wojahn, J.M., Tomas, F., and Mueller, R.S. (2018)
+Metatranscriptomics and amplicon sequencing reveal mutualisms in
+seagrass microbiomes. \emph{Front Microbiol} \textbf{9}: 388.
+
+\leavevmode\hypertarget{ref-Cucio2018}{}%
+Cúcio, C., Overmars, L., Engelen, A.H., and Muyzer, G. (2018)
+Metagenomic Analysis Shows the Presence of Bacteria Related to
+Free-Living Forms of Sulfur-Oxidizing Chemolithoautotrophic Symbionts in
+the Rhizosphere of the Seagrass Zostera marina. \emph{Front Mar Sci}
+\textbf{5}:
 
@@
 
+\leavevmode\hypertarget{ref-Korlevic2016}{}%
+Korlević, M., Šupraha, L., Ljubešić, Z., Henderiks, J., Ciglenečki, I.,
+Dautović, J., and Orlić, S. (2016) Bacterial diversity across a highly
+stratified ecosystem: A salt-wedge Mediterranean estuary. \emph{Syst
+Appl Microbiol} \textbf{39}: 398--408.
+
 \leavevmode\hypertarget{ref-Kozich2013}{}%
@@
 
+\leavevmode\hypertarget{ref-Kuo2001}{}%
+Kuo, J. and den Hartog, C. (2001) Seagrass taxonomy and identification
+key. In \emph{Global Seagrass Research Methods}. Short, F.T. and Coles,
+R.G. (eds). Amsterdam: Elsevier Science B.V., pp. 31--58.
+
+\leavevmode\hypertarget{ref-Lam2016}{}%
+Lam, D.W. and Lopez-Bautista, J.M. (2016) Complete chloroplast genome
+for \emph{Caulerpa racemosa} (Bryopsidales, Chlorophyta) and comparative
+analyses of siphonous green seaweed plastomes. \emph{Cymbella}
+\textbf{2}: 23--32.
+
+\leavevmode\hypertarget{ref-Leary2014}{}%
+Leary, D.H., Li, R.W., Hamdan, L.J., IV, W.J.H., Lebedev, N., Wang, Z.,
+et al. (2014) Integrated metagenomic and metaproteomic analyses of
+marine biofilm communities. \emph{Biofouling} \textbf{30}: 1211--1223.
+
 \leavevmode\hypertarget{ref-Li2015}{}%
@@
 
+\leavevmode\hypertarget{ref-Michelou2013}{}%
+Michelou, V.K., Caporaso, J.G., Knight, R., and Palumbi, S.R. (2013) The
+ecology of microbial communities associated with \emph{Macrocystis
+pyrifera}. \emph{PloS One} \textbf{8}: e67480.
+
+\leavevmode\hypertarget{ref-Morrissey2019}{}%
+Morrissey, K.L., Çavas, L., Willems, A., and De Clerck, O. (2019)
+Disentangling the influence of environment, host specificity and thallus
+differentiation on bacterial communities in siphonous green seaweeds.
+\emph{Front Microbiol} \textbf{10}: 717.
+
+\leavevmode\hypertarget{ref-Najdek2020}{}%
+Najdek, M., Korlević, M., Paliaga, P., Markovski, M., Ivančić, I.,
+Iveša, L., et al. (2020) Dynamics of environmental conditions during the
+decline of a \emph{Cymodocea nodosa} meadow. \emph{Biogeosciences}
+\textbf{17}: 3299--3315.
+
 \leavevmode\hypertarget{ref-Neuwirth2014}{}%
@@
 \emph{Front Microbiol} \textbf{7}: 773.
+
+\leavevmode\hypertarget{ref-Saito2019}{}%
+Saito, M.A., Bertrand, E.M., Duffy, M.E., Gaylord, D.A., Held, N.A.,
+Hervey, W.J., et al. (2019) Progress and Challenges in Ocean
+Metaproteomics and Proposed Best Practices for Data Sharing. \emph{J
+Proteome Res} \textbf{18}: 1461--1476.
 
@@
 
+\leavevmode\hypertarget{ref-Verlaque2003}{}%
+Verlaque, M., Durand, C., Huisman, J.M., Boudouresque, C.-F., and Le
+Parco, Y. (2003) On the identity and origin of the Mediterranean
+invasive \emph{Caulerpa} \emph{Racemosa} (Caulerpales, Chlorophyta).
+\emph{Eur J Phycol} \textbf{38}: 325--339.
+
 \leavevmode\hypertarget{ref-Weidner1996}{}%
@@
 
+\leavevmode\hypertarget{ref-Wilke2018}{}%
+Wilke, C.O. (2018) cowplot: Streamlined plot theme and plot annotations
+for 'ggplot2'.
+
+\leavevmode\hypertarget{ref-Williams2012}{}%
+Williams, T.J., Long, E., Evans, F., DeMaere, M.Z., Lauro, F.M.,
+Raftery, M.J., et al. (2012) A metaproteomic assessment of winter and
+summer bacterioplankton from Antarctic Peninsula coastal surface waters.
+\emph{ISME J} \textbf{6}: 1883--1900.
+
 \leavevmode\hypertarget{ref-Wisniewski2009}{}%
@@
 cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 4 December 2017 and stained with SYBR Green I.
-Scale bar in all images is 60 \si{\um}.\label{confocal_december}}\label{fig:unnamed-chunk-2}
+Scale bar in all images is 60 \si{\um}.\label{confocal_december}}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 cylindracea} surfaces from the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 19 June 2018 and stained with SYBR Green I.
-Scale bar in all images is 60 \si{\um}.\label{confocal_june}}\label{fig:unnamed-chunk-3}
+Scale bar in all images is 60 \si{\um}.\label{confocal_june}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 nodosa} and \textit{C.
-cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-4}
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 nodosa} and \textit{C.
-cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{chloroplast}}\label{fig:unnamed-chunk-5}
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{chloroplast}}\label{fig:unnamed-chunk-4}
 \end{figure}
 
+\newpage
+\begin{figure}[ht]
+
+{\centering \includegraphics[width=1\linewidth]{../results/figures/cog}
+
+}
+
+\caption{Relative contribution of each COG category to the total number of annotated coding sequences (A) from metagenomes and identified proteins (B) from metaproteomes associated with surfaces of two marine macrophytes (\textit{C.
+nodosa} and \textit{C.
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4/14 December 2017 and 19 June 2018).
+Total number of annotated coding sequences and identified proteins is given above the corresponding bar.\label{cog}}\label{fig:unnamed-chunk-5}
+\end{figure}
+
 \end{document}

### 2020-09-14 submission/supplementary.Rmd
Manuscript done.

@@
 
-\vspace{60mm}
+\vspace{20mm}
 Marino Korlević^1$*$^, Marsej Markovski^1^, Zihao Zhao^2^, Gerhard J. Herndl^2,3^, Mirjana Najdek^1^
-
-\vspace{40mm}
-
-^$*$^To whom correspondence should be addressed: marino.korlevic@irb.hr
 
@@
 
-3\. Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands
+3\. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands
 
+^$*$^To whom correspondence should be addressed:
+
+Marino Korlević
+
+G. Paliaga 5, 52210 Rovinj, Croatia
+
+Tel.: +385 52 804 768
+
+Fax: +385 52 804 780
+
+e-mail: marino.korlevic@irb.hr
+
+Running title: DNA and protein isolation from macrophyte surfaces
+
+\linenumbers
 \sisetup{mode=text}
@@
 ```{r nseq_notus}
+# 16S rRNA statistics
 # Loading OTU/sample data
@@
 of Sequences", "No.
-of OTUs"), caption="Sample ID, Station, Community Type, Sampling Date, No.
-of Sequences and No.
-of OTUs of each sample.
-No.
-of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\\label{nseq_notus}", align="c") %>%
+of OTUs"), caption="Sample ID, sampling station, community type, sampling date, number of sequences and number of OTUs of each sample.
+The number of sequences and OTUs was calculated after exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences.\\label{nseq_notus}", align="c") %>%
   kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
@@
 ```
+
+\newpage
+\blandscape
+```{r metagenomic_statistics}
+# Metagenomic statistics
+# Loading raw sequence statistics
+raw_45_R1 <- read_tsv("results/metagenomics/statistics/00raw_45_R1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+raw_45_R2 <- read_tsv("results/metagenomics/statistics/00raw_45_R2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+raw_47_R1 <- read_tsv("results/metagenomics/statistics/00raw_47_R1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+raw_47_R2 <- read_tsv("results/metagenomics/statistics/00raw_47_R2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+raw_61_R1 <- read_tsv("results/metagenomics/statistics/00raw_61_R1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+raw_61_R2 <- read_tsv("results/metagenomics/statistics/00raw_61_R2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+raw_63_R1 <- read_tsv("results/metagenomics/statistics/00raw_63_R1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+raw_63_R2 <- read_tsv("results/metagenomics/statistics/00raw_63_R2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+# Loading sequence statistics following AdapterRemoval
+truncated_45_R1 <- read_tsv("results/metagenomics/statistics/01truncated_45_pair1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+truncated_45_R2 <- read_tsv("results/metagenomics/statistics/01truncated_45_pair2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+truncated_47_R1 <- read_tsv("results/metagenomics/statistics/01truncated_47_pair1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+truncated_47_R2 <- read_tsv("results/metagenomics/statistics/01truncated_47_pair2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+truncated_61_R1 <- read_tsv("results/metagenomics/statistics/01truncated_61_pair1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+truncated_61_R2 <- read_tsv("results/metagenomics/statistics/01truncated_61_pair2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+truncated_63_R1 <- read_tsv("results/metagenomics/statistics/01truncated_63_pair1.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+truncated_63_R2 <- read_tsv("results/metagenomics/statistics/01truncated_63_pair2.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+# Loading sequence statistics following MEGAHIT
+contigs_45 <- read_tsv("results/metagenomics/statistics/03contigs_45.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:" | X1=="Main genome contig N/L50:") %>%
+  separate(X2, c("X2", "X3"), sep="/", fill="right")
+contigs_47 <- read_tsv("results/metagenomics/statistics/03contigs_47.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:" | X1=="Main genome contig N/L50:") %>%
+  separate(X2, c("X2", "X3"), sep="/", fill="right")
+contigs_61 <- read_tsv("results/metagenomics/statistics/03contigs_61.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:" | X1=="Main genome contig N/L50:") %>%
+  separate(X2, c("X2", "X3"), sep="/", fill="right")
+contigs_63 <- read_tsv("results/metagenomics/statistics/03contigs_63.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:" | X1=="Main genome contig N/L50:") %>%
+  separate(X2, c("X2", "X3"), sep="/", fill="right")
+
+# Loading sequence statistics following Prodigal
+cds_45 <- read_tsv("results/metagenomics/statistics/04cds_nucl_45.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+cds_47 <- read_tsv("results/metagenomics/statistics/04cds_nucl_47.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+cds_61 <- read_tsv("results/metagenomics/statistics/04cds_nucl_61.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+cds_63 <- read_tsv("results/metagenomics/statistics/04cds_nucl_63.txt", col_names=FALSE) %>%
+  filter(X1=="Main genome scaffold total:")
+
+# Loading sequence statistics following eggNOG mapper
+eggnog_45 <- read_tsv("results/metagenomics/statistics/07eggnog_45.txt", col_names=FALSE)
+eggnog_47 <- read_tsv("results/metagenomics/statistics/07eggnog_47.txt", col_names=FALSE)
+eggnog_61 <- read_tsv("results/metagenomics/statistics/07eggnog_61.txt", col_names=FALSE)
+eggnog_63 <- read_tsv("results/metagenomics/statistics/07eggnog_63.txt", col_names=FALSE)
+
+# Combining statistical data in a table
+metagenomic_statistics <- tibble(
+  ID=c(45, 47, 61, 63),
+  raw_R1=c(raw_45_R1$X2, raw_47_R1$X2, raw_61_R1$X2, raw_63_R1$X2),
+  raw_R2=c(raw_45_R2$X2, raw_47_R2$X2, raw_61_R2$X2, raw_63_R2$X2),
+  truncated_R1=c(truncated_45_R1$X2, truncated_47_R1$X2, truncated_61_R1$X2, truncated_63_R1$X2),
+  truncated_R2=c(truncated_45_R2$X2, truncated_47_R2$X2, truncated_61_R2$X2, truncated_63_R2$X2),
+  contigs=c(contigs_45$X2[1], contigs_47$X2[1], contigs_61$X2[1], contigs_63$X2[1]),
+  N50=c(contigs_45$X2[2], contigs_47$X2[2], contigs_61$X2[2], contigs_63$X2[2]),
+  L50=c(contigs_45$X3[2], contigs_47$X3[2], contigs_61$X3[2], contigs_63$X3[2]),
+  cds=c(cds_45$X2, cds_47$X2, cds_61$X2, cds_63$X2),
+  eggnog=c(eggnog_45$X2, eggnog_47$X2, eggnog_61$X2, eggnog_63$X2)
+  ) %>%
+  mutate(L50=str_replace(L50, " KB", "")) %>%
+  mutate(L50=str_replace(L50, "\\.", ""))
+
+# Loading metagenomic metadata
+metadata_metagenomics <- read_tsv("data/raw/metadata_metagenomics.csv")
+
+# Joining metagenomic metadata with metagenomic statistical data
+invisible(Sys.setlocale(locale="en_GB.utf8"))
+metadata_metagenomic_statistics <- inner_join(metadata_metagenomics, metagenomic_statistics, by=c("ID"="ID")) %>%
+  select(ID, station, date, raw_R1, contigs, N50, L50, cds, eggnog) %>%
+  mutate(type=station) %>%
+  select(ID, station, type, everything()) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  mutate(date=format(date, "%d %B %Y")) %>%
+  mutate(date=str_replace(date, "^0", "")) %>%
+  arrange(ID) %>%
+  mutate(type=str_replace(type, "^SCy$", "\\\\textit{Cymodocea nodosa} (Monospecific)")) %>%
+  mutate(type=str_replace(type, "^FCyM$", "\\\\textit{Cymodocea nodosa} (Mixed)")) %>%
+  mutate(type=str_replace(type, "^FCaM$", "\\\\textit{Caulerpa cylindracea} (Mixed)")) %>%
+  mutate(type=str_replace(type, "^FCa$", "\\\\textit{Caulerpa cylindracea} (Monospecific)")) %>%
+  mutate(station=str_replace(station, "^SCy$", "Saline")) %>%
+  mutate(station=str_replace(station, "^FCyM$", "Funtana")) %>%
+  mutate(station=str_replace(station, "^FCaM$", "Funtana")) %>%
+  mutate(station=str_replace(station, "^FCa$", "Funtana")) %>%
+  mutate_at(vars(c(raw_R1, contigs, N50, L50, cds, eggnog)), list(~as.numeric(.))) %>%
+  mutate(raw_R1=format(raw_R1, big.mark=","),
+         contigs=format(contigs, big.mark=","),
+         N50=format(N50, big.mark=","),
+         L50=format(L50, big.mark=","),
+         cds=format(cds, big.mark=","),
+         eggnog=format(eggnog, big.mark=","))
+
+# Generating the table
+kable(metadata_metagenomic_statistics, format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=linebreak(c("Sample ID", "Station", "Community Type", "Date", "No.
+of Raw\nSequence Pairs", "No.
+of\nContigs", "N50\\textsuperscript{*}", "L50 (bp)\\textsuperscript{*}", "No.
+of\nPredicted\nCDS", "No.
+of\nAnnotated\nCDS"), align="c"), caption="Sample ID, sampling station, community type, sampling date, number of raw sequence pairs, number of assembled contigs by MEGAHIT, N50 and L50 assembly statistics, number of predicted coding sequences (CDS) by Prodigal and number of eggNOGmapper annotated CDS.\\label{metagenomic_statistics}", align="c") %>%
+  kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
+  row_spec(0, bold=TRUE) %>%
+  column_spec(1, width="5em") %>%
+  footnote(symbol="The notation was preserved from the original output of BBTools stats.sh.")
+```
+
+\elandscape
+\newpage
+```{r metagenomic_taxonomy}
+# Metagenomic taxonomy
+# Loading summary taxonomy files (phylum)
+phylum_45 <- read_tsv("results/metagenomics/taxonomy/45.phylum.tsv") %>%
+  add_column(sample = 45, .before = "phylum") %>%
+  filter(!is.na(phylum)) %>%
+  mutate(cds_relative = cds / sum(cds) * 100) %>%
+  mutate(rpkm_relative = rpkm / sum(rpkm) * 100) %>%
+  arrange(desc(rpkm_relative)) %>%
+  top_n(10)
+phylum_47 <- read_tsv("results/metagenomics/taxonomy/47.phylum.tsv") %>%
+  add_column(sample = 47, .before = "phylum") %>%
+  filter(!is.na(phylum)) %>%
+  mutate(cds_relative = cds / sum(cds) * 100) %>%
+  mutate(rpkm_relative = rpkm / sum(rpkm) * 100) %>%
+  arrange(desc(rpkm_relative)) %>%
+  top_n(10)
+phylum_61 <- read_tsv("results/metagenomics/taxonomy/61.phylum.tsv") %>%
+  add_column(sample = 61, .before = "phylum") %>%
+  filter(!is.na(phylum)) %>%
+  mutate(cds_relative = cds / sum(cds) * 100) %>%
+  mutate(rpkm_relative = rpkm / sum(rpkm) * 100) %>%
+  arrange(desc(rpkm_relative)) %>%
+  top_n(10)
+phylum_63 <- read_tsv("results/metagenomics/taxonomy/63.phylum.tsv") %>%
+  add_column(sample = 63, .before = "phylum") %>%
+  filter(!is.na(phylum)) %>%
+  mutate(cds_relative = cds / sum(cds) * 100) %>%
+  mutate(rpkm_relative = rpkm / sum(rpkm) * 100) %>%
+  arrange(desc(rpkm_relative)) %>%
+  top_n(10)
+
+# Combining summary taxonomy files (phylum)
+metagenomic_phylum <- bind_rows(phylum_45, phylum_47, phylum_61, phylum_63)
+
+# Loading metagenomic metadata 
+metadata_metagenomics <- read_tsv("data/raw/metadata_metagenomics.csv")
+
+# Joining metagenomic metadata with metagenomic taxonomy data
+invisible(Sys.setlocale(locale="en_GB.utf8"))
+metadata_metagenomic_phylum <- inner_join(metadata_metagenomics, metagenomic_phylum, by=c("ID"="sample")) %>%
+  select(ID, station, date, phylum, cds, cds_relative, rpkm, rpkm_relative) %>%
+  mutate(type=station) %>%
+  select(ID, station, type, everything()) %>%
+  mutate(date=as.Date(date, "%d.%m.%Y")) %>%
+  mutate(date=format(date, "%d %B %Y")) %>%
+  mutate(date=str_replace(date, "^0", "")) %>%
+  arrange(ID) %>%
+  mutate(phylum=str_replace(phylum, "^Proteobacteria$", "\\\\textit{Proteobacteria}")) %>%
+  mutate(phylum=str_replace(phylum, "^Cyanobacteria$", "\\\\textit{Cyanobacteria}")) %>%
+  mutate(phylum=str_replace(phylum, "^Bacteroidetes$", "\\\\textit{Bacteroidetes}")) %>%
+  mutate(phylum=str_replace(phylum, "^Actinobacteria$", "\\\\textit{Actinobacteria}")) %>%
+  mutate(phylum=str_replace(phylum, "^Planctomycetes$", "\\\\textit{Planctomycetes}")) %>%
+  mutate(phylum=str_replace(phylum, "^Verrucomicrobia$", "\\\\textit{Verrucomicrobia}")) %>%
+  mutate(phylum=str_replace(phylum, "^Chloroflexi$", "\\\\textit{Chloroflexi}")) %>%
+  mutate(phylum=str_replace(phylum, "^Chlorophyta$", "\\\\textbf{Chlorophyta}")) %>%
+  mutate(phylum=str_replace(phylum, "^Streptophyta$", "\\\\textbf{Streptophyta}")) %>%  
+  mutate(cds=format(cds, big.mark=","),
+         cds_relative=format(round(cds_relative, 2), big.mark=',', nsmall=1),
+         rpkm=format(rpkm, big.mark=","),
+         rpkm_relative=format(round(rpkm_relative, 2), big.mark=',', nsmall=1))
+
+# Generating the table
+select(metadata_metagenomic_phylum, phylum, cds, cds_relative, rpkm, rpkm_relative) %>%
+kable(format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=linebreak(c("Phylum", "No.
+of CDS", "CDS (\\%)", "Summed RPKM", "RPKM (\\%)"), align="c"), caption="Phyla into which CDS were classifed, number and proportion of CDS in different phyla and sum of coding sequences' RPKM (Reads Per Kilobase Million) and their proportion in different phyla.
+Data are derived from sequenced metagenomes.
+Each metagenomic sample is labeled with sampling station, community type, sampling date and sample ID.
+For each sample top ten phyla based on RPKM were selected.
+CDS that were not successfully classified were excluded from the dataset.\\label{metagenomic_taxonomy}", align="c") %>%
+  kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9, full_width=TRUE) %>%
+  row_spec(0, bold=TRUE) %>%
+  group_rows("Funtana, \\\\textit{Cymodocea nodosa} (Monospecific), 14 December 2017 (sample ID: 45)", 1, 10, escape=FALSE, latex_gap_space="0.6 em") %>%
+  group_rows("Funtana, \\\\textit{Caulerpa cylindracea} (Monospecific), 14 December 2017 (sample ID: 47)", 11, 20, escape=FALSE, latex_gap_space="0.6 em") %>%
+  group_rows("Funtana, \\\\textit{Cymodocea nodosa} (Mixed), 19 June 2018 (sample ID: 61)", 21, 30, escape=FALSE, latex_gap_space="0.9 em") %>%
+  group_rows("Funtana, \\\\textit{Caulerpa cylindracea} (Monospecific), 19 June 2018 (sample ID: 63)", 31, 40, escape=FALSE, latex_gap_space="0.6 em")
+```

### 2020-09-14 submission/supplementary.tex
Manuscript done.

@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Supplementary Information},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1.0in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
-            pdftitle={Supplementary Information},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
 \usepackage{graphicx,grffile}
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{\textbf{Supplementary Information}}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-  \subtitle{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
-Surfaces}}
-  \author{}
-    \preauthor{}\postauthor{}
-    \date{}
-    \predate{}\postdate{}
-
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 \usepackage{times} % Times New Roman font
@@
 
+\usepackage{txfonts}
+
 \renewcommand{\figureautorefname}{Fig.}
@@
 
+\title{\textbf{Supplementary Information}}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
+Surfaces}}
+\author{}
+\date{\vspace{-2.5em}}
+
 \begin{document}
@@
 
-\vspace{60mm}
+\vspace{20mm}
 
@@
 Herndl\textsuperscript{2,3}, Mirjana Najdek\textsuperscript{1}
-
-\vspace{40mm}
-
-\textsuperscript{\(*\)}To whom correspondence should be addressed:
-\href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}
 
@@
 
+\textsuperscript{\(*\)}To whom correspondence should be addressed:
+
+Marino Korlević
+
+G. Paliaga 5, 52210 Rovinj, Croatia
+
+Tel.: +385 52 804 768
+
+Fax: +385 52 804 780
+
+e-mail:
+\href{mailto:marino.korlevic@irb.hr}{\nolinkurl{marino.korlevic@irb.hr}}
+
+Running title: DNA and protein isolation from macrophyte surfaces
+
+\linenumbers
 \sisetup{mode=text}
@@
 \begin{longtable}[t]{>{\centering\arraybackslash}p{6em}ccccc}
-\caption{\label{tab:nseq_notus}Sample ID, Station, Community Type, Sampling Date, No.
-of Sequences and No.
-of OTUs of each sample.
-No.
-of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\label{nseq_notus}}\\
+\caption{\label{tab:nseq_notus}Sample ID, sampling station, community type, sampling date, number of sequences and number of OTUs of each sample.
+The number of sequences and OTUs was calculated after exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences.\label{nseq_notus}}\\
 \toprule
@@
 \endfirsthead
-\caption[]{Sample ID, Station, Community Type, Sampling Date, No.
-of Sequences and No.
-of OTUs of each sample.
-No.
-of Sequences and OTUs was calculated after the exclusion of eukaryotic, chloroplast, mitochondrial and no relative sequences.\label{nseq_notus} \textit{(continued)}}\\
+\caption[]{Sample ID, sampling station, community type, sampling date, number of sequences and number of OTUs of each sample.
+The number of sequences and OTUs was calculated after exclusion of sequences without known relatives (no relative sequences) and eukaryotic, chloroplast and mitochondrial sequences.\label{nseq_notus} \textit{(continued)}}\\
 \toprule
@@
 
+\newpage
+\blandscape
+\begingroup\fontsize{9}{11}\selectfont
+
+\begin{longtable}[t]{>{\centering\arraybackslash}p{5em}ccccccccc}
+\caption{\label{tab:metagenomic_statistics}Sample ID, sampling station, community type, sampling date, number of raw sequence pairs, number of assembled contigs by MEGAHIT, N50 and L50 assembly statistics, number of predicted coding sequences (CDS) by Prodigal and number of eggNOGmapper annotated CDS.\label{metagenomic_statistics}}\\
+\toprule
+\textbf{Sample ID} & \textbf{Station} & \textbf{Community Type} & \textbf{Date} & \textbf{\makecell[c]{No.
+of Raw\\Sequence Pairs}} & \textbf{\makecell[c]{No.
+of\\Contigs}} & \textbf{N50\textsuperscript{*}} & \textbf{L50 (bp)\textsuperscript{*}} & \textbf{\makecell[c]{No.
+of\\Predicted\\CDS}} & \textbf{\makecell[c]{No.
+of\\Annotated\\CDS}}\\
+\midrule
+\endfirsthead
+\caption[]{Sample ID, sampling station, community type, sampling date, number of raw sequence pairs, number of assembled contigs by MEGAHIT, N50 and L50 assembly statistics, number of predicted coding sequences (CDS) by Prodigal and number of eggNOGmapper annotated CDS.\label{metagenomic_statistics} \textit{(continued)}}\\
+\toprule
+\textbf{Sample ID} & \textbf{Station} & \textbf{Community Type} & \textbf{Date} & \textbf{\makecell[c]{No.
+of Raw\\Sequence Pairs}} & \textbf{\makecell[c]{No.
+of\\Contigs}} & \textbf{N50\textsuperscript{*}} & \textbf{L50 (bp)\textsuperscript{*}} & \textbf{\makecell[c]{No.
+of\\Predicted\\CDS}} & \textbf{\makecell[c]{No.
+of\\Annotated\\CDS}}\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\multicolumn{10}{l}{\textsuperscript{*} The notation was preserved from the original output of BBTools stats.sh.}\\
+\endlastfoot
+45 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 14 December 2017 & 288,446,922 & 10,786,127 & 1,814,108 & 1,011 & 15,230,601 & 9,066,667\\
+47 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 14 December 2017 & 207,149,524 & 14,541,483 & 3,417,214 & 684 & 19,415,048 & 12,179,801\\
+61 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 19 June 2018 & 624,029,930 & 25,843,073 & 5,036,213 & 873 & 35,296,634 & 20,256,215\\
+63 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 19 June 2018 & 241,132,752 & 15,909,915 & 4,071,946 & 654 & 20,643,084 & 13,064,686\\*
+\end{longtable}
+\endgroup{}
+
+\elandscape
+\newpage
+\begingroup\fontsize{9}{11}\selectfont
+
+\begin{longtabu} to \linewidth {>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
+\caption{\label{tab:metagenomic_taxonomy}Phyla into which CDS were classifed, number and proportion of CDS in different phyla and sum of coding sequences' RPKM (Reads Per Kilobase Million) and their proportion in different phyla.
+Data are derived from sequenced metagenomes.
+Each metagenomic sample is labeled with sampling station, community type, sampling date and sample ID.
+For each sample top ten phyla based on RPKM were selected.
+CDS that were not successfully classified were excluded from the dataset.\label{metagenomic_taxonomy}}\\
+\toprule
+\textbf{Phylum} & \textbf{No. of CDS} & \textbf{CDS (\%)} & \textbf{Summed RPKM} & \textbf{RPKM (\%)}\\
+\midrule
+\endfirsthead
+\caption[]{Phyla into which CDS were classifed, number and proportion of CDS in different phyla and sum of coding sequences' RPKM (Reads Per Kilobase Million) and their proportion in different phyla.
+Data are derived from sequenced metagenomes.
+Each metagenomic sample is labeled with sampling station, community type, sampling date and sample ID.
+For each sample top ten phyla based on RPKM were selected.
+CDS that were not successfully classified were excluded from the dataset.\label{metagenomic_taxonomy} \textit{(continued)}}\\
+\toprule
+Phylum & No. of CDS & CDS (\%) & Summed RPKM & RPKM (\%)\\
+\midrule
+\endhead
+\
+\endfoot
+\bottomrule
+\endlastfoot
+\addlinespace[0.6 em]
+\multicolumn{5}{l}{\textbf{Funtana, \textit{Cymodocea nodosa} (Monospecific), 14 December 2017 (sample ID: 45)}}\\
+\hspace{1em}\textit{Proteobacteria} & 4,021,486 & 57.08 & 526,874.683 & 56.71\\
+\hspace{1em}\textit{Cyanobacteria} & 459,942 & 6.53 & 155,294.303 & 16.72\\
+\hspace{1em}\textit{Bacteroidetes} & 1,388,758 & 19.71 & 116,972.727 & 12.59\\
+\hspace{1em}\textit{Actinobacteria} & 298,264 & 4.23 & 38,176.240 & 4.11\\
+\hspace{1em}Bacillariophyta & 252,243 & 3.58 & 33,816.874 & 3.64\\
+\hspace{1em}\textbf{Streptophyta} & 91,316 & 1.30 & 15,839.692 & 1.71\\
+\hspace{1em}Rhodophyta & 76,937 & 1.09 & 15,298.496 & 1.65\\
+\hspace{1em}\textit{Planctomycetes} & 227,863 & 3.23 & 9,778.230 & 1.05\\
+\hspace{1em}\textit{Verrucomicrobia} & 39,912 & 0.57 & 2,142.129 & 0.23\\
+\hspace{1em}\textit{Chloroflexi} & 35,228 & 0.50 & 1,848.597 & 0.20\\
+\addlinespace[0.6 em]
+\multicolumn{5}{l}{\textbf{Funtana, \textit{Caulerpa cylindracea} (Monospecific), 14 December 2017 (sample ID: 47)}}\\
+\hspace{1em}\textit{Proteobacteria} & 5,384,137 & 61.24 & 493,882.897 & 57.93\\
+\hspace{1em}\textit{Bacteroidetes} & 1,187,188 & 13.50 & 77,090.557 & 9.04\\
+\hspace{1em}\textbf{Chlorophyta} & 13,745 & 0.16 & 70,249.251 & 8.24\\
+\hspace{1em}\textit{Actinobacteria} & 444,926 & 5.06 & 55,212.308 & 6.48\\
+\hspace{1em}\textit{Cyanobacteria} & 363,766 & 4.14 & 41,167.514 & 4.83\\
+\hspace{1em}\textit{Planctomycetes} & 502,248 & 5.71 & 27,522.230 & 3.23\\
+\hspace{1em}Bacillariophyta & 243,702 & 2.77 & 23,336.154 & 2.74\\
+\hspace{1em}\textit{Verrucomicrobia} & 178,424 & 2.03 & 10,152.046 & 1.19\\
+\hspace{1em}Porifera & 17,398 & 0.20 & 10,105.655 & 1.19\\
+\hspace{1em}Rhodophyta & 48,544 & 0.55 & 7,117.314 & 0.83\\
+\addlinespace[0.9 em]
+\multicolumn{5}{l}{\textbf{Funtana, \textit{Cymodocea nodosa} (Mixed), 19 June 2018 (sample ID: 61)}}\\
+\hspace{1em}\textit{Proteobacteria} & 8,185,781 & 55.21 & 573,484.714 & 65.33\\
+\hspace{1em}\textit{Bacteroidetes} & 2,226,547 & 15.02 & 78,921.488 & 8.99\\
+\hspace{1em}Bacillariophyta & 761,510 & 5.14 & 71,003.503 & 8.09\\
+\hspace{1em}\textit{Actinobacteria} & 655,055 & 4.42 & 43,153.463 & 4.92\\
+\hspace{1em}\textit{Planctomycetes} & 1,338,538 & 9.03 & 35,240.242 & 4.01\\
+\hspace{1em}\textit{Cyanobacteria} & 474,226 & 3.20 & 24,941.374 & 2.84\\
+\hspace{1em}\textit{Verrucomicrobia} & 371,837 & 2.51 & 19,794.186 & 2.25\\
+\hspace{1em}\textbf{Streptophyta} & 108,235 & 0.73 & 9,555.978 & 1.09\\
+\hspace{1em}\textit{Chloroflexi} & 150,402 & 1.01 & 3,438.376 & 0.39\\
+\hspace{1em}Mollusca & 50,750 & 0.34 & 1,768.748 & 0.20\\
+\addlinespace[0.6 em]
+\multicolumn{5}{l}{\textbf{Funtana, \textit{Caulerpa cylindracea} (Monospecific), 19 June 2018 (sample ID: 63)}}\\
+\hspace{1em}\textit{Proteobacteria} & 5,429,374 & 60.24 & 440,876.241 & 56.73\\
+\hspace{1em}\textbf{Chlorophyta} & 13,360 & 0.15 & 105,595.034 & 13.59\\
+\hspace{1em}\textit{Bacteroidetes} & 1,084,784 & 12.04 & 64,827.936 & 8.34\\
+\hspace{1em}\textit{Planctomycetes} & 794,696 & 8.82 & 43,962.118 & 5.66\\
+\hspace{1em}\textit{Actinobacteria} & 394,004 & 4.37 & 28,048.256 & 3.61\\
+\hspace{1em}\textit{Cyanobacteria} & 237,026 & 2.63 & 17,046.410 & 2.19\\
+\hspace{1em}\textit{Verrucomicrobia} & 203,253 & 2.26 & 12,033.107 & 1.55\\
+\hspace{1em}Bacillariophyta & 174,954 & 1.94 & 11,959.067 & 1.54\\
+\hspace{1em}\textit{Chloroflexi} & 175,514 & 1.95 & 8,952.527 & 1.15\\
+\hspace{1em}\textbf{Streptophyta} & 4,771 & 0.05 & 6,752.150 & 0.87\\*
+\end{longtabu}
+\endgroup{}
+
 \end{document}

### 2020-09-14 code/plot_cog.R

@@ -1,0 +1,212 @@
+#################################################################################################################
+# plot_cog.R
+# 
+# A script to plot the relative abundance of each COG functional category in each metagenomic and
+# metaproteomic sample.
+# Dependencies: data/results/metagenomics/cog/45.cog_categories.tsv
+#               data/results/metagenomics/cog/47.cog_categories.tsv
+#               data/results/metagenomics/cog/61.cog_categories.tsv
+#               data/results/metagenomics/cog/63.cog_categories.tsv  
+#               data/results/metaproteomics/metaproteomics_cog.tsv
+#               data/raw/cog_categories_colors.tsv
+#               
+# Produces: results/figures/cog.jpg
+#
+#################################################################################################################
+
+##############################################
+# COG functional categories in metagenomes
+##############################################
+
+# Loading COG input data and combining together
+cog_45 <- read_tsv("results/metagenomics/cog/45.cog_categories.tsv", col_names = c("COG Functional Category", "45"), skip = 1)
+cog_47 <- read_tsv("results/metagenomics/cog/47.cog_categories.tsv", col_names = c("COG Functional Category", "47"), skip = 1)
+cog_61 <- read_tsv("results/metagenomics/cog/61.cog_categories.tsv", col_names = c("COG Functional Category", "61"), skip = 1)
+cog_63 <- read_tsv("results/metagenomics/cog/63.cog_categories.tsv", col_names = c("COG Functional Category", "63"), skip = 1)
+
+cog <- full_join(cog_45, cog_47, by = c("COG Functional Category")) %>%
+  full_join(., cog_61, by = c("COG Functional Category")) %>%
+  full_join(., cog_63, by = c("COG Functional Category"))
+
+# Calculating relative abundances of each COG functional category
+cog_relative <- cog %>%
+  mutate(`COG Functional Category`, `COG Functional Category`=if_else(str_length(`COG Functional Category`) > 1,
[…]

### 2020-09-14 submission/manuscript.Rmd
Coauthor review.

@@
 ## Summary
-Studies of unculturable microbes often combine methods based on DNA, such as 16S rRNA sequencing and metagenomics, with methods that allow an insight into the metabolic status, such as metaproteomics.
-To apply these procedures to the microbial community inhabiting the surfaces of marine macrophytes it is advisable to perform, prior to the analysis, a selective DNA and protein isolation so that the host material, present in higher quantities, is not hampering the analysis.
+Studies of unculturable microbes often combine methods based on DNA, such as 16S rRNA sequencing and metagenomics, with methods that allow insight into the metabolic status, such as metaproteomics.
+To apply these techniques to the microbial community inhabiting the surfaces of marine macrophytes it is advisable to perform, prior to the analysis, a selective DNA and protein isolation so that the host material, present in higher quantities, is not hampering the analysis.
 Two protocols, for DNA and protein isolation, were adapted for selective extractions of DNA and proteins from epiphytic communities inhabiting the surfaces of two marine macrophytes: the seagrass *Cymodocea nodosa* and the macroalga *Caulerpa cylindracea*.
 Protocols showed an almost complete removal of the epiphytic community regardless of the sampling season, station, settlement or host species.
-The obtained DNA was suitable for metagenomic and 16S rRNA sequencing, while isolated proteins could be identified by mass spectrometry showing that protocols can be used in 16S rRNA, metagenomic and metaproteomic analysis.
-Low presence of host DNA and proteins, observed in isolated samples, indicated a selective nature of protocols.
+The obtained DNA was suitable for metagenomic and 16S rRNA sequencing, while isolated proteins could be identified by mass spectrometry, showing that protocols can be used in 16S rRNA, metagenomic and metaproteomic analysis.
+Low presence of host DNA and proteins, observed in isolated samples, indicated a selective nature of the protocols.
 Furthermore, the procedures are based on universally available laboratory chemicals making the protocols widely applicable.
@@
 An alternative to these procedures is a direct isolation of the targeted material by incubating macrophyte tissues in an extraction buffer.
-After the incubation, the undisrupted host tissue is removed and the isolation procedure continues omitting host material contaminations.
+After the incubation, the undisrupted host tissue is removed and the isolation procedure continues, omitting host material contaminations.
 To our knowledge the only procedure describing a direct and selective epiphytic DNA isolation from the surfaces of marine macrophytes was provided by @Burke2009.
@@
 
-To assess the removal efficiency of DNA and protein isolation procedures leaves and thalli were stained with SYBR Green I and examined under a confocal microscope before and after treatments were performed.
-To test the isolation protocols on different macrophyte species, DNA and proteins were isolated from the surfaces of two macrophytes, the seagrass *C.
-nodosa* and the macroalga *C.
-cylindracea*.
-*C.
-nodosa* was sampled in a monospecific meadow (Bay of Saline) and *C.
-cylindracea* invaded meadow (Bay of Funtana), while samples of *C.
-cylindracea* originated from a monospecific settlement and a *C.
-cylindracea* invaded *C.
-nodosa* meadow in the Bay of Funtana.
-In addition, the sampling was performed in two contrasting seasons, in December 2017 and in June 2018, to test the impact of seasonal conditions on the removal efficiency.
-Procedures developed for DNA and protein isolation showed an almost complete removal of the surface community of both, *C.
+To assess the removal efficiency of DNA and protein isolation procedures leaves and thalli were examined under a confocal microscope before and after treatments were performed.
+Developed procedures showed an almost complete removal of the surface community of both, *C.
 nodosa* and *C.
 cylindracea*.
-In addition, a similar removal efficiency was observed for communities sampled in December 2017 (\autoref{confocal_december}) and June 2018 (\autoref{confocal_june}).
+In addition, a similar removal efficiency was observed for communities sampled in contrasting seasons, December 2017 (\autoref{confocal_december}) and June 2018 (\autoref{confocal_june}).
 Also, no effect of station, settlement or isolation procedure (DNA or protein) on the removal efficiency was observed (Figs.
@@
 The study of marine macrophyte epiphytic communities using culture independent techniques, such as 16S rRNA analysis, metagenomics and metaproteomics, requires an initial step of biological material isolation.
-Methods that have been developed for selective isolation of epiphytic biological material could be divided into two groups: (i) procedures involving, prior to extraction, a cell detachment step such as shaking [@Gross2003; @Noges2010], scraping [@Uku2007] and ultrasonication [@Weidner1996; @Cai2014] and (ii) procedures involving a host tissue incubation aiming at direct lysis of epiphytic microbial cells [@Burke2009].
+Methods that have been developed for selective isolation of epiphytic biological material can be divided into two groups: (i) procedures involving, prior to extraction, a cell detachment step such as shaking [@Gross2003; @Noges2010], scraping [@Uku2007] and ultrasonication [@Weidner1996; @Cai2014] and (ii) procedures involving a host tissue incubation aiming at direct lysis of epiphytic microbial cells [@Burke2009].
 Protocols that include a cell detachment procedure do not provide complete cell removal [@Richter-Heitmann2016; @Cai2014] and in the case of probe ultrasonication tissue disruption can also be observed [@Richter-Heitmann2016].
-To avoid these problems a procedure for DNA isolation involving host tissue incubation in the extraction buffer and direct epiphytic cell lysis on the macrophyte surface was developed [@Burke2009].
-Though providing an almost complete cell removal and selective DNA isolation this protocol uses in the extraction buffer a rapid multienzyme cleaner (3M) not available worldwide and without a known composition [@Burke2009].
+Though, host tissue incubation procedures provide an almost complete cell removal and selective isolation, existing protocols like the one developed for DNA isolation by @Burke2009 use in the extraction buffer a rapid multienzyme cleaner (3M) not available worldwide and without a known composition [@Burke2009].
 To circumvent these problems we developed and tested two protocols for selective DNA and protein isolation from marine macrophyte epiphytic communities.
@@
 
-Metagenomic analysis reveals the potential metabolic capacity of microbial communities, while to obtain an insight into the metabolic status of uncultivated members, a metaproteomic approach is required [@Saito2019].
-The applied protocol for epipyhtic protein isolation followed by a metaproteomic analysis identified between `r format(min(metap_cog_num), big.mark=',')` and `r format(max(metap_cog_num), big.mark=',')` proteins that is higher than previously reported for e.g.
+To obtain an insight into the metabolic status of uncultivated members, a metaproteomic approach is required [@Saito2019].
+The applied protocol for epipyhtic protein isolation followed by a metaproteomic analysis identified between `r format(min(metap_cog_num), big.mark=',')` and `r format(max(metap_cog_num), big.mark=',')` proteins, which is higher than previously reported for e.g.
 soils [@Chourey2010; @Hultman2015], seawater [@Williams2012] and biofilms [@Leary2014].
@@
 cylindracea* were sampled in the same *C.
-nodosa* invaded meadow in the Bay Funtana and on a locality of only *C.
+nodosa* invaded meadow in the Bay of Funtana and on a locality of only *C.
 cylindracea* located in the proximity of the invaded meadow.

### 2020-09-14 submission/manuscript.tex
Coauthor review.

@@
 Studies of unculturable microbes often combine methods based on DNA,
-such as 16S rRNA sequencing and metagenomics, with methods that allow an
+such as 16S rRNA sequencing and metagenomics, with methods that allow
 insight into the metabolic status, such as metaproteomics. To apply
-these procedures to the microbial community inhabiting the surfaces of
+these techniques to the microbial community inhabiting the surfaces of
 marine macrophytes it is advisable to perform, prior to the analysis, a
@@
 suitable for metagenomic and 16S rRNA sequencing, while isolated
-proteins could be identified by mass spectrometry showing that protocols
-can be used in 16S rRNA, metagenomic and metaproteomic analysis. Low
-presence of host DNA and proteins, observed in isolated samples,
-indicated a selective nature of protocols. Furthermore, the procedures
-are based on universally available laboratory chemicals making the
-protocols widely applicable. Taken together, the adapted protocols
-ensure an almost complete removal of the macrophyte epiphytic community,
-are selective for microbes inhabiting macrophyte surfaces and are
-providing DNA and proteins applicable in 16S rRNA sequencing,
+proteins could be identified by mass spectrometry, showing that
+protocols can be used in 16S rRNA, metagenomic and metaproteomic
+analysis. Low presence of host DNA and proteins, observed in isolated
+samples, indicated a selective nature of the protocols. Furthermore, the
+procedures are based on universally available laboratory chemicals
+making the protocols widely applicable. Taken together, the adapted
+protocols ensure an almost complete removal of the macrophyte epiphytic
+community, are selective for microbes inhabiting macrophyte surfaces and
+are providing DNA and proteins applicable in 16S rRNA sequencing,
 metagenomics and metaproteomics.
@@
 the incubation, the undisrupted host tissue is removed and the isolation
-procedure continues omitting host material contaminations. To our
+procedure continues, omitting host material contaminations. To our
 knowledge the only procedure describing a direct and selective epiphytic
@@
 To assess the removal efficiency of DNA and protein isolation procedures
-leaves and thalli were stained with SYBR Green I and examined under a
-confocal microscope before and after treatments were performed. To test
-the isolation protocols on different macrophyte species, DNA and
-proteins were isolated from the surfaces of two macrophytes, the
-seagrass \emph{C. nodosa} and the macroalga \emph{C. cylindracea}.
-\emph{C. nodosa} was sampled in a monospecific meadow (Bay of Saline)
-and \emph{C. cylindracea} invaded meadow (Bay of Funtana), while samples
-of \emph{C. cylindracea} originated from a monospecific settlement and a
-\emph{C. cylindracea} invaded \emph{C. nodosa} meadow in the Bay of
-Funtana. In addition, the sampling was performed in two contrasting
-seasons, in December 2017 and in June 2018, to test the impact of
-seasonal conditions on the removal efficiency. Procedures developed for
-DNA and protein isolation showed an almost complete removal of the
-surface community of both, \emph{C. nodosa} and \emph{C. cylindracea}.
-In addition, a similar removal efficiency was observed for communities
-sampled in December 2017 (\autoref{confocal_december}) and June 2018
-(\autoref{confocal_june}). Also, no effect of station, settlement or
-isolation procedure (DNA or protein) on the removal efficiency was
-observed (Figs. \ref{confocal_december} and \ref{confocal_june}).
+leaves and thalli were examined under a confocal microscope before and
+after treatments were performed. Developed procedures showed an almost
+complete removal of the surface community of both, \emph{C. nodosa} and
+\emph{C. cylindracea}. In addition, a similar removal efficiency was
+observed for communities sampled in contrasting seasons, December 2017
+(\autoref{confocal_december}) and June 2018 (\autoref{confocal_june}).
+Also, no effect of station, settlement or isolation procedure (DNA or
+protein) on the removal efficiency was observed (Figs.
+\ref{confocal_december} and \ref{confocal_june}).
 
@@
 isolation. Methods that have been developed for selective isolation of
-epiphytic biological material could be divided into two groups: (i)
+epiphytic biological material can be divided into two groups: (i)
 procedures involving, prior to extraction, a cell detachment step such
@@
 probe ultrasonication tissue disruption can also be observed
-(Richter-Heitmann \emph{et al.}, 2016). To avoid these problems a
-procedure for DNA isolation involving host tissue incubation in the
-extraction buffer and direct epiphytic cell lysis on the macrophyte
-surface was developed (Burke \emph{et al.}, 2009). Though providing an
-almost complete cell removal and selective DNA isolation this protocol
-uses in the extraction buffer a rapid multienzyme cleaner (3M) not
-available worldwide and without a known composition (Burke \emph{et
-al.}, 2009). To circumvent these problems we developed and tested two
-protocols for selective DNA and protein isolation from marine macrophyte
-epiphytic communities.
+(Richter-Heitmann \emph{et al.}, 2016). Though, host tissue incubation
+procedures provide an almost complete cell removal and selective
+isolation, existing protocols like the one developed for DNA isolation
+by Burke \emph{et al.} (2009) use in the extraction buffer a rapid
+multienzyme cleaner (3M) not available worldwide and without a known
+composition (Burke \emph{et al.}, 2009). To circumvent these problems we
+developed and tested two protocols for selective DNA and protein
+isolation from marine macrophyte epiphytic communities.
 
@@
 
-Metagenomic analysis reveals the potential metabolic capacity of
-microbial communities, while to obtain an insight into the metabolic
-status of uncultivated members, a metaproteomic approach is required
-(Saito \emph{et al.}, 2019). The applied protocol for epipyhtic protein
-isolation followed by a metaproteomic analysis identified between 14,219
-and 16,449 proteins that is higher than previously reported for
-e.g.~soils (Chourey \emph{et al.}, 2010; Hultman \emph{et al.}, 2015),
-seawater (Williams \emph{et al.}, 2012) and biofilms (Leary \emph{et
-al.}, 2014). The functional annotation of identified proteins into COG
-functional categories showed that the protein isolation procedure can be
-used to assess the metabolic status of the macrophyte epiphytic
-community (Leary \emph{et al.}, 2014). Similar to metagenomes, the
-number of identified proteins, including their abundances, associated
-with Streptophyta in \emph{C. nodosa} samples were low indicating that
-the procedure is selective for epiphytic cell proteins. In addition, a
-higher number of identified proteins, and especially their abundances,
-associated with Chlorophyta was observed in \emph{C. cylindracea}
-samples. The cause of this elevated presence of Chlorophyta associated
-proteins can be, as in the case of the DNA isolation procedure,
-explained by the absence of individual cells in this siphonous alga
-(Coneva and Chitwood, 2015).
+To obtain an insight into the metabolic status of uncultivated members,
+a metaproteomic approach is required (Saito \emph{et al.}, 2019). The
+applied protocol for epipyhtic protein isolation followed by a
+metaproteomic analysis identified between 14,219 and 16,449 proteins,
+which is higher than previously reported for e.g.~soils (Chourey
+\emph{et al.}, 2010; Hultman \emph{et al.}, 2015), seawater (Williams
+\emph{et al.}, 2012) and biofilms (Leary \emph{et al.}, 2014). The
+functional annotation of identified proteins into COG functional
+categories showed that the protein isolation procedure can be used to
+assess the metabolic status of the macrophyte epiphytic community (Leary
+\emph{et al.}, 2014). Similar to metagenomes, the number of identified
+proteins, including their abundances, associated with Streptophyta in
+\emph{C. nodosa} samples were low indicating that the procedure is
+selective for epiphytic cell proteins. In addition, a higher number of
+identified proteins, and especially their abundances, associated with
+Chlorophyta was observed in \emph{C. cylindracea} samples. The cause of
+this elevated presence of Chlorophyta associated proteins can be, as in
+the case of the DNA isolation procedure, explained by the absence of
+individual cells in this siphonous alga (Coneva and Chitwood, 2015).
 
@@
 Thalli of \emph{C. cylindracea} were sampled in the same \emph{C.
-nodosa} invaded meadow in the Bay Funtana and on a locality of only
+nodosa} invaded meadow in the Bay of Funtana and on a locality of only
 \emph{C. cylindracea} located in the proximity of the invaded meadow.
@@
 \leavevmode\hypertarget{ref-Bengtsson2020}{}%
-Bengtsson, H. (2020) MatrixStats: Functions that apply to rows and
+Bengtsson, H. (2020) matrixStats: Functions that apply to rows and
 columns of matrices (and to vectors).
@@
 Cúcio, C., Overmars, L., Engelen, A.H., and Muyzer, G. (2018)
-Metagenomic Analysis Shows the Presence of Bacteria Related to
-Free-Living Forms of Sulfur-Oxidizing Chemolithoautotrophic Symbionts in
-the Rhizosphere of the Seagrass Zostera marina. \emph{Front Mar Sci}
-\textbf{5}:
+Metagenomic analysis shows the presence of bacteria related to
+free-living forms of sulfur-oxidizing chemolithoautotrophic symbionts in
+the rhizosphere of the seagrass \emph{Zostera marina}. \emph{Front Mar
+Sci} \textbf{5}: 171.
 
@@
 Saito, M.A., Bertrand, E.M., Duffy, M.E., Gaylord, D.A., Held, N.A.,
-Hervey, W.J., et al. (2019) Progress and Challenges in Ocean
-Metaproteomics and Proposed Best Practices for Data Sharing. \emph{J
+Hervey, W.J., et al. (2019) Progress and challenges in ocean
+metaproteomics and proposed best practices for data sharing. \emph{J
 Proteome Res} \textbf{18}: 1461--1476.

### 2020-09-14 submission/supplementary.Rmd
Coauthor review.

@@
 
-```{r out.width="80%", fig.align="center", fig.cap="Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes sampled in the Bay of Saline (\\textit{C.
-nodosa} [Monospecific Settlement]) and the Bay of Funatana (\\textit{C.
-nodosa} [Mixed Settlement] and \\textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) in two contrasting seasons.\\label{rarefaction}", fig.pos="H"}
+```{r out.width="80%", fig.align="center", fig.cap="Rarefaction curves of bacterial and archaeal communities from the surfaces of two marine macrophytes (\\textit{C.
+nodosa} and \\textit{C.
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{rarefaction}", fig.pos="H"}
 include_graphics("../results/figures/rarefaction.jpg")

### 2020-09-14 submission/supplementary.tex
Coauthor review.

@@
 
-\caption{Rarefaction curves of bacterial and archaeal communities from the surfaces of macrophytes sampled in the Bay of Saline (\textit{C.
-nodosa} [Monospecific Settlement]) and the Bay of Funatana (\textit{C.
-nodosa} [Mixed Settlement] and \textit{C.
-cylindracea} [Mixed and Monospecific Settlement]) in two contrasting seasons.\label{rarefaction}}\label{fig:unnamed-chunk-1}
+\caption{Rarefaction curves of bacterial and archaeal communities from the surfaces of two marine macrophytes (\textit{C.
+nodosa} and \textit{C.
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) in two contrasting seasons (4 December 2017 and 19 June 2018).\label{rarefaction}}\label{fig:unnamed-chunk-1}
 \end{figure}

### 2020-09-15 submission/manuscript.Rmd
Coauthor review #2.

@@
 In contrast to previously described methods this protocol enables an almost complete removal of the surface community and it was used for 16S rRNA gene clone library construction [@Burke2011a] and metagenome sequencing [@Burke2011].
-This method, thought providing a selective isolation procedure, uses a rapid multienzyme cleaner (3M) that is not available worldwide and without a known composition [@Burke2009].
+This method, though providing a selective isolation procedure, uses a rapid multienzyme cleaner (3M) that is not available worldwide and without a known composition [@Burke2009].
 Also to our knowledge, no selective isolation protocol for proteins from epiphytic communities inhabiting marine macrophytes was established.
@@
 For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Sequences classified as chloroplasts by SILVA were exported and reclassifed using mothur and the RDP (http://rdp.cme.msu.edu/) training set (version 16) reference files adapted for mothur [@Cole2014].
+Sequences classified as chloroplasts by SILVA were exported and reclassified using mothur and the RDP (http://rdp.cme.msu.edu/) training set (version 16) reference files adapted for mothur [@Cole2014].
 In comparison to SILVA, RDP allows a more detailed classification of chloroplast sequences.
-Pipeline data processing and visualization was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020; @Wilke2018].
-The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020).
 Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
@@
 
+### Data processing and visualization
+Processing and visualization  of 16S rRNA, metagenomic and metaproteomic data was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020; @Wilke2018].
+The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020).
+
 ### Confocal microscopy
@@
 **\autoref{chloroplast}.** \nameref{chloroplast}
+
+**\autoref{cog}.** \nameref{cog}
 

### 2020-09-15 submission/manuscript.tex
Coauthor review #2.

@@
 (Burke and Thomas \emph{et al.}, 2011) and metagenome sequencing (Burke
-and Peter Steinberg \emph{et al.}, 2011). This method, thought providing
+and Peter Steinberg \emph{et al.}, 2011). This method, though providing
 a selective isolation procedure, uses a rapid multienzyme cleaner (3M)
@@
 Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
-SILVA were exported and reclassifed using mothur and the RDP
+SILVA were exported and reclassified using mothur and the RDP
 (\url{http://rdp.cme.msu.edu/}) training set (version 16) reference
@@
 SILVA, RDP allows a more detailed classification of chloroplast
-sequences. Pipeline data processing and visualization was done using R
-(version 3.6.0) (R Core Team, 2019), package tidyverse (version 1.3.0)
-(Wickham \emph{et al.}, 2019) and multiple other packages (Xie, 2014,
-2015, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018; Wilke, 2018; Y.
-Xie, 2019b, 2019a; Allaire \emph{et al.}, 2019; Zhu, 2019; Bengtsson,
-2020). The detailed analysis procedure including the R Markdown file for
-this paper are available as a GitHub repository
-(\url{https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).
-Based on the ATCC MSA-1002 mock community included in the analysis a
-sequencing error rate of 0.009 \si{\percent} was determined, which is in
-line with previously reported values for next-generation sequencing data
-(Kozich \emph{et al.}, 2013; Schloss \emph{et al.}, 2016). In addition,
-the negative control processed together with the samples yielded only 2
-sequences after sequence quality curation.
+sequences. Based on the ATCC MSA-1002 mock community included in the
+analysis a sequencing error rate of 0.009 \si{\percent} was determined,
+which is in line with previously reported values for next-generation
+sequencing data (Kozich \emph{et al.}, 2013; Schloss \emph{et al.},
+2016). In addition, the negative control processed together with the
+samples yielded only 2 sequences after sequence quality curation.
 
@@
 
+\hypertarget{data-processing-and-visualization}{%
+\subsubsection{Data processing and
+visualization}\label{data-processing-and-visualization}}
+
+Processing and visualization of 16S rRNA, metagenomic and metaproteomic
+data was done using R (version 3.6.0) (R Core Team, 2019), package
+tidyverse (version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple
+other packages (Xie, 2014, 2015, 2019, 2020; Neuwirth, 2014; Xie
+\emph{et al.}, 2018; Wilke, 2018; Allaire \emph{et al.}, 2019; Zhu,
+2019; Bengtsson, 2020). The detailed analysis procedure including the R
+Markdown file for this paper are available as a GitHub repository
+(\url{https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).
+
 \hypertarget{confocal-microscopy}{%
@@
 
-\leavevmode\hypertarget{ref-Xie2019}{}%
-Xie, Y. (2019a) knitr: A general-purpose package for dynamic report
-generation in R.
-
 \leavevmode\hypertarget{ref-Xie2019a}{}%
-Xie, Y. (2019b) TinyTeX: A lightweight, cross-platform, and
+Xie, Y. (2019) TinyTeX: A lightweight, cross-platform, and
 easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
@@
 \textbf{\autoref{chloroplast}.} \nameref{chloroplast}
+
+\textbf{\autoref{cog}.} \nameref{cog}
 

### 2020-09-15 submission/supplementary.Rmd
Coauthor review #2.

@@
   row_spec(0, bold=TRUE) %>%
-  group_rows("Funtana, \\\\textit{Cymodocea nodosa} (Monospecific), 14 December 2017 (sample ID: 45)", 1, 10, escape=FALSE, latex_gap_space="0.6 em") %>%
+  group_rows("Funtana, \\\\textit{Cymodocea nodosa} (Mixed), 14 December 2017 (sample ID: 45)", 1, 10, escape=FALSE, latex_gap_space="0.6 em") %>%
   group_rows("Funtana, \\\\textit{Caulerpa cylindracea} (Monospecific), 14 December 2017 (sample ID: 47)", 11, 20, escape=FALSE, latex_gap_space="0.6 em") %>%

### 2020-09-15 submission/manuscript.Rmd
Added metagenomes ENA accession.

@@
 Sequence renaming, coverage information computing and metagenomic statistics calculations were performed using software tools from BBTools (https://jgi.doe.gov/data-and-tools/bbtools).
-Metagenomic sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers **TO BE ADDED**.
+Metagenomic sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers SAMEA6648795, SAMEA6648797, SAMEA6648809 and SAMEA6648811.
 

### 2020-09-15 submission/manuscript.tex
Added metagenomes ENA accession.

@@
 sequences obtained in this study have been deposited in the European
-Nucleotide Archive (ENA) at EMBL-EBI under accession numbers \textbf{TO
-BE ADDED}.
+Nucleotide Archive (ENA) at EMBL-EBI under accession numbers
+SAMEA6648795, SAMEA6648797, SAMEA6648809 and SAMEA6648811.
 

### 2020-09-16 submission/manuscript.Rmd
Pipeline optimization.

@@
 Processing and visualization  of 16S rRNA, metagenomic and metaproteomic data was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020; @Wilke2018].
-The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020).
+The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/MicrobesRovinj/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020).
 

### 2020-09-16 submission/manuscript.tex
Pipeline optimization.

@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[
-  12pt,
-]{article}
+\documentclass[12pt,]{article}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
   \fi
-% Use upquote if available, for straight quotes in verbatim environments
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
+\usepackage[margin=1.0in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
             pdftitle={Selective DNA and Protein Isolation from Marine Macrophyte Surfaces},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
-\usepackage[margin=1.0in]{geometry}
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\providecommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+
+  \title{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
+Surfaces}}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+    \author{}
+    \preauthor{}\postauthor{}
+    \date{}
+    \predate{}\postdate{}
+
 \usepackage{times} % Times New Roman font
@@
 \renewcommand{\figureautorefname}{Fig.}
-
-\title{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
-Surfaces}}
-\author{}
-\date{\vspace{-2.5em}}
 
@@
 community structure an Illumina sequencing of the V4 16S rRNA region was
-performed. Sequencing yielded a total of 336,937 sequences after quality
+performed. Sequencing yielded a total of 336,944 sequences after quality
 curation and exclusion of eukaryotic, mitochondrial and no relative
-sequences. The number of sequences classified as chloroplast was 97,328.
+sequences. The number of sequences classified as chloroplast was 97,334.
 After excluding these sequences the total number of retrieved reads
-dropped to 239,609, ranging from 22,587 to 52,958 sequences per sample
+dropped to 239,610, ranging from 22,596 to 52,930 sequences per sample
 (\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
@@
 (\autoref{supp-rarefaction}). Sequence clustering at a similarity level
-of 97 \si{\percent} yielded a total of 8,360 different OTUs. Taxonomic
+of 97 \si{\percent} yielded a total of 8,355 different OTUs. Taxonomic
 classification of reads allowed for a macrophyte associated epiphytic
@@
 \emph{Gammaproteobacteria} (11.6 ± 4.3 \si{\percent}),
-\emph{Desulfobacterota} (7.8 ± 7.4 \si{\percent}), \emph{Cyanobacteria}
+\emph{Desulfobacterota} (7.8 ± 7.5 \si{\percent}), \emph{Cyanobacteria}
 (6.5 ± 4.7 \si{\percent}) and \emph{Planctomycetota} (2.9 ± 1.7
@@
 Markdown file for this paper are available as a GitHub repository
-(\url{https://github.com/mkorlevic/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).
+(\url{https://github.com/MicrobesRovinj/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).
 

### 2020-09-16 submission/supplementary.tex
Coauthor review #2.
Pipeline optimization.

@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
   \fi
-% Use upquote if available, for straight quotes in verbatim environments
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
+\usepackage[margin=1.0in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
             pdftitle={Supplementary Information},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
-\usepackage[margin=1.0in]{geometry}
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
 \usepackage{graphicx,grffile}
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\providecommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+
+  \title{\textbf{Supplementary Information}}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+  \subtitle{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
+Surfaces}}
+  \author{}
+    \preauthor{}\postauthor{}
+    \date{}
+    \predate{}\postdate{}
+
 \usepackage{times} % Times New Roman font
@@
 \usepackage{xcolor}
-
-\title{\textbf{Supplementary Information}}
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
-\subtitle{\textbf{Selective DNA and Protein Isolation from Marine Macrophyte
-Surfaces}}
-\author{}
-\date{\vspace{-2.5em}}
 
@@
 \endlastfoot
-40 & Saline & \textit{Cymodocea nodosa} (Monospecific) & 4 December 2017 & 19,176 & 942\\
-41 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 4 December 2017 & 24,250 & 1,123\\
-42 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 4 December 2017 & 28,387 & 3,131\\
-43 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 4 December 2017 & 13,661 & 1,809\\
-61 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 19 June 2018 & 31,734 & 2,039\\
-62 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 19 June 2018 & 36,488 & 3,211\\
-63 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 19 June 2018 & 41,871 & 3,851\\*
+40 & Saline & \textit{Cymodocea nodosa} (Monospecific) & 4 December 2017 & 19,176 & 935\\
+41 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 4 December 2017 & 24,252 & 1,129\\
+42 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 4 December 2017 & 28,385 & 3,125\\
+43 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 4 December 2017 & 13,667 & 1,816\\
+61 & Funtana & \textit{Cymodocea nodosa} (Mixed) & 19 June 2018 & 31,736 & 2,050\\
+62 & Funtana & \textit{Caulerpa cylindracea} (Mixed) & 19 June 2018 & 36,499 & 3,216\\
+63 & Funtana & \textit{Caulerpa cylindracea} (Monospecific) & 19 June 2018 & 41,842 & 3,848\\*
 \end{longtable}
@@
 \addlinespace[0.6 em]
-\multicolumn{5}{l}{\textbf{Funtana, \textit{Cymodocea nodosa} (Monospecific), 14 December 2017 (sample ID: 45)}}\\
+\multicolumn{5}{l}{\textbf{Funtana, \textit{Cymodocea nodosa} (Mixed), 14 December 2017 (sample ID: 45)}}\\
 \hspace{1em}\textit{Proteobacteria} & 4,021,486 & 57.08 & 526,874.683 & 56.71\\

### 2020-11-16 submission/manuscript.Rmd
Coauthors review done.

@@
 \vspace{20mm}
-Marino Korlević^1$*$^, Marsej Markovski^1^, Zihao Zhao^2^, Gerhard J. Herndl^2,3^, Mirjana Najdek^1^
+Marino Korlević^1$*$^, Marsej Markovski^1^, Zihao Zhao^2^, Gerhard J. Herndl^2,3,4^, Mirjana Najdek^1^
 
@@
 3\. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands
+
+4\. Vienna Metabolomics Center, University of Vienna, Austria
 
@@
 \newpage
-\linenumbers
+[]: # (\linenumbers)
 \sisetup{mode=text}
@@
 ## Summary
-Studies of unculturable microbes often combine methods based on DNA, such as 16S rRNA sequencing and metagenomics, with methods that allow insight into the metabolic status, such as metaproteomics.
-To apply these techniques to the microbial community inhabiting the surfaces of marine macrophytes it is advisable to perform, prior to the analysis, a selective DNA and protein isolation so that the host material, present in higher quantities, is not hampering the analysis.
-Two protocols, for DNA and protein isolation, were adapted for selective extractions of DNA and proteins from epiphytic communities inhabiting the surfaces of two marine macrophytes: the seagrass *Cymodocea nodosa* and the macroalga *Caulerpa cylindracea*.
+Studies of unculturable microbes often combine methods based on DNA, such as 16S rRNA sequencing and metagenomics, with methods that allow insights into the metabolic status, such as metaproteomics.
+To apply these techniques to the microbial community inhabiting the surfaces of marine macrophytes it is advisable to perform a selective DNA and protein isolation prior to the analysis to avoid biases due to the host material being present in high quantities.
+Two protocols for DNA and protein isolation were adapted for selective extractions of DNA and proteins from epiphytic communities inhabiting the surfaces of two marine macrophytes, the seagrass *Cymodocea nodosa* and the macroalga *Caulerpa cylindracea*.
 Protocols showed an almost complete removal of the epiphytic community regardless of the sampling season, station, settlement or host species.
-The obtained DNA was suitable for metagenomic and 16S rRNA sequencing, while isolated proteins could be identified by mass spectrometry, showing that protocols can be used in 16S rRNA, metagenomic and metaproteomic analysis.
-Low presence of host DNA and proteins, observed in isolated samples, indicated a selective nature of the protocols.
-Furthermore, the procedures are based on universally available laboratory chemicals making the protocols widely applicable.
-Taken together, the adapted protocols ensure an almost complete removal of the macrophyte epiphytic community, are selective for microbes inhabiting macrophyte surfaces and are providing DNA and proteins applicable in 16S rRNA sequencing, metagenomics and metaproteomics.
+The obtained DNA was suitable for metagenomic and 16S rRNA sequencing, while isolated proteins could be identified by mass spectrometry.
+The low presence of host DNA and proteins in the samples indicated a high specificity of the protocols.
+The procedures are based on universally available laboratory chemicals making the protocols widely applicable.
+Taken together, the adapted protocols ensure an almost complete removal of the macrophyte epiphytic community.
+The procedures are selective for microbes inhabiting macrophyte surfaces and provide DNA and proteins applicable in 16S rRNA sequencing, metagenomics and metaproteomics.
 
@@
 ## Introduction
-Surfaces of marine macrophytes are inhabited by a diverse microbial community whose structure and function are poorly understood [@Egan2013].
-As less than 1 % of all prokaryotic species are culturable, to study these organisms, molecular methods such as 16S rRNA sequencing, metagenomics and metaproteomics are indispensable [@Amann1995; @Su2012].
-Applying these techniques requires an initial isolation step, with the purpose of obtaining high quality DNA and proteins.
+Surfaces of marine macrophytes are colonized by a diverse microbial community whose structure and function are poorly understood [@Egan2013].
+As less than 1 % of all prokaryotic species are culturable, molecular methods such as 16S rRNA sequencing, metagenomics and metaproteomics are indispensable to study these organisms [@Amann1995; @Su2012].
+Applying these techniques requires an initial isolation step with the purpose of obtaining high quality DNA and proteins.
 
-Biological material (i.e.
-proteins and DNA) from pelagic microbial communities is usually isolated by collecting cells onto filters and subsequently isolating the aimed compound [@Gilbert2009].
+Biological material (i.e., proteins and DNA) from pelagic microbial communities is usually isolated by collecting cells onto filters and subsequently isolating the target organisms or communities [@Gilbert2009].
 If a specific microbial size fraction is aimed sequential filtration is applied [@Massana1997; @Andersson2010].
-In contrast, obtaining biological materials from microorganism inhabiting surfaces requires either a cell detachment procedure prior to isolation or the host material is coextracted together with the targeted material.
-Methods for separating microbial cells form the host include shaking of host tissue [@Gross2003; @Noges2010], scraping of macrophyte surfaces [@Uku2007] or the application of ultrasonication [@Weidner1996; @Cai2014].
-It was shown that shaking alone is not sufficient to remove microbial cells, at least from plant root surfaces [@Richter-Heitmann2016].
-Manual separation methods, such as scraping and brushing, are time consuming and subjective, as the detachment efficiency depends on host tissue and the person performing the procedure [@Cai2014].
+In contrast, obtaining microorganisms associated to surfaces require either a cell detachment procedure prior to isolation or the host material is co-extracted with the target material.
+Methods for separating microbial cells from the host include shaking of host tissue [@Gross2003; @Noges2010], scraping of macrophyte surfaces [@Uku2007] or applying ultrasonication [@Weidner1996; @Cai2014].
+It was shown that shaking alone is not sufficient to remove microbial cells from surfaces, at least not from plant root surfaces [@Richter-Heitmann2016].
+Manual separation methods, such as scraping and brushing are time consuming and subjective, as the detachment efficiency depends on host tissue and the person performing the procedure [@Cai2014].
 Ultrasonication was proposed as an alternative method as it is providing better results in terms of detachment efficiency [@Cai2014; @Richter-Heitmann2016].
-The downside of this procedure is that complete cell removal was still not obtained and tissue disruption was observed especially after the application of probe ultrasonication [@Richter-Heitmann2016].
-An alternative to these cell detachment procedures is the isolation of targeted epiphytic compounds together with host materials [@Staufenberger2008; @Jiang2015].
+The downside of this procedure is that complete cell removal is still not obtained and tissue disruption was observed especially after the application of probe ultrasonication [@Richter-Heitmann2016].
+An alternative to these cell detachment procedures is the isolation of target epiphytic compounds together with host material [@Staufenberger2008; @Jiang2015].
 This procedure can lead to problems in the following processing steps such as mitochondrial and chloroplast 16S rRNA sequence contaminations from the host [@Longford2007; @Staufenberger2008].
@@
 
-An alternative to these procedures is a direct isolation of the targeted material by incubating macrophyte tissues in an extraction buffer.
-After the incubation, the undisrupted host tissue is removed and the isolation procedure continues, omitting host material contaminations.
-To our knowledge the only procedure describing a direct and selective epiphytic DNA isolation from the surfaces of marine macrophytes was provided by @Burke2009.
-In contrast to previously described methods this protocol enables an almost complete removal of the surface community and it was used for 16S rRNA gene clone library construction [@Burke2011a] and metagenome sequencing [@Burke2011].
-This method, though providing a selective isolation procedure, uses a rapid multienzyme cleaner (3M) that is not available worldwide and without a known composition [@Burke2009].
-Also to our knowledge, no selective isolation protocol for proteins from epiphytic communities inhabiting marine macrophytes was established.
+An alternative to these procedures is a direct isolation of the target material by incubating macrophyte tissues in an extraction buffer.
+After the incubation, the undisrupted host tissue is removed followed by the isolation procedure, omitting host material contaminations.
+To our knowledge, the only procedure describing a direct and selective epiphytic DNA isolation from the surfaces of marine macrophytes was described by @Burke2009.
+In contrast to previously described methods, this protocol enables an almost complete removal of the surface community.
+It was used for 16S rRNA gene clone library construction [@Burke2011a] and metagenome sequencing [@Burke2011].
+This method, although providing a selective isolation procedure, uses a rapid multi-enzyme cleaner (3M) that is not available worldwide and the chemical constituents are unknown [@Burke2009].
+Also to our knowledge, no selective isolation protocol to perform (meta)proteomics of epiphytic communities associated with marine macrophytes has been developed yet.
 
-In the present study, we adapted a protocol widely used for DNA isolation from filters [@Massana1997] and a protocol used for protein isolation from sediments [@Chourey2010; @Hultman2015] for selective extractions of DNA and proteins from epiphytic communities inhabiting the surfaces of two marine macrophytes: the seagrass *Cymodocea nodosa* and the macroalga *Caulerpa cylindracea*.
+In the present study, we adapted a protocol widely used for DNA isolation from filters [@Massana1997] and a protocol used for protein isolation from soils [@Chourey2010; @Hultman2015].
+These two adapted methods allowed for a selective extraction of DNA and proteins from epiphytic communities inhabiting the surfaces of two marine macrophytes, the seagrass *Cymodocea nodosa* and the macroalga *Caulerpa cylindracea*.
 In addition, we tested the removal efficiency of the protocol and the suitability of obtained DNA and proteins for 16S rRNA sequencing, metagenomics and metaproteomics.
@@
 
-To assess the removal efficiency of DNA and protein isolation procedures leaves and thalli were examined under a confocal microscope before and after treatments were performed.
-Developed procedures showed an almost complete removal of the surface community of both, *C.
+To assess the removal efficiency of the DNA and protein isolation procedures, leaves and thalli were examined under a confocal microscope before and after treatments were performed.
+The modified procedures resulted in an almost complete removal of the surface community of both, *C.
 nodosa* and *C.
 cylindracea*.
-In addition, a similar removal efficiency was observed for communities sampled in contrasting seasons, December 2017 (\autoref{confocal_december}) and June 2018 (\autoref{confocal_june}).
+In addition, a similar removal efficiency was observed for communities sampled in contrasting months, December 2017 (\autoref{confocal_december}) and June 2018 (\autoref{confocal_june}).
 Also, no effect of station, settlement or isolation procedure (DNA or protein) on the removal efficiency was observed (Figs.
@@
 
-To evaluate if the obtained DNA is suitable to determine the microbial community structure an Illumina sequencing of the V4 16S rRNA region was performed.
+To evaluate whether the obtained DNA is suitable to determine the composition of the microbial community Illumina sequencing of the V4 region of the 16S rRNA was performed.
 Sequencing yielded a total of `r number_sequences` sequences after quality curation and exclusion of eukaryotic, mitochondrial and no relative sequences.
-The number of sequences classified as chloroplast was `r number_chloroplast_sequences`.
-After excluding these sequences the total number of retrieved reads dropped to `r number_sequences_without_chloroplast`, ranging from `r min_max$min` to `r min_max$max` sequences per sample (\autoref{supp-nseq_notus}).
-Even when the highest sequencing effort was applied the rarefaction curves did not level off which is commonly observed in high-throughput 16S rRNA amplicon sequencing procedures (\autoref{supp-rarefaction}).
-Sequence clustering at a similarity level of 97 \si{\percent} yielded a total of `r no_otus` different OTUs.
-Taxonomic classification of reads allowed for a macrophyte associated epiphytic community determination that was mainly composed of: *Alphaproteobacteria* (`r filter(plot, taxon=="Alphaproteobacteria")$mean` ± `r filter(plot, taxon=="Alphaproteobacteria")$sd` \si{\percent}), *Bacteroidota* (`r filter(plot, taxon=="Bacteroidota")$mean` ± `r filter(plot, taxon=="Bacteroidota")$sd` \si{\percent}), *Gammaproteobacteria* (`r filter(plot, taxon=="Gammaproteobacteria")$mean` ± `r filter(plot, taxon=="Gammaproteobacteria")$sd` \si{\percent}), *Desulfobacterota* (`r filter(plot, taxon=="Desulfobacterota")$mean` ± `r filter(plot, taxon=="Desulfobacterota")$sd` \si{\percent}), *Cyanobacteria* (`r filter(plot, taxon=="Cyanobacteria")$mean` ± `r filter(plot, taxon=="Cyanobacteria")$sd` \si{\percent}) and *Planctomycetota* (`r filter(plot, taxon=="Planctomycetota")$mean` ± `r filter(plot, taxon=="Planctomycetota")$sd` \si{\percent}) (\autoref{community}).
+The number of sequences classified as chloroplasts was `r number_chloroplast_sequences`.
+After excluding these sequences the total number of retrieved reads was `r number_sequences_without_chloroplast`, ranging from `r min_max$min` to `r min_max$max` sequences per sample (\autoref{supp-nseq_notus}).
+Even when the highest sequencing effort was applied, the rarefaction curves did not level off which is commonly observed in high-throughput 16S rRNA amplicon sequencing (\autoref{supp-rarefaction}).
+Sequences clustering at a similarity level of 97 \si{\percent} yielded a total of `r no_otus` different OTUs.
+Taxonomic classification of reads revealed a macrophyte-associated epiphytic community that mainly composed of *Alphaproteobacteria* (`r filter(plot, taxon=="Alphaproteobacteria")$mean` ± `r filter(plot, taxon=="Alphaproteobacteria")$sd` \si{\percent}), *Bacteroidota* (`r filter(plot, taxon=="Bacteroidota")$mean` ± `r filter(plot, taxon=="Bacteroidota")$sd` \si{\percent}), *Gammaproteobacteria* (`r filter(plot, taxon=="Gammaproteobacteria")$mean` ± `r filter(plot, taxon=="Gammaproteobacteria")$sd` \si{\percent}), *Desulfobacterota* (`r filter(plot, taxon=="Desulfobacterota")$mean` ± `r filter(plot, taxon=="Desulfobacterota")$sd` \si{\percent}), *Cyanobacteria* (`r filter(plot, taxon=="Cyanobacteria")$mean` ± `r filter(plot, taxon=="Cyanobacteria")$sd` \si{\percent}) and *Planctomycetota* (`r filter(plot, taxon=="Planctomycetota")$mean` ± `r filter(plot, taxon=="Planctomycetota")$sd` \si{\percent}) (\autoref{community}).
 
@@
 We observed a high proportion of chloroplast sequences in all analyzed samples (`r filter(plot, taxon=="Chloroplast")$mean` ± `r filter(plot, taxon=="Chloroplast")$sd` \si{\percent}) (\autoref{community}).
-To determine if chloroplast sequences originate from hosts or eukaryotic epiphytic organisms, we exported SILVA classified chloroplast sequences and reclassified them using the RDP (Ribosomal Database Project) training set that allows for a more detailed chloroplast classification.
-The largest proportion of sequences were classified as Bacillariophyta (`r filter(chloroplast, taxon=="Bacillariophyta")$mean` ± `r filter(chloroplast, taxon=="Bacillariophyta")$sd` \si{\percent}) indicating that the DNA removal procedure did not coextract larger quantities of host DNA (\autoref{chloroplast}).
+To determine whether chloroplast sequences originate from the host or eukaryotic epiphytic organisms, we exported SILVA-classified chloroplast sequences and reclassified them using the RDP (Ribosomal Database Project) training set that allows for a more detailed chloroplast classification.
+The largest proportion of sequences was classified as Bacillariophyta (`r filter(chloroplast, taxon=="Bacillariophyta")$mean` ± `r filter(chloroplast, taxon=="Bacillariophyta")$sd` \si{\percent}) indicating that the DNA removal procedure resulted in only minor co-extracted quantities of host DNA (\autoref{chloroplast}).
 Chloroplast sequences classified as Streptophyta constituted `r streptophyta_16S$mean` ± `r streptophyta_16S$sd` \si{\percent} of all chloroplast sequences originating from *C.
@@
 
-To determine if the extracted DNA can be used for metagenomic sequencing four samples containing epiphytic DNA were selected and shotgun sequenced using an Illumina platform.
-Metagenomic sequencing yielded from `r format(min(metadata_metagenomic_statistics$raw_R1), big.mark=',')` to `r format(max(metadata_metagenomic_statistics$raw_R1), big.mark=',')` sequence pairs (\autoref{supp-metagenomic_statistics}).
+To determine whether the extracted DNA can be used for metagenomic sequencing, four samples containing epiphytic DNA were selected and shotgun sequenced using an Illumina platform.
+Metagenomic sequencing yielded between `r format(min(metadata_metagenomic_statistics$raw_R1), big.mark=',')` and `r format(max(metadata_metagenomic_statistics$raw_R1), big.mark=',')` sequence pairs (\autoref{supp-metagenomic_statistics}).
 Obtained sequences were successfully assembled into contigs whose L50 ranged from `r format(min(metadata_metagenomic_statistics$L50), big.mark=',')` to `r format(max(metadata_metagenomic_statistics$L50), big.mark=',')` bp.
-In addition, predicted coding sequences were successfully functionally annotated (`r format(min(metadata_metagenomic_statistics$eggnog), big.mark=',')` -- `r format(max(metadata_metagenomic_statistics$eggnog), big.mark=',')` annotated sequences; \autoref{cog}) and taxonomically classified.
+In addition, predicted coding sequences were functionally annotated (`r format(min(metadata_metagenomic_statistics$eggnog), big.mark=',')` -- `r format(max(metadata_metagenomic_statistics$eggnog), big.mark=',')` annotated sequences; \autoref{cog}) and taxonomically classified.
 Functional annotation allowed for an assessment of the relative contribution of each COG (Clusters of Orthologous Groups) functional category to the total number of annotated coding sequences (\autoref{cog}a).
-Functional categories containing the highest number of sequences were: `r cog_relative$"COG Functional Category"[3]` (`r cog_relative$Name[3]`), `r cog_relative$"COG Functional Category"[4]` (`r cog_relative$Name[4]`), `r cog_relative$"COG Functional Category"[5]` (`r cog_relative$Name[5]`), `r cog_relative$"COG Functional Category"[6]` (`r cog_relative$Name[6]`) and `r cog_relative$"COG Functional Category"[7]` (`r cog_relative$Name[7]`).
-If host DNA was coextracted with epiphytic it should be detected in larger proportions in sequenced metagenomes.
-Indeed, no higher proportions of coding sequences classified into phylum Streptophyta or Chlorophyta were detected (\autoref{supp-metagenomic_taxonomy}).
+Functional categories containing the highest number of sequences were `r cog_relative$"COG Functional Category"[3]` (`r cog_relative$Name[3]`), `r cog_relative$"COG Functional Category"[4]` (`r cog_relative$Name[4]`), `r cog_relative$"COG Functional Category"[5]` (`r cog_relative$Name[5]`), `r cog_relative$"COG Functional Category"[6]` (`r cog_relative$Name[6]`) and `r cog_relative$"COG Functional Category"[7]` (`r cog_relative$Name[7]`).
+If host DNA is co-extracted with epiphytes it should be detected in large proportions in sequenced metagenomes.
+However, no large proportions of coding sequences classified as Streptophyta and Chlorophyta were detected (\autoref{supp-metagenomic_taxonomy}).
 Sequenced metagenomic DNA originating from the surface of *C.
-nodosa* contained `r filter(metadata_metagenomic_phylum, label=="14/12-17 FCyM" & phylum=="Streptophyta")$cds_relative` \si{\percent} of coding sequences classified into the phylum Streptophyta in December 2017 and `r filter(metadata_metagenomic_phylum, label=="19/6-18 FCyM" & phylum=="Streptophyta")$cds_relative` \si{\percent} in June 2018.
+nodosa* contained `r filter(metadata_metagenomic_phylum, label=="14/12-17 FCyM" & phylum=="Streptophyta")$cds_relative` \si{\percent} of coding sequences classified as Streptophyta in December 2017 and `r filter(metadata_metagenomic_phylum, label=="19/6-18 FCyM" & phylum=="Streptophyta")$cds_relative` \si{\percent} in June 2018.
 Furthermore, the summed RPKM (Reads Per Kilobase Million) of these sequences constituted `r filter(metadata_metagenomic_phylum, label=="14/12-17 FCyM" & phylum=="Streptophyta")$rpkm_relative` \si{\percent} of total RPKM of all successfully classified sequences in December 2017 and `r filter(metadata_metagenomic_phylum, label=="19/6-18 FCyM" & phylum=="Streptophyta")$rpkm_relative` \si{\percent} in June 2018.
@@
 cylindracea*.
-Of all successfully classified coding sequences `r filter(metadata_metagenomic_phylum, label=="14/12-17 FCa" & phylum=="Chlorophyta")$cds_relative` \si{\percent} were classified into Chlorophyta in December 2017 and `r filter(metadata_metagenomic_phylum, label=="19/6-18 FCa" & phylum=="Chlorophyta")$cds_relative` \si{\percent} in June 2018.
-A relatively higher proportion of these sequences' RPKM than in the case of *C.
+Of all successfully classified coding sequences `r filter(metadata_metagenomic_phylum, label=="14/12-17 FCa" & phylum=="Chlorophyta")$cds_relative` \si{\percent} were classified as Chlorophyta in December 2017 and `r filter(metadata_metagenomic_phylum, label=="19/6-18 FCa" & phylum=="Chlorophyta")$cds_relative` \si{\percent} in June 2018.
+A relatively higher proportion of RPKM of these sequences than in the case of *C.
 nodosa* was observed, indicating a higher coextraction of host DNA in *C.
 cylindracea*.
-In December, the proportion of RPKM of sequences classified into Chlorophyta was `r filter(metadata_metagenomic_phylum, label=="14/12-17 FCa" & phylum=="Chlorophyta")$rpkm_relative` \si{\percent}, while in June 2018 it reached `r filter(metadata_metagenomic_phylum, label=="19/6-18 FCa" & phylum=="Chlorophyta")$rpkm_relative` \si{\percent}.
+In December, the proportion of RPKM of sequences classified as Chlorophyta was `r filter(metadata_metagenomic_phylum, label=="14/12-17 FCa" & phylum=="Chlorophyta")$rpkm_relative` \si{\percent}, while in June 2018 it reached `r filter(metadata_metagenomic_phylum, label=="19/6-18 FCa" & phylum=="Chlorophyta")$rpkm_relative` \si{\percent}.
 
-To evaluate if the procedure for protein extraction is suitable for metaproteomic analysis, obtained proteins were trypsin digested and sequenced using a mass spectrometer.
+To evaluate whether the procedure for protein extraction is suitable for metaproteomic analysis, obtained proteins were trypsin digested and sequenced using a mass spectrometer.
 Obtained MS/MS spectra were searched against a protein database from sequenced metagenomes.
@@
 In addition, successful identification of proteins allowed for an assessment of the relative contribution of each COG functional category to the total number of identified proteins (\autoref{cog}b).
-Functional categories containing the highest number of identified proteins were: `r metap_cog_relative$"COG Functional Category"[1]` (`r metap_cog_relative$Name[1]`), `r metap_cog_relative$"COG Functional Category"[2]` (`r metap_cog_relative$Name[2]`), `r metap_cog_relative$"COG Functional Category"[3]` (`r metap_cog_relative$Name[3]`), `r metap_cog_relative$"COG Functional Category"[4]` (`r metap_cog_relative$Name[4]`) and `r metap_cog_relative$"COG Functional Category"[5]` (`r metap_cog_relative$Name[5]`).
-Isolated proteins could derive from epiphytic organisms inhabiting the macrophyte surface or from macrophyte tissue underlying them.
-The contribution of proteins originating from host tissues was evaluated by identifying all the proteins predicted to belong to any taxonomic group within the phyla Streptophyta and Chlorophyta and by calculating their contribution to the number and abundance of all identified proteins.
+Functional categories containing the highest number of identified proteins were `r metap_cog_relative$"COG Functional Category"[1]` (`r metap_cog_relative$Name[1]`), `r metap_cog_relative$"COG Functional Category"[2]` (`r metap_cog_relative$Name[2]`), `r metap_cog_relative$"COG Functional Category"[3]` (`r metap_cog_relative$Name[3]`), `r metap_cog_relative$"COG Functional Category"[4]` (`r metap_cog_relative$Name[4]`) and `r metap_cog_relative$"COG Functional Category"[5]` (`r metap_cog_relative$Name[5]`).
+Isolated proteins could originate from epiphytic organisms inhabiting the macrophyte surface and/or from macrophyte tissue underlying them.
+The contribution of proteins originating from host tissues was evaluated by identifying all the proteins predicted to belong to any taxonomic group within the phyla Streptophyta and Chlorophyta and by calculating their contribution to the number and abundance (NAAF -- Normalized Abundance Area Factor) of all identified proteins.
 On average, proteins isolated from the surface of *C.
@@
 Chlorophyta associated proteins comprised `r filter(metap_cog_chlorophyta, cog_abundance=="cog" & host=="Ca")$mean`  ± `r filter(metap_cog_chlorophyta, cog_abundance=="cog" & host=="Ca")$sd` \si{\percent} of all identified proteins in *C.
-cylindracea*, contributing to `r filter(metap_cog_chlorophyta, cog_abundance=="abundance" & host=="Ca")$mean`  ± `r filter(metap_cog_chlorophyta, cog_abundance=="abundance" & host=="Ca")$sd` \si{\percent} of all protein abundances.
+cylindracea*, contributing `r filter(metap_cog_chlorophyta, cog_abundance=="abundance" & host=="Ca")$mean`  ± `r filter(metap_cog_chlorophyta, cog_abundance=="abundance" & host=="Ca")$sd` \si{\percent} to the total protein abundance.
 
 ## Discussion
-The study of marine macrophyte epiphytic communities using culture independent techniques, such as 16S rRNA analysis, metagenomics and metaproteomics, requires an initial step of biological material isolation.
-Methods that have been developed for selective isolation of epiphytic biological material can be divided into two groups: (i) procedures involving, prior to extraction, a cell detachment step such as shaking [@Gross2003; @Noges2010], scraping [@Uku2007] and ultrasonication [@Weidner1996; @Cai2014] and (ii) procedures involving a host tissue incubation aiming at direct lysis of epiphytic microbial cells [@Burke2009].
-Protocols that include a cell detachment procedure do not provide complete cell removal [@Richter-Heitmann2016; @Cai2014] and in the case of probe ultrasonication tissue disruption can also be observed [@Richter-Heitmann2016].
-Though, host tissue incubation procedures provide an almost complete cell removal and selective isolation, existing protocols like the one developed for DNA isolation by @Burke2009 use in the extraction buffer a rapid multienzyme cleaner (3M) not available worldwide and without a known composition [@Burke2009].
-To circumvent these problems we developed and tested two protocols for selective DNA and protein isolation from marine macrophyte epiphytic communities.
-
-To test if the developed DNA and protein isolation procedures could be applied on a variety of macrophyte species we selected *C.
+To test whether the developed DNA and protein isolation protocols efficiently detach microbes from the macrophyte surface we selected *C.
 nodosa* and *C.
-cylindracea* as representatives of seagrass and macroalgal species, on which the procedures were tested.
-These species especially differ morphologically.
+cylindracea* as representatives of seagrass and macroalgal species.
+These species differ morphologically.
 While *C.
 nodosa* leaves are flat, *C.
-cylindracea* thallus is composed of uneven surfaces [@Kuo2001; @Verlaque2003].
-The developed procedures showed an almost complete removal of epiphytic cells from the surfaces of both species comparable to the result of @Burke2009 and indicating that structural differences do not impact the removal efficiency.
-In addition, isolation protocols were tested in two contrasting season as it is known that macrophytes are harboring more algal epiphytes during autumn and winter [@Reyes2001].
-No differences in the removal efficiency was observed between seasons suggesting that protocols can be used on macrophyte samples retrieved throughout the year.
-Also, no removal differences were observed on samples derived from the same host but from different localities or settlements demonstrating the stability of the protocol in cell removal efficiency.
+cylindracea* thalli are characterized by an uneven surface [@Kuo2001; @Verlaque2003].
+The developed protocol led to an almost complete removal of epiphytic cells from the surfaces of both species comparable to the result of @Burke2009, indicating that structural differences do not impact the removal efficiency.
+In addition, isolation protocols were tested in two contrasting seasons, as it is known that macrophytes are harboring more algal epiphytes during autumn and winter [@Reyes2001].
+No differences in the removal efficiency was observed between seasons suggesting that these protocols can be used on macrophyte samples retrieved throughout the year.
+Also, no removal differences were observed on samples derived from the same host but from different locations.
 
-Successful amplification and sequencing of the V4 16S rRNA gene region proved that the isolated DNA can be used to estimate the microbial epiphytic diversity.
+Successful amplification and sequencing of the V4 region of the 16S rRNA gene proved that the isolated DNA can be used to estimate the microbial epiphytic diversity.
 Taxonomic groups detected in this step can also be often found in epiphytic communities associated with other macrophytes [@Morrissey2019; @Burke2011a].
@@
 These sequences can derive from host chloroplasts or from eukaryotic epiphytic chloroplast DNA.
-Although the proportion of obtained chloroplast 16S rRNA sequences in our samples was substantial they derived almost exclusively from eukaryotic epiphytes.
+Although the proportion of obtained chloroplast 16S rRNA sequences in our samples was substantial, they derived almost exclusively from eukaryotic epiphytes.
 High proportion of chloroplast 16S rRNA sequences in studies applying selective procedures that include direct cellular lysis on host surfaces were observed before [@Michelou2013].
-It is possible that chloroplast specific sequences even in these studies are originating from eukaryotic epiphytic cells and not host chloroplast.
+It is possible that chloroplast-specific sequences even in these studies originated from eukaryotic epiphytic cells and not from host chloroplasts.
 Indeed, it is common during 16S rRNA profiling of pelagic microbial communities to observe high proportions of chloroplast sequences [@Korlevic2016; @Gilbert2009].
@@
 
-Beside 16S rRNA sequencing high quality DNA is needed for metagenomics.
+High quality DNA is also needed for metagenomics.
 The obtained number of metagenomic sequences and assembly statistics were comparable to metagenomes and metatranscriptomes derived from similar surface associated communities [@Crump2018; @Cucio2018].
@@
 The proportion of coding sequences, including their RPKM, originating from *C.
-nodosa* metagenomes and classified as Streptophyta was low indicating a good selectivity of the isolation procedure towards epiphytic cells.
-In the case of DNA samples isolated from the surface of *C.
-cylindracea* the proportion of Chlorophyta coding sequences was also low but their RPKM was higher than in the case of *C.
+nodosa* metagenomes and classified as Streptophyta was low indicating that the isolation procedure was specific for epiphytic cells.
+DNA samples isolated from the surface of *C.
+cylindracea* exhibited a low proportion of Chlorophyta coding sequences, however, their RPKM was higher than in the samples originating from *C.
 nodosa*.
-One of the causes for this elevated RPKM of Chlorophyta sequences in *C.
-cylindracea* could lay in tissue structure differences between these two host species.
+One of the reasons for this elevated RPKM of Chlorophyta sequences in *C.
+cylindracea* could be the differences in the tissue structure between these two host species.
 While *C.
@@
 The absence of individual cells in *C.
-cylindracea* could cause a leakage of genetic material into the extraction buffer causing an elevated presence of host sequences in metagenomic data.
+cylindracea* could cause a leakage of genetic material into the extraction buffer causing an elevated presence of host sequences in the samples for metagenome analyses.
 
-To obtain an insight into the metabolic status of uncultivated members, a metaproteomic approach is required [@Saito2019].
-The applied protocol for epipyhtic protein isolation followed by a metaproteomic analysis identified between `r format(min(metap_cog_num), big.mark=',')` and `r format(max(metap_cog_num), big.mark=',')` proteins, which is higher than previously reported for e.g.
+To obtain insight into the metabolic status of uncultivated prokaryotes, a metaproteomic approach is required [@Saito2019].
+The applied protocol for epiphytic protein isolation followed by a metaproteomic analysis identified between `r format(min(metap_cog_num), big.mark=',')` and `r format(max(metap_cog_num), big.mark=',')` proteins, which is higher than previously reported for e.g.
 soils [@Chourey2010; @Hultman2015], seawater [@Williams2012] and biofilms [@Leary2014].
-The functional annotation of identified proteins into COG functional categories showed that the protein isolation procedure can be used to assess the metabolic status of the macrophyte epiphytic community [@Leary2014].
-Similar to metagenomes, the number of identified proteins, including their abundances, associated with Streptophyta in *C.
-nodosa* samples were low indicating that the procedure is selective for epiphytic cell proteins.
-In addition, a higher number of identified proteins, and especially their abundances, associated with Chlorophyta was observed in *C.
+The functional annotation of identified proteins into COG functional categories showed that the protein isolation protocol can be used to assess the metabolic status of the epiphytic community [@Leary2014].
+Similar to the results of the metagenomic analysis, the number and abundance of identified proteins affiliated to Streptophyta in *C.
+nodosa* samples were low, indicating that the procedure is selective for epiphytic cell proteins.
+In addition, a higher number and abundance of identified proteins associated with Chlorophyta were observed in *C.
 cylindracea* samples.
-The cause of this elevated presence of Chlorophyta associated proteins can be, as in the case of the DNA isolation procedure, explained by the absence of individual cells in this siphonous alga [@Coneva2015].
+The cause of this elevated presence of Chlorophyta-associated proteins can be, similar to the DNA isolation protocol, explained by the absence of individual cells in this siphonous alga [@Coneva2015].
 
-In conclusion, the developed protocols for DNA and protein isolation from macrophyte surfaces are providing an almost complete removal of the epiphytic community and are shown to ensure removal from both, *C.
+In conclusion, the developed protocols for DNA and protein isolation from macrophyte surfaces almost completely remove the epiphytic community from both, *C.
 nodosa* and *C.
 cylindracea*, in different seasons.
-Also, the obtained DNA and proteins are suitable for 16S rRNA sequencing, metagenomics and metaproteomics, while the obtained material contains low quantities of host DNA or proteins making the procedures epiphyte selective.
-Furthermore, the procedures are based on universally available laboratory chemicals making the protocols widely applicable.
+Also, the obtained DNA and proteins are suitable for 16S rRNA sequencing, metagenomics and metaproteomics analyses while the obtained material contains low quantities of host DNA and proteins making the protocols specific for epiphytes.
+Furthermore, the protocols are based on universally available laboratory chemicals hence, making them widely applicable.
 
@@
 nodosa* were sampled in a *C.
-nodosa* meadow in the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E) and in a *C.
+nodosa* meadow in the Bay of Saline, northern Adriatic Sea (\ang{45;7;5} N, \ang{13;37;20} E) and in a *C.
 nodosa* meadow invaded by *C.
-cylindracea* in the proximity of the village of Funtana (Bay of Funtana; \ang{45;10;39} N, \ang{13;35;42} E).
+cylindracea* in the Bay of Funtana, northern Adriatic Sea (\ang{45;10;39} N, \ang{13;35;42} E).
 Thalli of *C.
 cylindracea* were sampled in the same *C.
-nodosa* invaded meadow in the Bay of Funtana and on a locality of only *C.
+nodosa* invaded meadow in the Bay of Funtana and at a locality of only *C.
 cylindracea* located in the proximity of the invaded meadow.
 Leaves and thalli for 16S rRNA analysis, metagenomics and metaproteomics were collected in two contrasting seasons, on 4 December 2017 (16S rRNA analysis and metaproteomics), 14 December 2017 (metagenomics) and 18 June 2018 (16S rRNA analysis, metagenomics and metaproteomics).
-During spring 2018 the *C.
+During spring 2018, the *C.
 nodosa* meadow in the Bay of Saline decayed to an extent that no leaves could be retrieved [@Najdek2020].
 In addition, as not enough DNA for both metagenomic and 16S RNA analysis were obtained during the sampling on 4 December 2017, an additional sampling on 14 December 2017 was carried out in the Bay of Funtana.
-Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with site seawater.
+Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with seawater from this site.
 Upon arrival to the laboratory, *C.
@@
 This protocol was modified and adapted for microbial DNA isolation from macrophyte surfaces as described below.
-5 \si{\ml} of lysis buffer (40 \si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar} sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or 2 \si{\g} wet-weight of thalli.
-For every sample duplicate isolations were performed.
+Five (5) \si{\ml} of lysis buffer (40 \si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar} sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or 2 \si{\g} wet-weight of thalli.
+For every sample, duplicate isolations were performed.
 Lysozyme was added (final concentration 1 \si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius} for 30 \si{\minute}.
 Subsequently, proteinase K (final concentration 0.5 \si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were added and the samples were incubated at 55 \si{\degreeCelsius} for 2 \si{\hour}.
-Following the incubation, tubes were vortexed for 10 \si{\minute} and the mixture containing lyzed epiphytic cells was separated from host leaves or thalli by transferring the solution into a clean tube.
+Following the incubation, tubes were vortexed for 10 \si{\minute} and the mixture containing lysed epiphytic cells was separated from host leaves or thalli by transferring the solution into a clean tube.
 The lysate was extracted twice with a mixture of phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with chloroform:isoamyl alcohol (24:1).
-After each organic solvent mixture addition tubes were slightly vortexed and centrifuged at 4,500 × g for 10 \si{\minute}.
-Following each centrifugation aqueous phases were retrieved.
+After each addition of an organic solvent mixture, tubes were slightly vortexed and centrifuged at 4,500 × g for 10 \si{\minute}.
+Following each centrifugation the aqueous phases were retrieved.
 After the final extraction 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
-DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} °C and centrifuging at 16,000 × g and 4 \si{\degreeCelsius} for 20 \si{\minute}.
+DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 16,000 × g and 4 \si{\degreeCelsius} for 20 \si{\minute}.
 The pellet was washed twice with 1 \si{\ml} of chilled 70 \si{\percent} ethanol and centrifuged after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 10 \si{\minute}.
-After the first washing step duplicate pellets form the same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.
+After the first washing step duplicate pellets from the same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.
 The dried pellet was resuspended in 100 \si{\ul} of deionized water.
@@
 These primers contained on their 5' end a tagged sequence.
-Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5 Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template.
+Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template.
 Cycling conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}.
@@
 
-Obtained sequences were analyzed on a Life Science Compute Cluster (LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
+Obtained sequences were analyzed on the Life Science Compute Cluster (LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
 Individual sequences were assembled using MEGAHIT (version  1.1.2) [@Li2015] under default settings.
@@
 Sequence taxonomy classification was determined using the lowest common ancestor algorithm adapted from DIAMOND (version 0.8.36) [@Buchfink2015] and by searching against the NCBI non-redundant database (NR).
-To determine phylogeny the top 10 \si{\percent} hits with an e-value < 1 × 10^-5^ were used (\-\-top 10).
+To determine the phylogeny, the top 10 \si{\percent} hits with an e-value < 1 × 10^-5^ were used (\-\-top 10).
 Sequence renaming, coverage information computing and metagenomic statistics calculations were performed using software tools from BBTools (https://jgi.doe.gov/data-and-tools/bbtools).
@@
 This protocol was further modified and adapted for microbial protein isolation from macrophyte surfaces as described below.
-20 \si{\ml} of protein extraction buffer (4 \si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl [pH 8.0]) was added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of thalli.
+Twenty (20) \si{\ml} of protein extraction buffer (4 \si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl [pH 8.0]) was added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of thalli.
 The mixture was incubated in boiling water for 5 \si{\minute}, vortexed for 10 \si{\minute} and incubated again in boiling water for 5 \si{\minute}.
-After a brief vortex the lysate was transferred to a clean tube separating the host leaves or thalli from the mixture containing lyzed epiphytic cells.
+After a brief vortex, the lysate was transferred to a clean tube separating the host leaves or thalli from the mixture containing lysed epiphytic cells.
 Dithiothreitol (DTT; final concentration 24 \si{\milli\Molar}) was added and proteins were precipitated with chilled 100 \si{\percent} trichloroacetic acid (TCA; final concentration 20 \si{\percent}) overnight at \num{-20} °C.
@@
 During the first washing step the pellet was transferred to a clean 1.5 \si{\ml} tube.
-After each washing step samples were centrifuged at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}.
+After each washing step, samples were centrifuged at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}.
 Dried pellets were stored at \num{-80} °C until further analysis.
@@
 Isolated proteins were whole trypsin digested using the FASP (filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK) according to the manufacturer's instructions with small modifications [@Wisniewski2009].
-Before the solution was loaded to the column, protein pellets were solubilized in a urea sample buffer included in the kit amended with DTT (final concentration 100 \si{\milli\Molar}) for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to remove larger particles.
+Prior to loading the solution onto the column, protein pellets were solubilized in a urea sample buffer included in the kit amended with DTT (final concentration 100 \si{\milli\Molar}) for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to remove larger particles.
 The first washing step after protein solution loading was repeated twice.
-In addition, centrifugation steps were prolonged if the column was clogged.
+In addition, the centrifugation steps were prolonged if the column was clogged.
 Trypsin digestion was performed on column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
-The final filtrate containing peptides was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid, freezed at \num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent to the Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis.
-Peptides were resuspended in 1 \si{\percent} (final concentration) trifluoroacetic acid (TFA), desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
+The final filtrate containing peptides was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid (TFA), freezed at \num{-80} \si{\degreeCelsius}, lyophilized and sent to VIME -- Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis.
+Peptides were resuspended in 1 \si{\percent} (final concentration) TFA, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
 Obtained MS/MS spectra were searched against a protein database composed of combined sequenced metagenomes using SEQUEST-HT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
@@
 ### Data processing and visualization
-Processing and visualization  of 16S rRNA, metagenomic and metaproteomic data was done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020; @Wilke2018].
+Processing and visualization  of 16S rRNA, metagenomic and metaproteomic data were done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020; @Wilke2018].
 The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/MicrobesRovinj/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020).
@@
 In addition, nontreated leaves and thalli, washed three times in ASW to remove loosely attached microbial cells, were fixed in the same concentration of formaldehyde and used as a positive control.
-For long therm storage, fixed leaves and thalli were immersed in a mixture of phosphate-buffered saline (PBS) and ethanol (1:1) and stored at \num{-20} \si{\degreeCelsius}.
-Treated and untreated segemnts of leaves and thalli were stained in a 2 × solution of SYBR Green I and examined under a Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
+For-long term storage, fixed leaves and thalli were immersed in a mixture of phosphate-buffered saline (PBS) and ethanol (1:1) and stored at \num{-20} \si{\degreeCelsius}.
+Treated and untreated segments of leaves and thalli were stained in a 2 × solution of SYBR Green I and examined under a Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
 
 ## Acknowledgements
-This work was founded by the Croatian Science Foundation through the MICRO-SEAGRASS project (IP-2016-06-7118).
-We would like to thank Margareta Buterer for technical support, Paolo Paliaga for help during sampling, Lucija Horvat for technical support in confocal microscopy and Dušica Vujaklija for help in accessing the confocal microscope.
+This work was funded by the Croatian Science Foundation through the MICRO-SEAGRASS project (project number IP-2016-06-7118).
+ZZ and GJH were supported by the Austrian Science Fund (FWF) project ARTEMIS (project number P28781-B21).
+We would like to thank Margareta Buterer for technical support, Paolo Paliaga for help during sampling, Lucija Horvat for technical support in confocal microscopy, Dušica Vujaklija for help in accessing the confocal microscope and Sonja Tischler for technical support in mass spectrometry.
 In addition, we would like to thank the University Computing Center of the University of Zagreb for access to the computer cluster Isabella and CUBE -- Computational Systems Biology of the University of Vienna for access to the Life Science Compute Cluster (LiSC).
@@
 nodosa} and \\textit{C.
-cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 4 December 2017 and stained with SYBR Green I.
-Scale bar in all images is 60 \\si{\\um}.\\label{confocal_december}", fig.pos="ht"}
+cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) sampled on 4 December 2017 and stained with SYBR Green I.
+Scale bar at all images is 60 \\si{\\um}.\\label{confocal_december}", fig.pos="ht"}
 include_graphics("../results/images/confocal_december.jpg")
@@
 nodosa} and \\textit{C.
-cylindracea} surfaces from the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 19 June 2018 and stained with SYBR Green I.
-Scale bar in all images is 60 \\si{\\um}.\\label{confocal_june}", fig.pos="ht"}
+cylindracea} surfaces from the Bay of Funtana (mixed and monospecific settlements) sampled on 19 June 2018 and stained with SYBR Green I.
+Scale bar at all images is 60 \\si{\\um}.\\label{confocal_june}", fig.pos="ht"}
 include_graphics("../results/images/confocal_june.jpg")
@@
 \newpage
-```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant (> 1 \\si{\\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\\textit{C.
+```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\\textit{C.
 nodosa} and \\textit{C.
-cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{community}", fig.pos="ht"}
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{community}", fig.pos="ht"}
 include_graphics("../results/figures/community_bar_plot.jpg")
@@
 nodosa} and \\textit{C.
-cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{chloroplast}", fig.pos="ht"}
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{chloroplast}", fig.pos="ht"}
 include_graphics("../results/figures/chloroplast.jpg")
@@
 nodosa} and \\textit{C.
-cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4/14 December 2017 and 19 June 2018).
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4/14 December 2017 and 19 June 2018).
 Total number of annotated coding sequences and identified proteins is given above the corresponding bar.\\label{cog}", fig.pos="ht"}

### 2020-11-16 submission/manuscript.tex
Coauthors review done.

@@
 \usepackage{lineno}
+\renewcommand{\linenumberfont}{\normalfont\tiny}
 
@@
 Markovski\textsuperscript{1}, Zihao Zhao\textsuperscript{2}, Gerhard J.
-Herndl\textsuperscript{2,3}, Mirjana Najdek\textsuperscript{1}
+Herndl\textsuperscript{2,3,4}, Mirjana Najdek\textsuperscript{1}
 
@@
 
+4. Vienna Metabolomics Center, University of Vienna, Austria
+
 \textsuperscript{\(*\)}To whom correspondence should be addressed:
@@
 \newpage
-\linenumbers
+
 \sisetup{mode=text}
@@
 such as 16S rRNA sequencing and metagenomics, with methods that allow
-insight into the metabolic status, such as metaproteomics. To apply
+insights into the metabolic status, such as metaproteomics. To apply
 these techniques to the microbial community inhabiting the surfaces of
-marine macrophytes it is advisable to perform, prior to the analysis, a
-selective DNA and protein isolation so that the host material, present
-in higher quantities, is not hampering the analysis. Two protocols, for
-DNA and protein isolation, were adapted for selective extractions of DNA
-and proteins from epiphytic communities inhabiting the surfaces of two
-marine macrophytes: the seagrass \emph{Cymodocea nodosa} and the
+marine macrophytes it is advisable to perform a selective DNA and
+protein isolation prior to the analysis to avoid biases due to the host
+material being present in high quantities. Two protocols for DNA and
+protein isolation were adapted for selective extractions of DNA and
+proteins from epiphytic communities inhabiting the surfaces of two
+marine macrophytes, the seagrass \emph{Cymodocea nodosa} and the
 macroalga \emph{Caulerpa cylindracea}. Protocols showed an almost
@@
 suitable for metagenomic and 16S rRNA sequencing, while isolated
-proteins could be identified by mass spectrometry, showing that
-protocols can be used in 16S rRNA, metagenomic and metaproteomic
-analysis. Low presence of host DNA and proteins, observed in isolated
-samples, indicated a selective nature of the protocols. Furthermore, the
-procedures are based on universally available laboratory chemicals
-making the protocols widely applicable. Taken together, the adapted
-protocols ensure an almost complete removal of the macrophyte epiphytic
-community, are selective for microbes inhabiting macrophyte surfaces and
-are providing DNA and proteins applicable in 16S rRNA sequencing,
-metagenomics and metaproteomics.
+proteins could be identified by mass spectrometry. The low presence of
+host DNA and proteins in the samples indicated a high specificity of the
+protocols. The procedures are based on universally available laboratory
+chemicals making the protocols widely applicable. Taken together, the
+adapted protocols ensure an almost complete removal of the macrophyte
+epiphytic community. The procedures are selective for microbes
+inhabiting macrophyte surfaces and provide DNA and proteins applicable
+in 16S rRNA sequencing, metagenomics and metaproteomics.
 
@@
 
-Surfaces of marine macrophytes are inhabited by a diverse microbial
+Surfaces of marine macrophytes are colonized by a diverse microbial
 community whose structure and function are poorly understood (Egan
 \emph{et al.}, 2013). As less than 1 \% of all prokaryotic species are
-culturable, to study these organisms, molecular methods such as 16S rRNA
-sequencing, metagenomics and metaproteomics are indispensable (Amann
+culturable, molecular methods such as 16S rRNA sequencing, metagenomics
+and metaproteomics are indispensable to study these organisms (Amann
 \emph{et al.}, 1995; Su \emph{et al.}, 2012). Applying these techniques
-requires an initial isolation step, with the purpose of obtaining high
+requires an initial isolation step with the purpose of obtaining high
 quality DNA and proteins.
 
-Biological material (i.e.~proteins and DNA) from pelagic microbial
+Biological material (i.e., proteins and DNA) from pelagic microbial
 communities is usually isolated by collecting cells onto filters and
-subsequently isolating the aimed compound (Gilbert \emph{et al.}, 2009).
-If a specific microbial size fraction is aimed sequential filtration is
-applied (Massana \emph{et al.}, 1997; Andersson \emph{et al.}, 2010). In
-contrast, obtaining biological materials from microorganism inhabiting
-surfaces requires either a cell detachment procedure prior to isolation
-or the host material is coextracted together with the targeted material.
-Methods for separating microbial cells form the host include shaking of
+subsequently isolating the target organisms or communities (Gilbert
+\emph{et al.}, 2009). If a specific microbial size fraction is aimed
+sequential filtration is applied (Massana \emph{et al.}, 1997; Andersson
+\emph{et al.}, 2010). In contrast, obtaining microorganisms associated
+to surfaces require either a cell detachment procedure prior to
+isolation or the host material is co-extracted with the target material.
+Methods for separating microbial cells from the host include shaking of
 host tissue (Gross \emph{et al.}, 2003; Nõges \emph{et al.}, 2010),
-scraping of macrophyte surfaces (Uku \emph{et al.}, 2007) or the
-application of ultrasonication (Weidner \emph{et al.}, 1996; Cai
-\emph{et al.}, 2014). It was shown that shaking alone is not sufficient
-to remove microbial cells, at least from plant root surfaces
+scraping of macrophyte surfaces (Uku \emph{et al.}, 2007) or applying
+ultrasonication (Weidner \emph{et al.}, 1996; Cai \emph{et al.}, 2014).
+It was shown that shaking alone is not sufficient to remove microbial
+cells from surfaces, at least not from plant root surfaces
 (Richter-Heitmann \emph{et al.}, 2016). Manual separation methods, such
-as scraping and brushing, are time consuming and subjective, as the
+as scraping and brushing are time consuming and subjective, as the
 detachment efficiency depends on host tissue and the person performing
@@
 \emph{et al.}, 2016). The downside of this procedure is that complete
-cell removal was still not obtained and tissue disruption was observed
+cell removal is still not obtained and tissue disruption was observed
 especially after the application of probe ultrasonication
 (Richter-Heitmann \emph{et al.}, 2016). An alternative to these cell
-detachment procedures is the isolation of targeted epiphytic compounds
-together with host materials (Staufenberger \emph{et al.}, 2008; Jiang
+detachment procedures is the isolation of target epiphytic compounds
+together with host material (Staufenberger \emph{et al.}, 2008; Jiang
 \emph{et al.}, 2015). This procedure can lead to problems in the
@@
 
-An alternative to these procedures is a direct isolation of the targeted
+An alternative to these procedures is a direct isolation of the target
 material by incubating macrophyte tissues in an extraction buffer. After
-the incubation, the undisrupted host tissue is removed and the isolation
-procedure continues, omitting host material contaminations. To our
-knowledge the only procedure describing a direct and selective epiphytic
-DNA isolation from the surfaces of marine macrophytes was provided by
-Burke \emph{et al.} (2009). In contrast to previously described methods
-this protocol enables an almost complete removal of the surface
-community and it was used for 16S rRNA gene clone library construction
-(Burke and Thomas \emph{et al.}, 2011) and metagenome sequencing (Burke
-and Peter Steinberg \emph{et al.}, 2011). This method, though providing
-a selective isolation procedure, uses a rapid multienzyme cleaner (3M)
-that is not available worldwide and without a known composition (Burke
-\emph{et al.}, 2009). Also to our knowledge, no selective isolation
-protocol for proteins from epiphytic communities inhabiting marine
-macrophytes was established.
+the incubation, the undisrupted host tissue is removed followed by the
+isolation procedure, omitting host material contaminations. To our
+knowledge, the only procedure describing a direct and selective
+epiphytic DNA isolation from the surfaces of marine macrophytes was
+described by Burke \emph{et al.} (2009). In contrast to previously
+described methods, this protocol enables an almost complete removal of
+the surface community. It was used for 16S rRNA gene clone library
+construction (Burke and Thomas \emph{et al.}, 2011) and metagenome
+sequencing (Burke and Peter Steinberg \emph{et al.}, 2011). This method,
+although providing a selective isolation procedure, uses a rapid
+multi-enzyme cleaner (3M) that is not available worldwide and the
+chemical constituents are unknown (Burke \emph{et al.}, 2009). Also to
+our knowledge, no selective isolation protocol to perform
+(meta)proteomics of epiphytic communities associated with marine
+macrophytes has been developed yet.
 
@@
 isolation from filters (Massana \emph{et al.}, 1997) and a protocol used
-for protein isolation from sediments (Chourey \emph{et al.}, 2010;
-Hultman \emph{et al.}, 2015) for selective extractions of DNA and
-proteins from epiphytic communities inhabiting the surfaces of two
-marine macrophytes: the seagrass \emph{Cymodocea nodosa} and the
-macroalga \emph{Caulerpa cylindracea}. In addition, we tested the
-removal efficiency of the protocol and the suitability of obtained DNA
-and proteins for 16S rRNA sequencing, metagenomics and metaproteomics.
+for protein isolation from soils (Chourey \emph{et al.}, 2010; Hultman
+\emph{et al.}, 2015). These two adapted methods allowed for a selective
+extraction of DNA and proteins from epiphytic communities inhabiting the
+surfaces of two marine macrophytes, the seagrass \emph{Cymodocea nodosa}
+and the macroalga \emph{Caulerpa cylindracea}. In addition, we tested
+the removal efficiency of the protocol and the suitability of obtained
+DNA and proteins for 16S rRNA sequencing, metagenomics and
+metaproteomics.
 
@@
 
-To assess the removal efficiency of DNA and protein isolation procedures
-leaves and thalli were examined under a confocal microscope before and
-after treatments were performed. Developed procedures showed an almost
-complete removal of the surface community of both, \emph{C. nodosa} and
-\emph{C. cylindracea}. In addition, a similar removal efficiency was
-observed for communities sampled in contrasting seasons, December 2017
-(\autoref{confocal_december}) and June 2018 (\autoref{confocal_june}).
-Also, no effect of station, settlement or isolation procedure (DNA or
-protein) on the removal efficiency was observed (Figs.
-\ref{confocal_december} and \ref{confocal_june}).
-
-To evaluate if the obtained DNA is suitable to determine the microbial
-community structure an Illumina sequencing of the V4 16S rRNA region was
-performed. Sequencing yielded a total of 336,944 sequences after quality
-curation and exclusion of eukaryotic, mitochondrial and no relative
-sequences. The number of sequences classified as chloroplast was 97,334.
-After excluding these sequences the total number of retrieved reads
-dropped to 239,610, ranging from 22,596 to 52,930 sequences per sample
-(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
-applied the rarefaction curves did not level off which is commonly
-observed in high-throughput 16S rRNA amplicon sequencing procedures
-(\autoref{supp-rarefaction}). Sequence clustering at a similarity level
-of 97 \si{\percent} yielded a total of 8,355 different OTUs. Taxonomic
-classification of reads allowed for a macrophyte associated epiphytic
-community determination that was mainly composed of:
+To assess the removal efficiency of the DNA and protein isolation
+procedures, leaves and thalli were examined under a confocal microscope
+before and after treatments were performed. The modified procedures
+resulted in an almost complete removal of the surface community of both,
+\emph{C. nodosa} and \emph{C. cylindracea}. In addition, a similar
+removal efficiency was observed for communities sampled in contrasting
+months, December 2017 (\autoref{confocal_december}) and June 2018
+(\autoref{confocal_june}). Also, no effect of station, settlement or
+isolation procedure (DNA or protein) on the removal efficiency was
+observed (Figs. \ref{confocal_december} and \ref{confocal_june}).
+
+To evaluate whether the obtained DNA is suitable to determine the
+composition of the microbial community Illumina sequencing of the V4
+region of the 16S rRNA was performed. Sequencing yielded a total of
+336,944 sequences after quality curation and exclusion of eukaryotic,
+mitochondrial and no relative sequences. The number of sequences
+classified as chloroplasts was 97,334. After excluding these sequences
+the total number of retrieved reads was 239,610, ranging from 22,596 to
+52,930 sequences per sample (\autoref{supp-nseq_notus}). Even when the
+highest sequencing effort was applied, the rarefaction curves did not
+level off which is commonly observed in high-throughput 16S rRNA
+amplicon sequencing (\autoref{supp-rarefaction}). Sequences clustering
+at a similarity level of 97 \si{\percent} yielded a total of 8,355
+different OTUs. Taxonomic classification of reads revealed a
+macrophyte-associated epiphytic community that mainly composed of
 \emph{Alphaproteobacteria} (14.9 ± 3.5 \si{\percent}),
@@
 high proportion of chloroplast sequences in all analyzed samples (33.4 ±
-9.4 \si{\percent}) (\autoref{community}). To determine if chloroplast
-sequences originate from hosts or eukaryotic epiphytic organisms, we
-exported SILVA classified chloroplast sequences and reclassified them
-using the RDP (Ribosomal Database Project) training set that allows for
-a more detailed chloroplast classification. The largest proportion of
-sequences were classified as Bacillariophyta (89.7 ± 5.7 \si{\percent})
-indicating that the DNA removal procedure did not coextract larger
-quantities of host DNA (\autoref{chloroplast}). Chloroplast sequences
-classified as Streptophyta constituted 3.3 ± 2.8 \si{\percent} of all
-chloroplast sequences originating from \emph{C. nodosa} samples, while
-sequences classified as Chlorophyta comprised only 0.02 ± 0.01
-\si{\percent} of all chloroplast sequences associated with \emph{C.
-cylindracea} samples.
-
-To determine if the extracted DNA can be used for metagenomic sequencing
-four samples containing epiphytic DNA were selected and shotgun
-sequenced using an Illumina platform. Metagenomic sequencing yielded
-from 207,149,524 to 624,029,930 sequence pairs
+9.4 \si{\percent}) (\autoref{community}). To determine whether
+chloroplast sequences originate from the host or eukaryotic epiphytic
+organisms, we exported SILVA-classified chloroplast sequences and
+reclassified them using the RDP (Ribosomal Database Project) training
+set that allows for a more detailed chloroplast classification. The
+largest proportion of sequences was classified as Bacillariophyta (89.7
+± 5.7 \si{\percent}) indicating that the DNA removal procedure resulted
+in only minor co-extracted quantities of host DNA
+(\autoref{chloroplast}). Chloroplast sequences classified as
+Streptophyta constituted 3.3 ± 2.8 \si{\percent} of all chloroplast
+sequences originating from \emph{C. nodosa} samples, while sequences
+classified as Chlorophyta comprised only 0.02 ± 0.01 \si{\percent} of
+all chloroplast sequences associated with \emph{C. cylindracea} samples.
+
+To determine whether the extracted DNA can be used for metagenomic
+sequencing, four samples containing epiphytic DNA were selected and
+shotgun sequenced using an Illumina platform. Metagenomic sequencing
+yielded between 207,149,524 and 624,029,930 sequence pairs
 (\autoref{supp-metagenomic_statistics}). Obtained sequences were
 successfully assembled into contigs whose L50 ranged from 654 to 1,011
-bp. In addition, predicted coding sequences were successfully
-functionally annotated (9,066,667 -- 20,256,215 annotated sequences;
-\autoref{cog}) and taxonomically classified. Functional annotation
-allowed for an assessment of the relative contribution of each COG
-(Clusters of Orthologous Groups) functional category to the total number
-of annotated coding sequences (\autoref{cog}a). Functional categories
-containing the highest number of sequences were: C (Energy production
-and conversion), E (Amino acid transport and metabolism), M (Cell
-wall/membrane/envelope biogenesis), L (Replication, recombination and
-repair) and P (Inorganic ion transport and metabolism). If host DNA was
-coextracted with epiphytic it should be detected in larger proportions
-in sequenced metagenomes. Indeed, no higher proportions of coding
-sequences classified into phylum Streptophyta or Chlorophyta were
-detected (\autoref{supp-metagenomic_taxonomy}). Sequenced metagenomic
-DNA originating from the surface of \emph{C. nodosa} contained 1.3
-\si{\percent} of coding sequences classified into the phylum
-Streptophyta in December 2017 and 0.7 \si{\percent} in June 2018.
-Furthermore, the summed RPKM (Reads Per Kilobase Million) of these
-sequences constituted 1.7 \si{\percent} of total RPKM of all
-successfully classified sequences in December 2017 and 1.1 \si{\percent}
-in June 2018. Similar low proportions of host coding sequences were
-detected in metagenomic samples originating from the surfaces of
-\emph{C. cylindracea}. Of all successfully classified coding sequences
-0.2 \si{\percent} were classified into Chlorophyta in December 2017 and
-0.1 \si{\percent} in June 2018. A relatively higher proportion of these
-sequences' RPKM than in the case of \emph{C. nodosa} was observed,
-indicating a higher coextraction of host DNA in \emph{C. cylindracea}.
-In December, the proportion of RPKM of sequences classified into
-Chlorophyta was 8.2 \si{\percent}, while in June 2018 it reached 13.6
-\si{\percent}.
-
-To evaluate if the procedure for protein extraction is suitable for
+bp. In addition, predicted coding sequences were functionally annotated
+(9,066,667 -- 20,256,215 annotated sequences; \autoref{cog}) and
+taxonomically classified. Functional annotation allowed for an
+assessment of the relative contribution of each COG (Clusters of
+Orthologous Groups) functional category to the total number of annotated
+coding sequences (\autoref{cog}a). Functional categories containing the
+highest number of sequences were C (Energy production and conversion), E
+(Amino acid transport and metabolism), M (Cell wall/membrane/envelope
+biogenesis), L (Replication, recombination and repair) and P (Inorganic
+ion transport and metabolism). If host DNA is co-extracted with
+epiphytes it should be detected in large proportions in sequenced
+metagenomes. However, no large proportions of coding sequences
+classified as Streptophyta and Chlorophyta were detected
+(\autoref{supp-metagenomic_taxonomy}). Sequenced metagenomic DNA
+originating from the surface of \emph{C. nodosa} contained 1.3
+\si{\percent} of coding sequences classified as Streptophyta in December
+2017 and 0.7 \si{\percent} in June 2018. Furthermore, the summed RPKM
+(Reads Per Kilobase Million) of these sequences constituted 1.7
+\si{\percent} of total RPKM of all successfully classified sequences in
+December 2017 and 1.1 \si{\percent} in June 2018. Similar low
+proportions of host coding sequences were detected in metagenomic
+samples originating from the surfaces of \emph{C. cylindracea}. Of all
+successfully classified coding sequences 0.2 \si{\percent} were
+classified as Chlorophyta in December 2017 and 0.1 \si{\percent} in June
+2018. A relatively higher proportion of RPKM of these sequences than in
+the case of \emph{C. nodosa} was observed, indicating a higher
+coextraction of host DNA in \emph{C. cylindracea}. In December, the
+proportion of RPKM of sequences classified as Chlorophyta was 8.2
+\si{\percent}, while in June 2018 it reached 13.6 \si{\percent}.
+
+To evaluate whether the procedure for protein extraction is suitable for
 metaproteomic analysis, obtained proteins were trypsin digested and
@@
 (\autoref{cog}b). Functional categories containing the highest number of
-identified proteins were: C (Energy production and conversion), G
+identified proteins were C (Energy production and conversion), G
 (Carbohydrate transport and metabolism), P (Inorganic ion transport and
@@
 chaperones) and E (Amino acid transport and metabolism). Isolated
-proteins could derive from epiphytic organisms inhabiting the macrophyte
-surface or from macrophyte tissue underlying them. The contribution of
-proteins originating from host tissues was evaluated by identifying all
-the proteins predicted to belong to any taxonomic group within the phyla
-Streptophyta and Chlorophyta and by calculating their contribution to
-the number and abundance of all identified proteins. On average,
-proteins isolated from the surface of \emph{C. nodosa} contained 1.8 ±
-0.06 \si{\percent} of proteins associated with Streptophyta,
-contributing to 2.2 ± 0.8 \si{\percent} of total proteins. Similar to
-metagenomes, proteins associated with Chlorophyta contributed more to
-\emph{C. cylindracea} than proteins associated with Streptophyta to
-\emph{C. nodosa}. Chlorophyta associated proteins comprised 5.2 ± 0.06
-\si{\percent} of all identified proteins in \emph{C. cylindracea},
-contributing to 19.2 ± 1.5 \si{\percent} of all protein abundances.
+proteins could originate from epiphytic organisms inhabiting the
+macrophyte surface and/or from macrophyte tissue underlying them. The
+contribution of proteins originating from host tissues was evaluated by
+identifying all the proteins predicted to belong to any taxonomic group
+within the phyla Streptophyta and Chlorophyta and by calculating their
+contribution to the number and abundance (NAAF -- Normalized Abundance
+Area Factor) of all identified proteins. On average, proteins isolated
+from the surface of \emph{C. nodosa} contained 1.8 ± 0.06 \si{\percent}
+of proteins associated with Streptophyta, contributing to 2.2 ± 0.8
+\si{\percent} of total proteins. Similar to metagenomes, proteins
+associated with Chlorophyta contributed more to \emph{C. cylindracea}
+than proteins associated with Streptophyta to \emph{C. nodosa}.
+Chlorophyta associated proteins comprised 5.2 ± 0.06 \si{\percent} of
+all identified proteins in \emph{C. cylindracea}, contributing 19.2 ±
+1.5 \si{\percent} to the total protein abundance.
 
@@
 
-The study of marine macrophyte epiphytic communities using culture
-independent techniques, such as 16S rRNA analysis, metagenomics and
-metaproteomics, requires an initial step of biological material
-isolation. Methods that have been developed for selective isolation of
-epiphytic biological material can be divided into two groups: (i)
-procedures involving, prior to extraction, a cell detachment step such
-as shaking (Gross \emph{et al.}, 2003; Nõges \emph{et al.}, 2010),
-scraping (Uku \emph{et al.}, 2007) and ultrasonication (Weidner \emph{et
-al.}, 1996; Cai \emph{et al.}, 2014) and (ii) procedures involving a
-host tissue incubation aiming at direct lysis of epiphytic microbial
-cells (Burke \emph{et al.}, 2009). Protocols that include a cell
-detachment procedure do not provide complete cell removal (Cai \emph{et
-al.}, 2014; Richter-Heitmann \emph{et al.}, 2016) and in the case of
-probe ultrasonication tissue disruption can also be observed
-(Richter-Heitmann \emph{et al.}, 2016). Though, host tissue incubation
-procedures provide an almost complete cell removal and selective
-isolation, existing protocols like the one developed for DNA isolation
-by Burke \emph{et al.} (2009) use in the extraction buffer a rapid
-multienzyme cleaner (3M) not available worldwide and without a known
-composition (Burke \emph{et al.}, 2009). To circumvent these problems we
-developed and tested two protocols for selective DNA and protein
-isolation from marine macrophyte epiphytic communities.
-
-To test if the developed DNA and protein isolation procedures could be
-applied on a variety of macrophyte species we selected \emph{C. nodosa}
-and \emph{C. cylindracea} as representatives of seagrass and macroalgal
-species, on which the procedures were tested. These species especially
-differ morphologically. While \emph{C. nodosa} leaves are flat, \emph{C.
-cylindracea} thallus is composed of uneven surfaces (Kuo and den Hartog,
-2001; Verlaque \emph{et al.}, 2003). The developed procedures showed an
-almost complete removal of epiphytic cells from the surfaces of both
-species comparable to the result of Burke \emph{et al.} (2009) and
-indicating that structural differences do not impact the removal
-efficiency. In addition, isolation protocols were tested in two
-contrasting season as it is known that macrophytes are harboring more
-algal epiphytes during autumn and winter (Reyes and Sansón, 2001). No
-differences in the removal efficiency was observed between seasons
-suggesting that protocols can be used on macrophyte samples retrieved
-throughout the year. Also, no removal differences were observed on
-samples derived from the same host but from different localities or
-settlements demonstrating the stability of the protocol in cell removal
-efficiency.
-
-Successful amplification and sequencing of the V4 16S rRNA gene region
-proved that the isolated DNA can be used to estimate the microbial
+To test whether the developed DNA and protein isolation protocols
+efficiently detach microbes from the macrophyte surface we selected
+\emph{C. nodosa} and \emph{C. cylindracea} as representatives of
+seagrass and macroalgal species. These species differ morphologically.
+While \emph{C. nodosa} leaves are flat, \emph{C. cylindracea} thalli are
+characterized by an uneven surface (Kuo and den Hartog, 2001; Verlaque
+\emph{et al.}, 2003). The developed protocol led to an almost complete
+removal of epiphytic cells from the surfaces of both species comparable
+to the result of Burke \emph{et al.} (2009), indicating that structural
+differences do not impact the removal efficiency. In addition, isolation
+protocols were tested in two contrasting seasons, as it is known that
+macrophytes are harboring more algal epiphytes during autumn and winter
+(Reyes and Sansón, 2001). No differences in the removal efficiency was
+observed between seasons suggesting that these protocols can be used on
+macrophyte samples retrieved throughout the year. Also, no removal
+differences were observed on samples derived from the same host but from
+different locations.
+
+Successful amplification and sequencing of the V4 region of the 16S rRNA
+gene proved that the isolated DNA can be used to estimate the microbial
 epiphytic diversity. Taxonomic groups detected in this step can also be
@@
 epiphytic chloroplast DNA. Although the proportion of obtained
-chloroplast 16S rRNA sequences in our samples was substantial they
+chloroplast 16S rRNA sequences in our samples was substantial, they
 derived almost exclusively from eukaryotic epiphytes. High proportion of
@@
 that include direct cellular lysis on host surfaces were observed before
-(Michelou \emph{et al.}, 2013). It is possible that chloroplast specific
-sequences even in these studies are originating from eukaryotic
-epiphytic cells and not host chloroplast. Indeed, it is common during
-16S rRNA profiling of pelagic microbial communities to observe high
+(Michelou \emph{et al.}, 2013). It is possible that chloroplast-specific
+sequences even in these studies originated from eukaryotic epiphytic
+cells and not from host chloroplasts. Indeed, it is common during 16S
+rRNA profiling of pelagic microbial communities to observe high
 proportions of chloroplast sequences (Gilbert \emph{et al.}, 2009;
@@
 
-Beside 16S rRNA sequencing high quality DNA is needed for metagenomics.
-The obtained number of metagenomic sequences and assembly statistics
-were comparable to metagenomes and metatranscriptomes derived from
-similar surface associated communities (Crump \emph{et al.}, 2018; Cúcio
-\emph{et al.}, 2018). In addition, functional annotation of predicted
-coding sequences to COG functional categories showed that the obtained
-metagenomes can be used to determine the metabolic capacity of surface
-associated communities (Leary \emph{et al.}, 2014; Cúcio \emph{et al.},
-2018). The proportion of coding sequences, including their RPKM,
-originating from \emph{C. nodosa} metagenomes and classified as
-Streptophyta was low indicating a good selectivity of the isolation
-procedure towards epiphytic cells. In the case of DNA samples isolated
-from the surface of \emph{C. cylindracea} the proportion of Chlorophyta
-coding sequences was also low but their RPKM was higher than in the case
-of \emph{C. nodosa}. One of the causes for this elevated RPKM of
-Chlorophyta sequences in \emph{C. cylindracea} could lay in tissue
-structure differences between these two host species. While \emph{C.
-nodosa} leaves are composed of individual cells, the thallus of \emph{C.
+High quality DNA is also needed for metagenomics. The obtained number of
+metagenomic sequences and assembly statistics were comparable to
+metagenomes and metatranscriptomes derived from similar surface
+associated communities (Crump \emph{et al.}, 2018; Cúcio \emph{et al.},
+2018). In addition, functional annotation of predicted coding sequences
+to COG functional categories showed that the obtained metagenomes can be
+used to determine the metabolic capacity of surface associated
+communities (Leary \emph{et al.}, 2014; Cúcio \emph{et al.}, 2018). The
+proportion of coding sequences, including their RPKM, originating from
+\emph{C. nodosa} metagenomes and classified as Streptophyta was low
+indicating that the isolation procedure was specific for epiphytic
+cells. DNA samples isolated from the surface of \emph{C. cylindracea}
+exhibited a low proportion of Chlorophyta coding sequences, however,
+their RPKM was higher than in the samples originating from \emph{C.
+nodosa}. One of the reasons for this elevated RPKM of Chlorophyta
+sequences in \emph{C. cylindracea} could be the differences in the
+tissue structure between these two host species. While \emph{C. nodosa}
+leaves are composed of individual cells, the thallus of \emph{C.
 cylindracea} is, like in other siphonous algal species, composed of a
@@
 genetic material into the extraction buffer causing an elevated presence
-of host sequences in metagenomic data.
-
-To obtain an insight into the metabolic status of uncultivated members,
+of host sequences in the samples for metagenome analyses.
+
+To obtain insight into the metabolic status of uncultivated prokaryotes,
 a metaproteomic approach is required (Saito \emph{et al.}, 2019). The
-applied protocol for epipyhtic protein isolation followed by a
+applied protocol for epiphytic protein isolation followed by a
 metaproteomic analysis identified between 14,219 and 16,449 proteins,
@@
 functional annotation of identified proteins into COG functional
-categories showed that the protein isolation procedure can be used to
-assess the metabolic status of the macrophyte epiphytic community (Leary
-\emph{et al.}, 2014). Similar to metagenomes, the number of identified
-proteins, including their abundances, associated with Streptophyta in
-\emph{C. nodosa} samples were low indicating that the procedure is
-selective for epiphytic cell proteins. In addition, a higher number of
-identified proteins, and especially their abundances, associated with
-Chlorophyta was observed in \emph{C. cylindracea} samples. The cause of
-this elevated presence of Chlorophyta associated proteins can be, as in
-the case of the DNA isolation procedure, explained by the absence of
-individual cells in this siphonous alga (Coneva and Chitwood, 2015).
+categories showed that the protein isolation protocol can be used to
+assess the metabolic status of the epiphytic community (Leary \emph{et
+al.}, 2014). Similar to the results of the metagenomic analysis, the
+number and abundance of identified proteins affiliated to Streptophyta
+in \emph{C. nodosa} samples were low, indicating that the procedure is
+selective for epiphytic cell proteins. In addition, a higher number and
+abundance of identified proteins associated with Chlorophyta were
+observed in \emph{C. cylindracea} samples. The cause of this elevated
+presence of Chlorophyta-associated proteins can be, similar to the DNA
+isolation protocol, explained by the absence of individual cells in this
+siphonous alga (Coneva and Chitwood, 2015).
 
 In conclusion, the developed protocols for DNA and protein isolation
-from macrophyte surfaces are providing an almost complete removal of the
-epiphytic community and are shown to ensure removal from both, \emph{C.
-nodosa} and \emph{C. cylindracea}, in different seasons. Also, the
-obtained DNA and proteins are suitable for 16S rRNA sequencing,
-metagenomics and metaproteomics, while the obtained material contains
-low quantities of host DNA or proteins making the procedures epiphyte
-selective. Furthermore, the procedures are based on universally
-available laboratory chemicals making the protocols widely applicable.
+from macrophyte surfaces almost completely remove the epiphytic
+community from both, \emph{C. nodosa} and \emph{C. cylindracea}, in
+different seasons. Also, the obtained DNA and proteins are suitable for
+16S rRNA sequencing, metagenomics and metaproteomics analyses while the
+obtained material contains low quantities of host DNA and proteins
+making the protocols specific for epiphytes. Furthermore, the protocols
+are based on universally available laboratory chemicals hence, making
+them widely applicable.
 
@@
 Leaves of \emph{C. nodosa} were sampled in a \emph{C. nodosa} meadow in
-the Bay of Saline (\ang{45;7;5} N, \ang{13;37;20} E) and in a \emph{C.
-nodosa} meadow invaded by \emph{C. cylindracea} in the proximity of the
-village of Funtana (Bay of Funtana; \ang{45;10;39} N, \ang{13;35;42} E).
-Thalli of \emph{C. cylindracea} were sampled in the same \emph{C.
-nodosa} invaded meadow in the Bay of Funtana and on a locality of only
-\emph{C. cylindracea} located in the proximity of the invaded meadow.
-Leaves and thalli for 16S rRNA analysis, metagenomics and metaproteomics
-were collected in two contrasting seasons, on 4 December 2017 (16S rRNA
-analysis and metaproteomics), 14 December 2017 (metagenomics) and 18
-June 2018 (16S rRNA analysis, metagenomics and metaproteomics). During
-spring 2018 the \emph{C. nodosa} meadow in the Bay of Saline decayed to
-an extent that no leaves could be retrieved (Najdek \emph{et al.},
-2020). In addition, as not enough DNA for both metagenomic and 16S RNA
-analysis were obtained during the sampling on 4 December 2017, an
-additional sampling on 14 December 2017 was carried out in the Bay of
-Funtana. Leaves and thalli were collected by diving and transported to
-the laboratory in containers placed on ice and filled with site
-seawater. Upon arrival to the laboratory, \emph{C. nodosa} leaves were
-cut into sections of 1 -- 2 \si{\cm}, while \emph{C. cylindracea} thalli
-were cut into 5 -- 8 \si{\cm} long sections. Leaves and thalli were
-washed three times with sterile artificial seawater (ASW) to remove
-loosely attached microbial cells.
+the Bay of Saline, northern Adriatic Sea (\ang{45;7;5} N, \ang{13;37;20}
+E) and in a \emph{C. nodosa} meadow invaded by \emph{C. cylindracea} in
+the Bay of Funtana, northern Adriatic Sea (\ang{45;10;39} N,
+\ang{13;35;42} E). Thalli of \emph{C. cylindracea} were sampled in the
+same \emph{C. nodosa} invaded meadow in the Bay of Funtana and at a
+locality of only \emph{C. cylindracea} located in the proximity of the
+invaded meadow. Leaves and thalli for 16S rRNA analysis, metagenomics
+and metaproteomics were collected in two contrasting seasons, on 4
+December 2017 (16S rRNA analysis and metaproteomics), 14 December 2017
+(metagenomics) and 18 June 2018 (16S rRNA analysis, metagenomics and
+metaproteomics). During spring 2018, the \emph{C. nodosa} meadow in the
+Bay of Saline decayed to an extent that no leaves could be retrieved
+(Najdek \emph{et al.}, 2020). In addition, as not enough DNA for both
+metagenomic and 16S RNA analysis were obtained during the sampling on 4
+December 2017, an additional sampling on 14 December 2017 was carried
+out in the Bay of Funtana. Leaves and thalli were collected by diving
+and transported to the laboratory in containers placed on ice and filled
+with seawater from this site. Upon arrival to the laboratory, \emph{C.
+nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while \emph{C.
+cylindracea} thalli were cut into 5 -- 8 \si{\cm} long sections. Leaves
+and thalli were washed three times with sterile artificial seawater
+(ASW) to remove loosely attached microbial cells.
 
@@
 modified and adapted for microbial DNA isolation from macrophyte
-surfaces as described below. 5 \si{\ml} of lysis buffer (40
+surfaces as described below. Five (5) \si{\ml} of lysis buffer (40
 \si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar}
 sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or 2
-\si{\g} wet-weight of thalli. For every sample duplicate isolations were
-performed. Lysozyme was added (final concentration 1 \si{\mg\per\ml})
-and the mixture was incubated at 37 \si{\degreeCelsius} for 30
-\si{\minute}. Subsequently, proteinase K (final concentration 0.5
+\si{\g} wet-weight of thalli. For every sample, duplicate isolations
+were performed. Lysozyme was added (final concentration 1
+\si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius}
+for 30 \si{\minute}. Subsequently, proteinase K (final concentration 0.5
 \si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were
@@
 \si{\hour}. Following the incubation, tubes were vortexed for 10
-\si{\minute} and the mixture containing lyzed epiphytic cells was
+\si{\minute} and the mixture containing lysed epiphytic cells was
 separated from host leaves or thalli by transferring the solution into a
@@
 phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with
-chloroform:isoamyl alcohol (24:1). After each organic solvent mixture
-addition tubes were slightly vortexed and centrifuged at 4,500 × g for
-10 \si{\minute}. Following each centrifugation aqueous phases were
-retrieved. After the final extraction 1/10 of chilled 3 \si{\Molar}
+chloroform:isoamyl alcohol (24:1). After each addition of an organic
+solvent mixture, tubes were slightly vortexed and centrifuged at 4,500 ×
+g for 10 \si{\minute}. Following each centrifugation the aqueous phases
+were retrieved. After the final extraction 1/10 of chilled 3 \si{\Molar}
 sodium acetate (pH 5.2) was added. DNA was precipitated by adding 1
 volume of chilled isopropanol, incubating the mixtures overnight at
-\num{-20} °C and centrifuging at 16,000 × g and 4 \si{\degreeCelsius}
-for 20 \si{\minute}. The pellet was washed twice with 1 \si{\ml} of
-chilled 70 \si{\percent} ethanol and centrifuged after each washing step
-at 20,000 × g and 4 \si{\degreeCelsius} for 10 \si{\minute}. After the
-first washing step duplicate pellets form the same sample were pooled
-and transferred to a clean 1.5 \si{\ml} tube. The dried pellet was
-resuspended in 100 \si{\ul} of deionized water.
+\num{-20} \si{\degreeCelsius} and centrifuging at 16,000 × g and 4
+\si{\degreeCelsius} for 20 \si{\minute}. The pellet was washed twice
+with 1 \si{\ml} of chilled 70 \si{\percent} ethanol and centrifuged
+after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 10
+\si{\minute}. After the first washing step duplicate pellets from the
+same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.
+The dried pellet was resuspended in 100 \si{\ul} of deionized water.
 
@@
 contained on their 5' end a tagged sequence. Each sample was amplified
-in four parallel 25 \si{\ul} reactions of which each contained: 1 × Q5
-Reaction Buffer , 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml}
+in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5
+Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml}
 BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse
@@
 
-Obtained sequences were analyzed on a Life Science Compute Cluster
+Obtained sequences were analyzed on the Life Science Compute Cluster
 (LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
@@
 2015) and by searching against the NCBI non-redundant database (NR). To
-determine phylogeny the top 10 \si{\percent} hits with an e-value
+determine the phylogeny, the top 10 \si{\percent} hits with an e-value
 \textless{} 1 × 10\textsuperscript{-5} were used (-\/-top 10). Sequence
@@
 adapted for microbial protein isolation from macrophyte surfaces as
-described below. 20 \si{\ml} of protein extraction buffer (4
+described below. Twenty (20) \si{\ml} of protein extraction buffer (4
 \si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was
@@
 vortexed for 10 \si{\minute} and incubated again in boiling water for 5
-\si{\minute}. After a brief vortex the lysate was transferred to a clean
-tube separating the host leaves or thalli from the mixture containing
-lyzed epiphytic cells. Dithiothreitol (DTT; final concentration 24
-\si{\milli\Molar}) was added and proteins were precipitated with chilled
-100 \si{\percent} trichloroacetic acid (TCA; final concentration 20
-\si{\percent}) overnight at \num{-20} °C. Precipitated proteins were
-centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
-The obtained protein pellet was washed three times with chilled acetone.
-During the first washing step the pellet was transferred to a clean 1.5
-\si{\ml} tube. After each washing step samples were centrifuged at
-20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets
-were stored at \num{-80} °C until further analysis.
+\si{\minute}. After a brief vortex, the lysate was transferred to a
+clean tube separating the host leaves or thalli from the mixture
+containing lysed epiphytic cells. Dithiothreitol (DTT; final
+concentration 24 \si{\milli\Molar}) was added and proteins were
+precipitated with chilled 100 \si{\percent} trichloroacetic acid (TCA;
+final concentration 20 \si{\percent}) overnight at \num{-20} °C.
+Precipitated proteins were centrifuged at 10,000 × g and 4
+\si{\degreeCelsius} for 40 \si{\minute}. The obtained protein pellet was
+washed three times with chilled acetone. During the first washing step
+the pellet was transferred to a clean 1.5 \si{\ml} tube. After each
+washing step, samples were centrifuged at 20,000 × g and 4
+\si{\degreeCelsius} for 5 \si{\minute}. Dried pellets were stored at
+\num{-80} °C until further analysis.
 
@@
 according to the manufacturer's instructions with small modifications
-(Wiśniewski \emph{et al.}, 2009). Before the solution was loaded to the
+(Wiśniewski \emph{et al.}, 2009). Prior to loading the solution onto the
 column, protein pellets were solubilized in a urea sample buffer
@@
 remove larger particles. The first washing step after protein solution
-loading was repeated twice. In addition, centrifugation steps were
+loading was repeated twice. In addition, the centrifugation steps were
 prolonged if the column was clogged. Trypsin digestion was performed on
@@
 The final filtrate containing peptides was acidified with 1
-\si{\percent} (final concentration) trifluoroacetic acid, freezed at
-\num{-80} \si{\degreeCelsius} for 15 \si{\minute}, lyophilized and sent
-to the Vienna Metabolomics Center (University of Vienna) for
-metaproteomic analysis. Peptides were resuspended in 1 \si{\percent}
-(final concentration) trifluoroacetic acid (TFA), desalted using the
-Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the
-manufacturer's instructions and sequenced on a Q Exactive Hybrid
-Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
-Obtained MS/MS spectra were searched against a protein database composed
-of combined sequenced metagenomes using SEQUEST-HT engines and validated
-with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific,
-USA). The target-decoy approach was used to reduce the probability of
-false peptide identification. Results whose false discovery rate at the
-peptide level was \SI{< 1}{\percent} were kept. For protein
-identification a minimum of two peptides and one unique peptide were
-required. For protein quantification, a chromatographic peak area-based
-free quantitative method was applied.
+\si{\percent} (final concentration) trifluoroacetic acid (TFA), freezed
+at \num{-80} \si{\degreeCelsius}, lyophilized and sent to VIME -- Vienna
+Metabolomics Center (University of Vienna) for metaproteomic analysis.
+Peptides were resuspended in 1 \si{\percent} (final concentration) TFA,
+desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA)
+according to the manufacturer's instructions and sequenced on a Q
+Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher
+Scientific, USA). Obtained MS/MS spectra were searched against a protein
+database composed of combined sequenced metagenomes using SEQUEST-HT
+engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
+Fisher Scientific, USA). The target-decoy approach was used to reduce
+the probability of false peptide identification. Results whose false
+discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
+For protein identification a minimum of two peptides and one unique
+peptide were required. For protein quantification, a chromatographic
+peak area-based free quantitative method was applied.
 
@@
 Processing and visualization of 16S rRNA, metagenomic and metaproteomic
-data was done using R (version 3.6.0) (R Core Team, 2019), package
+data were done using R (version 3.6.0) (R Core Team, 2019), package
 tidyverse (version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple
@@
 loosely attached microbial cells, were fixed in the same concentration
-of formaldehyde and used as a positive control. For long therm storage,
+of formaldehyde and used as a positive control. For-long term storage,
 fixed leaves and thalli were immersed in a mixture of phosphate-buffered
 saline (PBS) and ethanol (1:1) and stored at \num{-20}
-\si{\degreeCelsius}. Treated and untreated segemnts of leaves and thalli
+\si{\degreeCelsius}. Treated and untreated segments of leaves and thalli
 were stained in a 2 × solution of SYBR Green I and examined under a
@@
 
-This work was founded by the Croatian Science Foundation through the
-MICRO-SEAGRASS project (IP-2016-06-7118). We would like to thank
-Margareta Buterer for technical support, Paolo Paliaga for help during
-sampling, Lucija Horvat for technical support in confocal microscopy and
-Dušica Vujaklija for help in accessing the confocal microscope. In
-addition, we would like to thank the University Computing Center of the
-University of Zagreb for access to the computer cluster Isabella and
-CUBE -- Computational Systems Biology of the University of Vienna for
-access to the Life Science Compute Cluster (LiSC).
+This work was funded by the Croatian Science Foundation through the
+MICRO-SEAGRASS project (project number IP-2016-06-7118). ZZ and GJH were
+supported by the Austrian Science Fund (FWF) project ARTEMIS (project
+number P28781-B21). We would like to thank Margareta Buterer for
+technical support, Paolo Paliaga for help during sampling, Lucija Horvat
+for technical support in confocal microscopy, Dušica Vujaklija for help
+in accessing the confocal microscope and Sonja Tischler for technical
+support in mass spectrometry. In addition, we would like to thank the
+University Computing Center of the University of Zagreb for access to
+the computer cluster Isabella and CUBE -- Computational Systems Biology
+of the University of Vienna for access to the Life Science Compute
+Cluster (LiSC).
 
@@
 nodosa} and \textit{C.
-cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 4 December 2017 and stained with SYBR Green I.
-Scale bar in all images is 60 \si{\um}.\label{confocal_december}}\label{fig:unnamed-chunk-1}
+cylindracea} surfaces from the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) sampled on 4 December 2017 and stained with SYBR Green I.
+Scale bar at all images is 60 \si{\um}.\label{confocal_december}}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 nodosa} and \textit{C.
-cylindracea} surfaces from the Bay of Funtana (Mixed and Monospecific Settlements) sampled on 19 June 2018 and stained with SYBR Green I.
-Scale bar in all images is 60 \si{\um}.\label{confocal_june}}\label{fig:unnamed-chunk-2}
+cylindracea} surfaces from the Bay of Funtana (mixed and monospecific settlements) sampled on 19 June 2018 and stained with SYBR Green I.
+Scale bar at all images is 60 \si{\um}.\label{confocal_june}}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant (> 1 \si{\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\textit{C.
+\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\textit{C.
 nodosa} and \textit{C.
-cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-3}
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{community}}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 nodosa} and \textit{C.
-cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{chloroplast}}\label{fig:unnamed-chunk-4}
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4 December 2017 and 19 June 2018).\label{chloroplast}}\label{fig:unnamed-chunk-4}
 \end{figure}
@@
 nodosa} and \textit{C.
-cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) and in two contrasting seasons (4/14 December 2017 and 19 June 2018).
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) and in two contrasting seasons (4/14 December 2017 and 19 June 2018).
 Total number of annotated coding sequences and identified proteins is given above the corresponding bar.\label{cog}}\label{fig:unnamed-chunk-5}

### 2020-11-16 submission/supplementary.Rmd
Coauthors review done.

@@
 
-\linenumbers
+[]: # (\linenumbers)
 \sisetup{mode=text}
@@
 nodosa} and \\textit{C.
-cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{rarefaction}", fig.pos="H"}
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) in two contrasting seasons (4 December 2017 and 19 June 2018).\\label{rarefaction}", fig.pos="H"}
 include_graphics("../results/figures/rarefaction.jpg")

### 2020-11-16 submission/supplementary.tex
Coauthors review done.

@@
 Netherlands
+
+4. Vienna Metabolomics Center, University of Vienna, Austria
 
@@
 
-\linenumbers
 \sisetup{mode=text}
@@
 nodosa} and \textit{C.
-cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (Mixed and Monospecific Settlements) in two contrasting seasons (4 December 2017 and 19 June 2018).\label{rarefaction}}\label{fig:unnamed-chunk-1}
+cylindracea}) sampled in the Bay of Saline and the Bay of Funtana (mixed and monospecific settlements) in two contrasting seasons (4 December 2017 and 19 June 2018).\label{rarefaction}}\label{fig:unnamed-chunk-1}
 \end{figure}

### 2020-11-16 code/plot_community_bar_plot.R

@@ -37,10 +37,10 @@ metadata <- read_tsv("data/raw/metadata.csv") %>%
 # Selection of groups for plotting
 plot <- filter(community, taxlevel==2 |
                  (taxlevel==4 & str_detect(taxon, "^Chloroplast$")) |
-                 (taxlevel==3 & str_detect(rankID, filter(community, str_detect(taxon, "^Proteobacteria$"))[[2]]))) %>%
+                 (taxlevel==3 & str_detect(rankID, paste0(filter(community, str_detect(taxon, "^Proteobacteria$"))$rankID, "\\.")))) %>%
   filter_at(6:ncol(.), any_vars(. >= 1)) %>%
   mutate_at(5:ncol(.), list(~case_when(taxon=="Cyanobacteria" ~ . - .[taxon=="Chloroplast"], TRUE ~ .))) %>%
-  mutate_at(5:ncol(.), list(~case_when(taxon=="Proteobacteria" ~ . - sum(.[taxlevel==3 & str_detect(rankID, filter(community, str_detect(taxon, "^Proteobacteria$"))[[2]])]), TRUE ~ .))) %>%
+  mutate_at(5:ncol(.), list(~case_when(taxon=="Proteobacteria" ~ . - sum(.[taxlevel==3 & str_detect(rankID, paste0(filter(community, str_detect(taxon, "^Proteobacteria$"))$rankID, "\\."))]), TRUE ~ .))) %>%
   mutate(taxon=str_replace(taxon, "Proteobacteria", "Other Proteobacteria")) %>%
   mutate(taxon=str_replace_all(taxon, c("unknown_unclassified"="No_Relative", "unknown"="No_Relative"))) %>%
   filter_at(6:ncol(.), any_vars(. >= 1)) %>%

### 2020-11-17 submission/manuscript.Rmd
Summary shortened.

@@
 ## Summary
-Studies of unculturable microbes often combine methods based on DNA, such as 16S rRNA sequencing and metagenomics, with methods that allow insights into the metabolic status, such as metaproteomics.
+Studies of unculturable microbes often combine methods such as 16S rRNA sequencing, metagenomics and metaproteomics.
 To apply these techniques to the microbial community inhabiting the surfaces of marine macrophytes it is advisable to perform a selective DNA and protein isolation prior to the analysis to avoid biases due to the host material being present in high quantities.
@@
 The obtained DNA was suitable for metagenomic and 16S rRNA sequencing, while isolated proteins could be identified by mass spectrometry.
-The low presence of host DNA and proteins in the samples indicated a high specificity of the protocols.
+Low presence of host DNA and proteins in the samples indicated a high specificity of the protocols.
 The procedures are based on universally available laboratory chemicals making the protocols widely applicable.

### 2020-11-17 submission/manuscript.tex

@@
 
-Studies of unculturable microbes often combine methods based on DNA,
-such as 16S rRNA sequencing and metagenomics, with methods that allow
-insights into the metabolic status, such as metaproteomics. To apply
-these techniques to the microbial community inhabiting the surfaces of
-marine macrophytes it is advisable to perform a selective DNA and
-protein isolation prior to the analysis to avoid biases due to the host
-material being present in high quantities. Two protocols for DNA and
-protein isolation were adapted for selective extractions of DNA and
-proteins from epiphytic communities inhabiting the surfaces of two
-marine macrophytes, the seagrass \emph{Cymodocea nodosa} and the
-macroalga \emph{Caulerpa cylindracea}. Protocols showed an almost
-complete removal of the epiphytic community regardless of the sampling
-season, station, settlement or host species. The obtained DNA was
-suitable for metagenomic and 16S rRNA sequencing, while isolated
-proteins could be identified by mass spectrometry. The low presence of
-host DNA and proteins in the samples indicated a high specificity of the
-protocols. The procedures are based on universally available laboratory
-chemicals making the protocols widely applicable. Taken together, the
-adapted protocols ensure an almost complete removal of the macrophyte
-epiphytic community. The procedures are selective for microbes
-inhabiting macrophyte surfaces and provide DNA and proteins applicable
-in 16S rRNA sequencing, metagenomics and metaproteomics.
+Studies of unculturable microbes often combine methods such as 16S rRNA
+sequencing, metagenomics and metaproteomics. To apply these techniques
+to the microbial community inhabiting the surfaces of marine macrophytes
+it is advisable to perform a selective DNA and protein isolation prior
+to the analysis to avoid biases due to the host material being present
+in high quantities. Two protocols for DNA and protein isolation were
+adapted for selective extractions of DNA and proteins from epiphytic
+communities inhabiting the surfaces of two marine macrophytes, the
+seagrass \emph{Cymodocea nodosa} and the macroalga \emph{Caulerpa
+cylindracea}. Protocols showed an almost complete removal of the
+epiphytic community regardless of the sampling season, station,
+settlement or host species. The obtained DNA was suitable for
+metagenomic and 16S rRNA sequencing, while isolated proteins could be
+identified by mass spectrometry. Low presence of host DNA and proteins
+in the samples indicated a high specificity of the protocols. The
+procedures are based on universally available laboratory chemicals
+making the protocols widely applicable. Taken together, the adapted
+protocols ensure an almost complete removal of the macrophyte epiphytic
+community. The procedures are selective for microbes inhabiting
+macrophyte surfaces and provide DNA and proteins applicable in 16S rRNA
+sequencing, metagenomics and metaproteomics.
 

### 2020-11-19 submission/manuscript.Rmd
Typos corrected.

@@
 The number of sequences classified as chloroplasts was `r number_chloroplast_sequences`.
-After excluding these sequences the total number of retrieved reads was `r number_sequences_without_chloroplast`, ranging from `r min_max$min` to `r min_max$max` sequences per sample (\autoref{supp-nseq_notus}).
+After excluding these sequences, the total number of retrieved reads was `r number_sequences_without_chloroplast`, ranging from `r min_max$min` to `r min_max$max` sequences per sample (\autoref{supp-nseq_notus}).
 Even when the highest sequencing effort was applied, the rarefaction curves did not level off which is commonly observed in high-throughput 16S rRNA amplicon sequencing (\autoref{supp-rarefaction}).
@@
 Primers used in this study, as in many other 16S rRNA amplicon procedures, also amplified chloroplast 16S rRNA genes.
-We observed a high proportion of chloroplast sequences in all analyzed samples (`r filter(plot, taxon=="Chloroplast")$mean` ± `r filter(plot, taxon=="Chloroplast")$sd` \si{\percent}) (\autoref{community}).
+We observed a high proportion of chloroplast sequences in all analysed samples (`r filter(plot, taxon=="Chloroplast")$mean` ± `r filter(plot, taxon=="Chloroplast")$sd` \si{\percent}) (\autoref{community}).
 To determine whether chloroplast sequences originate from the host or eukaryotic epiphytic organisms, we exported SILVA-classified chloroplast sequences and reclassified them using the RDP (Ribosomal Database Project) training set that allows for a more detailed chloroplast classification.
@@
 Obtained sequences were successfully assembled into contigs whose L50 ranged from `r format(min(metadata_metagenomic_statistics$L50), big.mark=',')` to `r format(max(metadata_metagenomic_statistics$L50), big.mark=',')` bp.
-In addition, predicted coding sequences were functionally annotated (`r format(min(metadata_metagenomic_statistics$eggnog), big.mark=',')` -- `r format(max(metadata_metagenomic_statistics$eggnog), big.mark=',')` annotated sequences; \autoref{cog}) and taxonomically classified.
+In addition, predicted coding sequences were functionally annotated (`r format(min(metadata_metagenomic_statistics$eggnog), big.mark=',')` -- `r format(max(metadata_metagenomic_statistics$eggnog), big.mark=',')` annotated sequences; \autoref{cog}a) and taxonomically classified.
 Functional annotation allowed for an assessment of the relative contribution of each COG (Clusters of Orthologous Groups) functional category to the total number of annotated coding sequences (\autoref{cog}a).
@@
 A relatively higher proportion of RPKM of these sequences than in the case of *C.
-nodosa* was observed, indicating a higher coextraction of host DNA in *C.
+nodosa* was observed, indicating a higher co-extraction of host DNA in *C.
 cylindracea*.
@@
 The developed protocol led to an almost complete removal of epiphytic cells from the surfaces of both species comparable to the result of @Burke2009, indicating that structural differences do not impact the removal efficiency.
-In addition, isolation protocols were tested in two contrasting seasons, as it is known that macrophytes are harboring more algal epiphytes during autumn and winter [@Reyes2001].
+In addition, isolation protocols were tested in two contrasting seasons, as it is known that macrophytes are harbouring more algal epiphytes during autumn and winter [@Reyes2001].
 No differences in the removal efficiency was observed between seasons suggesting that these protocols can be used on macrophyte samples retrieved throughout the year.
@@
 After the first washing step duplicate pellets from the same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.
-The dried pellet was resuspended in 100 \si{\ul} of deionized water.
+The dried pellet was re-suspended in 100 \si{\ul} of deionized water.
 
@@
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
-In the first PCR the 515F (5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3') primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016].
-These primers contained on their 5' end a tagged sequence.
+In the first PCR the 515F (5\prime-GTGYCAGCMGCCGCGGTAA-3\prime) and 806R (5\prime-GGACTACNVGGGTWTCTAAT-3\prime) primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016].
+These primers contained on their 5\prime end a tagged sequence.
 Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template.
@@
 
-Obtained sequences were analyzed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
+Obtained sequences were analysed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
 For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
@@
 
-Obtained sequences were analyzed on the Life Science Compute Cluster (LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
+Obtained sequences were analysed on the Life Science Compute Cluster (LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
 Individual sequences were assembled using MEGAHIT (version  1.1.2) [@Li2015] under default settings.
-Putative genes were predicted from contings longer than 200 bp using Prodigal (version  2.6.3) [@Hyatt2010] in metagenome mode (-p meta).
+Putative genes were predicted from contigs longer than 200 bp using Prodigal (version  2.6.3) [@Hyatt2010] in metagenome mode (-p meta).
 Abundances of predicted genes were expressed as Reads Per Kilobase Million (RPKM) and calculated using the BWA algorithm (version  0.7.16a) [@Li2009a].
@@
 The final filtrate containing peptides was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid (TFA), freezed at \num{-80} \si{\degreeCelsius}, lyophilized and sent to VIME -- Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis.
-Peptides were resuspended in 1 \si{\percent} (final concentration) TFA, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
+Peptides were re-suspended in 1 \si{\percent} (final concentration) TFA, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
 Obtained MS/MS spectra were searched against a protein database composed of combined sequenced metagenomes using SEQUEST-HT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
@@
 ### Confocal microscopy
-Host leaves and thalli from DNA and protein isolation steps were washed seven times in deionized water and fixed with formaldehyde (final concentration ~ 3 \si{\percent}).
-In addition, nontreated leaves and thalli, washed three times in ASW to remove loosely attached microbial cells, were fixed in the same concentration of formaldehyde and used as a positive control.
+Host leaves and thalli from DNA and protein isolation steps were washed seven times in deionized water and fixed with formaldehyde (final concentration ~3 \si{\percent}).
+In addition, non-treated leaves and thalli, washed three times in ASW to remove loosely attached microbial cells, were fixed in the same concentration of formaldehyde and used as a positive control.
 For-long term storage, fixed leaves and thalli were immersed in a mixture of phosphate-buffered saline (PBS) and ethanol (1:1) and stored at \num{-20} \si{\degreeCelsius}.
@@
 \newpage
-```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\\textit{C.
+```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$1 \\si{\\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\\textit{C.
 nodosa} and \\textit{C.

### 2020-11-19 submission/manuscript.Rmd
Format customization.

@@
 \newpage
-[]: # (\linenumbers)
+\linenumbers
 \sisetup{mode=text}
 \setlength\parindent{24pt}
+\doublespacing
 
@@
 
+\newpage
 ## Discussion
@@
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
-In the first PCR the 515F (5\prime-GTGYCAGCMGCCGCGGTAA-3\prime) and 806R (5\prime-GGACTACNVGGGTWTCTAAT-3\prime) primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016].
-These primers contained on their 5\prime end a tagged sequence.
+In the first PCR the 515F ($5'$-GTGYCAGCMGCCGCGGTAA-$3'$) and 806R ($5'$-GGACTACNVGGGTWTCTAAT-$3'$) primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016].
+These primers contained on their $5'$ end a tagged sequence.
 Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template.

### 2020-11-19 submission/manuscript.tex
Format customization.

@@
 
+\usepackage{setspace}
+
 \begin{document}
@@
 \newpage
-
+\linenumbers
 \sisetup{mode=text}
 \setlength\parindent{24pt}
+\doublespacing
 
@@
 mitochondrial and no relative sequences. The number of sequences
-classified as chloroplasts was 97,334. After excluding these sequences
+classified as chloroplasts was 97,334. After excluding these sequences,
 the total number of retrieved reads was 239,610, ranging from 22,596 to
@@
 procedures, also amplified chloroplast 16S rRNA genes. We observed a
-high proportion of chloroplast sequences in all analyzed samples (33.4 ±
+high proportion of chloroplast sequences in all analysed samples (33.4 ±
 9.4 \si{\percent}) (\autoref{community}). To determine whether
@@
 bp. In addition, predicted coding sequences were functionally annotated
-(9,066,667 -- 20,256,215 annotated sequences; \autoref{cog}) and
+(9,066,667 -- 20,256,215 annotated sequences; \autoref{cog}a) and
 taxonomically classified. Functional annotation allowed for an
@@
 the case of \emph{C. nodosa} was observed, indicating a higher
-coextraction of host DNA in \emph{C. cylindracea}. In December, the
+co-extraction of host DNA in \emph{C. cylindracea}. In December, the
 proportion of RPKM of sequences classified as Chlorophyta was 8.2
@@
 
+\newpage
+
 \hypertarget{discussion}{%
@@
 protocols were tested in two contrasting seasons, as it is known that
-macrophytes are harboring more algal epiphytes during autumn and winter
+macrophytes are harbouring more algal epiphytes during autumn and winter
 (Reyes and Sansón, 2001). No differences in the removal efficiency was
@@
 same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.
-The dried pellet was resuspended in 100 \si{\ul} of deionized water.
+The dried pellet was re-suspended in 100 \si{\ul} of deionized water.
 
@@
 amplified using a two-step PCR procedure. In the first PCR the 515F
-(5'-GTGYCAGCMGCCGCGGTAA-3') and 806R (5'-GGACTACNVGGGTWTCTAAT-3')
-primers from the Earth Microbiome Project
+(\(5'\)-GTGYCAGCMGCCGCGGTAA-\(3'\)) and 806R
+(\(5'\)-GGACTACNVGGGTWTCTAAT-\(3'\)) primers from the Earth Microbiome
+Project
 (\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
@@
 Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
-contained on their 5' end a tagged sequence. Each sample was amplified
-in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5
-Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml}
-BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse
-primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs,
-USA) and 5 \si{\ng} of DNA template. Cycling conditions were: initial
-denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of
-denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
-\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
-\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
-72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
-volumes were pooled and PCR products were purified using the GeneJET PCR
-Purification Kit (Thermo Fisher Scientific, USA) according to the
-manufacturer's instructions and following the protocol that included
-isopropanol addition for better small DNA fragment yield. The column was
-eluted in 30 \si{\ul} of deionized water. Purified PCR products were
-sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
-Martinsried, Germany. Before sequencing at IMGM, the second PCR
-amplification of the two-step PCR procedure was performed using primers
-targeting the tagged region incorporated in the first PCR. In addition,
-these primers contained adapter and sample-specific index sequences. The
-second PCR was carried out for 8 cycles. Beside samples, a positive and
-negative control were sequenced. A negative control was comprised of
-four parallel PCR reactions without DNA template, while for a positive
-control a mock community composed of evenly mixed DNA material
-originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
-used. Partial 16S rRNA sequences obtained in this study have been
-deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under
-accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794,
+contained on their \(5'\) end a tagged sequence. Each sample was
+amplified in four parallel 25 \si{\ul} reactions of which each contained
+1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7
+\si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of
+forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase
+(New England Biolabs, USA) and 5 \si{\ng} of DNA template. Cycling
+conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3
+\si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45
+\si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and
+elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an
+elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}. The four
+parallel reactions volumes were pooled and PCR products were purified
+using the GeneJET PCR Purification Kit (Thermo Fisher Scientific, USA)
+according to the manufacturer's instructions and following the protocol
+that included isopropanol addition for better small DNA fragment yield.
+The column was eluted in 30 \si{\ul} of deionized water. Purified PCR
+products were sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM
+Laboratories, Martinsried, Germany. Before sequencing at IMGM, the
+second PCR amplification of the two-step PCR procedure was performed
+using primers targeting the tagged region incorporated in the first PCR.
+In addition, these primers contained adapter and sample-specific index
+sequences. The second PCR was carried out for 8 cycles. Beside samples,
+a positive and negative control were sequenced. A negative control was
+comprised of four parallel PCR reactions without DNA template, while for
+a positive control a mock community composed of evenly mixed DNA
+material originating from 20 bacterial strains (ATCC MSA-1002, ATCC,
+USA) was used. Partial 16S rRNA sequences obtained in this study have
+been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI
+under accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794,
 SAMEA6648809 -- SAMEA6648811.
 
-Obtained sequences were analyzed on the computer cluster Isabella
+Obtained sequences were analysed on the computer cluster Isabella
 (University Computing Center, University of Zagreb) using mothur
@@
 
-Obtained sequences were analyzed on the Life Science Compute Cluster
+Obtained sequences were analysed on the Life Science Compute Cluster
 (LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
@@
 \emph{et al.}, 2015) under default settings. Putative genes were
-predicted from contings longer than 200 bp using Prodigal (version
-2.6.3) (Hyatt \emph{et al.}, 2010) in metagenome mode (-p meta).
-Abundances of predicted genes were expressed as Reads Per Kilobase
-Million (RPKM) and calculated using the BWA algorithm (version 0.7.16a)
-(Li and Durbin, 2009). All predicted genes were functionally annotated
-using the eggNOG-mapper (Huerta-Cepas \emph{et al.}, 2017) and eggNOG
-database (version 5.0) (Huerta-Cepas \emph{et al.}, 2019). Sequence
-taxonomy classification was determined using the lowest common ancestor
-algorithm adapted from DIAMOND (version 0.8.36) (Buchfink \emph{et al.},
-2015) and by searching against the NCBI non-redundant database (NR). To
-determine the phylogeny, the top 10 \si{\percent} hits with an e-value
-\textless{} 1 × 10\textsuperscript{-5} were used (-\/-top 10). Sequence
-renaming, coverage information computing and metagenomic statistics
-calculations were performed using software tools from BBTools
+predicted from contigs longer than 200 bp using Prodigal (version 2.6.3)
+(Hyatt \emph{et al.}, 2010) in metagenome mode (-p meta). Abundances of
+predicted genes were expressed as Reads Per Kilobase Million (RPKM) and
+calculated using the BWA algorithm (version 0.7.16a) (Li and Durbin,
+2009). All predicted genes were functionally annotated using the
+eggNOG-mapper (Huerta-Cepas \emph{et al.}, 2017) and eggNOG database
+(version 5.0) (Huerta-Cepas \emph{et al.}, 2019). Sequence taxonomy
+classification was determined using the lowest common ancestor algorithm
+adapted from DIAMOND (version 0.8.36) (Buchfink \emph{et al.}, 2015) and
+by searching against the NCBI non-redundant database (NR). To determine
+the phylogeny, the top 10 \si{\percent} hits with an e-value \textless{}
+1 × 10\textsuperscript{-5} were used (-\/-top 10). Sequence renaming,
+coverage information computing and metagenomic statistics calculations
+were performed using software tools from BBTools
 (\url{https://jgi.doe.gov/data-and-tools/bbtools}). Metagenomic
@@
 Metabolomics Center (University of Vienna) for metaproteomic analysis.
-Peptides were resuspended in 1 \si{\percent} (final concentration) TFA,
+Peptides were re-suspended in 1 \si{\percent} (final concentration) TFA,
 desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA)
@@
 seven times in deionized water and fixed with formaldehyde (final
-concentration \textasciitilde{} 3 \si{\percent}). In addition,
-nontreated leaves and thalli, washed three times in ASW to remove
+concentration \textasciitilde{}3 \si{\percent}). In addition,
+non-treated leaves and thalli, washed three times in ASW to remove
 loosely attached microbial cells, were fixed in the same concentration
@@
 Li, H. and Durbin, R. (2009) Fast and accurate short read alignment with
-BurrowsWheeler transform. \emph{Bioinformatics} \textbf{25}: 1754--1760.
+Burrows--Wheeler transform. \emph{Bioinformatics} \textbf{25}:
+1754--1760.
 
@@
 Parco, Y. (2003) On the identity and origin of the Mediterranean
-invasive \emph{Caulerpa} \emph{Racemosa} (Caulerpales, Chlorophyta).
-\emph{Eur J Phycol} \textbf{38}: 325--339.
+invasive \emph{Caulerpa racemosa} (Caulerpales, Chlorophyta). \emph{Eur
+J Phycol} \textbf{38}: 325--339.
 
@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\textit{C.
+\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$1 \si{\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\textit{C.
 nodosa} and \textit{C.

### 2020-11-19 submission/supplementary.tex
Format customization.

@@
 
-\hypertarget{supplementary-figures}{%
-\subsection{Supplementary Figures}\label{supplementary-figures}}
+\hypertarget{supplementary-figure}{%
+\subsection{Supplementary figure}\label{supplementary-figure}}
 
@@
 
-\hypertarget{supplementary-table}{%
-\subsection{Supplementary Table}\label{supplementary-table}}
+\hypertarget{supplementary-tables}{%
+\subsection{Supplementary tables}\label{supplementary-tables}}
 
@@
 \begin{longtable}[t]{>{\centering\arraybackslash}p{5em}ccccccccc}
-\caption{\label{tab:metagenomic_statistics}Sample ID, sampling station, community type, sampling date, number of raw sequence pairs, number of assembled contigs by MEGAHIT, N50 and L50 assembly statistics, number of predicted coding sequences (CDS) by Prodigal and number of eggNOGmapper annotated CDS.\label{metagenomic_statistics}}\\
+\caption{\label{tab:metagenomic_statistics}Sample ID, sampling station, community type, sampling date, number of raw sequence pairs, number of assembled contigs by MEGAHIT, N50 and L50 assembly statistics, number of predicted coding sequences (CDS) by Prodigal and number of eggNOG-mapper annotated CDS.\label{metagenomic_statistics}}\\
 \toprule
@@
 \endfirsthead
-\caption[]{Sample ID, sampling station, community type, sampling date, number of raw sequence pairs, number of assembled contigs by MEGAHIT, N50 and L50 assembly statistics, number of predicted coding sequences (CDS) by Prodigal and number of eggNOGmapper annotated CDS.\label{metagenomic_statistics} \textit{(continued)}}\\
+\caption[]{Sample ID, sampling station, community type, sampling date, number of raw sequence pairs, number of assembled contigs by MEGAHIT, N50 and L50 assembly statistics, number of predicted coding sequences (CDS) by Prodigal and number of eggNOG-mapper annotated CDS.\label{metagenomic_statistics} \textit{(continued)}}\\
 \toprule
@@
 \begin{longtabu} to \linewidth {>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
-\caption{\label{tab:metagenomic_taxonomy}Phyla into which CDS were classifed, number and proportion of CDS in different phyla and sum of coding sequences' RPKM (Reads Per Kilobase Million) and their proportion in different phyla.
+\caption{\label{tab:metagenomic_taxonomy}Phyla into which CDS were classified, number and proportion of CDS in different phyla and sum of coding sequences' RPKM (Reads Per Kilobase Million) and their proportion in different phyla.
 Data are derived from sequenced metagenomes.
-Each metagenomic sample is labeled with sampling station, community type, sampling date and sample ID.
+Each metagenomic sample is labelled with sampling station, community type, sampling date and sample ID.
 For each sample top ten phyla based on RPKM were selected.
@@
 \endfirsthead
-\caption[]{Phyla into which CDS were classifed, number and proportion of CDS in different phyla and sum of coding sequences' RPKM (Reads Per Kilobase Million) and their proportion in different phyla.
+\caption[]{Phyla into which CDS were classified, number and proportion of CDS in different phyla and sum of coding sequences' RPKM (Reads Per Kilobase Million) and their proportion in different phyla.
 Data are derived from sequenced metagenomes.
-Each metagenomic sample is labeled with sampling station, community type, sampling date and sample ID.
+Each metagenomic sample is labelled with sampling station, community type, sampling date and sample ID.
 For each sample top ten phyla based on RPKM were selected.

### 2020-11-20 submission/manuscript.Rmd
References customization.
Formatting text.
Units format.

@@
 After a brief vortex, the lysate was transferred to a clean tube separating the host leaves or thalli from the mixture containing lysed epiphytic cells.
-Dithiothreitol (DTT; final concentration 24 \si{\milli\Molar}) was added and proteins were precipitated with chilled 100 \si{\percent} trichloroacetic acid (TCA; final concentration 20 \si{\percent}) overnight at \num{-20} °C.
+Dithiothreitol (DTT; final concentration 24 \si{\milli\Molar}) was added and proteins were precipitated with chilled 100 \si{\percent} trichloroacetic acid (TCA; final concentration 20 \si{\percent}) overnight at \num{-20} \si{\degreeCelsius}.
 Precipitated proteins were centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
@@
 After each washing step, samples were centrifuged at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}.
-Dried pellets were stored at \num{-80} °C until further analysis.
+Dried pellets were stored at \num{-80} \si{\degreeCelsius} until further analysis.
 
@@
 ## References
+\interlinepenalty=10000
+\setlength{\emergencystretch}{3.5em}
 <div id="refs"></div>
@@
 
-## Figure Captions
+## Figure captions
 **\autoref{confocal_december}.** \nameref{confocal_december}

### 2020-11-20 submission/manuscript.tex
References customization.
Formatting text.
Unit format rendering.

@@
 precipitated with chilled 100 \si{\percent} trichloroacetic acid (TCA;
-final concentration 20 \si{\percent}) overnight at \num{-20} °C.
-Precipitated proteins were centrifuged at 10,000 × g and 4
-\si{\degreeCelsius} for 40 \si{\minute}. The obtained protein pellet was
-washed three times with chilled acetone. During the first washing step
-the pellet was transferred to a clean 1.5 \si{\ml} tube. After each
-washing step, samples were centrifuged at 20,000 × g and 4
+final concentration 20 \si{\percent}) overnight at \num{-20}
+\si{\degreeCelsius}. Precipitated proteins were centrifuged at 10,000 ×
+g and 4 \si{\degreeCelsius} for 40 \si{\minute}. The obtained protein
+pellet was washed three times with chilled acetone. During the first
+washing step the pellet was transferred to a clean 1.5 \si{\ml} tube.
+After each washing step, samples were centrifuged at 20,000 × g and 4
 \si{\degreeCelsius} for 5 \si{\minute}. Dried pellets were stored at
-\num{-80} °C until further analysis.
+\num{-80} \si{\degreeCelsius} until further analysis.
 
@@
 
+\interlinepenalty=10000 \setlength{\emergencystretch}{3.5em}
+
 \hypertarget{refs}{}
@@
 \hypertarget{figure-captions}{%
-\subsection{Figure Captions}\label{figure-captions}}
+\subsection{Figure captions}\label{figure-captions}}
 

### 2020-11-21 submission/manuscript.Rmd
Text formatting.

@@
 \newpage
-```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$1 \\si{\\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\\textit{C.
+```{r out.width="100%", fig.align="center", fig.cap="Taxonomic classification and relative contribution of the most abundant ($\\geq$ 1 \\si{\\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\\textit{C.
 nodosa} and \\textit{C.

### 2020-11-21 submission/manuscript.tex
Text formatting.

@@
 
-\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$1 \si{\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\textit{C.
+\caption{Taxonomic classification and relative contribution of the most abundant ($\geq$ 1 \si{\percent}) bacterial and archaeal sequences from surfaces of two marine macrophytes (\textit{C.
 nodosa} and \textit{C.

### 2020-11-21 submission/manuscript.Rmd
Text formatting.

@@
 The target-decoy approach was used to reduce the probability of false peptide identification.
-Results whose false discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
+Results whose false discovery rate at the peptide level was < \SI{1}{\percent} were kept.
 For protein identification a minimum of two peptides and one unique peptide were required.
@@
 
-## Figure captions
+## Figure legends
 **\autoref{confocal_december}.** \nameref{confocal_december}

### 2020-11-21 submission/supplementary.Rmd

@@
 
-## Supplementary Figures
+## Supplementary figure
 
@@
 \newpage
-## Supplementary Table
+## Supplementary tables
 
@@
 
-# Loading sequence statistics following eggNOG mapper
+# Loading sequence statistics following eggNOG-mapper
 eggnog_45 <- read_tsv("results/metagenomics/statistics/07eggnog_45.txt", col_names=FALSE)
@@
 of\nPredicted\nCDS", "No.
-of\nAnnotated\nCDS"), align="c"), caption="Sample ID, sampling station, community type, sampling date, number of raw sequence pairs, number of assembled contigs by MEGAHIT, N50 and L50 assembly statistics, number of predicted coding sequences (CDS) by Prodigal and number of eggNOGmapper annotated CDS.\\label{metagenomic_statistics}", align="c") %>%
+of\nAnnotated\nCDS"), align="c"), caption="Sample ID, sampling station, community type, sampling date, number of raw sequence pairs, number of assembled contigs by MEGAHIT, N50 and L50 assembly statistics, number of predicted coding sequences (CDS) by Prodigal and number of eggNOG-mapper annotated CDS.\\label{metagenomic_statistics}", align="c") %>%
   kable_styling(latex_options=c("hold_position", "repeat_header"), font_size=9) %>%
@@
 kable(format="latex", booktabs=TRUE, linesep="", longtable=TRUE, escape=FALSE, col.names=linebreak(c("Phylum", "No.
-of CDS", "CDS (\\%)", "Summed RPKM", "RPKM (\\%)"), align="c"), caption="Phyla into which CDS were classifed, number and proportion of CDS in different phyla and sum of coding sequences' RPKM (Reads Per Kilobase Million) and their proportion in different phyla.
+of CDS", "CDS (\\%)", "Summed RPKM", "RPKM (\\%)"), align="c"), caption="Phyla into which CDS were classified, number and proportion of CDS in different phyla and sum of coding sequences' RPKM (Reads Per Kilobase Million) and their proportion in different phyla.
 Data are derived from sequenced metagenomes.
-Each metagenomic sample is labeled with sampling station, community type, sampling date and sample ID.
+Each metagenomic sample is labelled with sampling station, community type, sampling date and sample ID.
 For each sample top ten phyla based on RPKM were selected.

### 2021-01-18 submission/manuscript.Rmd

@@
 A negative control was comprised of four parallel PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
-Partial 16S rRNA sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 -- SAMEA6648811.
+Partial 16S rRNA sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 -- SAMEA6648811 and SAMEA6648824.
 

### 2021-02-09 submission/manuscript.Rmd
Fromat text.

@@
 In addition, we tested the removal efficiency of the protocol and the suitability of obtained DNA and proteins for 16S rRNA sequencing, metagenomics and metaproteomics.
+
+\newpage
+## Materials and methods
+### Sampling
+Leaves of *C.
+nodosa* were sampled in a *C.
+nodosa* meadow in the Bay of Saline, northern Adriatic Sea (\ang{45;7;5} N, \ang{13;37;20} E) and in a *C.
+nodosa* meadow invaded by *C.
+cylindracea* in the Bay of Funtana, northern Adriatic Sea (\ang{45;10;39} N, \ang{13;35;42} E).
+Thalli of *C.
+cylindracea* were sampled in the same *C.
+nodosa* invaded meadow in the Bay of Funtana and at a locality of only *C.
+cylindracea* located in the proximity of the invaded meadow.
+Leaves and thalli for 16S rRNA analysis, metagenomics and metaproteomics were collected in two contrasting seasons, on 4 December 2017 (16S rRNA analysis and metaproteomics), 14 December 2017 (metagenomics) and 18 June 2018 (16S rRNA analysis, metagenomics and metaproteomics).
+During spring 2018, the *C.
+nodosa* meadow in the Bay of Saline decayed to an extent that no leaves could be retrieved [@Najdek2020].
+In addition, as not enough DNA for both metagenomic and 16S RNA analysis were obtained during the sampling on 4 December 2017, an additional sampling on 14 December 2017 was carried out in the Bay of Funtana.
+Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with seawater from this site.
+Upon arrival to the laboratory, *C.
+nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *C.
+cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections.
+Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.
+
+### DNA isolation
+The DNA was isolated according to the protocol for isolation from filters described in @Massana1997.
+This protocol was modified and adapted for microbial DNA isolation from macrophyte surfaces as described below.
+Five (5) \si{\ml} of lysis buffer (40 \si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar} sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or 2 \si{\g} wet-weight of thalli.
+For every sample, duplicate isolations were performed.
+Lysozyme was added (final concentration 1 \si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius} for 30 \si{\minute}.
+Subsequently, proteinase K (final concentration 0.5 \si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were added and the samples were incubated at 55 \si{\degreeCelsius} for 2 \si{\hour}.
+Following the incubation, tubes were vortexed for 10 \si{\minute} and the mixture containing lysed epiphytic cells was separated from host leaves or thalli by transferring the solution into a clean tube.
+The lysate was extracted twice with a mixture of phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with chloroform:isoamyl alcohol (24:1).
+After each addition of an organic solvent mixture, tubes were slightly vortexed and centrifuged at 4,500 × g for 10 \si{\minute}.
+Following each centrifugation the aqueous phases were retrieved.
+After the final extraction 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
+DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 16,000 × g and 4 \si{\degreeCelsius} for 20 \si{\minute}.
+The pellet was washed twice with 1 \si{\ml} of chilled 70 \si{\percent} ethanol and centrifuged after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 10 \si{\minute}.
+After the first washing step duplicate pellets from the same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.
+The dried pellet was re-suspended in 100 \si{\ul} of deionized water.
+
+### Illumina 16S rRNA sequencing
+An aliquot of isolated DNA was treated with RNase A (final concentration 200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}.
+The DNA concentration was determined using the Quant-iT PicoGreen dsDNA Assay Kit (Invitrogen, USA) according to the manufacturer’s instructions and diluted to 1 \si{\ng\per\ul}.
+The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
+In the first PCR the 515F ($5'$-GTGYCAGCMGCCGCGGTAA-$3'$) and 806R ($5'$-GGACTACNVGGGTWTCTAAT-$3'$) primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016].
+These primers contained on their $5'$ end a tagged sequence.
+Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template.
+Cycling conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}.
+The four parallel reactions volumes were pooled and PCR products were purified using the GeneJET PCR Purification Kit (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and following the protocol that included isopropanol addition for better small DNA fragment yield.
+The column was eluted in 30 \si{\ul} of deionized water.
+Purified PCR products were sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories, Martinsried, Germany.
+Before sequencing at IMGM, the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR.
+In addition, these primers contained adapter and sample-specific index sequences.
+The second PCR was carried out for 8 cycles.
+Beside samples, a positive and negative control were sequenced.
+A negative control was comprised of four parallel PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
+Partial 16S rRNA sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 -- SAMEA6648811 and SAMEA6648824.
+
+```{r error_negative_control}
+# Calculation of the error rate based on the ATCC MSA-1002 mock community
+error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
+  filter(numparents==1) %>%
+  mutate(mismatches_weight=weight*mismatches) %>%
+  mutate(total_weight=weight*total)
+error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
+
+# Calculation of the number of sequences in the negative control
+nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
+  filter(taxon=="Root") %>%
+  select(NC_1)
+```
+
+Obtained sequences were analysed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
+For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
+Sequences classified as chloroplasts by SILVA were exported and reclassified using mothur and the RDP (http://rdp.cme.msu.edu/) training set (version 16) reference files adapted for mothur [@Cole2014].
+In comparison to SILVA, RDP allows a more detailed classification of chloroplast sequences.
+Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
+In addition, the negative control processed together with the samples yielded only `r nc` sequences after sequence quality curation.
+
+### Metagenomics
+Four DNA samples were selected and sent on dry ice to IMGM Laboratories, Martinsried, Germany for metagenomic sequencing.
+DNA was purified using AMPure XP Beads (Beckman Coulter, USA) applying a bead:DNA ratio of 1:1 (v/v), quantified with a Qubit dsDNA HS Assay Kit (Thermo Fisher Scientific, USA) and check for integrity on a 1 \si{\percent} agarose gel.
+Metagenomic sequencing libraries were generated from 300 \si{\ng} of input DNA using a NEBNext Ultra II FS DNA Library Prep Kit for Illumina (New England Biolabs, USA) according to the manufacturer's instructions.
+Fragments were selected (500 -- 700 bp) using AMPure XP Beads, PCR enriched for 3 -- 5 cycles and quality controlled.
+Libraries generated from different DNA samples were pooled and sequenced on an Illumina NovaSeq 6000 sequencing system (2 × 250 bp).
+
+Obtained sequences were analysed on the Life Science Compute Cluster (LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
+Individual sequences were assembled using MEGAHIT (version  1.1.2) [@Li2015] under default settings.
+Putative genes were predicted from contigs longer than 200 bp using Prodigal (version  2.6.3) [@Hyatt2010] in metagenome mode (-p meta).
+Abundances of predicted genes were expressed as Reads Per Kilobase Million (RPKM) and calculated using the BWA algorithm (version  0.7.16a) [@Li2009a].
+All predicted genes were functionally annotated using the eggNOG-mapper [@Huerta-Cepas2017] and eggNOG database (version 5.0) [@Huerta-Cepas2019].
+Sequence taxonomy classification was determined using the lowest common ancestor algorithm adapted from DIAMOND (version 0.8.36) [@Buchfink2015] and by searching against the NCBI non-redundant database (NR).
+To determine the phylogeny, the top 10 \si{\percent} hits with an e-value < 1 × 10^-5^ were used (\-\-top 10).
+Sequence renaming, coverage information computing and metagenomic statistics calculations were performed using software tools from BBTools (https://jgi.doe.gov/data-and-tools/bbtools).
+Metagenomic sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers SAMEA6648795, SAMEA6648797, SAMEA6648809 and SAMEA6648811.
+
+### Protein isolation
+Proteins were isolated according to the protocol for protein isolation from soil described in @Chourey2010 and modified by @Hultman2015.
+This protocol was further modified and adapted for microbial protein isolation from macrophyte surfaces as described below.
+Twenty (20) \si{\ml} of protein extraction buffer (4 \si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl [pH 8.0]) was added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of thalli.
+The mixture was incubated in boiling water for 5 \si{\minute}, vortexed for 10 \si{\minute} and incubated again in boiling water for 5 \si{\minute}.
+After a brief vortex, the lysate was transferred to a clean tube separating the host leaves or thalli from the mixture containing lysed epiphytic cells.
+Dithiothreitol (DTT; final concentration 24 \si{\milli\Molar}) was added and proteins were precipitated with chilled 100 \si{\percent} trichloroacetic acid (TCA; final concentration 20 \si{\percent}) overnight at \num{-20} \si{\degreeCelsius}.
+Precipitated proteins were centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
+The obtained protein pellet was washed three times with chilled acetone.
+During the first washing step the pellet was transferred to a clean 1.5 \si{\ml} tube.
+After each washing step, samples were centrifuged at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}.
+Dried pellets were stored at \num{-80} \si{\degreeCelsius} until further analysis.
+
+### Metaproteomics
+Isolated proteins were whole trypsin digested using the FASP (filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK) according to the manufacturer's instructions with small modifications [@Wisniewski2009].
+Prior to loading the solution onto the column, protein pellets were solubilized in a urea sample buffer included in the kit amended with DTT (final concentration 100 \si{\milli\Molar}) for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to remove larger particles.
+The first washing step after protein solution loading was repeated twice.
+In addition, the centrifugation steps were prolonged if the column was clogged.
+Trypsin digestion was performed on column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
+The final filtrate containing peptides was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid (TFA), freezed at \num{-80} \si{\degreeCelsius}, lyophilized and sent to VIME -- Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis.
+Peptides were re-suspended in 1 \si{\percent} (final concentration) TFA, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
+Obtained MS/MS spectra were searched against a protein database composed of combined sequenced metagenomes using SEQUEST-HT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
+The target-decoy approach was used to reduce the probability of false peptide identification.
+Results whose false discovery rate at the peptide level was < \SI{1}{\percent} were kept.
+For protein identification a minimum of two peptides and one unique peptide were required.
+For protein quantification, a chromatographic peak area-based free quantitative method was applied.
+
+### Data processing and visualization
+Processing and visualization  of 16S rRNA, metagenomic and metaproteomic data were done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020; @Wilke2018].
+The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/MicrobesRovinj/Korlevic_SelectiveRemoval_FrontMicrobiol_2021).
+
+### Confocal microscopy
+Host leaves and thalli from DNA and protein isolation steps were washed seven times in deionized water and fixed with formaldehyde (final concentration ~3 \si{\percent}).
+In addition, non-treated leaves and thalli, washed three times in ASW to remove loosely attached microbial cells, were fixed in the same concentration of formaldehyde and used as a positive control.
+For-long term storage, fixed leaves and thalli were immersed in a mixture of phosphate-buffered saline (PBS) and ethanol (1:1) and stored at \num{-20} \si{\degreeCelsius}.
+Treated and untreated segments of leaves and thalli were stained in a 2 × solution of SYBR Green I and examined under a Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
 
@@
 
-\newpage
-## Experimental procedures
-### Sampling
-Leaves of *C.
-nodosa* were sampled in a *C.
-nodosa* meadow in the Bay of Saline, northern Adriatic Sea (\ang{45;7;5} N, \ang{13;37;20} E) and in a *C.
-nodosa* meadow invaded by *C.
-cylindracea* in the Bay of Funtana, northern Adriatic Sea (\ang{45;10;39} N, \ang{13;35;42} E).
-Thalli of *C.
-cylindracea* were sampled in the same *C.
-nodosa* invaded meadow in the Bay of Funtana and at a locality of only *C.
-cylindracea* located in the proximity of the invaded meadow.
-Leaves and thalli for 16S rRNA analysis, metagenomics and metaproteomics were collected in two contrasting seasons, on 4 December 2017 (16S rRNA analysis and metaproteomics), 14 December 2017 (metagenomics) and 18 June 2018 (16S rRNA analysis, metagenomics and metaproteomics).
-During spring 2018, the *C.
-nodosa* meadow in the Bay of Saline decayed to an extent that no leaves could be retrieved [@Najdek2020].
-In addition, as not enough DNA for both metagenomic and 16S RNA analysis were obtained during the sampling on 4 December 2017, an additional sampling on 14 December 2017 was carried out in the Bay of Funtana.
-Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with seawater from this site.
-Upon arrival to the laboratory, *C.
-nodosa* leaves were cut into sections of 1 -- 2 \si{\cm}, while *C.
-cylindracea* thalli were cut into 5 -- 8 \si{\cm} long sections.
-Leaves and thalli were washed three times with sterile artificial seawater (ASW) to remove loosely attached microbial cells.
-
-### DNA isolation
-The DNA was isolated according to the protocol for isolation from filters described in @Massana1997.
-This protocol was modified and adapted for microbial DNA isolation from macrophyte surfaces as described below.
-Five (5) \si{\ml} of lysis buffer (40 \si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar} sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or 2 \si{\g} wet-weight of thalli.
-For every sample, duplicate isolations were performed.
-Lysozyme was added (final concentration 1 \si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius} for 30 \si{\minute}.
-Subsequently, proteinase K (final concentration 0.5 \si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were added and the samples were incubated at 55 \si{\degreeCelsius} for 2 \si{\hour}.
-Following the incubation, tubes were vortexed for 10 \si{\minute} and the mixture containing lysed epiphytic cells was separated from host leaves or thalli by transferring the solution into a clean tube.
-The lysate was extracted twice with a mixture of phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with chloroform:isoamyl alcohol (24:1).
-After each addition of an organic solvent mixture, tubes were slightly vortexed and centrifuged at 4,500 × g for 10 \si{\minute}.
-Following each centrifugation the aqueous phases were retrieved.
-After the final extraction 1/10 of chilled 3 \si{\Molar} sodium acetate (pH 5.2) was added.
-DNA was precipitated by adding 1 volume of chilled isopropanol, incubating the mixtures overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 16,000 × g and 4 \si{\degreeCelsius} for 20 \si{\minute}.
-The pellet was washed twice with 1 \si{\ml} of chilled 70 \si{\percent} ethanol and centrifuged after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 10 \si{\minute}.
-After the first washing step duplicate pellets from the same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.
-The dried pellet was re-suspended in 100 \si{\ul} of deionized water.
-
-### Illumina 16S rRNA sequencing
-An aliquot of isolated DNA was treated with RNase A (final concentration 200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}.
-The DNA concentration was determined using the Quant-iT PicoGreen dsDNA Assay Kit (Invitrogen, USA) according to the manufacturer’s instructions and diluted to 1 \si{\ng\per\ul}.
-The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
-In the first PCR the 515F ($5'$-GTGYCAGCMGCCGCGGTAA-$3'$) and 806R ($5'$-GGACTACNVGGGTWTCTAAT-$3'$) primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016].
-These primers contained on their $5'$ end a tagged sequence.
-Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template.
-Cycling conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}.
-The four parallel reactions volumes were pooled and PCR products were purified using the GeneJET PCR Purification Kit (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and following the protocol that included isopropanol addition for better small DNA fragment yield.
-The column was eluted in 30 \si{\ul} of deionized water.
-Purified PCR products were sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories, Martinsried, Germany.
-Before sequencing at IMGM, the second PCR amplification of the two-step PCR procedure was performed using primers targeting the tagged region incorporated in the first PCR.
-In addition, these primers contained adapter and sample-specific index sequences.
-The second PCR was carried out for 8 cycles.
-Beside samples, a positive and negative control were sequenced.
-A negative control was comprised of four parallel PCR reactions without DNA template, while for a positive control a mock community composed of evenly mixed DNA material originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was used.
-Partial 16S rRNA sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 -- SAMEA6648811 and SAMEA6648824.
-
-```{r error_negative_control}
-# Calculation of the error rate based on the ATCC MSA-1002 mock community
-error <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.error.summary", na="") %>%
-  filter(numparents==1) %>%
-  mutate(mismatches_weight=weight*mismatches) %>%
-  mutate(total_weight=weight*total)
-error <- sum(error$mismatches_weight)/sum(error$total_weight)*100
-
-# Calculation of the number of sequences in the negative control
-nc <- read_tsv("data/mothur/raw.trim.contigs.good.unique.good.filter.unique.precluster.pick.nr_v138.wang.tax.summary") %>%
-  filter(taxon=="Root") %>%
-  select(NC_1)
-```
-
-Obtained sequences were analysed on the computer cluster Isabella (University Computing Center, University of Zagreb) using mothur (version 1.43.0) [@Schloss2009] according to the MiSeq Standard Operating Procedure (MiSeq SOP; https://mothur.org/wiki/MiSeq_SOP) [@Kozich2013] and recommendations given from the Riffomonas project to enhance data reproducibility (http://www.riffomonas.org/).
-For alignment and classification of sequences the SILVA SSU Ref NR 99 database (release 138; http://www.arb-silva.de) was used [@Quast2013; @Yilmaz2014].
-Sequences classified as chloroplasts by SILVA were exported and reclassified using mothur and the RDP (http://rdp.cme.msu.edu/) training set (version 16) reference files adapted for mothur [@Cole2014].
-In comparison to SILVA, RDP allows a more detailed classification of chloroplast sequences.
-Based on the ATCC MSA-1002 mock community included in the analysis a sequencing error rate of `r error` \si{\percent} was determined, which is in line with previously reported values for next-generation sequencing data [@Kozich2013; @Schloss2016].
-In addition, the negative control processed together with the samples yielded only `r nc` sequences after sequence quality curation.
-
-### Metagenomics
-Four DNA samples were selected and sent on dry ice to IMGM Laboratories, Martinsried, Germany for metagenomic sequencing.
-DNA was purified using AMPure XP Beads (Beckman Coulter, USA) applying a bead:DNA ratio of 1:1 (v/v), quantified with a Qubit dsDNA HS Assay Kit (Thermo Fisher Scientific, USA) and check for integrity on a 1 \si{\percent} agarose gel.
-Metagenomic sequencing libraries were generated from 300 \si{\ng} of input DNA using a NEBNext Ultra II FS DNA Library Prep Kit for Illumina (New England Biolabs, USA) according to the manufacturer's instructions.
-Fragments were selected (500 -- 700 bp) using AMPure XP Beads, PCR enriched for 3 -- 5 cycles and quality controlled.
-Libraries generated from different DNA samples were pooled and sequenced on an Illumina NovaSeq 6000 sequencing system (2 × 250 bp).
-
-Obtained sequences were analysed on the Life Science Compute Cluster (LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
-Individual sequences were assembled using MEGAHIT (version  1.1.2) [@Li2015] under default settings.
-Putative genes were predicted from contigs longer than 200 bp using Prodigal (version  2.6.3) [@Hyatt2010] in metagenome mode (-p meta).
-Abundances of predicted genes were expressed as Reads Per Kilobase Million (RPKM) and calculated using the BWA algorithm (version  0.7.16a) [@Li2009a].
-All predicted genes were functionally annotated using the eggNOG-mapper [@Huerta-Cepas2017] and eggNOG database (version 5.0) [@Huerta-Cepas2019].
-Sequence taxonomy classification was determined using the lowest common ancestor algorithm adapted from DIAMOND (version 0.8.36) [@Buchfink2015] and by searching against the NCBI non-redundant database (NR).
-To determine the phylogeny, the top 10 \si{\percent} hits with an e-value < 1 × 10^-5^ were used (\-\-top 10).
-Sequence renaming, coverage information computing and metagenomic statistics calculations were performed using software tools from BBTools (https://jgi.doe.gov/data-and-tools/bbtools).
-Metagenomic sequences obtained in this study have been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers SAMEA6648795, SAMEA6648797, SAMEA6648809 and SAMEA6648811.
-
-### Protein isolation
-Proteins were isolated according to the protocol for protein isolation from soil described in @Chourey2010 and modified by @Hultman2015.
-This protocol was further modified and adapted for microbial protein isolation from macrophyte surfaces as described below.
-Twenty (20) \si{\ml} of protein extraction buffer (4 \si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl [pH 8.0]) was added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of thalli.
-The mixture was incubated in boiling water for 5 \si{\minute}, vortexed for 10 \si{\minute} and incubated again in boiling water for 5 \si{\minute}.
-After a brief vortex, the lysate was transferred to a clean tube separating the host leaves or thalli from the mixture containing lysed epiphytic cells.
-Dithiothreitol (DTT; final concentration 24 \si{\milli\Molar}) was added and proteins were precipitated with chilled 100 \si{\percent} trichloroacetic acid (TCA; final concentration 20 \si{\percent}) overnight at \num{-20} \si{\degreeCelsius}.
-Precipitated proteins were centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40 \si{\minute}.
-The obtained protein pellet was washed three times with chilled acetone.
-During the first washing step the pellet was transferred to a clean 1.5 \si{\ml} tube.
-After each washing step, samples were centrifuged at 20,000 × g and 4 \si{\degreeCelsius} for 5 \si{\minute}.
-Dried pellets were stored at \num{-80} \si{\degreeCelsius} until further analysis.
-
-### Metaproteomics
-Isolated proteins were whole trypsin digested using the FASP (filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK) according to the manufacturer's instructions with small modifications [@Wisniewski2009].
-Prior to loading the solution onto the column, protein pellets were solubilized in a urea sample buffer included in the kit amended with DTT (final concentration 100 \si{\milli\Molar}) for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to remove larger particles.
-The first washing step after protein solution loading was repeated twice.
-In addition, the centrifugation steps were prolonged if the column was clogged.
-Trypsin digestion was performed on column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
-The final filtrate containing peptides was acidified with 1 \si{\percent} (final concentration) trifluoroacetic acid (TFA), freezed at \num{-80} \si{\degreeCelsius}, lyophilized and sent to VIME -- Vienna Metabolomics Center (University of Vienna) for metaproteomic analysis.
-Peptides were re-suspended in 1 \si{\percent} (final concentration) TFA, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
-Obtained MS/MS spectra were searched against a protein database composed of combined sequenced metagenomes using SEQUEST-HT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
-The target-decoy approach was used to reduce the probability of false peptide identification.
-Results whose false discovery rate at the peptide level was < \SI{1}{\percent} were kept.
-For protein identification a minimum of two peptides and one unique peptide were required.
-For protein quantification, a chromatographic peak area-based free quantitative method was applied.
-
-### Data processing and visualization
-Processing and visualization  of 16S rRNA, metagenomic and metaproteomic data were done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020; @Wilke2018].
-The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/MicrobesRovinj/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020).
-
-### Confocal microscopy
-Host leaves and thalli from DNA and protein isolation steps were washed seven times in deionized water and fixed with formaldehyde (final concentration ~3 \si{\percent}).
-In addition, non-treated leaves and thalli, washed three times in ASW to remove loosely attached microbial cells, were fixed in the same concentration of formaldehyde and used as a positive control.
-For-long term storage, fixed leaves and thalli were immersed in a mixture of phosphate-buffered saline (PBS) and ethanol (1:1) and stored at \num{-20} \si{\degreeCelsius}.
-Treated and untreated segments of leaves and thalli were stained in a 2 × solution of SYBR Green I and examined under a Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
-
 ## Acknowledgements

### 2021-02-09 submission/manuscript.tex
Text formatting.
Fromat text.

@@
 Surfaces of marine macrophytes are colonized by a diverse microbial
-community whose structure and function are poorly understood (Egan
-\emph{et al.}, 2013). As less than 1 \% of all prokaryotic species are
-culturable, molecular methods such as 16S rRNA sequencing, metagenomics
-and metaproteomics are indispensable to study these organisms (Amann
-\emph{et al.}, 1995; Su \emph{et al.}, 2012). Applying these techniques
-requires an initial isolation step with the purpose of obtaining high
-quality DNA and proteins.
+community whose structure and function are poorly understood (Egan et
+al., 2013). As less than 1 \% of all prokaryotic species are culturable,
+molecular methods such as 16S rRNA sequencing, metagenomics and
+metaproteomics are indispensable to study these organisms (Amann et al.,
+1995; Su et al., 2012). Applying these techniques requires an initial
+isolation step with the purpose of obtaining high quality DNA and
+proteins.
 
@@
 communities is usually isolated by collecting cells onto filters and
-subsequently isolating the target organisms or communities (Gilbert
-\emph{et al.}, 2009). If a specific microbial size fraction is aimed
-sequential filtration is applied (Massana \emph{et al.}, 1997; Andersson
-\emph{et al.}, 2010). In contrast, obtaining microorganisms associated
-to surfaces require either a cell detachment procedure prior to
-isolation or the host material is co-extracted with the target material.
-Methods for separating microbial cells from the host include shaking of
-host tissue (Gross \emph{et al.}, 2003; Nõges \emph{et al.}, 2010),
-scraping of macrophyte surfaces (Uku \emph{et al.}, 2007) or applying
-ultrasonication (Weidner \emph{et al.}, 1996; Cai \emph{et al.}, 2014).
-It was shown that shaking alone is not sufficient to remove microbial
-cells from surfaces, at least not from plant root surfaces
-(Richter-Heitmann \emph{et al.}, 2016). Manual separation methods, such
-as scraping and brushing are time consuming and subjective, as the
-detachment efficiency depends on host tissue and the person performing
-the procedure (Cai \emph{et al.}, 2014). Ultrasonication was proposed as
-an alternative method as it is providing better results in terms of
-detachment efficiency (Cai \emph{et al.}, 2014; Richter-Heitmann
-\emph{et al.}, 2016). The downside of this procedure is that complete
-cell removal is still not obtained and tissue disruption was observed
-especially after the application of probe ultrasonication
-(Richter-Heitmann \emph{et al.}, 2016). An alternative to these cell
-detachment procedures is the isolation of target epiphytic compounds
-together with host material (Staufenberger \emph{et al.}, 2008; Jiang
-\emph{et al.}, 2015). This procedure can lead to problems in the
-following processing steps such as mitochondrial and chloroplast 16S
-rRNA sequence contaminations from the host (Longford \emph{et al.},
-2007; Staufenberger \emph{et al.}, 2008). In addition, when performing
-metagenomics and metaproteomics host material can cause biased results
-towards more abundant host DNA and proteins.
+subsequently isolating the target organisms or communities (Gilbert et
+al., 2009). If a specific microbial size fraction is aimed sequential
+filtration is applied (Massana et al., 1997; Andersson et al., 2010). In
+contrast, obtaining microorganisms associated to surfaces require either
+a cell detachment procedure prior to isolation or the host material is
+co-extracted with the target material. Methods for separating microbial
+cells from the host include shaking of host tissue (Gross et al., 2003;
+Nõges et al., 2010), scraping of macrophyte surfaces (Uku et al., 2007)
+or applying ultrasonication (Weidner et al., 1996; Cai et al., 2014). It
+was shown that shaking alone is not sufficient to remove microbial cells
+from surfaces, at least not from plant root surfaces (Richter-Heitmann
+et al., 2016). Manual separation methods, such as scraping and brushing
+are time consuming and subjective, as the detachment efficiency depends
+on host tissue and the person performing the procedure (Cai et al.,
+2014). Ultrasonication was proposed as an alternative method as it is
+providing better results in terms of detachment efficiency (Cai et al.,
+2014; Richter-Heitmann et al., 2016). The downside of this procedure is
+that complete cell removal is still not obtained and tissue disruption
+was observed especially after the application of probe ultrasonication
+(Richter-Heitmann et al., 2016). An alternative to these cell detachment
+procedures is the isolation of target epiphytic compounds together with
+host material (Staufenberger et al., 2008; Jiang et al., 2015). This
+procedure can lead to problems in the following processing steps such as
+mitochondrial and chloroplast 16S rRNA sequence contaminations from the
+host (Longford et al., 2007; Staufenberger et al., 2008). In addition,
+when performing metagenomics and metaproteomics host material can cause
+biased results towards more abundant host DNA and proteins.
 
@@
 epiphytic DNA isolation from the surfaces of marine macrophytes was
-described by Burke \emph{et al.} (2009). In contrast to previously
-described methods, this protocol enables an almost complete removal of
-the surface community. It was used for 16S rRNA gene clone library
-construction (Burke and Thomas \emph{et al.}, 2011) and metagenome
-sequencing (Burke and Peter Steinberg \emph{et al.}, 2011). This method,
-although providing a selective isolation procedure, uses a rapid
-multi-enzyme cleaner (3M) that is not available worldwide and the
-chemical constituents are unknown (Burke \emph{et al.}, 2009). Also to
-our knowledge, no selective isolation protocol to perform
-(meta)proteomics of epiphytic communities associated with marine
-macrophytes has been developed yet.
+described by Burke et al. (2009). In contrast to previously described
+methods, this protocol enables an almost complete removal of the surface
+community. It was used for 16S rRNA gene clone library construction
+(Burke et al., 2011b) and metagenome sequencing (Burke et al., 2011a).
+This method, although providing a selective isolation procedure, uses a
+rapid multi-enzyme cleaner (3M) that is not available worldwide and the
+chemical constituents are unknown (Burke et al., 2009). Also to our
+knowledge, no selective isolation protocol to perform (meta)proteomics
+of epiphytic communities associated with marine macrophytes has been
+developed yet.
 
 In the present study, we adapted a protocol widely used for DNA
-isolation from filters (Massana \emph{et al.}, 1997) and a protocol used
-for protein isolation from soils (Chourey \emph{et al.}, 2010; Hultman
-\emph{et al.}, 2015). These two adapted methods allowed for a selective
-extraction of DNA and proteins from epiphytic communities inhabiting the
-surfaces of two marine macrophytes, the seagrass \emph{Cymodocea nodosa}
-and the macroalga \emph{Caulerpa cylindracea}. In addition, we tested
-the removal efficiency of the protocol and the suitability of obtained
-DNA and proteins for 16S rRNA sequencing, metagenomics and
-metaproteomics.
+isolation from filters (Massana et al., 1997) and a protocol used for
+protein isolation from soils (Chourey et al., 2010; Hultman et al.,
+2015). These two adapted methods allowed for a selective extraction of
+DNA and proteins from epiphytic communities inhabiting the surfaces of
+two marine macrophytes, the seagrass \emph{Cymodocea nodosa} and the
+macroalga \emph{Caulerpa cylindracea}. In addition, we tested the
+removal efficiency of the protocol and the suitability of obtained DNA
+and proteins for 16S rRNA sequencing, metagenomics and metaproteomics.
+
+\newpage
+
+\hypertarget{materials-and-methods}{%
+\subsection{Materials and methods}\label{materials-and-methods}}
+
+\hypertarget{sampling}{%
+\subsubsection{Sampling}\label{sampling}}
+
+Leaves of \emph{C. nodosa} were sampled in a \emph{C. nodosa} meadow in
+the Bay of Saline, northern Adriatic Sea (\ang{45;7;5} N, \ang{13;37;20}
+E) and in a \emph{C. nodosa} meadow invaded by \emph{C. cylindracea} in
+the Bay of Funtana, northern Adriatic Sea (\ang{45;10;39} N,
+\ang{13;35;42} E). Thalli of \emph{C. cylindracea} were sampled in the
+same \emph{C. nodosa} invaded meadow in the Bay of Funtana and at a
+locality of only \emph{C. cylindracea} located in the proximity of the
+invaded meadow. Leaves and thalli for 16S rRNA analysis, metagenomics
+and metaproteomics were collected in two contrasting seasons, on 4
+December 2017 (16S rRNA analysis and metaproteomics), 14 December 2017
+(metagenomics) and 18 June 2018 (16S rRNA analysis, metagenomics and
+metaproteomics). During spring 2018, the \emph{C. nodosa} meadow in the
+Bay of Saline decayed to an extent that no leaves could be retrieved
+(Najdek et al., 2020). In addition, as not enough DNA for both
+metagenomic and 16S RNA analysis were obtained during the sampling on 4
+December 2017, an additional sampling on 14 December 2017 was carried
+out in the Bay of Funtana. Leaves and thalli were collected by diving
+and transported to the laboratory in containers placed on ice and filled
+with seawater from this site. Upon arrival to the laboratory, \emph{C.
+nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while \emph{C.
+cylindracea} thalli were cut into 5 -- 8 \si{\cm} long sections. Leaves
+and thalli were washed three times with sterile artificial seawater
+(ASW) to remove loosely attached microbial cells.
+
+\hypertarget{dna-isolation}{%
+\subsubsection{DNA isolation}\label{dna-isolation}}
+
+The DNA was isolated according to the protocol for isolation from
+filters described in Massana et al. (1997). This protocol was modified
+and adapted for microbial DNA isolation from macrophyte surfaces as
+described below. Five (5) \si{\ml} of lysis buffer (40 \si{\milli\Molar}
+EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar} sucrose; pH 8.3)
+was added to 1 \si{\g} wet weight of leaves or 2 \si{\g} wet-weight of
+thalli. For every sample, duplicate isolations were performed. Lysozyme
+was added (final concentration 1 \si{\mg\per\ml}) and the mixture was
+incubated at 37 \si{\degreeCelsius} for 30 \si{\minute}. Subsequently,
+proteinase K (final concentration 0.5 \si{\mg\per\ml}) and SDS (final
+concentration 1 \si{\percent}) were added and the samples were incubated
+at 55 \si{\degreeCelsius} for 2 \si{\hour}. Following the incubation,
+tubes were vortexed for 10 \si{\minute} and the mixture containing lysed
+epiphytic cells was separated from host leaves or thalli by transferring
+the solution into a clean tube. The lysate was extracted twice with a
+mixture of phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once
+with chloroform:isoamyl alcohol (24:1). After each addition of an
+organic solvent mixture, tubes were slightly vortexed and centrifuged at
+4,500 × g for 10 \si{\minute}. Following each centrifugation the aqueous
+phases were retrieved. After the final extraction 1/10 of chilled 3
+\si{\Molar} sodium acetate (pH 5.2) was added. DNA was precipitated by
+adding 1 volume of chilled isopropanol, incubating the mixtures
+overnight at \num{-20} \si{\degreeCelsius} and centrifuging at 16,000 ×
+g and 4 \si{\degreeCelsius} for 20 \si{\minute}. The pellet was washed
+twice with 1 \si{\ml} of chilled 70 \si{\percent} ethanol and
+centrifuged after each washing step at 20,000 × g and 4
+\si{\degreeCelsius} for 10 \si{\minute}. After the first washing step
+duplicate pellets from the same sample were pooled and transferred to a
+clean 1.5 \si{\ml} tube. The dried pellet was re-suspended in 100
+\si{\ul} of deionized water.
+
+\hypertarget{illumina-16s-rrna-sequencing}{%
+\subsubsection{Illumina 16S rRNA
+sequencing}\label{illumina-16s-rrna-sequencing}}
+
+An aliquot of isolated DNA was treated with RNase A (final concentration
+200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
+concentration was determined using the Quant-iT PicoGreen dsDNA Assay
+Kit (Invitrogen, USA) according to the manufacturer's instructions and
+diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
+amplified using a two-step PCR procedure. In the first PCR the 515F
+(\(5'\)-GTGYCAGCMGCCGCGGTAA-\(3'\)) and 806R
+(\(5'\)-GGACTACNVGGGTWTCTAAT-\(3'\)) primers from the Earth Microbiome
+Project
+(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
+were used to amplify the target region (Caporaso et al., 2012; Apprill
+et al., 2015; Parada et al., 2016). These primers contained on their
+\(5'\) end a tagged sequence. Each sample was amplified in four parallel
+25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer,
+0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum
+Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of
+Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5
+\si{\ng} of DNA template. Cycling conditions were: initial denaturation
+at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation
+at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
+\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
+\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
+72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
+volumes were pooled and PCR products were purified using the GeneJET PCR
+Purification Kit (Thermo Fisher Scientific, USA) according to the
+manufacturer's instructions and following the protocol that included
+isopropanol addition for better small DNA fragment yield. The column was
+eluted in 30 \si{\ul} of deionized water. Purified PCR products were
+sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
+Martinsried, Germany. Before sequencing at IMGM, the second PCR
+amplification of the two-step PCR procedure was performed using primers
+targeting the tagged region incorporated in the first PCR. In addition,
+these primers contained adapter and sample-specific index sequences. The
+second PCR was carried out for 8 cycles. Beside samples, a positive and
+negative control were sequenced. A negative control was comprised of
+four parallel PCR reactions without DNA template, while for a positive
+control a mock community composed of evenly mixed DNA material
+originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
+used. Partial 16S rRNA sequences obtained in this study have been
+deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under
+accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794,
+SAMEA6648809 -- SAMEA6648811 and SAMEA6648824.
+
+Obtained sequences were analysed on the computer cluster Isabella
+(University Computing Center, University of Zagreb) using mothur
+(version 1.43.0) (Schloss et al., 2009) according to the MiSeq Standard
+Operating Procedure (MiSeq SOP; \url{https://mothur.org/wiki/MiSeq_SOP})
+(Kozich et al., 2013) and recommendations given from the Riffomonas
+project to enhance data reproducibility
+(\url{http://www.riffomonas.org/}). For alignment and classification of
+sequences the SILVA SSU Ref NR 99 database (release 138;
+\url{http://www.arb-silva.de}) was used (Quast et al., 2013; Yilmaz et
+al., 2014). Sequences classified as chloroplasts by SILVA were exported
+and reclassified using mothur and the RDP
+(\url{http://rdp.cme.msu.edu/}) training set (version 16) reference
+files adapted for mothur (Cole et al., 2014). In comparison to SILVA,
+RDP allows a more detailed classification of chloroplast sequences.
+Based on the ATCC MSA-1002 mock community included in the analysis a
+sequencing error rate of 0.009 \si{\percent} was determined, which is in
+line with previously reported values for next-generation sequencing data
+(Kozich et al., 2013; Schloss et al., 2016). In addition, the negative
+control processed together with the samples yielded only 2 sequences
+after sequence quality curation.
+
+\hypertarget{metagenomics}{%
+\subsubsection{Metagenomics}\label{metagenomics}}
+
+Four DNA samples were selected and sent on dry ice to IMGM Laboratories,
+Martinsried, Germany for metagenomic sequencing. DNA was purified using
+AMPure XP Beads (Beckman Coulter, USA) applying a bead:DNA ratio of 1:1
+(v/v), quantified with a Qubit dsDNA HS Assay Kit (Thermo Fisher
+Scientific, USA) and check for integrity on a 1 \si{\percent} agarose
+gel. Metagenomic sequencing libraries were generated from 300 \si{\ng}
+of input DNA using a NEBNext Ultra II FS DNA Library Prep Kit for
+Illumina (New England Biolabs, USA) according to the manufacturer's
+instructions. Fragments were selected (500 -- 700 bp) using AMPure XP
+Beads, PCR enriched for 3 -- 5 cycles and quality controlled. Libraries
+generated from different DNA samples were pooled and sequenced on an
+Illumina NovaSeq 6000 sequencing system (2 × 250 bp).
+
+Obtained sequences were analysed on the Life Science Compute Cluster
+(LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
+Individual sequences were assembled using MEGAHIT (version 1.1.2) (Li et
+al., 2015) under default settings. Putative genes were predicted from
+contigs longer than 200 bp using Prodigal (version 2.6.3) (Hyatt et al.,
+2010) in metagenome mode (-p meta). Abundances of predicted genes were
+expressed as Reads Per Kilobase Million (RPKM) and calculated using the
+BWA algorithm (version 0.7.16a) (Li and Durbin, 2009). All predicted
+genes were functionally annotated using the eggNOG-mapper (Huerta-Cepas
+et al., 2017) and eggNOG database (version 5.0) (Huerta-Cepas et al.,
+2019). Sequence taxonomy classification was determined using the lowest
+common ancestor algorithm adapted from DIAMOND (version 0.8.36)
+(Buchfink et al., 2015) and by searching against the NCBI non-redundant
+database (NR). To determine the phylogeny, the top 10 \si{\percent} hits
+with an e-value \textless{} 1 × 10\textsuperscript{-5} were used
+(-\/-top 10). Sequence renaming, coverage information computing and
+metagenomic statistics calculations were performed using software tools
+from BBTools (\url{https://jgi.doe.gov/data-and-tools/bbtools}).
+Metagenomic sequences obtained in this study have been deposited in the
+European Nucleotide Archive (ENA) at EMBL-EBI under accession numbers
+SAMEA6648795, SAMEA6648797, SAMEA6648809 and SAMEA6648811.
+
+\hypertarget{protein-isolation}{%
+\subsubsection{Protein isolation}\label{protein-isolation}}
+
+Proteins were isolated according to the protocol for protein isolation
+from soil described in Chourey et al. (2010) and modified by Hultman et
+al. (2015). This protocol was further modified and adapted for microbial
+protein isolation from macrophyte surfaces as described below. Twenty
+(20) \si{\ml} of protein extraction buffer (4 \si{\percent} SDS, 100
+\si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was added to 5 \si{\g} wet
+weight of leaves or 10 \si{\g} wet weight of thalli. The mixture was
+incubated in boiling water for 5 \si{\minute}, vortexed for 10
+\si{\minute} and incubated again in boiling water for 5 \si{\minute}.
+After a brief vortex, the lysate was transferred to a clean tube
+separating the host leaves or thalli from the mixture containing lysed
+epiphytic cells. Dithiothreitol (DTT; final concentration 24
+\si{\milli\Molar}) was added and proteins were precipitated with chilled
+100 \si{\percent} trichloroacetic acid (TCA; final concentration 20
+\si{\percent}) overnight at \num{-20} \si{\degreeCelsius}. Precipitated
+proteins were centrifuged at 10,000 × g and 4 \si{\degreeCelsius} for 40
+\si{\minute}. The obtained protein pellet was washed three times with
+chilled acetone. During the first washing step the pellet was
+transferred to a clean 1.5 \si{\ml} tube. After each washing step,
+samples were centrifuged at 20,000 × g and 4 \si{\degreeCelsius} for 5
+\si{\minute}. Dried pellets were stored at \num{-80} \si{\degreeCelsius}
+until further analysis.
+
+\hypertarget{metaproteomics}{%
+\subsubsection{Metaproteomics}\label{metaproteomics}}
+
+Isolated proteins were whole trypsin digested using the FASP
+(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
+according to the manufacturer's instructions with small modifications
+(Wiśniewski et al., 2009). Prior to loading the solution onto the
+column, protein pellets were solubilized in a urea sample buffer
+included in the kit amended with DTT (final concentration 100
+\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
+centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
+remove larger particles. The first washing step after protein solution
+loading was repeated twice. In addition, the centrifugation steps were
+prolonged if the column was clogged. Trypsin digestion was performed on
+column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
+The final filtrate containing peptides was acidified with 1
+\si{\percent} (final concentration) trifluoroacetic acid (TFA), freezed
+at \num{-80} \si{\degreeCelsius}, lyophilized and sent to VIME -- Vienna
+Metabolomics Center (University of Vienna) for metaproteomic analysis.
+Peptides were re-suspended in 1 \si{\percent} (final concentration) TFA,
+desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA)
+according to the manufacturer's instructions and sequenced on a Q
+Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher
+Scientific, USA). Obtained MS/MS spectra were searched against a protein
+database composed of combined sequenced metagenomes using SEQUEST-HT
+engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
+Fisher Scientific, USA). The target-decoy approach was used to reduce
+the probability of false peptide identification. Results whose false
+discovery rate at the peptide level was \textless{} \SI{1}{\percent}
+were kept. For protein identification a minimum of two peptides and one
+unique peptide were required. For protein quantification, a
+chromatographic peak area-based free quantitative method was applied.
+
+\hypertarget{data-processing-and-visualization}{%
+\subsubsection{Data processing and
+visualization}\label{data-processing-and-visualization}}
+
+Processing and visualization of 16S rRNA, metagenomic and metaproteomic
+data were done using R (version 3.6.0) (R Core Team, 2019), package
+tidyverse (version 1.3.0) (Wickham et al., 2019) and multiple other
+packages (Neuwirth, 2014; Xie, 2014, 2015, 2019, 2020; Wilke, 2018; Xie
+et al., 2018; Allaire et al., 2019; Zhu, 2019; Bengtsson, 2020). The
+detailed analysis procedure including the R Markdown file for this paper
+are available as a GitHub repository
+(\url{https://github.com/MicrobesRovinj/Korlevic_SelectiveRemoval_FrontMicrobiol_2021}).
+
+\hypertarget{confocal-microscopy}{%
+\subsubsection{Confocal microscopy}\label{confocal-microscopy}}
+
+Host leaves and thalli from DNA and protein isolation steps were washed
+seven times in deionized water and fixed with formaldehyde (final
+concentration \textasciitilde{}3 \si{\percent}). In addition,
+non-treated leaves and thalli, washed three times in ASW to remove
+loosely attached microbial cells, were fixed in the same concentration
+of formaldehyde and used as a positive control. For-long term storage,
+fixed leaves and thalli were immersed in a mixture of phosphate-buffered
+saline (PBS) and ethanol (1:1) and stored at \num{-20}
+\si{\degreeCelsius}. Treated and untreated segments of leaves and thalli
+were stained in a 2 × solution of SYBR Green I and examined under a
+Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
 
@@
 characterized by an uneven surface (Kuo and den Hartog, 2001; Verlaque
-\emph{et al.}, 2003). The developed protocol led to an almost complete
-removal of epiphytic cells from the surfaces of both species comparable
-to the result of Burke \emph{et al.} (2009), indicating that structural
-differences do not impact the removal efficiency. In addition, isolation
-protocols were tested in two contrasting seasons, as it is known that
-macrophytes are harbouring more algal epiphytes during autumn and winter
-(Reyes and Sansón, 2001). No differences in the removal efficiency was
-observed between seasons suggesting that these protocols can be used on
+et al., 2003). The developed protocol led to an almost complete removal
+of epiphytic cells from the surfaces of both species comparable to the
+result of Burke et al. (2009), indicating that structural differences do
+not impact the removal efficiency. In addition, isolation protocols were
+tested in two contrasting seasons, as it is known that macrophytes are
+harbouring more algal epiphytes during autumn and winter (Reyes and
+Sansón, 2001). No differences in the removal efficiency was observed
+between seasons suggesting that these protocols can be used on
 macrophyte samples retrieved throughout the year. Also, no removal
@@
 often found in epiphytic communities associated with other macrophytes
-(Burke and Thomas \emph{et al.}, 2011; Morrissey \emph{et al.}, 2019). A
-problem often encountered in studies focusing on epiphytic communities
-is the presence of large proportions of chloroplast 16S rRNA sequences
-in the pool of amplified molecules, especially if the epiphytic DNA was
-isolated without prior selection (Staufenberger \emph{et al.}, 2008).
-These sequences can derive from host chloroplasts or from eukaryotic
-epiphytic chloroplast DNA. Although the proportion of obtained
-chloroplast 16S rRNA sequences in our samples was substantial, they
-derived almost exclusively from eukaryotic epiphytes. High proportion of
-chloroplast 16S rRNA sequences in studies applying selective procedures
-that include direct cellular lysis on host surfaces were observed before
-(Michelou \emph{et al.}, 2013). It is possible that chloroplast-specific
-sequences even in these studies originated from eukaryotic epiphytic
-cells and not from host chloroplasts. Indeed, it is common during 16S
-rRNA profiling of pelagic microbial communities to observe high
-proportions of chloroplast sequences (Gilbert \emph{et al.}, 2009;
-Korlević \emph{et al.}, 2016). In addition, a very low proportion of
-chloroplast 16S rRNA sequences in samples originating from \emph{C.
-cylindracea} in comparison to \emph{C. nodosa} could be explained by the
-presence of three introns in the gene for 16S rRNA in some members of
-the genus \emph{Caulerpa} that could hamper the amplification process
-(Lam and Lopez-Bautista, 2016).
+(Burke et al., 2011b; Morrissey et al., 2019). A problem often
+encountered in studies focusing on epiphytic communities is the presence
+of large proportions of chloroplast 16S rRNA sequences in the pool of
+amplified molecules, especially if the epiphytic DNA was isolated
+without prior selection (Staufenberger et al., 2008). These sequences
+can derive from host chloroplasts or from eukaryotic epiphytic
+chloroplast DNA. Although the proportion of obtained chloroplast 16S
+rRNA sequences in our samples was substantial, they derived almost
+exclusively from eukaryotic epiphytes. High proportion of chloroplast
+16S rRNA sequences in studies applying selective procedures that include
+direct cellular lysis on host surfaces were observed before (Michelou et
+al., 2013). It is possible that chloroplast-specific sequences even in
+these studies originated from eukaryotic epiphytic cells and not from
+host chloroplasts. Indeed, it is common during 16S rRNA profiling of
+pelagic microbial communities to observe high proportions of chloroplast
+sequences (Gilbert et al., 2009; Korlević et al., 2016). In addition, a
+very low proportion of chloroplast 16S rRNA sequences in samples
+originating from \emph{C. cylindracea} in comparison to \emph{C. nodosa}
+could be explained by the presence of three introns in the gene for 16S
+rRNA in some members of the genus \emph{Caulerpa} that could hamper the
+amplification process (Lam and Lopez-Bautista, 2016).
 
@@
 metagenomes and metatranscriptomes derived from similar surface
-associated communities (Crump \emph{et al.}, 2018; Cúcio \emph{et al.},
-2018). In addition, functional annotation of predicted coding sequences
-to COG functional categories showed that the obtained metagenomes can be
-used to determine the metabolic capacity of surface associated
-communities (Leary \emph{et al.}, 2014; Cúcio \emph{et al.}, 2018). The
-proportion of coding sequences, including their RPKM, originating from
-\emph{C. nodosa} metagenomes and classified as Streptophyta was low
-indicating that the isolation procedure was specific for epiphytic
-cells. DNA samples isolated from the surface of \emph{C. cylindracea}
-exhibited a low proportion of Chlorophyta coding sequences, however,
-their RPKM was higher than in the samples originating from \emph{C.
-nodosa}. One of the reasons for this elevated RPKM of Chlorophyta
-sequences in \emph{C. cylindracea} could be the differences in the
-tissue structure between these two host species. While \emph{C. nodosa}
-leaves are composed of individual cells, the thallus of \emph{C.
-cylindracea} is, like in other siphonous algal species, composed of a
-single large multinucleate cell (Coneva and Chitwood, 2015). The absence
-of individual cells in \emph{C. cylindracea} could cause a leakage of
-genetic material into the extraction buffer causing an elevated presence
-of host sequences in the samples for metagenome analyses.
+associated communities (Crump et al., 2018; Cúcio et al., 2018). In
+addition, functional annotation of predicted coding sequences to COG
+functional categories showed that the obtained metagenomes can be used
+to determine the metabolic capacity of surface associated communities
+(Leary et al., 2014; Cúcio et al., 2018). The proportion of coding
+sequences, including their RPKM, originating from \emph{C. nodosa}
+metagenomes and classified as Streptophyta was low indicating that the
+isolation procedure was specific for epiphytic cells. DNA samples
+isolated from the surface of \emph{C. cylindracea} exhibited a low
+proportion of Chlorophyta coding sequences, however, their RPKM was
+higher than in the samples originating from \emph{C. nodosa}. One of the
+reasons for this elevated RPKM of Chlorophyta sequences in \emph{C.
+cylindracea} could be the differences in the tissue structure between
+these two host species. While \emph{C. nodosa} leaves are composed of
+individual cells, the thallus of \emph{C. cylindracea} is, like in other
+siphonous algal species, composed of a single large multinucleate cell
+(Coneva and Chitwood, 2015). The absence of individual cells in \emph{C.
+cylindracea} could cause a leakage of genetic material into the
+extraction buffer causing an elevated presence of host sequences in the
+samples for metagenome analyses.
 
 To obtain insight into the metabolic status of uncultivated prokaryotes,
-a metaproteomic approach is required (Saito \emph{et al.}, 2019). The
-applied protocol for epiphytic protein isolation followed by a
-metaproteomic analysis identified between 14,219 and 16,449 proteins,
-which is higher than previously reported for e.g.~soils (Chourey
-\emph{et al.}, 2010; Hultman \emph{et al.}, 2015), seawater (Williams
-\emph{et al.}, 2012) and biofilms (Leary \emph{et al.}, 2014). The
-functional annotation of identified proteins into COG functional
-categories showed that the protein isolation protocol can be used to
-assess the metabolic status of the epiphytic community (Leary \emph{et
-al.}, 2014). Similar to the results of the metagenomic analysis, the
+a metaproteomic approach is required (Saito et al., 2019). The applied
+protocol for epiphytic protein isolation followed by a metaproteomic
+analysis identified between 14,219 and 16,449 proteins, which is higher
+than previously reported for e.g.~soils (Chourey et al., 2010; Hultman
+et al., 2015), seawater (Williams et al., 2012) and biofilms (Leary et
+al., 2014). The functional annotation of identified proteins into COG
+functional categories showed that the protein isolation protocol can be
+used to assess the metabolic status of the epiphytic community (Leary et
+al., 2014). Similar to the results of the metagenomic analysis, the
 number and abundance of identified proteins affiliated to Streptophyta
@@
 them widely applicable.
-
-\newpage
-
-\hypertarget{experimental-procedures}{%
-\subsection{Experimental procedures}\label{experimental-procedures}}
-
-\hypertarget{sampling}{%
-\subsubsection{Sampling}\label{sampling}}
-
-Leaves of \emph{C. nodosa} were sampled in a \emph{C. nodosa} meadow in
-the Bay of Saline, northern Adriatic Sea (\ang{45;7;5} N, \ang{13;37;20}
-E) and in a \emph{C. nodosa} meadow invaded by \emph{C. cylindracea} in
-the Bay of Funtana, northern Adriatic Sea (\ang{45;10;39} N,
-\ang{13;35;42} E). Thalli of \emph{C. cylindracea} were sampled in the
-same \emph{C. nodosa} invaded meadow in the Bay of Funtana and at a
-locality of only \emph{C. cylindracea} located in the proximity of the
-invaded meadow. Leaves and thalli for 16S rRNA analysis, metagenomics
-and metaproteomics were collected in two contrasting seasons, on 4
-December 2017 (16S rRNA analysis and metaproteomics), 14 December 2017
-(metagenomics) and 18 June 2018 (16S rRNA analysis, metagenomics and
-metaproteomics). During spring 2018, the \emph{C. nodosa} meadow in the
-Bay of Saline decayed to an extent that no leaves could be retrieved
-(Najdek \emph{et al.}, 2020). In addition, as not enough DNA for both
-metagenomic and 16S RNA analysis were obtained during the sampling on 4
-December 2017, an additional sampling on 14 December 2017 was carried
-out in the Bay of Funtana. Leaves and thalli were collected by diving
-and transported to the laboratory in containers placed on ice and filled
-with seawater from this site. Upon arrival to the laboratory, \emph{C.
-nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while \emph{C.
-cylindracea} thalli were cut into 5 -- 8 \si{\cm} long sections. Leaves
-and thalli were washed three times with sterile artificial seawater
-(ASW) to remove loosely attached microbial cells.
-
-\hypertarget{dna-isolation}{%
-\subsubsection{DNA isolation}\label{dna-isolation}}
-
-The DNA was isolated according to the protocol for isolation from
-filters described in Massana \emph{et al.} (1997). This protocol was
-modified and adapted for microbial DNA isolation from macrophyte
-surfaces as described below. Five (5) \si{\ml} of lysis buffer (40
-\si{\milli\Molar} EDTA, 50 \si{\milli\Molar} Tris-HCl, 0.75 \si{\Molar}
-sucrose; pH 8.3) was added to 1 \si{\g} wet weight of leaves or 2
-\si{\g} wet-weight of thalli. For every sample, duplicate isolations
-were performed. Lysozyme was added (final concentration 1
-\si{\mg\per\ml}) and the mixture was incubated at 37 \si{\degreeCelsius}
-for 30 \si{\minute}. Subsequently, proteinase K (final concentration 0.5
-\si{\mg\per\ml}) and SDS (final concentration 1 \si{\percent}) were
-added and the samples were incubated at 55 \si{\degreeCelsius} for 2
-\si{\hour}. Following the incubation, tubes were vortexed for 10
-\si{\minute} and the mixture containing lysed epiphytic cells was
-separated from host leaves or thalli by transferring the solution into a
-clean tube. The lysate was extracted twice with a mixture of
-phenol:chloroform:isoamyl alcohol (25:24:1; pH 8) and once with
-chloroform:isoamyl alcohol (24:1). After each addition of an organic
-solvent mixture, tubes were slightly vortexed and centrifuged at 4,500 ×
-g for 10 \si{\minute}. Following each centrifugation the aqueous phases
-were retrieved. After the final extraction 1/10 of chilled 3 \si{\Molar}
-sodium acetate (pH 5.2) was added. DNA was precipitated by adding 1
-volume of chilled isopropanol, incubating the mixtures overnight at
-\num{-20} \si{\degreeCelsius} and centrifuging at 16,000 × g and 4
-\si{\degreeCelsius} for 20 \si{\minute}. The pellet was washed twice
-with 1 \si{\ml} of chilled 70 \si{\percent} ethanol and centrifuged
-after each washing step at 20,000 × g and 4 \si{\degreeCelsius} for 10
-\si{\minute}. After the first washing step duplicate pellets from the
-same sample were pooled and transferred to a clean 1.5 \si{\ml} tube.
-The dried pellet was re-suspended in 100 \si{\ul} of deionized water.
-
-\hypertarget{illumina-16s-rrna-sequencing}{%
-\subsubsection{Illumina 16S rRNA
-sequencing}\label{illumina-16s-rrna-sequencing}}
-
-An aliquot of isolated DNA was treated with RNase A (final concentration
-200 \si{\ug\per\ml}) for 2 \si{\hour} at 37 \si{\degreeCelsius}. The DNA
-concentration was determined using the Quant-iT PicoGreen dsDNA Assay
-Kit (Invitrogen, USA) according to the manufacturer's instructions and
-diluted to 1 \si{\ng\per\ul}. The V4 region of the 16S rRNA gene was
-amplified using a two-step PCR procedure. In the first PCR the 515F
-(\(5'\)-GTGYCAGCMGCCGCGGTAA-\(3'\)) and 806R
-(\(5'\)-GGACTACNVGGGTWTCTAAT-\(3'\)) primers from the Earth Microbiome
-Project
-(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
-were used to amplify the target region (Caporaso \emph{et al.}, 2012;
-Apprill \emph{et al.}, 2015; Parada \emph{et al.}, 2016). These primers
-contained on their \(5'\) end a tagged sequence. Each sample was
-amplified in four parallel 25 \si{\ul} reactions of which each contained
-1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7
-\si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of
-forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase
-(New England Biolabs, USA) and 5 \si{\ng} of DNA template. Cycling
-conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3
-\si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45
-\si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and
-elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an
-elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}. The four
-parallel reactions volumes were pooled and PCR products were purified
-using the GeneJET PCR Purification Kit (Thermo Fisher Scientific, USA)
-according to the manufacturer's instructions and following the protocol
-that included isopropanol addition for better small DNA fragment yield.
-The column was eluted in 30 \si{\ul} of deionized water. Purified PCR
-products were sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM
-Laboratories, Martinsried, Germany. Before sequencing at IMGM, the
-second PCR amplification of the two-step PCR procedure was performed
-using primers targeting the tagged region incorporated in the first PCR.
-In addition, these primers contained adapter and sample-specific index
-sequences. The second PCR was carried out for 8 cycles. Beside samples,
-a positive and negative control were sequenced. A negative control was
-comprised of four parallel PCR reactions without DNA template, while for
-a positive control a mock community composed of evenly mixed DNA
-material originating from 20 bacterial strains (ATCC MSA-1002, ATCC,
-USA) was used. Partial 16S rRNA sequences obtained in this study have
-been deposited in the European Nucleotide Archive (ENA) at EMBL-EBI
-under accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794,
-SAMEA6648809 -- SAMEA6648811.
-
-Obtained sequences were analysed on the computer cluster Isabella
-(University Computing Center, University of Zagreb) using mothur
-(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
-Standard Operating Procedure (MiSeq SOP;
-\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
-and recommendations given from the Riffomonas project to enhance data
-reproducibility (\url{http://www.riffomonas.org/}). For alignment and
-classification of sequences the SILVA SSU Ref NR 99 database (release
-138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
-Yilmaz \emph{et al.}, 2014). Sequences classified as chloroplasts by
-SILVA were exported and reclassified using mothur and the RDP
-(\url{http://rdp.cme.msu.edu/}) training set (version 16) reference
-files adapted for mothur (Cole \emph{et al.}, 2014). In comparison to
-SILVA, RDP allows a more detailed classification of chloroplast
-sequences. Based on the ATCC MSA-1002 mock community included in the
-analysis a sequencing error rate of 0.009 \si{\percent} was determined,
-which is in line with previously reported values for next-generation
-sequencing data (Kozich \emph{et al.}, 2013; Schloss \emph{et al.},
-2016). In addition, the negative control processed together with the
-samples yielded only 2 sequences after sequence quality curation.
-
-\hypertarget{metagenomics}{%
-\subsubsection{Metagenomics}\label{metagenomics}}
-
-Four DNA samples were selected and sent on dry ice to IMGM Laboratories,
-Martinsried, Germany for metagenomic sequencing. DNA was purified using
-AMPure XP Beads (Beckman Coulter, USA) applying a bead:DNA ratio of 1:1
-(v/v), quantified with a Qubit dsDNA HS Assay Kit (Thermo Fisher
-Scientific, USA) and check for integrity on a 1 \si{\percent} agarose
-gel. Metagenomic sequencing libraries were generated from 300 \si{\ng}
-of input DNA using a NEBNext Ultra II FS DNA Library Prep Kit for
-Illumina (New England Biolabs, USA) according to the manufacturer's
-instructions. Fragments were selected (500 -- 700 bp) using AMPure XP
-Beads, PCR enriched for 3 -- 5 cycles and quality controlled. Libraries
-generated from different DNA samples were pooled and sequenced on an
-Illumina NovaSeq 6000 sequencing system (2 × 250 bp).
-
-Obtained sequences were analysed on the Life Science Compute Cluster
-(LiSC) (CUBE -- Computational Systems Biology, University of Vienna).
-Individual sequences were assembled using MEGAHIT (version 1.1.2) (Li
-\emph{et al.}, 2015) under default settings. Putative genes were
-predicted from contigs longer than 200 bp using Prodigal (version 2.6.3)
-(Hyatt \emph{et al.}, 2010) in metagenome mode (-p meta). Abundances of
-predicted genes were expressed as Reads Per Kilobase Million (RPKM) and
-calculated using the BWA algorithm (version 0.7.16a) (Li and Durbin,
-2009). All predicted genes were functionally annotated using the
-eggNOG-mapper (Huerta-Cepas \emph{et al.}, 2017) and eggNOG database
-(version 5.0) (Huerta-Cepas \emph{et al.}, 2019). Sequence taxonomy
-classification was determined using the lowest common ancestor algorithm
-adapted from DIAMOND (version 0.8.36) (Buchfink \emph{et al.}, 2015) and
-by searching against the NCBI non-redundant database (NR). To determine
-the phylogeny, the top 10 \si{\percent} hits with an e-value \textless{}
-1 × 10\textsuperscript{-5} were used (-\/-top 10). Sequence renaming,
-coverage information computing and metagenomic statistics calculations
-were performed using software tools from BBTools
-(\url{https://jgi.doe.gov/data-and-tools/bbtools}). Metagenomic
-sequences obtained in this study have been deposited in the European
-Nucleotide Archive (ENA) at EMBL-EBI under accession numbers
-SAMEA6648795, SAMEA6648797, SAMEA6648809 and SAMEA6648811.
-
-\hypertarget{protein-isolation}{%
-\subsubsection{Protein isolation}\label{protein-isolation}}
-
-Proteins were isolated according to the protocol for protein isolation
-from soil described in Chourey \emph{et al.} (2010) and modified by
-Hultman \emph{et al.} (2015). This protocol was further modified and
-adapted for microbial protein isolation from macrophyte surfaces as
-described below. Twenty (20) \si{\ml} of protein extraction buffer (4
-\si{\percent} SDS, 100 \si{\milli\Molar} Tris-HCl {[}pH 8.0{]}) was
-added to 5 \si{\g} wet weight of leaves or 10 \si{\g} wet weight of
-thalli. The mixture was incubated in boiling water for 5 \si{\minute},
-vortexed for 10 \si{\minute} and incubated again in boiling water for 5
-\si{\minute}. After a brief vortex, the lysate was transferred to a
-clean tube separating the host leaves or thalli from the mixture
-containing lysed epiphytic cells. Dithiothreitol (DTT; final
-concentration 24 \si{\milli\Molar}) was added and proteins were
-precipitated with chilled 100 \si{\percent} trichloroacetic acid (TCA;
-final concentration 20 \si{\percent}) overnight at \num{-20}
-\si{\degreeCelsius}. Precipitated proteins were centrifuged at 10,000 ×
-g and 4 \si{\degreeCelsius} for 40 \si{\minute}. The obtained protein
-pellet was washed three times with chilled acetone. During the first
-washing step the pellet was transferred to a clean 1.5 \si{\ml} tube.
-After each washing step, samples were centrifuged at 20,000 × g and 4
-\si{\degreeCelsius} for 5 \si{\minute}. Dried pellets were stored at
-\num{-80} \si{\degreeCelsius} until further analysis.
-
-\hypertarget{metaproteomics}{%
-\subsubsection{Metaproteomics}\label{metaproteomics}}
-
-Isolated proteins were whole trypsin digested using the FASP
-(filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK)
-according to the manufacturer's instructions with small modifications
-(Wiśniewski \emph{et al.}, 2009). Prior to loading the solution onto the
-column, protein pellets were solubilized in a urea sample buffer
-included in the kit amended with DTT (final concentration 100
-\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
-centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
-remove larger particles. The first washing step after protein solution
-loading was repeated twice. In addition, the centrifugation steps were
-prolonged if the column was clogged. Trypsin digestion was performed on
-column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
-The final filtrate containing peptides was acidified with 1
-\si{\percent} (final concentration) trifluoroacetic acid (TFA), freezed
-at \num{-80} \si{\degreeCelsius}, lyophilized and sent to VIME -- Vienna
-Metabolomics Center (University of Vienna) for metaproteomic analysis.
-Peptides were re-suspended in 1 \si{\percent} (final concentration) TFA,
-desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA)
-according to the manufacturer's instructions and sequenced on a Q
-Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher
-Scientific, USA). Obtained MS/MS spectra were searched against a protein
-database composed of combined sequenced metagenomes using SEQUEST-HT
-engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
-Fisher Scientific, USA). The target-decoy approach was used to reduce
-the probability of false peptide identification. Results whose false
-discovery rate at the peptide level was \SI{< 1}{\percent} were kept.
-For protein identification a minimum of two peptides and one unique
-peptide were required. For protein quantification, a chromatographic
-peak area-based free quantitative method was applied.
-
-\hypertarget{data-processing-and-visualization}{%
-\subsubsection{Data processing and
-visualization}\label{data-processing-and-visualization}}
-
-Processing and visualization of 16S rRNA, metagenomic and metaproteomic
-data were done using R (version 3.6.0) (R Core Team, 2019), package
-tidyverse (version 1.3.0) (Wickham \emph{et al.}, 2019) and multiple
-other packages (Xie, 2014, 2015, 2019, 2020; Neuwirth, 2014; Xie
-\emph{et al.}, 2018; Wilke, 2018; Allaire \emph{et al.}, 2019; Zhu,
-2019; Bengtsson, 2020). The detailed analysis procedure including the R
-Markdown file for this paper are available as a GitHub repository
-(\url{https://github.com/MicrobesRovinj/Korlevic_SelectiveRemoval_EnvironMicrobiol_2020}).
-
-\hypertarget{confocal-microscopy}{%
-\subsubsection{Confocal microscopy}\label{confocal-microscopy}}
-
-Host leaves and thalli from DNA and protein isolation steps were washed
-seven times in deionized water and fixed with formaldehyde (final
-concentration \textasciitilde{}3 \si{\percent}). In addition,
-non-treated leaves and thalli, washed three times in ASW to remove
-loosely attached microbial cells, were fixed in the same concentration
-of formaldehyde and used as a positive control. For-long term storage,
-fixed leaves and thalli were immersed in a mixture of phosphate-buffered
-saline (PBS) and ethanol (1:1) and stored at \num{-20}
-\si{\degreeCelsius}. Treated and untreated segments of leaves and thalli
-were stained in a 2 × solution of SYBR Green I and examined under a
-Leica TCS SP8 X FLIM confocal microscope (Leica Microsystems, Germany).
 
@@
 \leavevmode\hypertarget{ref-Allaire2019}{}%
-Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
-A., et al. (2019) rmarkdown: Dynamic documents for R.
+Allaire, J. J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
+A., et al. (2019). \emph{rmarkdown: Dynamic documents for R}.
 
 \leavevmode\hypertarget{ref-Amann1995}{}%
-Amann, R.I., Ludwig, W., and Schleifer, K.H. (1995) Phylogenetic
+Amann, R. I., Ludwig, W., and Schleifer, K. H. (1995). Phylogenetic
 identification and \emph{in situ} detection of individual microbial
-cells without cultivation. \emph{Microbiol Rev} \textbf{59}: 143--169.
+cells without cultivation. \emph{Microbiol. Rev.} 59, 143--169.
 
 \leavevmode\hypertarget{ref-Andersson2010}{}%
-Andersson, A.F., Riemann, L., and Bertilsson, S. (2010) Pyrosequencing
+Andersson, A. F., Riemann, L., and Bertilsson, S. (2010). Pyrosequencing
 reveals contrasting seasonal dynamics of taxa within Baltic Sea
-bacterioplankton communities. \emph{ISME J} \textbf{4}: 171--181.
+bacterioplankton communities. \emph{ISME J.} 4, 171--181.
+doi:\href{https://doi.org/10.1038/ismej.2009.108}{10.1038/ismej.2009.108}.
 
 \leavevmode\hypertarget{ref-Apprill2015}{}%
-Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015) Minor
+Apprill, A., McNally, S., Parsons, R., and Weber, L. (2015). Minor
 revision to V4 region SSU rRNA 806R gene primer greatly increases
-detection of SAR11 bacterioplankton. \emph{Aquat Microb Ecol}
-\textbf{75}: 129--137.
+detection of SAR11 bacterioplankton. \emph{Aquat. Microb. Ecol.} 75,
+129--137. doi:\href{https://doi.org/10.3354/ame01753}{10.3354/ame01753}.
 
 \leavevmode\hypertarget{ref-Bengtsson2020}{}%
-Bengtsson, H. (2020) matrixStats: Functions that apply to rows and
-columns of matrices (and to vectors).
+Bengtsson, H. (2020). \emph{matrixStats: Functions that apply to rows
+and columns of matrices (and to vectors)}.
 
 \leavevmode\hypertarget{ref-Buchfink2015}{}%
-Buchfink, B., Xie, C., and Huson, D.H. (2015) Fast and sensitive protein
-alignment using DIAMOND. \emph{Nat Methods} \textbf{12}: 59--60.
+Buchfink, B., Xie, C., and Huson, D. H. (2015). Fast and sensitive
+protein alignment using DIAMOND. \emph{Nat. Methods} 12, 59--60.
+doi:\href{https://doi.org/10.1038/nmeth.3176}{10.1038/nmeth.3176}.
 
 \leavevmode\hypertarget{ref-Burke2009}{}%
-Burke, C., Kjelleberg, S., and Thomas, T. (2009) Selective extraction of
-bacterial DNA from the surfaces of macroalgae. \emph{Appl Environ
-Microbiol} \textbf{75}: 252--256.
+Burke, C., Kjelleberg, S., and Thomas, T. (2009). Selective extraction
+of bacterial DNA from the surfaces of macroalgae. \emph{Appl. Environ.
+Microbiol.} 75, 252--256.
+doi:\href{https://doi.org/10.1128/AEM.01630-08}{10.1128/AEM.01630-08}.
 
@@
 Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
-(2011) Bacterial community assembly based on functional genes rather
-than species. \emph{Proc Natl Acad Sci U S A} \textbf{108}:
-14288--14293.
+(2011a). Bacterial community assembly based on functional genes rather
+than species. \emph{Proc. Natl. Acad. Sci. U. S. A.} 108, 14288--14293.
+doi:\href{https://doi.org/10.1073/pnas.1101591108}{10.1073/pnas.1101591108}.
 
@@
 Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
-(2011) Composition, uniqueness and variability of the epiphytic
+(2011b). Composition, uniqueness and variability of the epiphytic
 bacterial community of the green alga \emph{Ulva australis}. \emph{ISME
-J} \textbf{5}: 590--600.
+J.} 5, 590--600.
+doi:\href{https://doi.org/10.1038/ismej.2010.164}{10.1038/ismej.2010.164}.
 
 \leavevmode\hypertarget{ref-Cai2014}{}%
-Cai, X., Gao, G., Yang, J., Tang, X., Dai, J., Chen, D., and Song, Y.
-(2014) An ultrasonic method for separation of epiphytic microbes from
-freshwater submerged macrophytes. \emph{J Basic Microbiol} \textbf{54}:
+Cai, X., Gao, G., Yang, J., Tang, X., Dai, J., Chen, D., et al. (2014).
+An ultrasonic method for separation of epiphytic microbes from
+freshwater submerged macrophytes. \emph{J. Basic Microbiol.} 54,
 758--761.
+doi:\href{https://doi.org/10.1002/jobm.201300041}{10.1002/jobm.201300041}.
 
 \leavevmode\hypertarget{ref-Caporaso2012}{}%
-Caporaso, J.G., Lauber, C.L., Walters, W.A., Berg-Lyons, D., Huntley,
-J., Fierer, N., et al. (2012) Ultra-high-throughput microbial community
-analysis on the Illumina HiSeq and MiSeq platforms. \emph{ISME J}
-\textbf{6}: 1621--1624.
+Caporaso, J. G., Lauber, C. L., Walters, W. A., Berg-Lyons, D., Huntley,
+J., Fierer, N., et al. (2012). Ultra-high-throughput microbial community
+analysis on the Illumina HiSeq and MiSeq platforms. \emph{ISME J.} 6,
+1621--1624.
+doi:\href{https://doi.org/10.1038/ismej.2012.8}{10.1038/ismej.2012.8}.
 
 \leavevmode\hypertarget{ref-Chourey2010}{}%
-Chourey, K., Jansson, J., VerBerkmoes, N., Shah, M., Chavarria, K.L.,
-Tom, L.M., et al. (2010) Direct cellular lysis/protein extraction
-protocol for soil metaproteomics. \emph{J Proteome Res} \textbf{9}:
-6615--6622.
+Chourey, K., Jansson, J., VerBerkmoes, N., Shah, M., Chavarria, K. L.,
+Tom, L. M., et al. (2010). Direct cellular lysis/protein extraction
+protocol for soil metaproteomics. \emph{J. Proteome Res.} 9, 6615--6622.
+doi:\href{https://doi.org/10.1021/pr100787q}{10.1021/pr100787q}.
 
 \leavevmode\hypertarget{ref-Cole2014}{}%
-Cole, J.R., Wang, Q., Fish, J.A., Chai, B., McGarrell, D.M., Sun, Y., et
-al. (2014) Ribosomal Database Project: Data and tools for high
-throughput rRNA analysis. \emph{Nucleic Acids Res} \textbf{42}:
-D633--D642.
+Cole, J. R., Wang, Q., Fish, J. A., Chai, B., McGarrell, D. M., Sun, Y.,
+et al. (2014). Ribosomal Database Project: Data and tools for high
+throughput rRNA analysis. \emph{Nucleic Acids Res} 42, D633--D642.
+doi:\href{https://doi.org/10.1093/nar/gkt1244}{10.1093/nar/gkt1244}.
 
 \leavevmode\hypertarget{ref-Coneva2015}{}%
-Coneva, V. and Chitwood, D.H. (2015) Plant architecture without
+Coneva, V., and Chitwood, D. H. (2015). Plant architecture without
 multicellularity: Quandaries over patterning and the soma-germline
-divide in siphonous algae. \emph{Front Plant Sci} \textbf{6}: 287.
+divide in siphonous algae. \emph{Front. Plant Sci.} 6, 287.
+doi:\href{https://doi.org/10.3389/fpls.2015.00287}{10.3389/fpls.2015.00287}.
 
 \leavevmode\hypertarget{ref-Crump2018}{}%
-Crump, B.C., Wojahn, J.M., Tomas, F., and Mueller, R.S. (2018)
+Crump, B. C., Wojahn, J. M., Tomas, F., and Mueller, R. S. (2018).
 Metatranscriptomics and amplicon sequencing reveal mutualisms in
-seagrass microbiomes. \emph{Front Microbiol} \textbf{9}: 388.
+seagrass microbiomes. \emph{Front. Microbiol.} 9, 388.
+doi:\href{https://doi.org/10.3389/fmicb.2018.00388}{10.3389/fmicb.2018.00388}.
 
 \leavevmode\hypertarget{ref-Cucio2018}{}%
-Cúcio, C., Overmars, L., Engelen, A.H., and Muyzer, G. (2018)
+Cúcio, C., Overmars, L., Engelen, A. H., and Muyzer, G. (2018).
 Metagenomic analysis shows the presence of bacteria related to
 free-living forms of sulfur-oxidizing chemolithoautotrophic symbionts in
-the rhizosphere of the seagrass \emph{Zostera marina}. \emph{Front Mar
-Sci} \textbf{5}: 171.
+the rhizosphere of the seagrass \emph{Zostera marina}. \emph{Front. Mar.
+Sci.} 5, 171.
+doi:\href{https://doi.org/10.3389/fmars.2018.00171}{10.3389/fmars.2018.00171}.
 
@@
 Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
-Thomas, T. (2013) The seaweed holobiont: Understanding seaweed-bacteria
-interactions. \emph{FEMS Microbiol Rev} \textbf{37}: 462--476.
+Thomas, T. (2013). The seaweed holobiont: Understanding seaweed-bacteria
+interactions. \emph{FEMS Microbiol. Rev.} 37, 462--476.
+doi:\href{https://doi.org/10.1111/1574-6976.12011}{10.1111/1574-6976.12011}.
 
 \leavevmode\hypertarget{ref-Gilbert2009}{}%
-Gilbert, J.A., Field, D., Swift, P., Newbold, L., Oliver, A., Smyth, T.,
-et al. (2009) The seasonal structure of microbial communities in the
-Western English Channel. \emph{Environ Microbiol} \textbf{11}:
-3132--3139.
+Gilbert, J. A., Field, D., Swift, P., Newbold, L., Oliver, A., Smyth,
+T., et al. (2009). The seasonal structure of microbial communities in
+the Western English Channel. \emph{Environ. Microbiol.} 11, 3132--3139.
+doi:\href{https://doi.org/10.1111/j.1462-2920.2009.02017.x}{10.1111/j.1462-2920.2009.02017.x}.
 
 \leavevmode\hypertarget{ref-Gross2003}{}%
-Gross, E.M., Feldbaum, C., and Graf, A. (2003) Epiphyte biomass and
+Gross, E. M., Feldbaum, C., and Graf, A. (2003). Epiphyte biomass and
 elemental composition on submersed macrophytes in shallow eutrophic
-lakes. \emph{Hydrobiologia} \textbf{506-509}: 559--565.
+lakes. \emph{Hydrobiologia} 506-509, 559--565.
+doi:\href{https://doi.org/10.1023/B:HYDR.0000008538.68268.82}{10.1023/B:HYDR.0000008538.68268.82}.
 
 \leavevmode\hypertarget{ref-Huerta-Cepas2017}{}%
-Huerta-Cepas, J., Forslund, K., Coelho, L.P., Szklarczyk, D., Jensen,
-L.J., von Mering, C., and Bork, P. (2017) Fast genome-wide functional
-annotation through orthology assignment by eggNOG-mapper. \emph{Mol Biol
-Evol} \textbf{34}: 2115--2122.
+Huerta-Cepas, J., Forslund, K., Coelho, L. P., Szklarczyk, D., Jensen,
+L. J., von Mering, C., et al. (2017). Fast genome-wide functional
+annotation through orthology assignment by eggNOG-mapper. \emph{Mol.
+Biol. Evol.} 34, 2115--2122.
+doi:\href{https://doi.org/10.1093/molbev/msx148}{10.1093/molbev/msx148}.
 
@@
 Huerta-Cepas, J., Szklarczyk, D., Heller, D., Hernández-Plaza, A.,
-Forslund, S.K., Cook, H., et al. (2019) eggNOG 5.0: A hierarchical,
+Forslund, S. K., Cook, H., et al. (2019). eggNOG 5.0: A hierarchical,
 functionally and phylogenetically annotated orthology resource based on
-5090 organisms and 2502 viruses. \emph{Nucleic Acids Res} \textbf{47}:
+5090 organisms and 2502 viruses. \emph{Nucleic Acids Res} 47,
 D309--D314.
+doi:\href{https://doi.org/10.1093/nar/gky1085}{10.1093/nar/gky1085}.
 
 \leavevmode\hypertarget{ref-Hultman2015}{}%
-Hultman, J., Waldrop, M.P., Mackelprang, R., David, M.M., McFarland, J.,
-Blazewicz, S.J., et al. (2015) Multi-omics of permafrost, active layer
-and thermokarst bog soil microbiomes. \emph{Nature} \textbf{521}:
-208--212.
+Hultman, J., Waldrop, M. P., Mackelprang, R., David, M. M., McFarland,
+J., Blazewicz, S. J., et al. (2015). Multi-omics of permafrost, active
+layer and thermokarst bog soil microbiomes. \emph{Nature} 521, 208--212.
+doi:\href{https://doi.org/10.1038/nature14238}{10.1038/nature14238}.
 
 \leavevmode\hypertarget{ref-Hyatt2010}{}%
-Hyatt, D., Chen, G.-L., LoCascio, P.F., Land, M.L., Larimer, F.W., and
-Hauser, L.J. (2010) Prodigal: Prokaryotic gene recognition and
+Hyatt, D., Chen, G.-L., LoCascio, P. F., Land, M. L., Larimer, F. W.,
+and Hauser, L. J. (2010). Prodigal: Prokaryotic gene recognition and
 translation initiation site identification. \emph{BMC Bioinformatics}
-\textbf{11}: 119.
+11, 119.
+doi:\href{https://doi.org/10.1186/1471-2105-11-119}{10.1186/1471-2105-11-119}.
 
@@
 Jiang, Y.-F., Ling, J., Dong, J.-D., Chen, B., Zhang, Y.-Y., Zhang,
-Y.-Z., and Wang, Y.-S. (2015) Illumina-based analysis the microbial
-diversity associated with \emph{Thalassia hemprichii} in Xincun Bay,
-South China Sea. \emph{Ecotoxicology} \textbf{24}: 1548--56.
+Y.-Z., et al. (2015). Illumina-based analysis the microbial diversity
+associated with \emph{Thalassia hemprichii} in Xincun Bay, South China
+Sea. \emph{Ecotoxicology} 24, 1548--56.
+doi:\href{https://doi.org/10.1007/s10646-015-1511-z}{10.1007/s10646-015-1511-z}.
 
@@
 Korlević, M., Šupraha, L., Ljubešić, Z., Henderiks, J., Ciglenečki, I.,
-Dautović, J., and Orlić, S. (2016) Bacterial diversity across a highly
-stratified ecosystem: A salt-wedge Mediterranean estuary. \emph{Syst
-Appl Microbiol} \textbf{39}: 398--408.
+Dautović, J., et al. (2016). Bacterial diversity across a highly
+stratified ecosystem: A salt-wedge Mediterranean estuary. \emph{Syst.
+Appl. Microbiol.} 39, 398--408.
+doi:\href{https://doi.org/10.1016/j.syapm.2016.06.006}{10.1016/j.syapm.2016.06.006}.
 
 \leavevmode\hypertarget{ref-Kozich2013}{}%
-Kozich, J.J., Westcott, S.L., Baxter, N.T., Highlander, S.K., and
-Schloss, P.D. (2013) Development of a dual-index sequencing strategy and
-curation pipeline for analyzing amplicon sequence data on the MiSeq
-Illumina sequencing platform. \emph{Appl Environ Microbiol} \textbf{79}:
+Kozich, J. J., Westcott, S. L., Baxter, N. T., Highlander, S. K., and
+Schloss, P. D. (2013). Development of a dual-index sequencing strategy
+and curation pipeline for analyzing amplicon sequence data on the MiSeq
+Illumina sequencing platform. \emph{Appl. Environ. Microbiol.} 79,
 5112--5120.
+doi:\href{https://doi.org/10.1128/AEM.01043-13}{10.1128/AEM.01043-13}.
 
 \leavevmode\hypertarget{ref-Kuo2001}{}%
-Kuo, J. and den Hartog, C. (2001) Seagrass taxonomy and identification
-key. In \emph{Global Seagrass Research Methods}. Short, F.T. and Coles,
-R.G. (eds). Amsterdam: Elsevier Science B.V., pp. 31--58.
+Kuo, J., and den Hartog, C. (2001). ``Seagrass taxonomy and
+identification key,'' in \emph{Global Seagrass Research Methods}, eds.
+F. T. Short and R. G. Coles (Amsterdam: Elsevier Science B.V.), 31--58.
 
 \leavevmode\hypertarget{ref-Lam2016}{}%
-Lam, D.W. and Lopez-Bautista, J.M. (2016) Complete chloroplast genome
-for \emph{Caulerpa racemosa} (Bryopsidales, Chlorophyta) and comparative
-analyses of siphonous green seaweed plastomes. \emph{Cymbella}
-\textbf{2}: 23--32.
+Lam, D. W., and Lopez-Bautista, J. M. (2016). Complete chloroplast
+genome for \emph{Caulerpa racemosa} (Bryopsidales, Chlorophyta) and
+comparative analyses of siphonous green seaweed plastomes.
+\emph{Cymbella} 2, 23--32.
 
 \leavevmode\hypertarget{ref-Leary2014}{}%
-Leary, D.H., Li, R.W., Hamdan, L.J., IV, W.J.H., Lebedev, N., Wang, Z.,
-et al. (2014) Integrated metagenomic and metaproteomic analyses of
-marine biofilm communities. \emph{Biofouling} \textbf{30}: 1211--1223.
+Leary, D. H., Li, R. W., Hamdan, L. J., IV, W. J. H., Lebedev, N., Wang,
+Z., et al. (2014). Integrated metagenomic and metaproteomic analyses of
+marine biofilm communities. \emph{Biofouling} 30, 1211--1223.
+doi:\href{https://doi.org/10.1080/08927014.2014.977267}{10.1080/08927014.2014.977267}.
 
 \leavevmode\hypertarget{ref-Li2015}{}%
-Li, D., Liu, C.-M., Luo, R., Sadakane, K., and Lam, T.-W. (2015)
+Li, D., Liu, C.-M., Luo, R., Sadakane, K., and Lam, T.-W. (2015).
 MEGAHIT: An ultra-fast single-node solution for large and complex
 metagenomics assembly via succinct de Bruijn graph.
-\emph{Bioinformatics} \textbf{31}: 1674--1676.
+\emph{Bioinformatics} 31, 1674--1676.
+doi:\href{https://doi.org/10.1093/bioinformatics/btv033}{10.1093/bioinformatics/btv033}.
 
 \leavevmode\hypertarget{ref-Li2009a}{}%
-Li, H. and Durbin, R. (2009) Fast and accurate short read alignment with
-Burrows--Wheeler transform. \emph{Bioinformatics} \textbf{25}:
-1754--1760.
+Li, H., and Durbin, R. (2009). Fast and accurate short read alignment
+with Burrows--Wheeler transform. \emph{Bioinformatics} 25, 1754--1760.
+doi:\href{https://doi.org/10.1093/bioinformatics/btp324}{10.1093/bioinformatics/btp324}.
 
@@
 Longford, S., Tujula, N., Crocetti, G., Holmes, A., Holmström, C.,
-Kjelleberg, S., et al. (2007) Comparisons of diversity of bacterial
-communities associated with three sessile marine eukaryotes. \emph{Aquat
-Microb Ecol} \textbf{48}: 217--229.
+Kjelleberg, S., et al. (2007). Comparisons of diversity of bacterial
+communities associated with three sessile marine eukaryotes.
+\emph{Aquat. Microb. Ecol.} 48, 217--229.
+doi:\href{https://doi.org/10.3354/ame048217}{10.3354/ame048217}.
 
 \leavevmode\hypertarget{ref-Massana1997}{}%
-Massana, R., Murray, A.E., Preston, C.M., and DeLong, E.F. (1997)
+Massana, R., Murray, A. E., Preston, C. M., and DeLong, E. F. (1997).
 Vertical distribution and phylogenetic characterization of marine
-planktonic \emph{Archaea} in the Santa Barbara Channel. \emph{Appl
-Environ Microbiol} \textbf{63}: 50--56.
+planktonic \emph{Archaea} in the Santa Barbara Channel. \emph{Appl.
+Environ. Microbiol.} 63, 50--56.
 
 \leavevmode\hypertarget{ref-Michelou2013}{}%
-Michelou, V.K., Caporaso, J.G., Knight, R., and Palumbi, S.R. (2013) The
-ecology of microbial communities associated with \emph{Macrocystis
-pyrifera}. \emph{PloS One} \textbf{8}: e67480.
+Michelou, V. K., Caporaso, J. G., Knight, R., and Palumbi, S. R. (2013).
+The ecology of microbial communities associated with \emph{Macrocystis
+pyrifera}. \emph{PloS One} 8, e67480.
+doi:\href{https://doi.org/10.1371/journal.pone.0067480}{10.1371/journal.pone.0067480}.
 
 \leavevmode\hypertarget{ref-Morrissey2019}{}%
-Morrissey, K.L., Çavas, L., Willems, A., and De Clerck, O. (2019)
+Morrissey, K. L., Çavas, L., Willems, A., and De Clerck, O. (2019).
 Disentangling the influence of environment, host specificity and thallus
 differentiation on bacterial communities in siphonous green seaweeds.
-\emph{Front Microbiol} \textbf{10}: 717.
+\emph{Front. Microbiol.} 10, 717.
+doi:\href{https://doi.org/10.3389/fmicb.2019.00717}{10.3389/fmicb.2019.00717}.
 
@@
 Najdek, M., Korlević, M., Paliaga, P., Markovski, M., Ivančić, I.,
-Iveša, L., et al. (2020) Dynamics of environmental conditions during the
-decline of a \emph{Cymodocea nodosa} meadow. \emph{Biogeosciences}
-\textbf{17}: 3299--3315.
+Iveša, L., et al. (2020). Dynamics of environmental conditions during
+the decline of a \emph{Cymodocea nodosa} meadow. \emph{Biogeosciences}
+17, 3299--3315.
+doi:\href{https://doi.org/10.5194/bg-17-3299-2020}{10.5194/bg-17-3299-2020}.
 
 \leavevmode\hypertarget{ref-Neuwirth2014}{}%
-Neuwirth, E. (2014) RColorBrewer: ColorBrewer palettes.
+Neuwirth, E. (2014). \emph{RColorBrewer: ColorBrewer palettes}.
 
 \leavevmode\hypertarget{ref-Noges2010}{}%
-Nõges, T., Luup, H., and Feldmann, T. (2010) Primary production of
+Nõges, T., Luup, H., and Feldmann, T. (2010). Primary production of
 aquatic macrophytes and their epiphytes in two shallow lakes (Peipsi and
-Võrtsjärv) in Estonia. \emph{Aquat Ecol} \textbf{44}: 83--92.
+Võrtsjärv) in Estonia. \emph{Aquat. Ecol.} 44, 83--92.
+doi:\href{https://doi.org/10.1007/s10452-009-9249-4}{10.1007/s10452-009-9249-4}.
 
 \leavevmode\hypertarget{ref-Parada2016}{}%
-Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
+Parada, A. E., Needham, D. M., and Fuhrman, J. A. (2016). Every base
 matters: Assessing small subunit rRNA primers for marine microbiomes
 with mock communities, time series and global field samples.
-\emph{Environ Microbiol} \textbf{18}: 1403--1414.
+\emph{Environ. Microbiol.} 18, 1403--1414.
+doi:\href{https://doi.org/10.1111/1462-2920.13023}{10.1111/1462-2920.13023}.
 
@@
 Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
-et al. (2013) The SILVA ribosomal RNA gene database project: Improved
-data processing and web-based tools. \emph{Nucleic Acids Res}
-\textbf{41}: D590--D596.
+et al. (2013). The SILVA ribosomal RNA gene database project: Improved
+data processing and web-based tools. \emph{Nucleic Acids Res.} 41,
+D590--D596.
+doi:\href{https://doi.org/10.1093/nar/gks1219}{10.1093/nar/gks1219}.
 
 \leavevmode\hypertarget{ref-RCoreTeam2019}{}%
-R Core Team (2019) R: A language and environment for statistical
-computing, Vienna, Austria: R Foundation for Statistical Computing.
+R Core Team (2019). \emph{R: A language and environment for statistical
+computing}. Vienna, Austria: R Foundation for Statistical Computing.
 
 \leavevmode\hypertarget{ref-Reyes2001}{}%
-Reyes, J. and Sansón, M. (2001) Biomass and production of the epiphytes
-on the leaves of \emph{Cymodocea nodosa} in the Canary Islands.
-\emph{Bot Mar} \textbf{44}: 307--313.
+Reyes, J., and Sansón, M. (2001). Biomass and production of the
+epiphytes on the leaves of \emph{Cymodocea nodosa} in the Canary
+Islands. \emph{Bot. Mar.} 44, 307--313.
+doi:\href{https://doi.org/10.1515/BOT.2001.039}{10.1515/BOT.2001.039}.
 
 \leavevmode\hypertarget{ref-Richter-Heitmann2016}{}%
-Richter-Heitmann, T., Eickhorst, T., Knauth, S., Friedrich, M.W., and
-Schmidt, H. (2016) Evaluation of strategies to separate root-associated
+Richter-Heitmann, T., Eickhorst, T., Knauth, S., Friedrich, M. W., and
+Schmidt, H. (2016). Evaluation of strategies to separate root-associated
 microbial communities: A crucial choice in rhizobiome research.
-\emph{Front Microbiol} \textbf{7}: 773.
+\emph{Front. Microbiol.} 7, 773.
+doi:\href{https://doi.org/10.3389/fmicb.2016.00773}{10.3389/fmicb.2016.00773}.
 
 \leavevmode\hypertarget{ref-Saito2019}{}%
-Saito, M.A., Bertrand, E.M., Duffy, M.E., Gaylord, D.A., Held, N.A.,
-Hervey, W.J., et al. (2019) Progress and challenges in ocean
-metaproteomics and proposed best practices for data sharing. \emph{J
-Proteome Res} \textbf{18}: 1461--1476.
+Saito, M. A., Bertrand, E. M., Duffy, M. E., Gaylord, D. A., Held, N.
+A., Hervey, W. J., et al. (2019). Progress and challenges in ocean
+metaproteomics and proposed best practices for data sharing. \emph{J.
+Proteome Res.} 18, 1461--1476.
+doi:\href{https://doi.org/10.1021/acs.jproteome.8b00761}{10.1021/acs.jproteome.8b00761}.
 
 \leavevmode\hypertarget{ref-Schloss2016}{}%
-Schloss, P.D., Jenior, M.L., Koumpouras, C.C., Westcott, S.L., and
-Highlander, S.K. (2016) Sequencing 16S rRNA gene fragments using the
-PacBio SMRT DNA sequencing system. \emph{PeerJ} \textbf{4}: e1869.
+Schloss, P. D., Jenior, M. L., Koumpouras, C. C., Westcott, S. L., and
+Highlander, S. K. (2016). Sequencing 16S rRNA gene fragments using the
+PacBio SMRT DNA sequencing system. \emph{PeerJ} 4, e1869.
+doi:\href{https://doi.org/10.7717/peerj.1869}{10.7717/peerj.1869}.
 
 \leavevmode\hypertarget{ref-Schloss2009}{}%
-Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
-Hollister, E.B., et al. (2009) Introducing mothur: Open-source,
+Schloss, P. D., Westcott, S. L., Ryabin, T., Hall, J. R., Hartmann, M.,
+Hollister, E. B., et al. (2009). Introducing mothur: Open-source,
 platform-independent, community-supported software for describing and
-comparing microbial communities. \emph{Appl Environ Microbiol}
-\textbf{75}: 7537--7541.
+comparing microbial communities. \emph{Appl. Environ. Microbiol.} 75,
+7537--7541.
+doi:\href{https://doi.org/10.1128/AEM.01541-09}{10.1128/AEM.01541-09}.
 
 \leavevmode\hypertarget{ref-Staufenberger2008}{}%
-Staufenberger, T., Thiel, V., Wiese, J., and Imhoff, J.F. (2008)
+Staufenberger, T., Thiel, V., Wiese, J., and Imhoff, J. F. (2008).
 Phylogenetic analysis of bacteria associated with \emph{Laminaria
-saccharina}. \emph{FEMS Microbiol Ecol} \textbf{64}: 65--77.
+saccharina}. \emph{FEMS Microbiol. Ecol.} 64, 65--77.
+doi:\href{https://doi.org/10.1111/j.1574-6941.2008.00445.x}{10.1111/j.1574-6941.2008.00445.x}.
 
 \leavevmode\hypertarget{ref-Su2012}{}%
-Su, C., Lei, L., Duan, Y., Zhang, K.-Q., and Yang, J. (2012)
+Su, C., Lei, L., Duan, Y., Zhang, K.-Q., and Yang, J. (2012).
 Culture-independent methods for studying environmental microorganisms:
-Methods, application, and perspective. \emph{Appl Microbiol Biotechnol}
-\textbf{93}: 993--1003.
+Methods, application, and perspective. \emph{Appl. Microbiol.
+Biotechnol.} 93, 993--1003.
+doi:\href{https://doi.org/10.1007/s00253-011-3800-7}{10.1007/s00253-011-3800-7}.
 
 \leavevmode\hypertarget{ref-Uku2007}{}%
-Uku, J., Björk, M., Bergman, B., and Díez, B. (2007) Characterization
+Uku, J., Björk, M., Bergman, B., and Díez, B. (2007). Characterization
 and comparison of prokaryotic epiphytes associated with three East
-African seagrasses. \emph{J Phycol} \textbf{43}: 768--779.
+African seagrasses. \emph{J. Phycol.} 43, 768--779.
+doi:\href{https://doi.org/10.1111/j.1529-8817.2007.00371.x}{10.1111/j.1529-8817.2007.00371.x}.
 
 \leavevmode\hypertarget{ref-Verlaque2003}{}%
-Verlaque, M., Durand, C., Huisman, J.M., Boudouresque, C.-F., and Le
-Parco, Y. (2003) On the identity and origin of the Mediterranean
-invasive \emph{Caulerpa racemosa} (Caulerpales, Chlorophyta). \emph{Eur
-J Phycol} \textbf{38}: 325--339.
+Verlaque, M., Durand, C., Huisman, J. M., Boudouresque, C.-F., and Le
+Parco, Y. (2003). On the identity and origin of the Mediterranean
+invasive \emph{Caulerpa racemosa} (Caulerpales, Chlorophyta). \emph{Eur.
+J. Phycol.} 38, 325--339.
+doi:\href{https://doi.org/10.1080/09670260310001612592}{10.1080/09670260310001612592}.
 
 \leavevmode\hypertarget{ref-Weidner1996}{}%
-Weidner, S., Arnold, W., and Puhler, A. (1996) Diversity of uncultured
+Weidner, S., Arnold, W., and Puhler, A. (1996). Diversity of uncultured
 microorganisms associated with the seagrass \emph{Halophila stipulacea}
 estimated by restriction fragment length polymorphism analysis of
-PCR-amplified 16S rRNA genes. \emph{Appl Environ Microbiol} \textbf{62}:
+PCR-amplified 16S rRNA genes. \emph{Appl. Environ. Microbiol.} 62,
 766--771.
@@
 \leavevmode\hypertarget{ref-Wickham2019}{}%
-Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François,
-R., et al. (2019) Welcome to the tidyverse. \emph{J Open Source Softw}
-\textbf{4}: 1686.
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
+François, R., et al. (2019). Welcome to the tidyverse. \emph{J. Open
+Source Softw.} 4, 1686.
+doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}.
 
 \leavevmode\hypertarget{ref-Wilke2018}{}%
-Wilke, C.O. (2018) cowplot: Streamlined plot theme and plot annotations
-for 'ggplot2'.
+Wilke, C. O. (2018). \emph{cowplot: Streamlined plot theme and plot
+annotations for 'ggplot2'}.
 
 \leavevmode\hypertarget{ref-Williams2012}{}%
-Williams, T.J., Long, E., Evans, F., DeMaere, M.Z., Lauro, F.M.,
-Raftery, M.J., et al. (2012) A metaproteomic assessment of winter and
+Williams, T. J., Long, E., Evans, F., DeMaere, M. Z., Lauro, F. M.,
+Raftery, M. J., et al. (2012). A metaproteomic assessment of winter and
 summer bacterioplankton from Antarctic Peninsula coastal surface waters.
-\emph{ISME J} \textbf{6}: 1883--1900.
+\emph{ISME J.} 6, 1883--1900.
+doi:\href{https://doi.org/10.1038/ismej.2012.28}{10.1038/ismej.2012.28}.
 
 \leavevmode\hypertarget{ref-Wisniewski2009}{}%
-Wiśniewski, J.R., Zougman, A., Nagaraj, N., and Mann, M. (2009)
-Universal sample preparation method for proteome analysis. \emph{Nat
-Methods} \textbf{6}: 359--362.
+Wiśniewski, J. R., Zougman, A., Nagaraj, N., and Mann, M. (2009).
+Universal sample preparation method for proteome analysis. \emph{Nat.
+Methods} 6, 359--362.
+doi:\href{https://doi.org/10.1038/nmeth.1322}{10.1038/nmeth.1322}.
+
+\leavevmode\hypertarget{ref-Xie2014}{}%
+Xie, Y. (2014). `` knitr: A comprehensive tool for reproducible research
+in R,'' in \emph{Implementing Reproducible Computational Research}, eds.
+V. Stodden, F. Leisch, and R. D. Peng (New York: Chapman and Hall/CRC),
+3--32.
 
 \leavevmode\hypertarget{ref-Xie2015}{}%
-Xie, Y. (2015) Dynamic Documents with R and knitr, 2nd ed. Boca Raton,
-Florida: Chapman and Hall/CRC.
-
-\leavevmode\hypertarget{ref-Xie2014}{}%
-Xie, Y. (2014) knitr: A comprehensive tool for reproducible research in
-R. In \emph{Implementing Reproducible Computational Research}. Stodden,
-V., Leisch, F., and Peng, R.D. (eds). New York: Chapman and Hall/CRC,
-pp. 3--32.
+Xie, Y. (2015). \emph{Dynamic Documents with R and knitr}. 2nd ed. Boca
+Raton, Florida: Chapman and Hall/CRC.
 
 \leavevmode\hypertarget{ref-Xie2019a}{}%
-Xie, Y. (2019) TinyTeX: A lightweight, cross-platform, and
+Xie, Y. (2019). TinyTeX: A lightweight, cross-platform, and
 easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
-\textbf{40}: 30--32.
+40, 30--32.
 
 \leavevmode\hypertarget{ref-Xie2020}{}%
-Xie, Y. (2020) tinytex: Helper functions to install and maintain 'TeX
-Live', and compile 'LaTeX' documents.
+Xie, Y. (2020). \emph{tinytex: Helper functions to install and maintain
+'TeX Live', and compile 'LaTeX' documents}.
 
 \leavevmode\hypertarget{ref-Xie2018}{}%
-Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
-Definitive Guide, 1st ed. Boca Raton, Florida: Chapman and Hall/CRC.
+Xie, Y., Allaire, J. J., and Grolemund, G. (2018). \emph{R Markdown: The
+Definitive Guide}. 1st ed. Boca Raton, Florida: Chapman and Hall/CRC.
 
 \leavevmode\hypertarget{ref-Yilmaz2014}{}%
-Yilmaz, P., Parfrey, L.W., Yarza, P., Gerken, J., Pruesse, E., Quast,
-C., et al. (2014) The SILVA and "All-Species Living Tree Project (LTP)"
-taxonomic frameworks. \emph{Nucleic Acids Res} \textbf{42}: D643--D648.
+Yilmaz, P., Parfrey, L. W., Yarza, P., Gerken, J., Pruesse, E., Quast,
+C., et al. (2014). The SILVA and "All-Species Living Tree Project (LTP)"
+taxonomic frameworks. \emph{Nucleic Acids Res.} 42, D643--D648.
+doi:\href{https://doi.org/10.1093/nar/gkt1209}{10.1093/nar/gkt1209}.
 
 \leavevmode\hypertarget{ref-Zhu2019}{}%
-Zhu, H. (2019) kableExtra: Construct complex table with 'kable' and pipe
-syntax.
+Zhu, H. (2019). \emph{kableExtra: Construct complex table with 'kable'
+and pipe syntax}.
 
@@
 
-\hypertarget{figure-captions}{%
-\subsection{Figure captions}\label{figure-captions}}
+\hypertarget{figure-legends}{%
+\subsection{Figure legends}\label{figure-legends}}
 

### 2021-03-25 submission/manuscript.tex
Format text.
Review done.

@@
 
-\hypertarget{summary}{%
-\subsection{Summary}\label{summary}}
+\hypertarget{abstract}{%
+\subsection{Abstract}\label{abstract}}
 
@@
 and transported to the laboratory in containers placed on ice and filled
-with seawater from this site. Upon arrival to the laboratory, \emph{C.
+with seawater from the site. Upon arrival to the laboratory, \emph{C.
 nodosa} leaves were cut into sections of 1 -- 2 \si{\cm}, while \emph{C.
@@
 (\(5'\)-GGACTACNVGGGTWTCTAAT-\(3'\)) primers from the Earth Microbiome
-Project
-(\url{http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/})
+Project (\url{https://earthmicrobiome.org/protocols-and-standards/16s/})
 were used to amplify the target region (Caporaso et al., 2012; Apprill
@@
 25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer,
-0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum
+0.2 \si{\milli\Molar} dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum
 Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of
@@
 (Wiśniewski et al., 2009). Prior to loading the solution onto the
-column, protein pellets were solubilized in a urea sample buffer
-included in the kit amended with DTT (final concentration 100
-\si{\milli\Molar}) for 45 \si{\minute} at room temperature and
-centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to
-remove larger particles. The first washing step after protein solution
-loading was repeated twice. In addition, the centrifugation steps were
-prolonged if the column was clogged. Trypsin digestion was performed on
-column filters at 37 \si{\degreeCelsius} overnight for 18 \si{\hour}.
-The final filtrate containing peptides was acidified with 1
-\si{\percent} (final concentration) trifluoroacetic acid (TFA), freezed
-at \num{-80} \si{\degreeCelsius}, lyophilized and sent to VIME -- Vienna
-Metabolomics Center (University of Vienna) for metaproteomic analysis.
-Peptides were re-suspended in 1 \si{\percent} (final concentration) TFA,
-desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA)
-according to the manufacturer's instructions and sequenced on a Q
-Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher
-Scientific, USA). Obtained MS/MS spectra were searched against a protein
-database composed of combined sequenced metagenomes using SEQUEST-HT
-engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo
-Fisher Scientific, USA). The target-decoy approach was used to reduce
-the probability of false peptide identification. Results whose false
-discovery rate at the peptide level was \textless{} \SI{1}{\percent}
-were kept. For protein identification a minimum of two peptides and one
-unique peptide were required. For protein quantification, a
-chromatographic peak area-based free quantitative method was applied.
+column, protein pellets were solubilized in urea sample buffer included
+in the kit amended with DTT (final concentration 100 \si{\milli\Molar})
+for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g
+for 2 -- 5 \si{\minute} at room temperature to remove larger particles.
+The first washing step after protein solution loading was repeated
+twice. In addition, the centrifugation steps were prolonged if the
+column was clogged. Trypsin digestion was performed on column filters at
+37 \si{\degreeCelsius} overnight for 18 \si{\hour}. The final filtrate
+containing peptides was acidified with 1 \si{\percent} (final
+concentration) trifluoroacetic acid (TFA), freezed at \num{-80}
+\si{\degreeCelsius}, lyophilized and sent to VIME -- Vienna Metabolomics
+Center (University of Vienna) for metaproteomic analysis. Peptides were
+re-suspended in 1 \si{\percent} (final concentration) TFA, desalted
+using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to
+the manufacturer's instructions and sequenced on a Q Exactive Hybrid
+Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
+Obtained MS/MS spectra were searched against a protein database composed
+of combined sequenced metagenomes obtained in this study using
+SEQUEST-HT engines and validated with Percolator in Proteome Discoverer
+2.1 (Thermo Fisher Scientific, USA). The target-decoy approach was used
+to reduce the probability of false peptide identification. Results whose
+false discovery rate at the peptide level was \textless{}
+\SI{1}{\percent} were kept. For protein identification a minimum of two
+peptides and one unique peptide were required. For protein
+quantification, a chromatographic peak area-based free quantitative
+method was applied.
 
@@
 tidyverse (version 1.3.0) (Wickham et al., 2019) and multiple other
-packages (Neuwirth, 2014; Xie, 2014, 2015, 2019, 2020; Wilke, 2018; Xie
-et al., 2018; Allaire et al., 2019; Zhu, 2019; Bengtsson, 2020). The
+packages (Neuwirth, 2014; Xie, 2014, 2015, 2019a, 2019b; Wilke, 2018;
+Xie et al., 2018; Allaire et al., 2019; Zhu, 2019; Bengtsson, 2020). The
 detailed analysis procedure including the R Markdown file for this paper
@@
 region of the 16S rRNA was performed. Sequencing yielded a total of
-336,944 sequences after quality curation and exclusion of eukaryotic,
+292,888 sequences after quality curation and exclusion of eukaryotic,
 mitochondrial and no relative sequences. The number of sequences
-classified as chloroplasts was 97,334. After excluding these sequences,
-the total number of retrieved reads was 239,610, ranging from 22,596 to
-52,930 sequences per sample (\autoref{supp-nseq_notus}). Even when the
+classified as chloroplasts was 97,331. After excluding these sequences,
+the total number of retrieved reads was 195,557, ranging from 13,667 to
+41,842 sequences per sample (\autoref{supp-nseq_notus}). Even when the
 highest sequencing effort was applied, the rarefaction curves did not
@@
 different OTUs. Taxonomic classification of reads revealed a
-macrophyte-associated epiphytic community that mainly composed of
+macrophyte-associated epiphytic community mainly composed of
 \emph{Alphaproteobacteria} (14.9 ± 3.5 \si{\percent}),
@@
 pelagic microbial communities to observe high proportions of chloroplast
-sequences (Gilbert et al., 2009; Korlević et al., 2016). In addition, a
-very low proportion of chloroplast 16S rRNA sequences in samples
-originating from \emph{C. cylindracea} in comparison to \emph{C. nodosa}
-could be explained by the presence of three introns in the gene for 16S
-rRNA in some members of the genus \emph{Caulerpa} that could hamper the
-amplification process (Lam and Lopez-Bautista, 2016).
+sequences (Gilbert et al., 2009; Korlević et al., 2016). On of the
+solutions to further reduce chloroplast 16S rRNA sequence contamination
+is to use primers that minimise the amplification of these reads if the
+sequencing and study design allow it (Hanshew et al., 2013). In
+addition, a very low proportion of chloroplast 16S rRNA sequences in
+samples originating from \emph{C. cylindracea} in comparison to \emph{C.
+nodosa} could be explained by the presence of three introns in the gene
+for 16S rRNA in some members of the genus \emph{Caulerpa} that could
+hamper the amplification process (Lam and Lopez-Bautista, 2016).
 
@@
 et al. (2014). Ribosomal Database Project: Data and tools for high
-throughput rRNA analysis. \emph{Nucleic Acids Res} 42, D633--D642.
+throughput rRNA analysis. \emph{Nucleic Acids Res.} 42, D633--D642.
 doi:\href{https://doi.org/10.1093/nar/gkt1244}{10.1093/nar/gkt1244}.
@@
 
+\leavevmode\hypertarget{ref-Hanshew2013}{}%
+Hanshew, A. S., Mason, C. J., Raffa, K. F., and Currie, C. R. (2013).
+Minimization of chloroplast contamination in 16S rRNA gene
+pyrosequencing of insect herbivore bacterial communities. \emph{J.
+Microbiol. Methods} 95, 149--155.
+doi:\href{https://doi.org/10.1016/j.mimet.2013.08.007}{10.1016/j.mimet.2013.08.007}.
+
 \leavevmode\hypertarget{ref-Huerta-Cepas2017}{}%
@@
 functionally and phylogenetically annotated orthology resource based on
-5090 organisms and 2502 viruses. \emph{Nucleic Acids Res} 47,
+5090 organisms and 2502 viruses. \emph{Nucleic Acids Res.} 47,
 D309--D314.
@@
 and Hauser, L. J. (2010). Prodigal: Prokaryotic gene recognition and
-translation initiation site identification. \emph{BMC Bioinformatics}
-11, 119.
+translation initiation site identification. \emph{BMC Bioinf.} 11, 119.
 doi:\href{https://doi.org/10.1186/1471-2105-11-119}{10.1186/1471-2105-11-119}.
@@
 associated with \emph{Thalassia hemprichii} in Xincun Bay, South China
-Sea. \emph{Ecotoxicology} 24, 1548--56.
+Sea. \emph{Ecotoxicology} 24, 1548--1556.
 doi:\href{https://doi.org/10.1007/s10646-015-1511-z}{10.1007/s10646-015-1511-z}.
@@
 \leavevmode\hypertarget{ref-Leary2014}{}%
-Leary, D. H., Li, R. W., Hamdan, L. J., IV, W. J. H., Lebedev, N., Wang,
-Z., et al. (2014). Integrated metagenomic and metaproteomic analyses of
-marine biofilm communities. \emph{Biofouling} 30, 1211--1223.
+Leary, D. H., Li, R. W., Hamdan, L. J., Judson Hervey IV, W., Lebedev,
+N., Wang, Z., et al. (2014). Integrated metagenomic and metaproteomic
+analyses of marine biofilm communities. \emph{Biofouling} 30,
+1211--1223.
 doi:\href{https://doi.org/10.1080/08927014.2014.977267}{10.1080/08927014.2014.977267}.
@@
 The ecology of microbial communities associated with \emph{Macrocystis
-pyrifera}. \emph{PloS One} 8, e67480.
+pyrifera}. \emph{PLoS One} 8, e67480.
 doi:\href{https://doi.org/10.1371/journal.pone.0067480}{10.1371/journal.pone.0067480}.
@@
 \leavevmode\hypertarget{ref-Xie2014}{}%
-Xie, Y. (2014). `` knitr: A comprehensive tool for reproducible research
+Xie, Y. (2014). ``knitr: A comprehensive tool for reproducible research
 in R,'' in \emph{Implementing Reproducible Computational Research}, eds.
@@
 \leavevmode\hypertarget{ref-Xie2019a}{}%
-Xie, Y. (2019). TinyTeX: A lightweight, cross-platform, and
+Xie, Y. (2019a). TinyTeX: A lightweight, cross-platform, and
 easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
@@
 
-\leavevmode\hypertarget{ref-Xie2020}{}%
-Xie, Y. (2020). \emph{tinytex: Helper functions to install and maintain
+\leavevmode\hypertarget{ref-Xie2019b}{}%
+Xie, Y. (2019b). \emph{tinytex: Helper functions to install and maintain
 'TeX Live', and compile 'LaTeX' documents}.

### 2021-03-25 submission/manuscript.Rmd

@@
 
-## Summary
+## Abstract
 Studies of unculturable microbes often combine methods such as 16S rRNA sequencing, metagenomics and metaproteomics.
@@
 In addition, as not enough DNA for both metagenomic and 16S RNA analysis were obtained during the sampling on 4 December 2017, an additional sampling on 14 December 2017 was carried out in the Bay of Funtana.
-Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with seawater from this site.
+Leaves and thalli were collected by diving and transported to the laboratory in containers placed on ice and filled with seawater from the site.
 Upon arrival to the laboratory, *C.
@@
 The V4 region of the 16S rRNA gene was amplified using a two-step PCR procedure.
-In the first PCR the 515F ($5'$-GTGYCAGCMGCCGCGGTAA-$3'$) and 806R ($5'$-GGACTACNVGGGTWTCTAAT-$3'$) primers from the Earth Microbiome Project (http://press.igsb.anl.gov/earthmicrobiome/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016].
+In the first PCR the 515F ($5'$-GTGYCAGCMGCCGCGGTAA-$3'$) and 806R ($5'$-GGACTACNVGGGTWTCTAAT-$3'$) primers from the Earth Microbiome Project (https://earthmicrobiome.org/protocols-and-standards/16s/) were used to amplify the target region [@Caporaso2012; @Apprill2015; @Parada2016].
 These primers contained on their $5'$ end a tagged sequence.
-Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} of dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template.
+Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template.
 Cycling conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}.
@@
 Isolated proteins were whole trypsin digested using the FASP (filter-aided sample preparation) Protein Digestion Kit (Expedeon, UK) according to the manufacturer's instructions with small modifications [@Wisniewski2009].
-Prior to loading the solution onto the column, protein pellets were solubilized in a urea sample buffer included in the kit amended with DTT (final concentration 100 \si{\milli\Molar}) for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to remove larger particles.
+Prior to loading the solution onto the column, protein pellets were solubilized in urea sample buffer included in the kit amended with DTT (final concentration 100 \si{\milli\Molar}) for 45 \si{\minute} at room temperature and centrifuged at 20,000 × g for 2 -- 5 \si{\minute} at room temperature to remove larger particles.
 The first washing step after protein solution loading was repeated twice.
@@
 Peptides were re-suspended in 1 \si{\percent} (final concentration) TFA, desalted using the Pierce C18 Tips (Thermo Fisher Scientific, USA) according to the manufacturer's instructions and sequenced on a Q Exactive Hybrid Quadrupole-Orbitrap Mass Spectrometer (Thermo Fisher Scientific, USA).
-Obtained MS/MS spectra were searched against a protein database composed of combined sequenced metagenomes using SEQUEST-HT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
+Obtained MS/MS spectra were searched against a protein database composed of combined sequenced metagenomes obtained in this study using SEQUEST-HT engines and validated with Percolator in Proteome Discoverer 2.1 (Thermo Fisher Scientific, USA).
 The target-decoy approach was used to reduce the probability of false peptide identification.
@@
 ### Data processing and visualization
-Processing and visualization  of 16S rRNA, metagenomic and metaproteomic data were done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2020; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020; @Wilke2018].
+Processing and visualization  of 16S rRNA, metagenomic and metaproteomic data were done using R (version 3.6.0) [@RCoreTeam2019],  package tidyverse (version 1.3.0) [@Wickham2019] and multiple other packages [@Xie2014; @Xie2015; @Xie2019a; @Xie2019b; @Allaire2019; @Xie2018; @Zhu2019; @Neuwirth2014; @Bengtsson2020; @Wilke2018].
 The detailed analysis procedure including the R Markdown file for this paper are available as a GitHub repository (https://github.com/MicrobesRovinj/Korlevic_SelectiveRemoval_FrontMicrobiol_2021).
@@
 number_sequences <- filter(community, taxon=="Archaea" | taxon=="Bacteria") %>%
-  select(total) %>%
+  select(`40` : `63`) %>%
   sum()
@@
 number_chloroplast_sequences <- filter(community, taxon=="Chloroplast") %>%
-  select(total) %>%
+  select(`40` : `63`) %>%
   sum()
@@
 # Calculating the minimal and maximal number of sequences
-min_max <- filter(community, taxon=="Archaea" | taxon=="Bacteria") %>%
+min_max <- filter(community, taxon=="Archaea" | taxon=="Bacteria" | taxon=="Chloroplast") %>%
+  mutate_at(6:ncol(.), list(~case_when(taxon=="Bacteria" ~ . - .[taxon=="Chloroplast"], TRUE ~ .))) %>%
+  filter(taxon!="Chloroplast") %>%
   summarise_at(6:ncol(.), sum) %>%
@@
 Sequences clustering at a similarity level of 97 \si{\percent} yielded a total of `r no_otus` different OTUs.
-Taxonomic classification of reads revealed a macrophyte-associated epiphytic community that mainly composed of *Alphaproteobacteria* (`r filter(plot, taxon=="Alphaproteobacteria")$mean` ± `r filter(plot, taxon=="Alphaproteobacteria")$sd` \si{\percent}), *Bacteroidota* (`r filter(plot, taxon=="Bacteroidota")$mean` ± `r filter(plot, taxon=="Bacteroidota")$sd` \si{\percent}), *Gammaproteobacteria* (`r filter(plot, taxon=="Gammaproteobacteria")$mean` ± `r filter(plot, taxon=="Gammaproteobacteria")$sd` \si{\percent}), *Desulfobacterota* (`r filter(plot, taxon=="Desulfobacterota")$mean` ± `r filter(plot, taxon=="Desulfobacterota")$sd` \si{\percent}), *Cyanobacteria* (`r filter(plot, taxon=="Cyanobacteria")$mean` ± `r filter(plot, taxon=="Cyanobacteria")$sd` \si{\percent}) and *Planctomycetota* (`r filter(plot, taxon=="Planctomycetota")$mean` ± `r filter(plot, taxon=="Planctomycetota")$sd` \si{\percent}) (\autoref{community}).
+Taxonomic classification of reads revealed a macrophyte-associated epiphytic community mainly composed of *Alphaproteobacteria* (`r filter(plot, taxon=="Alphaproteobacteria")$mean` ± `r filter(plot, taxon=="Alphaproteobacteria")$sd` \si{\percent}), *Bacteroidota* (`r filter(plot, taxon=="Bacteroidota")$mean` ± `r filter(plot, taxon=="Bacteroidota")$sd` \si{\percent}), *Gammaproteobacteria* (`r filter(plot, taxon=="Gammaproteobacteria")$mean` ± `r filter(plot, taxon=="Gammaproteobacteria")$sd` \si{\percent}), *Desulfobacterota* (`r filter(plot, taxon=="Desulfobacterota")$mean` ± `r filter(plot, taxon=="Desulfobacterota")$sd` \si{\percent}), *Cyanobacteria* (`r filter(plot, taxon=="Cyanobacteria")$mean` ± `r filter(plot, taxon=="Cyanobacteria")$sd` \si{\percent}) and *Planctomycetota* (`r filter(plot, taxon=="Planctomycetota")$mean` ± `r filter(plot, taxon=="Planctomycetota")$sd` \si{\percent}) (\autoref{community}).
 
@@
 Indeed, it is common during 16S rRNA profiling of pelagic microbial communities to observe high proportions of chloroplast sequences [@Korlevic2016; @Gilbert2009].
+On of the solutions to further reduce chloroplast 16S rRNA sequence contamination is to use primers that minimise the amplification of these reads if the sequencing and study design allow it [@Hanshew2013].
 In addition, a very low proportion of chloroplast 16S rRNA sequences in samples originating from *C.

### 2021-03-25 submission/manuscript.Rmd
Review done.

@@
 These primers contained on their $5'$ end a tagged sequence.
-Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template.
+Each sample was amplified in four parallel 25 \si{\ul} reactions of which each contained 1 × Q5 Reaction Buffer, 0.2 \si{\milli\Molar} dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum Albumin), 0.2 \si{\micro\Molar} forward and reverse primers, 0.5 U of Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng} of DNA template.
 Cycling conditions were: initial denaturation at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50 \si{\degreeCelsius} for 60 \si{\s} and elongation at 72 \si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at 72 \si{\degreeCelsius} for 10 \si{\minute}.

### 2021-03-25 submission/manuscript.tex

@@
 0.2 \si{\milli\Molar} dNTPmix, 0.7 \si{\mg\per\ml} BSA (Bovine Serum
-Albumin), 0.2 \si{\micro\Molar} of forward and reverse primers, 0.5 U of
-Q5 High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5
-\si{\ng} of DNA template. Cycling conditions were: initial denaturation
-at 94 \si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation
-at 94 \si{\degreeCelsius} for 45 \si{\s}, annealing at 50
-\si{\degreeCelsius} for 60 \si{\s} and elongation at 72
-\si{\degreeCelsius} for 90 \si{\s}, finalized by an elongation step at
-72 \si{\degreeCelsius} for 10 \si{\minute}. The four parallel reactions
-volumes were pooled and PCR products were purified using the GeneJET PCR
-Purification Kit (Thermo Fisher Scientific, USA) according to the
-manufacturer's instructions and following the protocol that included
-isopropanol addition for better small DNA fragment yield. The column was
-eluted in 30 \si{\ul} of deionized water. Purified PCR products were
-sent for Illumina MiSeq sequencing (2 × 250 bp) at IMGM Laboratories,
-Martinsried, Germany. Before sequencing at IMGM, the second PCR
-amplification of the two-step PCR procedure was performed using primers
-targeting the tagged region incorporated in the first PCR. In addition,
-these primers contained adapter and sample-specific index sequences. The
-second PCR was carried out for 8 cycles. Beside samples, a positive and
-negative control were sequenced. A negative control was comprised of
-four parallel PCR reactions without DNA template, while for a positive
-control a mock community composed of evenly mixed DNA material
-originating from 20 bacterial strains (ATCC MSA-1002, ATCC, USA) was
-used. Partial 16S rRNA sequences obtained in this study have been
-deposited in the European Nucleotide Archive (ENA) at EMBL-EBI under
-accession numbers SAMEA6786270, SAMEA6648792 -- SAMEA6648794,
-SAMEA6648809 -- SAMEA6648811 and SAMEA6648824.
+Albumin), 0.2 \si{\micro\Molar} forward and reverse primers, 0.5 U of Q5
+High-Fidelity DNA Polymerase (New England Biolabs, USA) and 5 \si{\ng}
+of DNA template. Cycling conditions were: initial denaturation at 94
+\si{\degreeCelsius} for 3 \si{\minute}, 20 cycles of denaturation at 94
+\si{\degreeCelsius} for 45 \si{\s}, annealing at 50 \si{\degreeCelsius}
+for 60 \si{\s} and elongation at 72 \si{\degreeCelsius} for 90 \si{\s},
+finalized by an elongation step at 72 \si{\degreeCelsius} for 10
+\si{\minute}. The four parallel reactions volumes were pooled and PCR
+products were purified using the GeneJET PCR Purification Kit (Thermo
+Fisher Scientific, USA) according to the manufacturer's instructions and
+following the protocol that included isopropanol addition for better
+small DNA fragment yield. The column was eluted in 30 \si{\ul} of
+deionized water. Purified PCR products were sent for Illumina MiSeq
+sequencing (2 × 250 bp) at IMGM Laboratories, Martinsried, Germany.
+Before sequencing at IMGM, the second PCR amplification of the two-step
+PCR procedure was performed using primers targeting the tagged region
+incorporated in the first PCR. In addition, these primers contained
+adapter and sample-specific index sequences. The second PCR was carried
+out for 8 cycles. Beside samples, a positive and negative control were
+sequenced. A negative control was comprised of four parallel PCR
+reactions without DNA template, while for a positive control a mock
+community composed of evenly mixed DNA material originating from 20
+bacterial strains (ATCC MSA-1002, ATCC, USA) was used. Partial 16S rRNA
+sequences obtained in this study have been deposited in the European
+Nucleotide Archive (ENA) at EMBL-EBI under accession numbers
+SAMEA6786270, SAMEA6648792 -- SAMEA6648794, SAMEA6648809 -- SAMEA6648811
+and SAMEA6648824.
 

### 2021-03-25 submission/manuscript.Rmd
Review done.

@@
 Indeed, it is common during 16S rRNA profiling of pelagic microbial communities to observe high proportions of chloroplast sequences [@Korlevic2016; @Gilbert2009].
-On of the solutions to further reduce chloroplast 16S rRNA sequence contamination is to use primers that minimise the amplification of these reads if the sequencing and study design allow it [@Hanshew2013].
+One of the solutions to further reduce chloroplast 16S rRNA sequence contamination is to use primers that minimise the amplification of these reads if the sequencing and study design allow it [@Hanshew2013].
 In addition, a very low proportion of chloroplast 16S rRNA sequences in samples originating from *C.

### 2021-03-25 submission/manuscript.tex

@@
 pelagic microbial communities to observe high proportions of chloroplast
-sequences (Gilbert et al., 2009; Korlević et al., 2016). On of the
+sequences (Gilbert et al., 2009; Korlević et al., 2016). One of the
 solutions to further reduce chloroplast 16S rRNA sequence contamination

### 2021-05-10 submission/manuscript.Rmd
Manuscript production.

@@
 
-3\. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research, Utrecht University, The Netherlands
+3\. Department of Marine Microbiology and Biogeochemistry, Royal Netherlands Institute for Sea Research (NIOZ), Utrecht University, The Netherlands
 
@@
 
-Running title: DNA and protein isolation from macrophyte surfaces
+Running title: Selective isolation from macrophyte surfaces
 

### 2021-05-10 submission/manuscript.tex

@@
 
-3. NIOZ, Department of Marine Microbiology and Biogeochemistry, Royal
-Netherlands Institute for Sea Research, Utrecht University, The
+3. Department of Marine Microbiology and Biogeochemistry, Royal
+Netherlands Institute for Sea Research (NIOZ), Utrecht University, The
 Netherlands

### 2021-05-10 submission/supplementary.Rmd

@@
 ---
-title: "**Supplementary Information**"
+title: "**Supplementary Material**"
 subtitle: "**Selective DNA and Protein Isolation from Marine Macrophyte Surfaces**"

### 2021-05-10 submission/supplementary.tex

@@
 \hypersetup{unicode=true,
-            pdftitle={Supplementary Information},
+            pdftitle={Supplementary Material},
             pdfborder={0 0 0},
@@
 
-  \title{\textbf{Supplementary Information}}
+  \title{\textbf{Supplementary Material}}
     \pretitle{\vspace{\droptitle}\centering\huge}
